### 2018-10-02 tex/main.tex
Added tex doucment from overleav

\documentclass[journal=esthag,manuscript=article]{achemso}

% additional packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{array,multirow,graphicx}
\graphicspath{{./../figures/}}
\usepackage{lineno}
\linenumbers
% macros

% authors
\author{Jonathan G. V. Ström}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
\author{Yijun Yao}
\affiliation[Zhejiang University]{Zhejiang University, Hangzhou, China}
\author{Eric M. Suuberg}
\email{eric_suuberg@brown.edu}
\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}

% title
\title{Transient Variability In Vapor Intrusion And The Factors That Influence It}

% keywords
\abbreviations{VI}
\keywords{VI, Preferential pathways, Temporal transient variability}

\begin{document}

\begin{abstract}
Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
We consider the source of the variability at VI sites located near Hill Air Force Base and Naval Air Station North Island using statistical analysis methods and three-dimensional subsurface computational fluid dynamics modeling.
The results suggest that an order of magnitude variation in indoor air contaminant concentrations may be expected at ”normal” VI sites where preferential pathways do not play a role, whereas three or more orders of magnitude can be observed at sites characterized by preferential pathways.
A CFD modeling sensitivity analysis reveals that it is not only the presence of contaminant vapor in a preferential pathway that is required to see large observed variations, but there must also exist a permeable region beneath the structure.
Large temporal fluctuations in indoor air contaminant concentrations may be observed where no preferential pathways exist, but this requires a particular combination of large fluctuations in pressure driving force and high soil permeability.
\end{abstract}

\section{Introduction}

Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017}.
VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures. VI has been a recognized problem for some time, but many aspects remain poorly understood, particularly with respect to the causes of large temporal transients in indoor air concentrations.
There is uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}. \par

Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
All have shown that these concentrations varied significantly with time - orders of magnitude on the timescale of a day or days.
\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}. \par

In one instance the source of the variation was clearly established during the study; at the ASU house a field drain pipe (or “land drain”), which connected to a sewer system, was discovered beneath the house, and careful isolation of this source led to a clear conclusion that this preferential pathway significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway to an area beneath a structure would always be expected to lead to large fluctuations in indoor air contaminant concentrations. \par

Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house\cite{mchugh_evidence_2017}.
That site was also characterized by large indoor air contaminant concentration fluctuations.
Sewer lines have been generally implicated as VI source at several sites to date\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
A Danish study estimates that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
In some of these cases\cite{pennell_sewer_2013,riis_vapor_2010}, the sewer provided a pathway for direct entry of contaminant into the living space.
While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant to the soil beneath a structure. \par

It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
One example is a site at NAS North Island at which no preferential pathways have been identified.
Instead, a building at this site is characterized by significant variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
It is believed that this is the origin of the fluctuations at that site. \par

This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of in soil preferential pathways.
In this work, the latter scenarios are referred to as ”normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at  such normal  VI  sites vs. those characterized by preferential pathways.
The conditions required for preferential pathways to be significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.

\section{Methods}

\subsection{Statistical Analysis Of Field Data}

To begin to characterize transient behavior in indoor air contaminant concentrations, actual datasets are analyzed to establish common levels of variability at VI sites.
For this purpose, the datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
This paper relies on statistical analysis of published field data, and readers are referred to the original works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}. \par

The ASU house and data were obtained over a period of a few years.
During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” operation.
This caused greater than normal advective flow from the subsurface into the house, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
This period of CPM testing is considered separately from the otherwise ”natural” VI conditions in the analysis.
Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining the dataset, noting that during some of the testing, this pathway was deliberately cut off, resulting in what we have termed “normal” VI conditions in which the main source of contaminant was believed to be groundwater. \par

The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to large internal pressure fluctuations, much more extensive than those typically recorded at the ASU house during normal operations.
Additionally, the underlying soil at NAS North Island is sandy and more permeable than that at the ASU site, which, as will be shown, contributes to the indoor air contaminant concentrations being more sensitive to pressure fluctuations\cite{hosangadi_high-frequency_2017}. \par

Likewise, the Indianapolis site investigation spanned over a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
The goal of which is to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above, and therefore only the period before the installation of this system was considered in the analysis.
It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this PP removed, making it difficult to assess how significant the PP was at this site, regardless it is of interest to analyze this site due to its wealth of data. \par

The typical variation in indoor air contaminant concentrations with time will first be considered below in the case of the ASU house during ”natural”, (i.e. non-CPM conditions), in the case of the NAS North Island site over the entire available dataset, and for the Indianapolis case we consider the variations before the installation of the SSD system.
The deviations from the mean TCE (Chloroform and PCE at the Indianapolis site) in indoor air concentration, as well as the indoor-outdoor pressure differential associated with these concentration fluctuations, were examined, and both univariate and bivariate kernel density estimations (KDE) were constructed.
KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (that there is statistical fluctuation associated with each sampling event). In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
The distributions of the individual parameters and the relationship between them will be examined using the KDE method.

\subsection{Modeling Work}
In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
This model was implemented in a finite element solver, COMSOL Multiphysics.
In the present work, there has been an addition of a preferential pathway to the ”standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
In this modeling work, only the vadose zone soil domain is directly modeled.

The modeled structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab that spans the perimeter of the house.
All of the contaminant leaving the indoor air space is assumed does so via air exchange with the ambient.
The indoor control volume is assumed to consist of only of the basement, assumed as having a total volume of $300 \; \mathrm{m^3}$.
Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of certain other features can be.
The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy clay (except as noted).
Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, which is allows the contaminant concentration in the pipe to be parameterized.
The groundwater beneath the structure is assumed to be homogeneously contaminated with trichloroethylene (TCE) as a prototypical contaminant.
The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
The ground surface and the pipe are both assumed to be sources of air to the soil domain.
Both are assumed to be at reference atmospheric pressure.
Vapor transport in the soil is governed by Richard’s equation, a modified version  of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
The transport of vapor contaminant in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
The equations and the boundary conditions are given in Table \ref{tbl:eqns-bc-parameters}.

\begin{table}[htb!]
  \centering
  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}
  \label{tbl:eqns-bc-parameters}
  \bigskip
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Governing equations
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subcaption{Governing equations}
  \begin{tabular}{l l}
    \toprule
    % Indoor air space equation
    Unsteady-CSTR                 & $V\frac{d u}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - u A_e V$ \\
    % Richard's equation
    Richard's equation            & $\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$ \\
    % Millington-Quirk equation
    Millington-Quirk              & $D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$ \\
    % Advection-diffusion equation
    Advection-diffusion equation  & $\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$ \\
    % van Genuchten's equations
    \multirow{3}{*}{van Genuchten equations}     & $\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$ \\
                                  & $\theta_g = \theta_t - \theta_w$ \\
                                  & $k_r = (1 - \mathrm{Se})^{l} [1 - (\mathrm{Se}^{-m})^m]^2$ \\
                                  & $m = 1 - 1/n$ \\
    \bottomrule
  \end{tabular}
  \bigskip
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Boundary conditions
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subcaption{Boundary conditions}
  \begin{tabular}{l l l}
    \toprule
    \textbf{Boundary}          & \textbf{Richard's equation}      &   \textbf{Advection-diffusion equation} \\
    % Foundation crack
    At foundation crack  & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$                            & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
    % Groundwater source
    At groundwater source &  N/A & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$ \\
    % Ground surface
    At ground surface      & $p = 0 \; \mathrm{(Pa)}$  & $c = 0 \; \mathrm{(\mu g/m^3)}$ \\
    % Preferential pathway
    Exit of preferential pathway  & $p = 0 \; \mathrm{(Pa)}$  & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$ \\
    \bottomrule
  \end{tabular}
  \bigskip
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Soil input parameters
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subcaption{Soil \& gravel properties\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
  \begin{tabular}{l l l l l l l}
    \toprule
    % Descriptions
    Soil & $\text{Permeability} \; \mathrm{(m^2)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$  & $\theta_s$  & $\theta_r$  & $\alpha \; \mathrm{(1/m)}$  & $n$ \\
    % Gravel
    Gravel     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
    % Sand
    Sand     & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053        & 3.5       & 3.2 \\
    % Sandy clay
    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
    \bottomrule
  \end{tabular}
  \bigskip
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % TCE input parameters
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subcaption{Trichloroethylene (diluted in air) properties\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
  \begin{tabular}{l l l l l l}
    \toprule
    $D_\mathrm{air} \; \mathrm{(m^2/h)}$  & $D_\mathrm{water} \; \mathrm{(m^2/h)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$ & $\mathrm{Viscosity} \; \mathrm{(Pa \cdot s)}$  & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
    $2.47 \cdot 10^{-2}$  & $3.67 \cdot 10^{-6}$  & 1.614 & $1.86 \cdot 10^{-5}$  & 0.403 & 131.39 \\
    \bottomrule
  \end{tabular}
  \bigskip
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % Building input parameters
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \subcaption{Building properties}
  \begin{tabular}{l l l}
    \toprule
    $V_\mathrm{base} \; \mathrm{(m^3)}$  & $L_\mathrm{slab} \; \mathrm{(cm)}$  & $A_e \; \mathrm{(1/hr)}$ \\
    %
    300  &  15  & 0.5 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Drivers For Indoor Air Contaminant Variability}

\subsubsection{The Role Of Building Depressurization/Pressurization}

% 2D KDE plot of deviation from mean TCE in indoor air and pressure difference from the ASU house and North Island NAS
\begin{figure}[htb!]
  \caption{Probability density plots of the deviation from the mean concentration along with the indoor-outdoor pressure difference at the ASU house both before and after the preferential pathway was closed and at North Island NAS.}
  \label{fig:2d_kde}
  \includegraphics[width=\textwidth]{pressure_vs_indoor_concentration.pdf}
\end{figure}

Figure \ref{fig:2d_kde} show the result of the statistical analysis of the field data from the ASU and NAS North Island sites.  On the axes, the univariate KDE’s are shown.
The y-margin shows the distribution of the log-10 deviation from the mean TCE concentration in indoor air and the x-margin shows the distribution of indoor-outdoor pressure differences.
A negative pressure difference indicates that the structure is underpressurized, and thus air flows into the building, and a positive pressure difference indicates the opposite.
Thus the curves on the axis margins are the distribution functions of the respective measured variable values, considered independently.
They show the distributions of the results from the many different sampling events.
The distribution functions of pressure fluctuations are relatively simpler than those for the indoor air concentration measurements, and the pressure distribution functions for the pressure difference measurements at the ASU house are much simpler and narrower than that for comparable measurements at NAS North Island (see the distribution functions on the top margin of the figure).
Comparing the narrow red and green distribution functions for the ASU house results shows that whether the valve to the preferential pathway (PP) is open or not, roughly the same indoor-outdoor pressure differentials were seen.
This is not at all surprising, given that the existence of a preferential pathway buried beneath the foundation slab would not be expected to influence the indoor-outdoor pressure differential.
But when comparing the ASU house pressure differentials with those reported for the NAS North Island site (the blue distribution function), it is immediately apparent how much larger the pressure swings are at the latter site.
On this basis, if one associates contaminant entry rate with pressure differential alone, the NAS North Island site would be expected to show a much larger fluctuation in indoor air contaminant concentration than would the ASU house (but of course the preferential pathway changes the situation, as described below).

Turning to the univariate distributions for indoor air contaminant concentrations, the results show that even with the preferential pathway closed off, the results from the ASU house clearly show variations of an order of magnitude.
The construction of the KDE is naturally such that the overall distribution gives a mean of what is characterized as "$1.0 \mu$" in the figure, but around that mean the distribution of all measured concentrations falls within an order of magnitude of the mean.
The implication is that the most probable result from a single concentration sampling event will be a value around the mean, but the possibility that a single sampling event might give a result that is an order of magnitude higher or lower than a second independent sampling event cannot be ruled out, even under non-preferential pathway circumstances.
Thus, to the extent that the ASU house serves as a prototype for “normal” VI scenarios (when the preferential pathway was closed), it is only when any particular sampling event provides a result that is an order of magnitude below some action threshold that one can be highly confident that further sampling will not change this outcome.
The more sampling events that are involved, such that a population mean can confidently be estimated, the less likely that any future sampling event would give a value that is more than about a half an order of magnitude beyond that mean.
These are of course very rough rules of thumb, based upon this analysis of this data set.

When the preferential pathway beneath the ASU house is open, a very different picture emerges.
The calculated univariate distribution function takes on a much more complicated, somewhat bimodal character.
It should be recalled that the mean of all of the results remains, as before, at $"1.0 \mu"$.
The red distribution curve shows the origin of what has been reported as the “three orders of magnitude” variability in the indoor air contaminant concentrations.
On the one hand, there is a significant probability that any one measured concentration can be lower, up to even an order of magnitude below, the mean of the whole data set.
On the other hand, there is a spread to the upside of the mean by two orders of magnitude.
So in other words, the existence of an open preferential pathway in this case offers little guidance on what might be expected from any single sampling event.
This could easily have been inferred from the already raw published data from this site, but this analysis offers a statistical picture that suggests that there are distinctly different operational situations that give rise to different peaks in the observed distributions.
This, in turn, points to different operational parameters that have not yet been fully captured.

There is inherently greater spread in the ASU house data when the preferential pathway is open.
The variations in a direction greater than the mean show much greater spread, and arguably the choice of a Gaussian kernel is not necessarily the best.
The bivariate analysis below will shed more light on this point.

The univariate KDE distributions form the NAS North Island site show a distinctly bimodal distribution (the blue curve).
There is a population of sampling results above the mean and another below the mean.
Given that there was no influence of preferential pathways at this site, these results immediately flag the possibility that one is dealing with sample concentration populations that are strongly influenced by the other variable displayed on this plot - the indoor-outdoor pressure difference.
When the interior of the building is depressurized relative to ambient, there is a tendency towards indoor air contaminant concentrations higher than the mean, and when the building is pressurized, the values tend to fall below the mean.
This immediately suggests the examination of bivariate distributions, which are also shown on Figure \ref{fig:2d_kde}, in the center of the plot.

The center of Figure \ref{fig:2d_kde} shows the bivariate KDEs constructed based on the relationship between the two aforementioned variables.
The color-coding indicates the source of the dataset, with green again representing the ASU house dataset after the preferential pathway was closed, red is for before the pathway was closed, and blue the NAS North Island dataset.
The Pearson’s r-values and p-value for the relation between TCE in indoor air and indoor-outdoor pressure difference for each dataset are also shown.
The bivariate distributions in Figure \ref{fig:2d_kde} show even more starkly the differences between the ASU house datasets and the NAS North Island dataset.
We believe that this is primarily due to the significant different indoor-outdoor pressure differences existing at these two sites.
Again, NAS North Island showed a much larger spread of indoor-outdoor pressure differential values, seen in the distributions on the top x-margin; the ASU house pressure difference distributions were quite Gaussian with a variation of just a few Pascals, whereas the NAS North Island distribution is rather more flat and includes a much larger span of values.

The bivariate distributions make clear one aspect of the ASU house results.
The “outlier” points at almost two orders of magnitude beyond the mean are associated with the existence of a preferential pathway and an unusually high indoor-outdoor pressure differential.
But for that particular combination of circumstances, the variability in indoor air contaminant concentrations, given existence of the preferential pathway, would have been “only” two orders of magnitude.
Comparing the green and red bivariate distributions, it is clear that the existence of the preferential pathway increases the potential spread of measured concentration values.
But beyond this, the correlation statistics show fairly convincingly that the pressure driving force is connected with the observed indoor air contaminant concentrations (even leaving out the few points at high depressurization), but that the correlation between indoor air contaminant concentration and pressure driving force is weak under “normal” VI conditions, where the preferential pathway was closed off.
This strongly suggests that under the latter conditions, advective entry was not a significant driver, and that ordinary diffusive processes were playing a role in bringing contaminant into the house from the area beneath the slab.
The above results reaffirm the problem that the potential existence of preferential pathways can cause for VI site investigations, i.e. they can greatly reduce the reliability of any single sampling event, requiring a larger number of samples to properly assess the site.
The existence of preferential pathways to the vicinity of a structure can enhance the effect of other variables, such as small pressure differences.
In this case, any single sampling event under “natural” conditions might give a value that is two orders of magnitude different from another sampling event.
Conservatism in that instance would suggest that any sampling results within two orders of magnitude of an action threshold would require re-sampling.
While the present results are taken from a single site, there is every reason to believe that this site can be representative of a large number of real sites.

Turning to the bivariate results for the NAS North Island site, the bivariate distributions leave little doubt that indoor-outdoor pressure differential is a key driver for measured indoor air contaminant concentrations.
The correlation statistics confirm the visual conclusion.
There is little surprise in the results, in that higher extents of depressurization lead to more contaminant entry.
Note that but for the portions of the blue contours at high and low pressure differences, the distribution of indoor air concentration values, relative to a mean, would be close to those from the ASU house.
In other words, at least some of the high temporal variability at NAS North Island comes from the rather extreme range of pressure differentials that exist there.
This supports a conclusion that under “normal” VI conditions, absent a preferential pathway, with the normally encountered pressure differentials of a few Pa, temporal variations in measured indoor air contaminant concentrations of an order of magnitude must be expected.
Clearly it would be useful to have a few more datasets from different sites to support such a general conclusion.

\begin{figure}[htb!]
  \caption{Probability density plots of the deviation from the mean Chloroform and Tetrachloroethene concentrations along with the indoor-outdoor pressure difference at the Indianapolis site}
  \label{fig:2d_kde_indianapolis}
  \includegraphics[width=\textwidth]{pressure_vs_indoor_concentration_422.pdf}
\end{figure}

Figure \ref{fig:2d_kde_indianapolis} shows the same as Figure \ref{fig:2d_kde} but at the Indianapolis site, and the variations of Chloroform and PCE are considered instead of TCE.
The concentration of both Chloroform and PCE are still roughly equally log-normally distributed within half an order of magnitude above and below the mean, similar to the variation observed at the ASU house after the land drain was closed, as well as to the North Island site, if one would consider comparable indoor-outdoor pressure differences.
The distribution of pressure difference at the Indianapolis site is between that of ASU and North Island, with the structure there typically being more depressurized than ASU but less than North Island.
The relationship between pressure difference and IACC is much stronger at the Indianapolis site than it was at the ASU house, but less so than at North Island.
Given what is known about the soil around Indianapolis site, this is somewhat surprising, as it is not significantly different from the ASU site (which North Island is).
This may be evidence of the interaction with the PP that exists at the site, although the extreme variations in IACC cannot be observed here.

\subsubsection{The Role Of Air Exchange Rate}

% Plots showing the relative minor contribution to TCE in indoor air variance fluctuating air exchange rate has
\begin{figure}[htb!]
  \caption{Comparison between the recorded  and the calculated TCE in indoor air at the ASU house, assuming constant air exchange rate. \ref{fig:asu_exchange_rate_analysis_overview} shows the TCE in indoor air across time as well as the exchange rate. \ref{fig:asu_exchange_rate_analysis_violin_plot} shows the distribution of these values for three periods.}
  \label{fig:asu_exchange_rate_analysis}
  % "Overview" plot
  \begin{subfigure}{\textwidth}
    \caption{ }
    \label{fig:asu_exchange_rate_analysis_overview}
    \includegraphics[width=\textwidth]{tce_indoor_air_exchange_rate_impact.png}
  \end{subfigure}
  % Violinplot figure
  \begin{subfigure}{0.75\textwidth}
    \caption{ }
    \label{fig:asu_exchange_rate_analysis_violin_plot}
    \includegraphics[width=\textwidth]{violin_plot_exchange_rate_impact.png}
  \end{subfigure}
\end{figure}

Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion.
Exploring the impact of fluctuations in air exchange rate is thus also important for the development of sampling strategies.
At the ASU house, the contaminant entry rate into the structure (called emission rate in that study) as well as the building air exchange rate were recorded across time.
This allows for exploration of the impact that fluctuating air exchange rate has on the final IACC by solving the CST equation, using the recorded contaminant entry rate as one input, and assuming a constant air exchange rate.
The rationale is to show the extent of variability in IACC due to factors other than fluctuations in air exchange rates.
We have taken this factor out of consideration by assuming a single constant exchange rate, characteristic of the whole period.

In Figure \ref{fig:asu_exchange_rate_analysis_overview}, the measured IACC and air exchange rate across time are shown, with the predicted IACC assuming a constant air exchange rate.
The median air exchange rate value for the pre-CPM period was chosen as this seemed to be the most representative value for this time period.
From Figure \ref{fig:asu_exchange_rate_analysis_overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC assuming constant and actual fluctuating exchange rates are quite similar.
In other words, in this instance it is unlikely that the variations in IACC were driven by fluctuations in air exchange rate.

This is more clearly shown in Figure \ref{fig:asu_exchange_rate_analysis_violin_plot}, where KDEs of the IACC are constructed assuming the constant and fluctuating air exchange rate cases.
This figure shows the relative probability of being at a particular TCE in indoor air concentration at the ASU house.
These are compared to each other for 1. the full measurement period, 2. the non-CPM periods and 3. the CPM period.
Looking at the full period, it does not appear as though the two cases of steady and fluctuating air exchange rate are comparable, the distribution functions for IACC are different and offset.
The curves to the different sides of the vertical line for the "full" measurement period represent the probability distributions for the indicated values of IACC.
It appears as though assuming a constant air exchange rate shifts the IACC probability distribution to higher values.
The reason for this are more apparent when considering the CPM and non-CPM periods separately.

Considering only the CPM period, assuming a constant air exchange rate does not influence the shape of the IACC distribution, but assuming constant exchange rate shifts the IACC to a higher values.
This is because the actual air exchange rate for this CPM period is significantly higher than the assumed median value, and much more contaminant is actually exchanged with the exterior.
The shape of the IACC distribution is not very different from that which takes the air exchange fluctuations into account however, indicating that fluctuating exchange rate is not a major contributor to observed IACC variation for this period (the distinction being drawn between variation and absolute levels of IACC).

The non-CPM period is the condition under which field practitioners would normally collect samples at a VI site.
It is clear here that there is only a very minor difference made in assuming a constant vs. real fluctating air exchange rate, again indicating that account for the minor fluctuations of air exchange rate's impact on variations in IACC is not important.
\citeauthor{rackes_time-averaged_2016}\cite{rackes_time-averaged_2016} drew similar conclusions regarding the relationship between IACC and air exchange rate.

\subsection{The Influence of Preferential Pathways}

% Point is:
% If preferential pathways exist, then roughly 1-3 orders of magnitude variability are possible.

% Definition: A preferential pathway is a pathway that allows for contaminant vapor to enter a structure directly or close to the building - circumventing much of the resistance to transport in the soil.

% How do I deal with this? How do I know if preferential pathways may be an issue?

% Scenario: Preferential pathway directly connected to the building.
% - Identify sewer & plumbing fixtures.
% - Drains and sumps may be sources.

% Scenario: Preferential pathway in the near slab area.
% - Permeable sub-base necessary.
% - May have a limited “sphere of influence”.
% -- Limit likely not reached for a normal house.
% -- Limit more likely to be reached for larger structures.

% Relevant for both: Check local manhole/sewer for presence of contaminant - likely source for the preferential pathway.

\subsubsection{Conclusions From Steady-State Modeling}

Clearly, PPs pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover; even at a well studied site like the ASU house it took years to discover the PP.
Therefore, there is a need to consider other indications that a PP is present.

% Sensitivity analysis plot showing when a PP is relevant
\begin{figure}[htb!]
  \caption{Attenuation factor relative to contaminant vapor in equilibrium with groundwater as a function of indoor-outdoor pressure difference. The effects of a preferential pathway, that is either contaminated or uncontaminated as well as that of having a gravel sub-base vs. uniform sandy clay soil are considered.}
  \label{fig:model_results_sensitivity_analysis}
  % Upper left
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_0}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_0.png}
  \end{subfigure}
  % Upper right
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_1}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_1.png}
  \end{subfigure}
  % Middle left
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_2}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_2.png}
  \end{subfigure}
  % Middle right
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_3}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_3.png}
  \end{subfigure}
  % Lower left
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_4}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_4.png}
  \end{subfigure}
  % Lower right
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:model_results_sensitivity_analysis_5}
    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_5.png}
  \end{subfigure}
\end{figure}

One way has been suggested by the situation at the ASU house, where there clearly was a PP containing contaminant vapor, and the PP exited into a gravel (permeable) sub-base.
This situation can be examined using the VI model introduced earlier by examining how IACC is impacted by various factors over a range of indoor-outdoor pressure difference values.
Instead of representing IACC as absolute concentration, we now non-dimensionalize with respect to the vapor in equilibrium with the groundwater contaminant concentration.
This non-dimensionalized property is commonly called the attenuation factor and is denoted by $\alpha_\mathrm{gw}$.
The result of a sensitivity analysis relative to several factors is seen in Figure \ref{fig:model_results_sensitivity_analysis}.
These examine scenarios in which the indoor-outdoor pressure difference is a driver for VI.

In this figure, all the graphs in the left column show results for a scenario, in which the house has a gravel sub-base, with the remaining surrounding soil assumed to be sandy clay.
All the right column graphs features only sandy clay soil, directly in contact with the slab.
Figure \ref{fig:model_results_sensitivity_analysis_0} considers an uncontaminated PP ($\chi = 0$), i.e. the PP delivers clean air to beneath the slab.
In \ref{fig:model_results_sensitivity_analysis_1} there is no PP present at all, giving a reference state for a "normal" VI scenario.
\ref{fig:model_results_sensitivity_analysis_2} features a contaminated PP ($\chi = 1$), i.e. the PP delivers contaminant vapor at a concentration characteristic of underlying groundwater.

In terms of conditions that may lead to significant temporal transients in $\alpha_\mathrm{gw}$, driven by indoor-outdoor pressure difference, most of the cases in this analysis show negative results.
There are only two combinations of factors where significant changes in $\alpha_\mathrm{gw}$ accompany the changes in indoor-outdoor pressure difference, both involving the permeable (gravel) sub-base into which a PP enters.
In \ref{fig:model_results_sensitivity_analysis_2}, the PP is filled with contaminant vapor whereas figure \ref{fig:model_results_sensitivity_analysis} involves the PP only delivering "clean" air to the gravel sub-base.
When the building is overpressurized (positive indoor-outdoor pressure difference), the two cases exhibit similar behavior, i.e. $\alpha_\mathrm{gw}$ is low as expected.

When the building is underpressurized, the uncontaminated PP case involves barely any increase in $\alpha_\mathrm{gw}$ while there is a very significant increase in $\alpha_\mathrm{gw}$ in the contaminated PP case.
For the contaminated PP scenario, this is quite easy to understand, as the extra air flow provided by the PP into the gravel sub-base very effectively disperses contaminant vapor across the sub-base and the contaminant entry rate is increased.

When the PP is provides clean air, the increase in air flow into the sub-base with depressurization is the same.
However, the resulting increase in air entry rate does not bring in more contaminant as clean air is being drawn in.
This means that most of the contaminant entry rate is still only controlled by the usual process of diffusion through the soil, giving a result that is actually similar to that for the "no PP and with gravel sub-base" scenario.
Thus, for a PP to be a significant contributor to contaminant entry at a VI site the conditions such as featured in the bottom left case are required, i.e. a permeable sub-base and the PP must be a source of contaminant.

Various techniques for determining the potential for sewer gas entry into a house are described by \citeauthor{nielsen_remediation_2017-1}\cite{nielsen_remediation_2017-1}.
Collecting contaminant vapor samples from nearby manholes is also recommended, however one should be aware that contaminant vapors can potentially travel very long distances in a sewer system\cite{mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.

% Models specific to ASU & North Island
\begin{figure}[htb!]
  \caption{Attenuation factor vs. indoor-outdoor pressure difference when considering VI scenarios similar to the ASU house and the North Island NAS site (\ref{fig:ss_asu} and \ref{fig:ss_north_island} respectively).}
  \label{fig:model_results_asu_north_island}
  % ASU
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:ss_asu}
    \includegraphics[width=\textwidth]{ss_asu.png}
  \end{subfigure}
  % North Island
  \begin{subfigure}[t]{0.45\textwidth}
    \caption{ }
    \label{fig:ss_north_island}
    \includegraphics[width=\textwidth]{ss_north_island.png}
  \end{subfigure}
\end{figure}

At the ASU house, after the PP was closed, there was still roughly an order of magnitude variation in IACC, the reasons for which Figure \ref{fig:model_results_sensitivity_analysis} does not address.
Likewise, there was more than two orders of magnitude observed at the North Island site over the indoor-outdoor pressure difference ranges simulated despite there not being a PP present.
Therefore, it is of interest to examine these sites a bit further using the model.

The foundation crack at the ASU house seems by all accounts to be quite small, roughly 180x1 cm and located close to the PP\cite{guo_vapor_2015}.
This crack is smaller than the one assumed in the model used in Figure \ref{fig:model_results_sensitivity_analysis}.
The influence of indoor-outdoor pressure difference is explored for a case in which a smaller crack area is assumed.
We also consider the effects of having a significantly lower air exchange rate, 0.1 vs. the regular 0.5 per hour.

The results of the case shown in Figure \ref{fig:model_results_sensitivity_analysis_2} is shown in Figure \ref{fig:ss_asu}.
Even though there is an increase in $\alpha_\mathrm{gw}$ with increased underpressurization, it is no where near the roughly one order of magnitude recorded at the ASU house after the PP was closed.
So while crack dimensions can clearly influence $\alpha_\mathrm{gw}$, they cannot be used to infer a bigger role for pressure difference in temporal variations in $\alpha_\mathrm{gw}$ or IACC.
The influence of air exchange rate on $\alpha_\mathrm{gw}$ predictions is also shown in Figure \ref{fig:model_results_asu_north_island}.
Clearly, the value of air exchange rate at steady-state conditions will only influence absolute concentrations in the structure.
Ultimately, the reason for the transient variability in IACC during the post-CPM period at the ASU house remains elusive, as a Pearson's r analysis between IACC and various factors do not yield any explanation.

In the case of the North Island site, variations in IACC are clearly not related to the existence of a PP.
The soil underlying the North Island site is more sand-like\cite{hosangadi_high-frequency_2017}.
The significance of this is that sand soil is much more permeable than sandy clay soil, which means that $\alpha_\mathrm{gw}$ will be much more sensitive to changes in indoor-outdoor pressure difference.
Thus, another "normal" VI scenario was run but this time with sand soil instead of sandy clay, the results of which may be seen in Figure \ref{fig:ss_north_island}.
Under these conditions, $\alpha_\mathrm{gw}$ changes significantly with pressurization, spanning roughly two orders of magnitude.
This is consistent with what was observed at North Island, as there the span of IACC values was roughly the same across this range of indoor-outdoor pressure difference values.
This simulation, together with the Pearson $r = -0.64$ for North Island in Figure \ref{fig:2d_kde} suggest that under conditions where the soil around a structure is relatively permeable, indoor-outdoor pressure difference can be a significant driving forces for transient changes in IACC.
Therefore, recording the indoor-outdoor pressure difference at a site (in particular if the soil is permeable) may offer insight into the potential for transient changes in IACC (some good methods for doing this are suggested by \citeauthor{nielsen_remediation_2017-1}\cite{nielsen_remediation_2017-1}).

% Transient part
\subsubsection{Transient Modeling}

Analyzing the impact of a PP using steady-state simulations such as those in Figures \ref{fig:model_results_sensitivity_analysis} and does not revealing about how quickly IACC can change in the presence of a PP.
A transient simulation of a VI scenario that is characterized by a PP has been performed, where only the indoor-outdoor pressure difference was temporally changed.
The pressure difference is changed in a sinusoidal fashion, defined by \eqref{eq:p_transient},
\begin{equation}
  \label{eq:p_transient}
  p_\mathrm{in/out} = 10 \sin{(2 \pi t)} - 0.5
\end{equation}
where $t$ is given in days and $p_\mathrm{in/out}$ is the indoor-outdoor pressure difference in Pa.
Figure \ref{fig:transient_simulation} shows the results of this simulation.

\begin{figure}
  \caption{Transient response of TCE in indoor air (as attenuation factor) and in the surrounding soil \& gravel sub-base for a PP VI scenario subject to sinusoidal indoor-outdoor pressure fluctuations.
  \ref{fig:alpha_response} shows the attenuation factor and indoor-outdoor pressure across time.
  Figures \ref{fig:soil_30_hr} \& \ref{fig:soil_42_hr} show the contaminant concentration (normalized to the contaminant vapor in equilibrium with groundwater) in the surrounding soil \& gravel sub-base at different depths and times.}
  \label{fig:transient_simulation}
  % Transient response figure
  \begin{subfigure}[t]{0.75\textwidth}
    \caption{ }
    \label{fig:alpha_response}
    \includegraphics[width=\textwidth]{alpha_transient.png}
  \end{subfigure}
  % Soil t = 30 hr ...
  \begin{subfigure}[c]{0.45\textwidth}
    \caption{t = 30 hr}
    \label{fig:soil_30_hr}
    \includegraphics[width=\textwidth]{tri_layer_soil_concs1.png}
  \end{subfigure}
  % Soil t = 42 hr ...
  \begin{subfigure}[c]{0.45\textwidth}
    \caption{t = 42 hr}
    \label{fig:soil_42_hr}
    \includegraphics[width=\textwidth]{tri_layer_soil_concs2.png}
  \end{subfigure}
  % Color map
  \begin{subfigure}[c]{0.05\textwidth}
    \phantomcaption{ }
    %\label{fig:cmap}
    \includegraphics[width=\textwidth]{cmap.png}
  \end{subfigure}
\end{figure}

Figure \ref{fig:alpha_response} shows the IACC response, given as attenuation factor relative to contaminant vapor in equilibrium with groundwater, and the indoor-outdoor pressure difference with time.
What is apparent is that $\alpha_\mathrm{gw}$ closely follows the indoor-outdoor pressure difference, with the max./min. $\alpha_\mathrm{gw}$ being reached roughly 0.5 hours after the indoor-outdoor pressure difference reaches its max./min..
The reader should be aware that the assumption that the indoor air space can be modeled as a CST overestimates how dynamic this process is as a real house probably would not be perfectly mixed on as a short time scale as this.
Regardless, this suggests that VI sites characterized by PPs are highly dynamic and IACC's may change quickly in response to changing indoor-outdoor pressure differences, consistent with what was observed at the ASU house\cite{holton_temporal_2013}.
Again, the dynamics of the indoor-outdoor pressure difference at a VI site can offer insight into the potential for transient behavior of IACC.

Figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} show the contaminant concentration (normalized to the contaminant vapor in equilibrium with groundwater) in the soil and gravel sub-base at three different depths at times of 30 and 42 hr respectively corresponding to roughly the maximum in overpressurization and depressurization respectively.
The top layer is 3 m above groundwater and immediately under the house foundation (indicated by the black square).
The middle and bottom layers are 2 and 1 m above groundwater respectively.
The black streamlines show the contaminant transport flow paths.
The region closest to the PP is expanded in the circular insets.

The impact of the PP on the contaminant vapor concentration in the subsurface is apparent in these figures, as one easily sees how (especially) the gravel sub-base is filled up.
The contaminant vapor concentration is the highest closest to the PP (which is located at the edge of the gravel sub-base) and decreases with increased distance from the PP.
Note that a very small amount of contaminant vapor leaves the gravel sub-base, as the sandy clay has a relatively large resistance to transport compared with gravel.
This makes it difficult to detect the contribution of the PP outside the foundation footprint.
However, this may not always be true, as vapor contaminant from the PP would disperse more easily into more permeable soils.

It should also be noted that the vapor contaminant emanating from the PP in this model is relatively high concentration; the vapor contaminant concentration is equal to that in equilibrium with the groundwater source.
This does not always seem to be the case, as VI sites that have uncovered PPs find that the vapor contaminant emanating from the PP can be lower than that\cite{guo_vapor_2015}.
Therefore, it may be much more difficult than it may appear in Figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} to localize a vapor contaminant hotspot as a way to determining the presence of a PP at a particular VI site.

The insert circles in figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} also show that for the simulated time period, only a small amount of the contaminant enters the structure (as evident by the blue low concentration zone around the perimeter), yet a very large increase in $\alpha_\mathrm{gw}$ occurs.
The concentration in the rest of the sub-base is unchanged.
This suggests that once a PP has deposited contaminant underneath a structure, it may persist for a long time period, as pointed out by \citeauthor{guo_vapor_2015}\cite{guo_vapor_2015}.

\begin{acknowledgement}
This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).
\end{acknowledgement}

\begin{table}[htb!]
  \caption{Abbreviations \& symbols}
  \begin{tabular}{l l}
  \toprule
  \textbf{Abbreviation or symbol}     & \textbf{Explanation} \\
  % General abbreviations
  VI                                  & Vapor intrusion \\
  PP                                  & Preferential pathway \\
  $L_\mathrm{slab}$                   & Thickness of the foundation concrete slab \\
  % Richard's equation
  $p$                                 & Pressure \\
  % Unsteady-cstr
  $u$                                 & TCE in indoor air concentration \\
  $c$                                 & TCE in soil gas concentration \\
  % Diffusion stuff
  $D_\mathrm{air}$                    & Diffusion of TCE in air \\
  $D_\mathrm{water}$                  & Diffusion of TCE in water \\
  $D_\mathrm{eff}$                    & Effective Diffusion of TCE in soil \\
  $K_H$                               & Henry's Law constant \\
  % van Genuchten stuff
  $\theta_g, \theta_w\;\&\;\theta_s$  & Vapor, water filled \& saturated soil porosity \\
  $\theta_r$                          & Residual moisture porosity \\
  $\alpha,\; l,\; n\;\&\; m$          & van Genuchten parameters \\
  $\mathrm{Se}$                       & Soil moisture saturation \\
  $k_r$                               & Relative permeability \\
  $\alpha_\mathrm{gw}$                & Attenuation factor \\
  \bottomrule
  \end{tabular}
\end{table}

\bibliography{Zotero}

\end{document}

### 2018-10-02 python/ternary-soil-cpm-pressure.py
Set up ternary plot

@@ -1,0 +1,74 @@
+import numpy as np
+import plotly.graph_objs as go
+from plotly.offline import download_plotlyjs, init_notebook_mode, plot, iplot
+import plotly.io as pio
+
+import json
+from urllib import request
+
+url = 'https://gist.githubusercontent.com/davenquinn/988167471993bc2ece29/raw/f38d9cb3dd86e315e237fde5d65e185c39c931c2/data.json'
+response = request.urlopen(url).read()
+data = json.loads(response)
+
+colors = ['#8dd3c7','#ffffb3','#bebada',
+          '#fb8072','#80b1d3','#fdb462',
+          '#b3de69','#fccde5','#d9d9d9',
+          '#bc80bd','#ccebc5','#ffed6f'];
+
+
+pressure = {
+    'sand': '8.0',
+    'loamy sand': '45',
+    'sandy loam': '34',
+    'sandy clay loam': '234',
+    'sandy clay': '223',
+    'clay': '44',
+    'clay loam': '34',
+    'silty clay': '43',
+    'silty clay loam': '12',
+    'silty loam': '13',
+    'silt': '87',
+    'loam': '78',
+    }
+
+
+# generate a,b and c from JSON data..
+traces = []
+color_iter = iter(colors)
+for i in data.keys():
+    trace = dict(text=i,
+        type='scatterternary',
+        a=[ k['clay'] for k in data[i] ],
+        b=[ k['sand'] for k in data[i] ],
+        c=[ k['silt'] for k in data[i] ],
+        mode='lines',
+        line=dict(color='#444'),
+        fill='toself',
+        fillcolor=next(color_iter),
+        name = pressure[i],
+        showlegend = True,
+    )
+    traces.append(trace)
+
+layout = {
+    'title': 'Soil types',
+    'ternary':
+        {'sum':100,
+         'aaxis':{'title': 'clay', 'ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' },
+         'baxis':{'title': 'sand', 'ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' },
+         'caxis':{'title': 'silt','ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' }},
+    'showlegend': False,
+    'annotations': [{
+      'showarrow': False,
+      'text': 'Clay',
+        'x': 0.5,
[…]

### 2018-10-02 python/radar-soil-cpm-pressure.py
Made soil pressure plot

@@ -1,0 +1,50 @@
+import plotly.graph_objs as go
+import plotly.offline as py
+import plotly.io as pio
+import pandas as pd
+
+df = pd.read_csv('./data/cpm-pressure-study/cpm-pressure-soil-type.csv')
+
+soils = (
+    'Sand',
+    'Loamy Sand',
+    'Sandy Loam',
+    'Sandy Clay Loam',
+    'Loam',
+    'Silt Loam',
+    'Clay Loam',
+    'Silty Clay Loam',
+    'Silty Clay',
+    'Silt',
+    'Sandy Clay',
+    'Clay',
+)
+
+
+data = [
+    go.Scatterpolar(
+      r = df['Indoor/outdoor pressure difference'].values*-1,
+      theta = soils,
+      fill = 'toself',
+      name = 'Indoor/outdoor'
+    ),
+    go.Scatterpolar(
+      r = df['Sub-slab pressure difference'].values*-1,
+      theta = soils,
+      fill = 'toself',
+      name = 'Across slab'
+    )
+]
+
+layout = go.Layout(
+  polar = dict(
+    radialaxis = dict(
+      visible = True,
+      #range = [0, 1954.3]
+    )
+  ),
+  showlegend = True
+)
+
+fig = go.Figure(data=data, layout=layout)
+pio.write_image(fig, 'figures/radar-soil-cpm-pressure.pdf')

### 2018-10-02 python/kde-iacc-pressure/pp.py
Moved kde plpt here

@@ -1,0 +1,141 @@
+import numpy as np
+import sqlite3
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from scipy import stats
+import matplotlib.patches as mpatches
+
+
+db_asu = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/asu_house.db')
+db_ind = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/indianapolis.db')
+
+asu = pd.read_sql_query("SELECT day, pressure_difference AS Pressure, tce_emission_rate, building_flow_rate FROM parameters;", db_asu).interpolate(limit_area='inside',limit=250).dropna()
+asu['Concentration'] = asu['tce_emission_rate']/asu['building_flow_rate']
+cpm_start, cpm_end = 780.0, 1157.0
+asu_pre_cpm = asu[(asu['day'] < cpm_start) & (asu['Pressure'] < 7.5)].copy().dropna()
+asu_pre_cpm['Pressure'] = -1.0*asu_pre_cpm['Pressure'].dropna()
+
+
+ind_chloro = pd.read_sql_query(
+    "SELECT \
+        O.StopDate AS StopDate, \
+        AVG(C.Value) AS Concentration, \
+        AVG(P.Value) AS Pressure \
+    FROM \
+        Observation_Status_Data AS O, \
+        VOC_Data_SRI_8610_Onsite_GC AS C, \
+        Differential_Pressure_Data AS P \
+    WHERE \
+        O.Value = 'not yet installed' AND \
+        C.StopDate = O.StopDate AND \
+        C.Location = '422BaseS' AND \
+        P.StopDate = O.StopDate AND \
+        C.Variable = 'Chloroform' AND \
+        P.Variable = 'Basement.Vs.Exterior' AND \
+        P.Location = '422' \
+    GROUP BY \
+        O.StopDate \
+    ;", db_ind)
+
+ind_pce = pd.read_sql_query(
+    "SELECT \
+        O.StopDate AS StopDate, \
+        AVG(C.Value) AS Concentration, \
+        AVG(P.Value) AS Pressure \
+    FROM \
+        Observation_Status_Data AS O, \
+        VOC_Data_SRI_8610_Onsite_GC AS C, \
+        Differential_Pressure_Data AS P \
+    WHERE \
+        O.Value = 'not yet installed' AND \
[…]

### 2018-10-03 tex/cover-letter/cover_letter.tex
Added the journal cover letter

\documentclass{letter}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\signature{Jonathan Ström}
\address{School of Engineering \\ Brown University \\ Providence 02912}
\begin{document}

\begin{letter}{Editor of Environmental Science \& Technology \\
1155 Sixteenth Street N.W. \\
Washington, DC 20036}
\opening{Dear Sir or Madam:}

I am pleased to submit this manuscript entitled "Modeling Variation in Indoor Air Contaminant Concentration \& Impact of Preferential Contaminant Pathways on Vapor Intrusion" for review and possible publication in Environmental Science \& Technology.
This is original work, and has not been previously submitted to any other journal or conference proceedings.

Using numerical models and published field data from two vapor intrusion (VI) sites, we explore the causes of the temporal variation in indoor air contaminant concentrations in VI impacted houses, a poorly understood phenomena, and especially the role of preferential contaminant pathways in this.

We found under which conditions one may expect to observe significant changes in indoor air concentrations, and when one should be especially be concerned with preferential pathways.
These finding help understand and explain the temporal variation that has been observed at many VI sites, and giving practical guidance for assessment of these sites, something that we expect will be of great interest to your readership.

All of the work presented is either created by us or properly referenced, including the TOC art which adhere to the TOC art policy and is the prescribed size of 240 by 135 points.

Thank you for considering this manuscript for publication.
\closing{Sincerely,}

\end{letter}
\end{document}

### 2018-10-03 python/kde-iacc-pressure/nas-asu-house.py
Renamed files

@@ -1,0 +1,112 @@
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+import sqlite3
+from scipy import stats
+import matplotlib.patches as mpatches
+
+
+# databases
+db_asu = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/asu_house.db')
+db_nas = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/north_island_nas.db')
+
+# loads asu data, interpolates and drops NAs
+asu = pd.read_sql_query("SELECT day, pressure_difference AS Pressure, tce_emission_rate, building_flow_rate FROM parameters;", db_asu).interpolate(limit_area='inside',limit=250).dropna()
+
+# loads nas data, interpolates and drops NAs
+nas = pd.read_sql_query("SELECT tce_indoor_air AS Concentration, pressure_difference AS Pressure FROM north_island_nas;", db_nas).interpolate(limit_area='inside',limit=250).dropna()
+
+
+asu['Concentration'] = asu['tce_emission_rate']/asu['building_flow_rate']
+
+cpm_start, cpm_end = 780.0, 1157.0
+
+asu_pre_cpm = asu[(asu['day'] < cpm_start) & (asu['Pressure'] < 7.5)].copy().dropna()
+asu_post_cpm = asu[asu['day'] > cpm_end].copy().dropna()
+
+asu_pre_cpm.name = "ASU House, PP open"
+asu_post_cpm.name = "ASU House, PP closed"
+nas.name = "North Island"
+
+dfs = []
+alpha = 0.6
+cmaps = ["Blues", "Reds", "Greens"]
+
+df = pd.DataFrame({'x': [], 'y': []})
+g = sns.JointGrid(x="x", y="y", data=df) # empty joint grid
+label_patches = []
+for i, df in enumerate([nas, asu_pre_cpm, asu_post_cpm]):
+    print("Dataset: %s" % df.name)
+
+    df['Concentration'] = df['Concentration'].apply(pd.to_numeric, errors='ignore').apply(lambda x: np.log10(x)).replace([np.inf, -np.inf], np.nan).dropna()
+    df['Concentration'] -= df['Concentration'].mean()
+    df['Pressure'] = -1.0*df['Pressure'].dropna()
+
+    #if df.name is "ASU House, PP open":
+    #    df = df.loc[df['Pressure'] > -7.5]
[…]

### 2018-10-03 python/advection-threshold/radar-plot.py
Made plotting function

@@ -1,0 +1,62 @@
+import plotly.graph_objs as go
+import plotly.offline as py
+import plotly.io as pio
+import pandas as pd
+
+data_dir = './data/advection-threshold/'
+figures_dir = './figures/advection-threshold/'
+
+def radar_plot(dfs, filename):
+    # soil list
+    soils = (
+        'Sand',
+        'Loamy Sand',
+        'Sandy Loam',
+        'Sandy Clay Loam',
+        'Loam',
+        'Silt Loam',
+        'Clay Loam',
+        'Silty Clay Loam',
+        'Silty Clay',
+        'Silt',
+        'Sandy Clay',
+        'Clay',
+    )
+
+    data = []
+    for df in dfs:
+        data.append(
+            go.Scatterpolar(
+                r = df*-1,
+                theta = soils,
+                fill = 'toself',
+                name = df.name
+            )
+        )
+
+    layout = go.Layout(
+      polar = dict(
+        radialaxis = dict(
+          visible = True,
+        )
+      ),
+      showlegend = True
+    )
+
+    fig = go.Figure(data=data, layout=layout)
+    pio.write_image(
+        fig,
+        file = figures_dir + filename + '.pdf',
+        #format='pdf',
+        )
+    return
+
+df = pd.read_csv(data_dir+'cpm-pressure-soil-type.csv')
+radar_plot(
+    filename = 'indoor-outdoor-and-sub-slab-pressure-difference',
+
+    dfs = (
+        df['Indoor/outdoor pressure difference'],
+        df['Sub-slab pressure difference'],
+    ),
+)

### 2018-10-03 python/advection-threshold/radar-soil-cpm-pressure.py
Name changes

@@ -1,50 +1,0 @@
-import plotly.graph_objs as go
-import plotly.offline as py
-import plotly.io as pio
-import pandas as pd
-
-df = pd.read_csv('./data/cpm-pressure-study/cpm-pressure-soil-type.csv')
-
-soils = (
-    'Sand',
-    'Loamy Sand',
-    'Sandy Loam',
-    'Sandy Clay Loam',
-    'Loam',
-    'Silt Loam',
-    'Clay Loam',
-    'Silty Clay Loam',
-    'Silty Clay',
-    'Silt',
-    'Sandy Clay',
-    'Clay',
-)
-
-
-data = [
-    go.Scatterpolar(
-      r = df['Indoor/outdoor pressure difference'].values*-1,
-      theta = soils,
-      fill = 'toself',
-      name = 'Indoor/outdoor'
-    ),
-    go.Scatterpolar(
-      r = df['Sub-slab pressure difference'].values*-1,
-      theta = soils,
-      fill = 'toself',
-      name = 'Across slab'
-    )
-]
-
-layout = go.Layout(
-  polar = dict(
-    radialaxis = dict(
-      visible = True,
-      #range = [0, 1954.3]
-    )
-  ),
-  showlegend = True
-)
-
-fig = go.Figure(data=data, layout=layout)
-pio.write_image(fig, 'figures/radar-soil-cpm-pressure.pdf')

### 2018-10-03 python/advection-threshold/radar-plot.py
Added support for pivoted pandas dataframes

@@ -2,7 +2,7 @@ import plotly.graph_objs as go
 import plotly.offline as py
 import plotly.io as pio
 import pandas as pd
-
+import matplotlib.pyplot as plt
 data_dir = './data/advection-threshold/'
 figures_dir = './figures/advection-threshold/'
 
@@ -24,13 +24,14 @@ def radar_plot(dfs, filename):
     )
 
     data = []
-    for df in dfs:
+    for i in range(len(df.columns)):
+        print(i)
         data.append(
             go.Scatterpolar(
-                r = df*-1,
+                r = df.values[:,i]*-1,
                 theta = soils,
                 fill = 'toself',
-                name = df.name
+                name = "%s = %1.1f" % (df.columns.name, df.columns.values[i])
             )
         )
 
@@ -47,16 +48,23 @@ def radar_plot(dfs, filename):
     pio.write_image(
         fig,
         file = figures_dir + filename + '.pdf',
-        #format='pdf',
         )
     return
 
-df = pd.read_csv(data_dir+'cpm-pressure-soil-type.csv')
-radar_plot(
-    filename = 'indoor-outdoor-and-sub-slab-pressure-difference',
+df = pd.read_csv(data_dir+'perimeter-ck-gravel-sub-base.csv', header=4)
 
+print(df.pivot(index='Soil type', columns='L', values='Indoor/outdoor pressure difference').index.values)
+
+radar_plot(
+    filename='test',
+    dfs = [df.pivot(index='Soil type', columns='L', values='Indoor/outdoor pressure difference')]
+)
+"""
+radar_plot(
+    filename = 'perimeter-ck-uniform-soil',
     dfs = (
         df['Indoor/outdoor pressure difference'],
-        df['Sub-slab pressure difference'],
+        df['Indoor/outdoor pressure difference'] - df['Sub-slab pressure difference'],
     ),
 )
+"""

### 2018-10-03 python/advection-threshold/radar-plot.py
Added support for pivoted dataframes

@@ -59,12 +59,3 @@ radar_plot(
     filename='test',
     dfs = [df.pivot(index='Soil type', columns='L', values='Indoor/outdoor pressure difference')]
 )
-"""
-radar_plot(
-    filename = 'perimeter-ck-uniform-soil',
-    dfs = (
-        df['Indoor/outdoor pressure difference'],
-        df['Indoor/outdoor pressure difference'] - df['Sub-slab pressure difference'],
-    ),
-)
-"""

### 2018-10-04 python/advection-threshold/radar-plot.py
Generated radar plots

@@ -1,12 +1,13 @@
 import plotly.graph_objs as go
 import plotly.offline as py
 import plotly.io as pio
+import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 data_dir = './data/advection-threshold/'
 figures_dir = './figures/advection-threshold/'
 
-def radar_plot(dfs, filename):
+def radar_plot(df, filename):
     # soil list
     soils = (
         'Sand',
@@ -25,7 +26,7 @@ def radar_plot(dfs, filename):
 
     data = []
     for i in range(len(df.columns)):
-        print(i)
+
         data.append(
             go.Scatterpolar(
                 r = df.values[:,i]*-1,
@@ -51,11 +52,13 @@ def radar_plot(dfs, filename):
         )
     return
 
-df = pd.read_csv(data_dir+'perimeter-ck-gravel-sub-base.csv', header=4)
 
-print(df.pivot(index='Soil type', columns='L', values='Indoor/outdoor pressure difference').index.values)
-
-radar_plot(
-    filename='test',
-    dfs = [df.pivot(index='Soil type', columns='L', values='Indoor/outdoor pressure difference')]
-)
+for file in ('perimeter-ck-gravel-sub-base', 'perimeter-ck-uniform-soil'):
+    df = pd.read_csv(data_dir+file+'.csv', header=4)
+    for col in ('GW depth','Base depth','Crack width'):
+        filename = '%s-%s' % (file, col.lower().replace(' ','-'))
+        print('Generating %s plot' % filename)
+        radar_plot(
+            filename=filename,
+            df = df.pivot_table(index='Soil type', columns=col, values='Indoor/outdoor pressure difference', aggfunc=np.median),
+        )

### 2018-10-04 python/advection-threshold/radar-plot.py
Moved to cpm repo

@@ -1,64 +1,0 @@
-import plotly.graph_objs as go
-import plotly.offline as py
-import plotly.io as pio
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-data_dir = './data/advection-threshold/'
-figures_dir = './figures/advection-threshold/'
-
-def radar_plot(df, filename):
-    # soil list
-    soils = (
-        'Sand',
-        'Loamy Sand',
-        'Sandy Loam',
-        'Sandy Clay Loam',
-        'Loam',
-        'Silt Loam',
-        'Clay Loam',
-        'Silty Clay Loam',
-        'Silty Clay',
-        'Silt',
-        'Sandy Clay',
-        'Clay',
-    )
-
-    data = []
-    for i in range(len(df.columns)):
-
-        data.append(
-            go.Scatterpolar(
-                r = df.values[:,i]*-1,
-                theta = soils,
-                fill = 'toself',
-                name = "%s = %1.1f" % (df.columns.name, df.columns.values[i])
-            )
-        )
-
-    layout = go.Layout(
-      polar = dict(
-        radialaxis = dict(
-          visible = True,
-        )
-      ),
-      showlegend = True
-    )
-
-    fig = go.Figure(data=data, layout=layout)
-    pio.write_image(
-        fig,
-        file = figures_dir + filename + '.pdf',
-        )
-    return
-
-
-for file in ('perimeter-ck-gravel-sub-base', 'perimeter-ck-uniform-soil'):
-    df = pd.read_csv(data_dir+file+'.csv', header=4)
-    for col in ('GW depth','Base depth','Crack width'):
-        filename = '%s-%s' % (file, col.lower().replace(' ','-'))
-        print('Generating %s plot' % filename)
-        radar_plot(
-            filename=filename,
-            df = df.pivot_table(index='Soil type', columns=col, values='Indoor/outdoor pressure difference', aggfunc=np.median),
-        )

### 2018-10-04 python/advection-threshold/ternary-plot.py

@@ -1,74 +1,0 @@
-import numpy as np
-import plotly.graph_objs as go
-from plotly.offline import download_plotlyjs, init_notebook_mode, plot, iplot
-import plotly.io as pio
-
-import json
-from urllib import request
-
-url = 'https://gist.githubusercontent.com/davenquinn/988167471993bc2ece29/raw/f38d9cb3dd86e315e237fde5d65e185c39c931c2/data.json'
-response = request.urlopen(url).read()
-data = json.loads(response)
-
-colors = ['#8dd3c7','#ffffb3','#bebada',
-          '#fb8072','#80b1d3','#fdb462',
-          '#b3de69','#fccde5','#d9d9d9',
-          '#bc80bd','#ccebc5','#ffed6f'];
-
-
-pressure = {
-    'sand': '8.0',
-    'loamy sand': '45',
-    'sandy loam': '34',
-    'sandy clay loam': '234',
-    'sandy clay': '223',
-    'clay': '44',
-    'clay loam': '34',
-    'silty clay': '43',
-    'silty clay loam': '12',
-    'silty loam': '13',
-    'silt': '87',
-    'loam': '78',
-    }
-
-
-# generate a,b and c from JSON data..
-traces = []
-color_iter = iter(colors)
-for i in data.keys():
-    trace = dict(text=i,
-        type='scatterternary',
-        a=[ k['clay'] for k in data[i] ],
-        b=[ k['sand'] for k in data[i] ],
-        c=[ k['silt'] for k in data[i] ],
-        mode='lines',
-        line=dict(color='#444'),
-        fill='toself',
-        fillcolor=next(color_iter),
-        name = pressure[i],
-        showlegend = True,
-    )
-    traces.append(trace)
-
-layout = {
-    'title': 'Soil types',
-    'ternary':
-        {'sum':100,
-         'aaxis':{'title': 'clay', 'ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' },
-         'baxis':{'title': 'sand', 'ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' },
-         'caxis':{'title': 'silt','ticksuffix':'%', 'min': 0.01, 'linewidth':2, 'ticks':'outside' }},
-    'showlegend': False,
-    'annotations': [{
-      'showarrow': False,
-      'text': 'Clay',
-        'x': 0.5,
[…]

### 2018-10-04 tex/main.tex
Updated figure dirs, names and labels

@@
 \usepackage{array,multirow,graphicx}
-\graphicspath{{./../figures/}}
+\graphicspath{{./../figures/},{./../figures/air-exchange-rate/},{./../figures/kde-iacc-pressure/}}
 \usepackage{lineno}
@@
   \caption{Probability density plots of the deviation from the mean concentration along with the indoor-outdoor pressure difference at the ASU house both before and after the preferential pathway was closed and at North Island NAS.}
-  \label{fig:2d_kde}
-  \includegraphics[width=\textwidth]{pressure_vs_indoor_concentration.pdf}
+  \label{fig:kde-asu-nas}
+  \includegraphics[width=\textwidth]{asu-nas.pdf}
 \end{figure}
 
-Figure \ref{fig:2d_kde} show the result of the statistical analysis of the field data from the ASU and NAS North Island sites.  On the axes, the univariate KDE’s are shown.
+Figure \ref{fig:kde-asu-nas} show the result of the statistical analysis of the field data from the ASU and NAS North Island sites.  On the axes, the univariate KDE’s are shown.
 The y-margin shows the distribution of the log-10 deviation from the mean TCE concentration in indoor air and the x-margin shows the distribution of indoor-outdoor pressure differences.
@@
 When the interior of the building is depressurized relative to ambient, there is a tendency towards indoor air contaminant concentrations higher than the mean, and when the building is pressurized, the values tend to fall below the mean.
-This immediately suggests the examination of bivariate distributions, which are also shown on Figure \ref{fig:2d_kde}, in the center of the plot.
+This immediately suggests the examination of bivariate distributions, which are also shown on Figure \ref{fig:kde-asu-nas}, in the center of the plot.
 
-The center of Figure \ref{fig:2d_kde} shows the bivariate KDEs constructed based on the relationship between the two aforementioned variables.
+The center of Figure \ref{fig:kde-asu-nas} shows the bivariate KDEs constructed based on the relationship between the two aforementioned variables.
 The color-coding indicates the source of the dataset, with green again representing the ASU house dataset after the preferential pathway was closed, red is for before the pathway was closed, and blue the NAS North Island dataset.
 The Pearson’s r-values and p-value for the relation between TCE in indoor air and indoor-outdoor pressure difference for each dataset are also shown.
-The bivariate distributions in Figure \ref{fig:2d_kde} show even more starkly the differences between the ASU house datasets and the NAS North Island dataset.
+The bivariate distributions in Figure \ref{fig:kde-asu-nas} show even more starkly the differences between the ASU house datasets and the NAS North Island dataset.
 We believe that this is primarily due to the significant different indoor-outdoor pressure differences existing at these two sites.
@@
   \caption{Probability density plots of the deviation from the mean Chloroform and Tetrachloroethene concentrations along with the indoor-outdoor pressure difference at the Indianapolis site}
-  \label{fig:2d_kde_indianapolis}
-  \includegraphics[width=\textwidth]{pressure_vs_indoor_concentration_422.pdf}
+  \label{fig:kde-indianapolis}
+  \includegraphics[width=\textwidth]{indianapolis-422.pdf}
 \end{figure}
 
-Figure \ref{fig:2d_kde_indianapolis} shows the same as Figure \ref{fig:2d_kde} but at the Indianapolis site, and the variations of Chloroform and PCE are considered instead of TCE.
+Figure \ref{fig:kde-indianapolis} shows the same as Figure \ref{fig:kde-asu-nas} but at the Indianapolis site, and the variations of Chloroform and PCE are considered instead of TCE.
 The concentration of both Chloroform and PCE are still roughly equally log-normally distributed within half an order of magnitude above and below the mean, similar to the variation observed at the ASU house after the land drain was closed, as well as to the North Island site, if one would consider comparable indoor-outdoor pressure differences.
@@
     \label{fig:asu_exchange_rate_analysis_overview}
-    \includegraphics[width=\textwidth]{tce_indoor_air_exchange_rate_impact.png}
+    \includegraphics[width=\textwidth]{asu-iacc-ae-overview.png}
   \end{subfigure}
@@
     \label{fig:asu_exchange_rate_analysis_violin_plot}
-    \includegraphics[width=\textwidth]{violin_plot_exchange_rate_impact.png}
+    \includegraphics[width=\textwidth]{violin-asu-iacc.png}
   \end{subfigure}
@@
 This is consistent with what was observed at North Island, as there the span of IACC values was roughly the same across this range of indoor-outdoor pressure difference values.
-This simulation, together with the Pearson $r = -0.64$ for North Island in Figure \ref{fig:2d_kde} suggest that under conditions where the soil around a structure is relatively permeable, indoor-outdoor pressure difference can be a significant driving forces for transient changes in IACC.
+This simulation, together with the Pearson $r = -0.64$ for North Island in Figure \ref{fig:kde-asu-nas} suggest that under conditions where the soil around a structure is relatively permeable, indoor-outdoor pressure difference can be a significant driving forces for transient changes in IACC.
 Therefore, recording the indoor-outdoor pressure difference at a site (in particular if the soil is permeable) may offer insight into the potential for transient changes in IACC (some good methods for doing this are suggested by \citeauthor{nielsen_remediation_2017-1}\cite{nielsen_remediation_2017-1}).
@@
 
-\bibliography{Zotero}
+\bibliography{library}
 

### 2018-10-04 tex/main.tex
Wrote about the KDE plots

@@
 \usepackage{array,multirow,graphicx}
+\usepackage{comment}
 \graphicspath{{./../figures/},{./../figures/air-exchange-rate/},{./../figures/kde-iacc-pressure/}}
@@
 \subsection{Drivers For Indoor Air Contaminant Variability}
+% KDE figures
+\begin{figure}[htb!]
+  \caption{KDE analysis of IACC dependence on indoor/outdoor pressure difference at the ASU house and North Island site (\ref{fig:kde-asu-nas}) and the Indianapolis site (\ref{fig:kde-indianapolis}).
+p-values and Pearson's r-values shown for each dataset.}
+  \label{fig:kde-analysis}
+  % ASU/North Island plot
+  \begin{subfigure}{\textwidth}
+    \centering
+    \caption{Period before and after the PP was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
+    \label{fig:kde-asu-nas}
+    \includegraphics[height=0.4\textheight,keepaspectratio]{asu-nas.pdf}
+  \end{subfigure}
+  % Indianapolis plot
+  \begin{subfigure}{\textwidth}
+    \centering
+    \caption{Chloroform and PCE considered separeately at Indianapolis duplex 422.}
+    \label{fig:kde-indianapolis}
+    \includegraphics[height=0.4\textheight,keepaspectratio]{indianapolis-422.pdf}
+  \end{subfigure}
+\end{figure}
+% Introduction
+Out of the factors that influence the IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
+Three well studied VI sites are considered, the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
+% Univariate IACC description
+The IACC at these sites vary significantly, therefore some means of comparing them all to each other is necessary.
+Additionally, the focus in this paper is the driver for variations in IACC, and reprentation of the variations between the different sites must also be achieved, specifically variations on the scale of order of magnitudes is of the greatest interest, as these are the largest hinderance to proper site IACC characterization.
+To achieve these goals, the log-10 deviation from the mean IACC ($\mu$) within each dataset is calculated, e.g. $10\mu$ indicates the IACC is an order of magnitude above the mean IACC for the dataset.
+On the y-margin in figures \ref{fig:kde-asu-nas} and \ref{fig:kde-indianapolis}, the univariate KDE distribution of deviation from the mean may be seen. \par
+% Univariate pressure description
+The indoor/outdoor pressure differences are simply taken as it, and their univariate KDE distributions are shown on the x-margin.
+A negative pressure difference indicate that the building is underpressurized relative to outdoor.\par
+% Bivariate description
+The relationship between the deviation from mean IACC, and indoor/outdoor pressure difference may be seen in the central portion of each figure, in the form of a bivariate KDE distribution.
+The p-value and Pearson's r-value for each bivariate distribution is shown in the legend.\par
+% Sub-figure description (details about sites)
+In Figure \ref{fig:kde-asu-nas} the North Island and ASU house datasets are plotted, with blue representing the North Island site, with red and green representing the ASU house.
+The ASU house dataset is considered in two different parts, the first (red) is data belonging to the time period before the PP was discovered, the second (green) is after the PP was sealed off.
+At both of these sites, TCE was the only contaminant considered.
+The IACC of Chloroform and PCE the 422 part of the Indianapolis duplex are considered in Figure \ref{fig:kde-indianapolis}.\par
+% North Island discussion
+From Figure \ref{fig:kde-asu-nas}, it is apparent that the three datasets differ significantly from each other.
+The North Island site exhibit the greatest variation in both pressure difference and IACC, in fact, looking at the univariate pressure distribution - it is almost entirely flat and spanning from extremly high and low values.
+A likely explation for this is the poor condition of the structure, rendering it highly susceptible to envivornmental influnces.\par
 
+To accompany this, there is significant variation in IACC, that has a relativly strong link with pressure difference ($r = -0.71$), although there is clearly a distinct peak (around $0.5\mu$) that is primally associated with the smaller negative and positive pressure differences ($~-4 \leq \Delta p \leq 8$).
+This demonstrates that at this particular site, the very large variations (an order of magnitude or more) are primarily influenced by the unusually large pressure difference fluctations.
+Likely, this relationship is amplified by the fact that the soil surrounding the structure is sandy and therefore relatively permeable - increasing the influence of pressure differences.
+However, despite all this, pressure differences can only partly explain the variations observed at this site.\par
+% ASU discussion
+The two ASU datasets are not only significantly different from the North Island dataset, but also different from each other.
+Consider the dataset before the PP was discovered (red) and one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit rarely.)
+This is more variation than was observed at North Island, and yet the distribution of pressure differences is narrow, mostly varying by a few Pa around zero (~$\pm 2$ Pa), and with $r = -0.50$, suggests that pressure difference is not insignificant but still a weak predictor for the variations in IACC.\par
+
+The picture changes completely after the PP was closed off, where a similar distribution of pressure differences is observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
+This clearly shows just how much variance the PP at this site contributed with, and how critical it is to assess whether a PP is present at a particular site in general, lest potentially face unacceptable uncertainty in determining the relevant contaminant exposure.
+It is also clear that in the absence of a PP, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are here insignficant in determining variation in IACC.
+This also suggests that the PP at the ASU house acted not only as a PP for contaminant vapors, but for air in general - allowing the impermeable soil surrounding the house to be circumvented.
+PPs, and the factors that influnce them will be explored later in this article.\par
+% Indianapolis discussion
+The Indianapolis duplex also featured a PP, but despite this, there is significantly less variance in IACC of both Chloroform and PCE.
+The exact reason for this is not known, but considering that the PP at the Indianapolis site was a sanitary sewer, and at ASU the PP was a foundation drain system, it is not inconceivable that their dynamics are quite different.
+Nevertheless, both PCE and Chloroform are more or less log-normally distributed, with (mostly) less than an order of magnitude variance around the mean, indicating that the varation in IACC are species independent.\par
+
+The pressure difference distribution exhibit some bimodal characteristics, with a larger variation than was observed at the ASU site.
+This may explain why the r-values for both PCE and Chloroform, $r=-0.51$ and $r=-0.53$ respectively, are so similar to the ASU site before the PP there was closed; the larger pressure differences, which occured more often than ASU, increase contaminant entry rates to the extent that a PP was needed to "achieve" at the ASU site.\par
+% Summary discussion
+All of these data clearly suggest that there is significant variation in IACC that cannot be easily explained by variations in indoor/outdoor pressure difference - even with a PP.
+It is also clear that PP may contribute with unacceptable levels of variations, and screening for PPs should be part of any site investiation.
+It may also be prudent, even in the absence of PP, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air.
+E.g. if a single sample is less than an order of magnitude from some target limit, it may be justified to perform subsequent samples to better assess the relevant exposure.
+These data also call the role of advection in VI into question, as "normal" pressure difference ranges are such a poor predictor of IACC variation, which is the opposite of what one would expect if advection was a dominant transport mechanism.\par
+
+\begin{comment}
 \subsubsection{The Role Of Building Depressurization/Pressurization}
@@
 This may be evidence of the interaction with the PP that exists at the site, although the extreme variations in IACC cannot be observed here.
+\end{comment}
 
 \subsubsection{The Role Of Air Exchange Rate}
-
-% Plots showing the relative minor contribution to TCE in indoor air variance fluctuating air exchange rate has
+% Air exchange rate plots
 \begin{figure}[htb!]
   \caption{Comparison between the recorded  and the calculated TCE in indoor air at the ASU house, assuming constant air exchange rate.
-\ref{fig:asu_exchange_rate_analysis_overview} shows the TCE in indoor air across time as well as the exchange rate.
-\ref{fig:asu_exchange_rate_analysis_violin_plot} shows the distribution of these values for three periods.}
-  \label{fig:asu_exchange_rate_analysis}
+\ref{fig:asu-iacc-ae-overview} shows the TCE in indoor air across time as well as the exchange rate.
+\ref{fig:violin-iacc-aet} shows the distribution of these values for three periods.}
+  \label{fig:ae-analysis}
   % "Overview" plot
@@
     \caption{ }
-    \label{fig:asu_exchange_rate_analysis_overview}
+    \label{fig:asu-iacc-ae-overview}
     \includegraphics[width=\textwidth]{asu-iacc-ae-overview.png}
@@
     \caption{ }
-    \label{fig:asu_exchange_rate_analysis_violin_plot}
+    \label{fig:violin-iacc-aet}
     \includegraphics[width=\textwidth]{violin-asu-iacc.png}
   \end{subfigure}
-\end{figure}
-
-Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion.
+\end{figure}\
+Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion, and another highly dynamic process.
 Exploring the impact of fluctuations in air exchange rate is thus also important for the development of sampling strategies.
@@
 
-In Figure \ref{fig:asu_exchange_rate_analysis_overview}, the measured IACC and air exchange rate across time are shown, with the predicted IACC assuming a constant air exchange rate.
+In Figure \ref{fig:asu-iacc-ae-overview}, the measured IACC and air exchange rate across time are shown, with the predicted IACC assuming a constant air exchange rate.
 The median air exchange rate value for the pre-CPM period was chosen as this seemed to be the most representative value for this time period.
-From Figure \ref{fig:asu_exchange_rate_analysis_overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC assuming constant and actual fluctuating exchange rates are quite similar.
+From Figure \ref{fig:asu-iacc-ae-overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC assuming constant and actual fluctuating exchange rates are quite similar.
 In other words, in this instance it is unlikely that the variations in IACC were driven by fluctuations in air exchange rate.
 
-This is more clearly shown in Figure \ref{fig:asu_exchange_rate_analysis_violin_plot}, where KDEs of the IACC are constructed assuming the constant and fluctuating air exchange rate cases.
+This is more clearly shown in Figure \ref{fig:violin-iacc-aet}, where KDEs of the IACC are constructed assuming the constant and fluctuating air exchange rate cases.
 This figure shows the relative probability of being at a particular TCE in indoor air concentration at the ASU house.

### 2018-10-04 python/kde-iacc-pressure/indianapolis.py
Changed y-axis label

@@ -1,0 +1,203 @@
+import numpy as np
+import sqlite3
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from scipy import stats
+import matplotlib.patches as mpatches
+
+
+
+class Observation:
+    def __init__(self,select,where,grouby):
+        return
+    def opt_select(self,value):
+        return
+class IndoorAir:
+    def __init__(self,method,compound,location,groupby):
+        return
+
+class Indianapolis:
+    def __init__(self):
+        # loads database
+        self.db = sqlite3.connect('/home/jonathan/lib/vapor-intrusion-dbs/indianapolis.db')
+        #self.db = sqlite3.connect('D:\Dropbox\Research\indianapolis.db')
+
+        return
+
+    def ssd(self,status='not yet installed'):
+        select = "O.StopDate AS StopDate"
+        table = "Observation_Status_Data AS O"
+        where = "O.Value = \"%s\"" % status
+        return select, table, where
+
+    def pressure(self,location='422',type='Basement.Vs.Exterior',avg=True):
+        if avg is True:
+            select = "AVG(P.Value) AS Pressure"
+        else:
+            select = "P.Value AS Pressure"
+        table = "Differential_Pressure_Data AS P"
+        where = "P.Location = \"%s\"" % location + " AND " + "P.Variable = \"%s\"" % type
+
+        return select, table, where
+
+    def concentration(self,location='422BaseS',compound='Chloroform',avg=True):
+        if avg is True:
+            select = "AVG(C.Value) AS Concentration"
+        else:
+            select = "C.Value AS Concentration"
+        table = "VOC_Data_SRI_8610_Onsite_GC AS C"
+        where = "C.Location = \"%s\"" % location + " AND " + "C.Variable = \"%s\"" % compound
+
+        return select, table, where
+
+
+
+    def sql_query(self,methods):
+        select = "SELECT "
+        tables = "FROM "
+        wheres = "WHERE "
+
+
+        for method in methods:
+            select += method[0] + ", "
+            tables += method[1] + ", "
+            wheres += method[2] + " AND "
+
+        select = select.rstrip(", ")
+        tables = tables.rstrip(", ")
[…]

### 2018-10-04 tex/main.tex
Fixed spelling errors

@@
 \subsection{Drivers For Indoor Air Contaminant Variability}
+\subsubsection{The Role Of Building Depressurization/Pressurization}
 % KDE figures
@@
 % Introduction
-Out of the factors that influence the IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
-Three well studied VI sites are considered, the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
+Out of the factors that influence IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
+Three well-studied VI sites are considered, the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
 % Univariate IACC description
-The IACC at these sites vary significantly, therefore some means of comparing them all to each other is necessary.
-Additionally, the focus in this paper is the driver for variations in IACC, and reprentation of the variations between the different sites must also be achieved, specifically variations on the scale of order of magnitudes is of the greatest interest, as these are the largest hinderance to proper site IACC characterization.
+The absolute IACC at these sites vary significantly, therefore some means of comparing them all to each other is necessary.
+Additionally, the focus in this section is the driver for variations in IACC, and representation of the variations between the different sites must also be achieved, specifically variations on the scale of order of magnitudes is of the greatest interest, as these are the largest hinderance to proper site IACC characterization.
 To achieve these goals, the log-10 deviation from the mean IACC ($\mu$) within each dataset is calculated, e.g. $10\mu$ indicates the IACC is an order of magnitude above the mean IACC for the dataset.
@@
 % Sub-figure description (details about sites)
-In Figure \ref{fig:kde-asu-nas} the North Island and ASU house datasets are plotted, with blue representing the North Island site, with red and green representing the ASU house.
-The ASU house dataset is considered in two different parts, the first (red) is data belonging to the time period before the PP was discovered, the second (green) is after the PP was sealed off.
+In Figure \ref{fig:kde-asu-nas} the North Island and ASU house datasets are plotted, with blue representing the North Island site, and with red and green representing the ASU house.
+The ASU house dataset is split into two different parts, the first (red) is data from the time period before the PP was discovered, the second (green) is from after the PP was sealed off.
 At both of these sites, TCE was the only contaminant considered.
-The IACC of Chloroform and PCE the 422 part of the Indianapolis duplex are considered in Figure \ref{fig:kde-indianapolis}.\par
+The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex are considered in Figure \ref{fig:kde-indianapolis}.\par
 % North Island discussion
@@
 \begin{comment}
-\subsubsection{The Role Of Building Depressurization/Pressurization}
 

### 2018-10-08 tex/supporting-information/esthag_supporting.tex
Added supporting information

\documentclass[journal=esthag,manuscript=suppinfo]{achemso}

% additional packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{array,multirow,graphicx}
\graphicspath{{../../Figures/}}
% macros

% authors
\author{Jonathan G. V. Ström}
\affiliation[Brown University]{Brown University, School of Engineering}
\author{Yijun Yao}
\affiliation[Zhejiang University]{Zhejiang University}
\author{Eric M. Suuberg}
\email{eric_suuberg@brown.edu}
\affiliation[Brown University]{Brown University, School of Engineering}

% title
\title{Modeling the Impact of Preferential Contaminant Pathways on Vapor Intrusion}

% keywords
\abbreviations{VI}
\keywords{Vapor intrusion, Preferential pathways}

\begin{document}

\subsection{Supporting Information Content:}

\begin{itemize}
  \item x pages: S1 - S6
  \item 4 figures: Figure S1 - Figure S4
  \item 0 tables
\end{itemize}

\section{Method}
The van Genuchten equation\cite{Genuchten1980a}:
\begin{equation}
  \mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m} \label{eq:saturation}
\end{equation}
Se is the soil saturation, $\theta_w$, $\theta_r$, $\theta_t$ are the water filled, residual and total porosity of the soil.
The vapor or gas filled porosity may be calculated by $\theta_g = \theta_t - \theta_w$
The local soil saturation depends on the parameters $\alpha$, $n$, $m= 1 - 1/n$ and the elevation above groundwater $z$.

Convection-diffusion equation in porous media:
\begin{equation}
  \frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c \label{eq:conv-diff}
\end{equation}
Where $c_w$ and $c$ are the liquid phase and vapor phase contaminant concentration, $D_\mathrm{eff}$ is
the effective diffusion coefficient and $\vec{u}$ is the soil gas velocity, as determined by Darcy’s Law.
It is important to recall that the soil gas flow velocity is not significantly influenced by the contaminant itself, because the contaminant concentrations are normally so low that the soil gas is largely air.
The above equation is presented in an unsteady form, reflecting the fact that there may be holdup of the contaminant vapor in both the vapor phase as well as in the soil moisture phase. \par

Richards' equation\cite{Richards1931a}:
\begin{equation}
  \nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0 \label{eq:richards-equation}
\end{equation}
The soil gas flow is driven by a pressure gradient $\nabla p$ and is proportional to the modified permeability of the soil $k_r \kappa_s$ and inversely proportional to the vapor viscosity $\mu$.
The modified permeability is the saturated fluid permeability $\kappa_s$ corrected using the relative permeability $k_r$ defined by:
\begin{equation}
  k_r = (1 - \mathrm{Se})^{0.5} [1 - (\mathrm{Se}^{-m})^m]^2 \label{eq:k_r}
\end{equation}

Millington-Quirk equation for effective diffusion coefficient in porous media\cite{Millington1961a}.
\begin{equation}
  D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2} \label{eq:millington-quirk}
\end{equation}
$D_\mathrm{air}$ and $D_\mathrm{water}$ are the diffusion coefficient of the contaminant in air and water respectively and $K_H$ is the dimensionless Henry’s Law constant. \par

The boundary condition for contaminant transport through the foundation crack:
\begin{align}
  j_\mathrm{ck} &= u_\mathrm{ck} c_\mathrm{ck} - \frac{u_\mathrm{ck} (c_\mathrm{ck} - c_\mathrm{in})}{1 - \exp{(u_\mathrm{ck}L_\mathrm{slab}/D_\mathrm{air})}} \label{eq:j_ck} \\
  n_\mathrm{ck} &= \int_{A_\mathrm{ck}} j_\mathrm{ck} dA \label{eq:n_ck}
\end{align}
Where $u_\mathrm{ck}$ and $c_\mathrm{ck}$ are the flow velocity and contaminant concentration at the soil side of the crack, respectively, $L_\mathrm{slab}$ is the thickness of the foundation slab (here, 15 cm), $D_\mathrm{air}$ is the diffusion coefficient of the contaminant in air, $j_\mathrm{ck}$ is the flux in the crack and $A$ represents the crack cross-sectional area.

The assumed values for the various parameters mentioned above are shown in Table \ref{tbl:data}.
\begin{table}[!htb]
  \centering
  \caption{Contaminant and soil properties\cite{Abreu2012a,U.S.-Environmental-Protection-Agency2004a}.} \label{tbl:data}
  \begin{tabular}{l r r r r r r}
  \toprule
    Soil & $\text{Permeability} \; [\mathrm{m^2}]$  & $\mathrm{Density} \; [\mathrm{kg/m^3}]$  & $\theta_s$  & $\theta_r$  & $\alpha \; [\mathrm{1/m}]$  & $n$ \\
    \hline
    Gravel\cite{Dan2012a}     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
    Sand          & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053       & 3.5       & 3.2 \\
    Loamy Sand    & $1.6 \cdot 10^{-12}$  & 1430    & 0.39        & 0.049       & 3.5       & 1.7 \\
    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
                  &                       &         &             &             &           & \\
    \multicolumn{7}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
                  &                       &         &             &             &           & \\
    \multicolumn{7}{c}{Contaminant: Trichloroethylene (Diluted in air)} \\
    \multicolumn{1}{r}{$D_\mathrm{air} \; [\mathrm{m^2/h}]$}  & $D_\mathrm{water} \; [\mathrm{m^2/h}]$  & $\mathrm{Density} \; [\mathrm{kg/m^3}]$ & $\mathrm{Viscosity} \; [\mathrm{Pa \cdot s}]$  & $K_H$ & $M \; [\mathrm{g/mol}]$ &  \\
    \hline
    \multicolumn{1}{r}{$2.47 \cdot 10^{-2}$}  & $3.67 \cdot 10^{-6}$  & 1.614 & $1.86 \cdot 10^{-5}$  & 0.403 & 131.39  &  \\
    \bottomrule
    \multicolumn{7}{c}{$\theta_t=$ soil total porosity, $\theta_g=$ gas filled porosity, $\theta_w=$ water filled porosity}
  \end{tabular}
\end{table}

\subsection{Transient Behavior Under "Normal" VI Scenarios}

If the air exchange rate $A_e$ is varied sinusoidally across a day between the high and low values as are realistic in a residential setting, one may explore the potential impact of this variable on transient changes in $c_\mathrm{in}$ while keeping $n_\mathrm{ck}$ constant.
The range of air exchange rate values was taken to be roughly the same as was observed at the ASU house, where air exchange rates varied from 5 to 24 per day with the majority of the time value being between 10 to 20 per day\cite{Holton2013a,Holton2015a}, this is relatively representative of normal residential houses\cite{Murray1995a}.
As may be seen in Figure \ref{fig:cstr_ae}, $c_\mathrm{in}$ responds quickly with a change in $A_e$, but the change is significantly lower than the few orders of magnitude that has been observed to occur at the ASU house on timescales of a day.
The modest magnitude of the change is consistent with earlier conclusions on this point\cite{Shen2016a}. \par

Now instead $n_\mathrm{ck}$ may be varied and $A_e$ kept constant.
It may be seen in Figure \ref{fig:cstr_nck} that $c_\mathrm{in}$ quickly responds to changes in $n_\mathrm{ck}$ and the increase in $c_\mathrm{in}$ is proportional to the increase in $n_\mathrm{ck}$, indicating, not surprisingly, that steep changes in this parameter can be the driving force for increases in $c_\mathrm{in}$. \par

\begin{figure}[!htb]
  \caption[Varying air exchange and contaminant entry rate impact on indoor air concentration]{Varying inputs of the indoor air contaminant concentration calculation. (\subref{fig:cstr_ae}) considers the impact of varying air exchange rate with constant contaminant entry rate. (\subref{fig:cstr_nck}) considers the impact of varying contaminant entry rate with constant air exchange rate.}
  \begin{subfigure}[b]{0.9\textwidth}
    \caption{ } \label{fig:cstr_ae}
    \includegraphics[width=\textwidth]{cstr_ae.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.9\textwidth}
    \caption{ } \label{fig:cstr_nck}
    \includegraphics[width=\textwidth]{cstr_nck.png}
  \end{subfigure}
\end{figure}

In Figure \ref{fig:stack_effect} the effects of building pressurization/depressurization are considered.
In order to maximize the impact of pressure swings, the building is assumed to have a gravel sub-base and to be surrounded by sand soil.
For an already depressurized building, further depressurization only moderately increases $c_\mathrm{in}$.
Overpressurization of a building however leads to significantly lower $c_\mathrm{in}$, as contaminant entry rates are significantly decreased.
However, the full extent of this decrease is not shown in the figure; the y-axis was purposely truncated in order to highlight the effects of increased depressurization. \par

\begin{figure}[!htb]
  \centering
  \caption[Effects of house depressurization]{A side view of steady-state concentration profiles of "normal" VI scenarios featuring a gravel sub-base with sand soil, with different stack effect magnitudes. Color indicate contaminant vapor mole fraction and are logarithmic. The attenuation from groundwater source is given for each figure.} \label{fig:stack_effect}
  % top figure
  \begin{subfigure}{0.9\textwidth}
    \centering
    \caption{ } \label{fig:stack_effect1}
    \includegraphics[width=\textwidth]{c_in_p_ck.png}
  \end{subfigure}
  % left bottom digure
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \caption{-5 Pa, $\alpha = 9.1 \cdot 10^{-6}$} \label{fig:stack_effect2}
    \includegraphics[width=\textwidth]{soil_concentration_right_side1.png}
  \end{subfigure}
  % right bottom figure
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \caption{-20 Pa, $\alpha = 1.6 \cdot 10^{-5}$} \label{fig:stack_effect3}
    \includegraphics[width=\textwidth]{soil_concentration_right_side2.png}
  \end{subfigure}
\end{figure}

Figure \ref{fig:diffusion} illustrates the effective diffusion coefficient for the three different soil types used, ranging from soils allowing for faster diffusion (sand) to soils that show slower vapor diffusion (sandy clay).
Notice that diffusion is the slowest through the water filled soil, such as just above groundwater, in the capillary fringe, providing a significant resistance to contaminant vapor diffusion.
The characteristic time scale for transport by diffusion is given by $L^2/D$, where $L$ is the characteristic length scale, which in this case is of order meters (based on the source-slab separation).
Even with only a 1 m separation and a high $D_\mathrm{eff} = 10^{-6} \; \mathrm{m^2/s}$, response times can only be be as short as weeks.
Therefore, fluctuations in groundwater contaminant concentrations cannot be the underlying cause of the observed rapid transients in indoor air contaminant concentrations. \par

\begin{figure}[!htb]
  \caption[Soil moisture \& contaminant transport speed through soil]{In (\subref{fig:soil_properties}) the van Genuchten soil moisture content of various soils as a function of elevation above groundwater along with the corresponding effective diffusion of TCE in those soils. The diffusive propagation of TCE through sand soil is shown in (\subref{fig:contaminant_propagation})}\label{fig:diffusion}
  \centering
  \begin{subfigure}[t]{0.65\textwidth}
    \centering
    \caption{ } \label{fig:soil_properties}
    \includegraphics[width=\textwidth]{d_eff.png}
  \end{subfigure}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \caption{ } \label{fig:contaminant_propagation}
    \includegraphics[width=\textwidth]{c_soil_transient.png}
  \end{subfigure}
\end{figure}

\subsection{The Preferential Pathway as a Source of Contaminant Beneath a Structure}

Air flow from the preferential pathway is important for the distribution of vapor contaminant in the sub-base and the soil, and ultimately for the entry rate of contaminant into a building.
To highlight this, a scenario of a house with a gravel sub-base, surrounded by loamy sand soil with contaminated preferential pathway that exits into the gravel sub-base is considered.
This scenario is solved at steady-state twice - first with both advection-diffusion enabled (like all other calculations presented) and the second with advection disabled, i.e. diffusion only. \par

The result of this may be seen in Figure \ref{fig:adv-diff_vs_diff}, where (\subref{fig:advection_diffusion}) shows the advection-diffusion result, i.e. the regular result and (\subref{fig:diffusion_only}) the diffusion only result.
Just from observation it is clear that there is a massively different soil contaminant vapor concentration profile, showing that existence of a diffusing preferential pathway will have a very modest impact on VI.
Contaminant needs to be blown in to have a major effect.

\begin{figure}[!htb]
  \centering
  \caption[Importance of advection for the impact of a preferential pathway]{The impact of advection for the distribution of contaminant vapor from a preferential pathway. The soil vapor contaminant concentration in the loamy sand soil and gravel sub-base, a short distance away from the preferential pathway shown. (\subref{fig:advection_diffusion}) shows the regular advection and diffusion case and (\subref{fig:diffusion_only}) shows the same scenario but advection disabled.} \label{fig:adv-diff_vs_diff}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \caption{Advection enabled} \label{fig:advection_diffusion}
    \includegraphics[width=\textwidth]{steady-state_preferential_pathway2.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \caption{Advection disabled} \label{fig:diffusion_only}
    \includegraphics[width=\textwidth]{steady-state_preferential_pathway_diff_only2.png}
  \end{subfigure}
\end{figure}

\newpage
\bibliography{../library}

\end{document}

### 2018-10-08 python/preferential-pathway-sensitivity/sensitivity-analysis.py
Finished new graph, comparing sensitivity-modeling data to field

@@ -1,0 +1,102 @@
+import numpy as np
+import pandas as pd
+import pandas.tools.plotting as pdplt
+import matplotlib.pyplot as plt
+from matplotlib import cm
+import seaborn as sns
+import sqlite3
+data_dir = './data/preferential-pathway-sensitivity/'
+db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
+
+"""
+Prepares model data
+"""
+# no PP data
+no_pp = pd.read_csv(
+    data_dir + 'param-no-preferential-pathway.csv',
+    header=4,
+)
+no_pp['PP'] = pd.Series(np.repeat('No', len(no_pp['p'])))
+# PP data
+pp = pd.read_csv(
+    data_dir + 'param-preferential-pathway.csv',
+    header=4,
+)
+pp['PP'] = pd.Series(np.repeat('Yes', len(no_pp['p'])))
+df = pd.concat([pp, no_pp],axis=0,sort=True)
+# adds friendlier tags for gravel sub-base present (SB) and if the PP is
+# contaminated (chi)
+df['Gravel sub-base'] = pd.Series(df['SB'].apply(lambda x: 'Yes' if x == 1 else 'No'))
+df['Contaminated PP'] = pd.Series(df['chi'].apply(lambda x: 'Yes' if x == 1 else 'No'))
+# data from permeable model, unknown if I want to add it,
+# might be able to concat at it and assign SB = Sand or whatever
+permamble = pd.read_csv(
+    data_dir + 'param-permeable-soil.csv',
+    header=4,
+)
+"""
+Prepares ASU data for comparison 
+"""
+db_asu = sqlite3.connect(db_dir + 'asu_house.db')
+asu = pd.read_sql_query(
+    "SELECT \
+        day, \
+        pressure_difference AS Pressure, \
+        tce_emission_rate, \
+        building_flow_rate, tce_groundwater \
+    FROM \
+        parameters; \
+    ", db_asu,
+).interpolate(limit_area='inside',limit=250).dropna()
+# removes data that is outside model data range
+asu['Pressure'] = -1.0*asu['Pressure']
+asu = asu[(asu['Pressure'] < 10) & (asu['Pressure'] > -10)]
+# removes cpm period from dataset
+cpm_start, cpm_end = 780.0, 1157.0
+asu = asu[(asu['day'] < cpm_start) | (asu['day'] > cpm_end)]
[…]

### 2018-10-08 python/kde-iacc-pressure/nas-asu-house.py
Added option to plot pressure deviation from mean (still bit buggy)

@@ -5,11 +5,12 @@ import matplotlib.pyplot as plt
 import sqlite3
 from scipy import stats
 import matplotlib.patches as mpatches
+db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
 
 
 # databases
-db_asu = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/asu_house.db')
-db_nas = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/north_island_nas.db')
+db_asu = sqlite3.connect(db_dir+'asu_house.db')
+db_nas = sqlite3.connect(db_dir+'north_island_nas.db')
 
 # loads asu data, interpolates and drops NAs
 asu = pd.read_sql_query("SELECT day, pressure_difference AS Pressure, tce_emission_rate, building_flow_rate FROM parameters;", db_asu).interpolate(limit_area='inside',limit=250).dropna()
@@ -33,80 +34,85 @@ dfs = []
 alpha = 0.6
 cmaps = ["Blues", "Reds", "Greens"]
 
-df = pd.DataFrame({'x': [], 'y': []})
-g = sns.JointGrid(x="x", y="y", data=df) # empty joint grid
+
 label_patches = []
-for i, df in enumerate([nas, asu_pre_cpm, asu_post_cpm]):
-    print("Dataset: %s" % df.name)
 
-    df['Concentration'] = df['Concentration'].apply(pd.to_numeric, errors='ignore').apply(lambda x: np.log10(x)).replace([np.inf, -np.inf], np.nan).dropna()
-    df['Concentration'] -= df['Concentration'].mean()
-    df['Pressure'] = -1.0*df['Pressure'].dropna()
+for dp in [False]:
+    df = pd.DataFrame({'x': [], 'y': []})
+    g = sns.JointGrid(x="x", y="y", data=df) # empty joint grid
+    for i, df in enumerate([nas, asu_pre_cpm, asu_post_cpm]):
+        print("Dataset: %s" % df.name)
 
-    #if df.name is "ASU House, PP open":
-    #    df = df.loc[df['Pressure'] > -7.5]
-    #    df.name = "ASU House, PP open"
-
-    r, p = stats.pearsonr(
-        df['Pressure'],
-        df['Concentration'],
-    )
-    # bivariate plot
-    sns.kdeplot(
-        df['Pressure'],
-        df['Concentration'],
-        cmap=cmaps[i],
-        shade=True,
[…]

### 2018-10-08 python/transient-response/param-period.py
Added param period plot (wip)

@@ -1,0 +1,8 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+
+df = pd.read_csv('./data/transient-response/param-period.csv', header=4)
+
+print(df.pivot( columns='period', values='Attenuation factor'))

### 2018-10-12 python/air-exchange-rate/asu_exchange_rate_analysis.py
Updated figures to .pdf instead of .png

@@ -2,7 +2,6 @@ import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
 import sqlite3
-from natu import units
 import seaborn as sns
 
 # returns the error between the predicted and measured TCE in indoor air
@@ -94,7 +93,7 @@ def plot_comparison(data,title=None,save_figure=False):
     plt.tight_layout()
     # if a figure name is given, the figure is saved
     if isinstance(save_figure, str):
-        fig.savefig('../Figures/%s.png' % save_figure, dpi=300)
+        fig.savefig('./figures/air-exchange-rate/%s.pdf' % save_figure, dpi=300)
         return
     # otherwise it will just show the plot
     else:
@@ -102,7 +101,7 @@ def plot_comparison(data,title=None,save_figure=False):
         return
 
 # connects to the database
-db = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/asu_house.db')
+db = sqlite3.connect('/home/jonathan/lib/vapor-intrusion-dbs/asu_house.db')
 data = pd.read_sql_query("SELECT * FROM parameters;", db)
 data = data.interpolate(limit_area='inside',limit=250) # 250 limit for pretty "real" dataset, 1500 for fuller plot
 
@@ -116,9 +115,11 @@ data = get_error(data)
 error_table = get_error_table(data)
 
 M = 131.4
-data['tce_indoor_air_estimation'] *= M*units.g/units.ug # ug/m3
-data['tce_indoor_air_prediction'] *= M*units.g/units.ug # ug/m3
-data['exchange_rate'] *= units.d/units.s # 1/day
+g_to_ug = 1e6
+s_to_day = 3600.0*24.0
+data['tce_indoor_air_estimation'] *= M*g_to_ug # ug/m3
+data['tce_indoor_air_prediction'] *= M*g_to_ug # ug/m3
+data['exchange_rate'] *= s_to_day # 1/day
 
 
 
@@ -186,5 +187,5 @@ ax.set_yticklabels(['0.001', '0.01', '0.1', '1.0', '10.0', '100.0'])
 
 ax.set_title('Fluctating vs. constant air exchange rate\'s impact\non distribution of TCE in indoor air')
 
-plt.savefig('../Figures/violin_plot_exchange_rate_impact.png', dpi=300)
+plt.savefig('./figures/air-exchange-rate/violin-asu-iacc.pdf', dpi=300)
 fig.clf()

### 2018-10-12 tex/main.tex
Added new figures

@@
 \usepackage{comment}
-\graphicspath{{./../figures/},{./../figures/air-exchange-rate/},{./../figures/kde-iacc-pressure/}}
+\graphicspath{
+  {./../figures/},
+  {./../figures/air-exchange-rate/},
+  {./../figures/kde-iacc-pressure/},
+  {./../figures/preferential-pathway-sensitivity/},
+  {./../figures/transient-response/}
+}
 \usepackage{lineno}
@@
     \label{fig:kde-asu-nas}
-    \includegraphics[height=0.4\textheight,keepaspectratio]{asu-nas.pdf}
+    \includegraphics[height=0.4\textheight,keepaspectratio]{nas-asu-house-iacc-deviating.pdf}
   \end{subfigure}
@@
   \end{subfigure}
-\end{figure}\
+\end{figure}
 Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion, and another highly dynamic process.
@@
 \subsubsection{Conclusions From Steady-State Modeling}
+
+\begin{figure}[htb!]
+  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for a model featuring a PP (\ref{fig:ss-sensitivity-analysis-pp}) and without a PP (\ref{fig:ss-sensitivity-analysis-no-pp}).
+Results compared with field data from ASU house.}
+  \label{fig:ss-sensitivity-analysis}
+  % "Overview" plot
+  \begin{subfigure}{0.55\textwidth}
+    \caption{PP present. Sensitivity to the presence of a gravel sub-base and contamination in the PP considered.}
+    \label{fig:ss-sensitivity-analysis-pp}
+    \includegraphics[width=\textwidth]{yes-pp.pdf}
+  \end{subfigure}
+  % Violinplot figure
+  \begin{subfigure}{0.55\textwidth}
+    \caption{No PP preseent. Sensitivitiy to the presence of a gravel sub-base considered.}
+    \label{fig:ss-sensitivity-analysis-no-pp}
+    \includegraphics[width=\textwidth]{no-pp.pdf}
+  \end{subfigure}
+\end{figure}
+
+\begin{comment}
 
@@
 
+\end{comment}
+
 \begin{acknowledgement}

### 2018-10-12 tex/main.tex
Updated text

@@
     \label{fig:asu-iacc-ae-overview}
-    \includegraphics[width=\textwidth]{asu-iacc-ae-overview.png}
+    \includegraphics[width=\textwidth]{tce_indoor_air_exchange_rate_impact.pdf}
   \end{subfigure}
@@
     \label{fig:violin-iacc-aet}
-    \includegraphics[width=\textwidth]{violin-asu-iacc.png}
+    \includegraphics[width=\textwidth]{violin-asu-iacc.pdf}
   \end{subfigure}
@@
 \begin{figure}[htb!]
-  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for a model featuring a PP (\ref{fig:ss-sensitivity-analysis-pp}) and without a PP (\ref{fig:ss-sensitivity-analysis-no-pp}).
+  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for cases featuring a PP (\ref{fig:ss-sensitivity-analysis-pp}) and without a PP (\ref{fig:ss-sensitivity-analysis-no-pp}).
 Results compared with field data from ASU house.}
@@
   \begin{subfigure}{0.55\textwidth}
-    \caption{No PP preseent. Sensitivitiy to the presence of a gravel sub-base considered.}
+    \caption{No PP present. Sensitivitiy to the presence of a gravel sub-base considered.}
     \label{fig:ss-sensitivity-analysis-no-pp}
@@
 \end{figure}
+
+Clearly, PPs pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover; even at a well studied site like the ASU house it took years to discover the PP.
+Therefore, there is a need to consider other indications that a PP is present or a potential issue.\par
+
+To investigate this, inspiration is taken from the ASU house, where there clearly was a PP containing contaminant vapor, which exited into a gravel (permeable) sub-base.
+The impact of these features is simulated by examining the IACC at different indoor/outdoor pressure differences for each combination of cases.
+Instead of representing IACC as absolute concentration, we now non-dimensionalize with respect to the vapor in equilibrium with the groundwater contaminant concentration.
+This non-dimensionalized property is commonly called the attenuation factor and is denoted by $\alpha_\mathrm{gw}$.
+The result of a steady-state sensitivity analysis relative to several factors is seen in Figure \ref{fig:ss-sensitivity-analysis}.\par
+
+Figure \ref{fig:ss-sensitivity-analysis-pp} deal with cases where a PP is present and, the impact of having a gravel vs.
+soil sub-base and vapor contaminant vs.
+clean air in the PP ($\chi=1$ and $\chi=0$ respectively) are considered.
+Figure \ref{fig:ss-sensitivity-analysis-no-pp} considers the cases absent a PP, for reference, but still considers the impact of a gravel vs. soil sub-base.
+To validate these results, the IACC at the ASU house (as $\alpha_\mathrm{gw}$) vs. indoor/outdoor pressure difference are also plotted.
+Only data from period before the PP was discovered, and after the PP was closed are considered in each respective figure (\ref{fig:ss-sensitivity-analysis-pp} and \ref{fig:ss-sensitivity-analysis-no-pp}).
+To improve visibility of the ASU data, it's average $\alpha_\mathrm{gw}$ are aggregated into up to 40 evenly spaced bins on the x-axis.\par
+
+Begining with Figure \ref{fig:ss-sensitivity-analysis-pp} it immediately clear that absent a gravel sub-base, the impact of a PP is minimal, as the orange and blue lines show.
+Since in the model, it is assumed that the soil type is sandy clay, which is relatively impermeable, there is simply too much resistance to advection for changes in pressurization to matter.
+Additionally, whether the PP is filled with contaminant vapors or not does matter; without a permeable sub-base, the transport of contaminant vapor in the sub-base is too inhibited to be effectively transported into the building.\par
+
+Moving on to the case of when a gravel sub-base is present, and the PP is filled with contaminant vapors ($\chi=1$).
+In this scenario the PP has a massive impact on $\alpha_\mathrm{gw}$, spanning more than four orders of magnitude.
+This is explained by the increased role of advective transport under these conditions.
+The gravel sub-base allows vapor to flow more easily, and the PP is able to circumvent the large resistance to transport due to the surrounding soil, realizing the advective potential offered by the gravel sub-base.
+When the building is underpressurized under these conditions, a significant amount of contaminant vapor is able to be pulled from the PP, leading to the large increase of $\alpha_\mathrm{gw}$.\par
+
+On the other hand, due to the increased advective potential, overpressurization more effectively prevents contaminant vapors from entering the structure, and most of the contaminant entering the building does so through diffusion.
+This also explains why there is no difference between the uncontaminted and contaminted PP cases (red and green) when the building is overpressurized.
+But these two cases (red and green) differ signficant when the building is underpressurized, obviously because there simply is no additional contaminant being supplied by the PP, $\alpha_\mathrm{gw}$ thus limits to whatever diffuses from the contaminted groundwater source.\par
+
+It is interesting to note that $\alpha_\mathrm{gw}$ is higher with an uncontaminted PP (green) than the cases without a gravel sub-base.
+One could expect that pulling so much clean air into the sub-base region would significantly dilute the contaminant vapors and if anything cause a decrease in $\alpha_\mathrm{gw}$ relative to blue and orange.
+However, the PP is small relative to the sub-base (10 cm diameter pipe), thus impacting the contaminant concentration in a limited region of the sub-base, but still allows for higher advection in the entire sub-base region, leading to a $\alpha_\mathrm{gw}$ plateau that is limited by contaminant transport from the groundwater source.\par
+
+These case most resembling the ASU site (with an open PP) is the red case, and fits the data fairly well, especially for small under-/overpressurization values.
+The most significant deviations occur as the under-/overpressurization becomes larger.
+However, the model is not intended to replicate the ASU data perfectly, but rather investigate the factors make a PP the most impactful in a general sense, and perfect simulation is not to be expected.
+This is especially reflected in the choice of vapor contaminant concentration in the PP, with it either being equal to the groundwater contaminant vapor concentration ($\chi=1$) or clean ($\chi=0$), neither of which is likely to be true at the ASU site.
+The idea with picking $\chi=1$ and $\chi=0$ is to give a span of $\alpha_\mathrm{gw}$ values that one can may expect a PP to fall between.
+Thus, this helps explain why $\alpha_\mathrm{gw}$ is overpredicted when the structure is increasingly underpressurized - the contaminant vapor in the PP at the ASU house was lower than the groundwater contaminant vapor concentration.
+This does not explain why underprediction occurs at increasing overpressurization, but it is likely due to the steady-state nature of the simulations.
+Regardless, the simulation are able to capture the general trends well, with not too significant deviation, giving validity to the cases simulated.\par
+
+In Figure \ref{fig:ss-sensitivity-analysis-no-pp} cases where a PP is absent are considered, i.e. a "normal" VI scenario, which may be considered reference scenarios.
+Here the only cases considered is whether there is a gravel or soil sub-base.
+It is apparent that without a PP, the indoor/outdoor pressure difference changes do not cause any signficant change in $\alpha_\mathrm{gw}$, due to the resistance to advective transport in the surrounding soil - showing the importance of the PP for increasing advection.
+Furthermore, the gravel sub-base does increase the overall $\alpha_\mathrm{gw}$, partly due to slightly higher advective potential but also due to the contaminant vapors more easily diffuse in the more permeable sub-base.
+The data from the ASU house after the PP there had been close agrees well with these model predictions, clearly again demonstrating the role of PP in increasing advective potential at a site.
+It should also be stated that if the surrounding soil was more permeable, e.g. sand, there would be more of a relationship between $\alpha_\mathrm{gw}$ and pressurization.\par
+
+Based on the simulated cases in Figure \ref{fig:ss-sensitivity-analysis} it can be concluded that a PP can be a signficant contributor to VI, but only when some conditions are met.
+\begin{enumerate}
+  \item The PP has to supply additional vapor contaminant.
+  \item There must be a source of increased advective potential, which may be due to the PP itself, or other site features or simply due to more permeable surrounding soil.
+  \item A permeable sub-base must exist to realize this advective potential.
+\end{enumerate}
+In the absense of one or all of these conditions, a PP is unlikely to be a signficant contributor to VI.
+In lieu of actually uncovering an actual PP at a site, it may be easier for a site investigator to look for the above criteria.
+Strategies for screening for PP is also discussed by \citeauthor{nielsen_remediation_2017}\cite{nielsen_remediation_2017}. \par
 

### 2018-10-18 python/transient-response/cstr-reponse.py
Created framework for analysing CSTR reponse times

@@ -1,43 +1,52 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from scipy.integrate import odeint
+from scipy.integrate import LSODA
 import sqlite3
 import pandas as pd
-"""
+
 # constants
-Ae = 0.5/3600.0 # air exchange rate 1/s
-V = 300.0 # basement volume m3
-D = 7.4e-6 # diffusivity of TCE in air m2/s
-L = 15.0e-2 # foundation slab thickness m
+Ae = 0.5 # air exchange rate 1/s
+V = 1.0 # basement volume m3
+data_dir = './data/preferential-pathway-sensitivity/'
 
-# variables
-w = lambda t: 10.0**(-6.0 + 2.0*np.sin(2.0*np.pi*t/period))
-c = 1.0e-2 # concentration assumed constant
 
-# functions
-Pe = lambda w: w*L/(2.0*D)
-n = lambda w, c: w*c/(1.0 - np.exp(-Pe(w)))
+pp = pd.read_csv(
+    data_dir + 'param-preferential-pathway.csv',
+    header=4,
+)
 
-def dudt(u, t):
+print(pp)
 
-    w = 10.0**(-6.0 + 2.0*np.sin(2.0*np.pi*t/period))
-    Pe = w*L/(2.0*D)
-    n = w*c/(1.0 - np.exp(-Pe))
-    dudt =  n/V - u*Ae
+def dudt(t, u):
+    dudt =  n(t)/V - u*Ae
     return dudt
 
-#dudt = lambda u, t: n(w,c)/V - u*Ae
+
+n_max, n0 = 10.0, 1.0
+y_max, y0 = n_max/V/Ae, n0/V/Ae
+
+t0 = 0.0
+tau = 50.0
+solver = LSODA(dudt, t0, [y0], tau)
+
+t_eq = 10.0
+t, y = [], []
+n = lambda t: n_max if (t <= t_eq) else n0
+while solver.y < 0.99*y_max:
+    t.append(solver.t)
+    y.append(solver.y)
+    solver.step()
+t_eq = solver.t
+while solver.y > 1.01*y0:
+    t.append(solver.t)
+    y.append(solver.y)
+    solver.step()
+t_org = solver.t
 
 
-# parameters
-period = 3600.0 # 1 hour period
+plt.plot(t, y)
 
-t = np.linspace(0,period)
[…]

### 2018-10-18 python/transient-response/param-period.py
Corrected KDE plotting

@@ -110,13 +110,13 @@ asu['Attenuation factor'] = asu['Concentration']/asu['tce_groundwater']
 
 
 #asu['Attenuation factor'] = asu['Attenuation factor'].apply(np.log10)
-asu['Day'] *= 24.0*3600.0
+asu['Day'] *= 24.0
 
 
 asu['dcdt'] = pd.Series(asu['Attenuation factor'].diff()/asu['Day'].diff())
 asu['dcdt'] = asu['dcdt'].apply(np.log10)
 
-asu['dcdt'] = asu['dcdt'].apply(np.abs)
+#asu['dcdt'] = asu['dcdt'].apply(np.abs)
 
 """
 asu[asu['PP'] == 'Open'][['Day','Attenuation factor']].apply(np.diff).apply(np.abs).plot(
@@ -150,11 +150,11 @@ def dp(t, period):
 df = pd.read_csv('./data/transient-response/param-period2.csv', header=4)
 df['Pressure'] = dp(df['Time'], df['period'])
 
-
+df['Time'] /= 3600.0
 df['dcdt'] = pd.Series(df['Attenuation factor'].diff()/df['Time'].diff())
 df['dcdt'] = df['dcdt'].apply(np.log10)
 
-df['dcdt'] = df['dcdt'].apply(np.abs)
+#df['dcdt'] = df['dcdt'].apply(np.abs)
 
 pivoted_df = df.pivot(
     index='Time',
@@ -164,7 +164,7 @@ pivoted_df = df.pivot(
     method='piecewise_polynomial',
 )
 
-
+"""
 pivoted_df.plot(
     y='dcdt',
     kind='kde',
@@ -172,11 +172,16 @@ pivoted_df.plot(
 )
 
 """
-sns.kdeplot(
-    pivoted_df['dcdt'],
-    ax=ax,
-)
-"""
-ax.set_ylim([0,2])
+
+print(df.period.unique())
+
+for period in df.period.unique():
+
+    sns.kdeplot(
+        df[df.period == period]['dcdt'],
+        ax=ax,
+        label='Period = %i' % period,
+    )
+
 
 plt.show()

### 2018-10-18 tex/main.tex
Updated some of Eric changes

--- 2018-10-12 tex/main.tex
@@
 Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the variability at VI sites located near Hill Air Force Base and Naval Air Station North Island using statistical analysis methods and three-dimensional subsurface computational fluid dynamics modeling.
+We consider the source of the variability at VI sites located near Hill Air Force Base in Utah, an EPA experimental house in Indianapolis, IN and Naval Air Station North Island in California using statistical analysis methods and three-dimensional subsurface computational fluid dynamics modeling.
 The results suggest that an order of magnitude variation in indoor air contaminant concentrations may be expected at ”normal” VI sites where preferential pathways do not play a role, whereas three or more orders of magnitude can be observed at sites characterized by preferential pathways.
-A CFD modeling sensitivity analysis reveals that it is not only the presence of contaminant vapor in a preferential pathway that is required to see large observed variations, but there must also exist a permeable region beneath the structure.
+A computational fluid dynamics modeling sensitivity analysis reveals that it is not only the presence of contaminant vapor in a preferential pathway that is required to see large observed variations, but there must also exist a permeable region beneath the structure.
 Large temporal fluctuations in indoor air contaminant concentrations may be observed where no preferential pathways exist, but this requires a particular combination of large fluctuations in pressure driving force and high soil permeability.
@@
 
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017}.
+Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
 VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
@@
 
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house\cite{mchugh_evidence_2017}.
-That site was also characterized by large indoor air contaminant concentration fluctuations.
-Sewer lines have been generally implicated as VI source at several sites to date\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study estimates that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
-In some of these cases\cite{pennell_sewer_2013,riis_vapor_2010}, the sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant to the soil beneath a structure. \par
+Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
+That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
+Sewer lines have been generally implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
+A Danish study estimate that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
+Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
+In some of these cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
+While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant via the soil beneath a structure. \par
 
@@
 One example is a site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the fluctuations at that site. \par
+Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
+It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site. \par
 
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of in soil preferential pathways.
+This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
 In this work, the latter scenarios are referred to as ”normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
 Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at  such normal  VI  sites vs. those characterized by preferential pathways.
-The conditions required for preferential pathways to be significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.
+The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.
 
@@
 
-The ASU house and data were obtained over a period of a few years.
+The ASU house data were obtained over a period of a few years.
 During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” operation.
@@
 
-Likewise, the Indianapolis site investigation spanned over a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
-The goal of which is to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above, and therefore only the period before the installation of this system was considered in the analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this PP removed, making it difficult to assess how significant the PP was at this site, regardless it is of interest to analyze this site due to its wealth of data.
-\par
+Likewise, the Indianapolis site investigation spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
+The goal of the SSD testing was to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above.
+Only the period before the installation of this system was considered in the present analysis.
+It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this preferential pathway removed, making it difficult to assess how significant the role of the preferential pathway was at this site.
+Regardless of this it is of interest to consider the data from this site due to how extensive and complete the data collection was. \par
 
@@
 non-CPM conditions), in the case of the NAS North Island site over the entire available dataset, and for the Indianapolis case we consider the variations before the installation of the SSD system.
-The deviations from the mean TCE (Chloroform and PCE at the Indianapolis site) in indoor air concentration, as well as the indoor-outdoor pressure differential associated with these concentration fluctuations, were examined, and both univariate and bivariate kernel density estimations (KDE) were constructed.
+The deviations in indoor air concentration from the mean TCE (and Chloroform and PCE at the Indianapolis site) values, as well as the indoor-outdoor pressure differentials associated with these concentrations were examined.
+Both univariate and bivariate kernel density estimations (KDE) were constructed.
 KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (that there is statistical fluctuation associated with each sampling event).
+This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (and that there are statistical fluctuations associated with each sampling event).
 In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
@@
 In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
-This model was implemented in a finite element solver, COMSOL Multiphysics.
+This model was implemented in a finite element solver package, COMSOL Multiphysics.
 In the present work, there has been an addition of a preferential pathway to the ”standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-In this modeling work, only the vadose zone soil domain is directly modeled.
+As in the earlier studies, only the vadose zone soil domain is directly modeled.
 
 The modeled structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab that spans the perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed does so via air exchange with the ambient.
+The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the perimeter of the house.
+All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
 The indoor control volume is assumed to consist of only of the basement, assumed as having a total volume of $300 \; \mathrm{m^3}$.
 Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of certain other features can be.
+\par
+
 The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy clay (except as noted).
 Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
-The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
-The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, which is allows the contaminant concentration in the pipe to be parameterized.
+\par
+
+Where relevant The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
+The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, allowing the contaminant concentration in the pipe to be parameterized.
+\par
+
 The groundwater beneath the structure is assumed to be homogeneously contaminated with trichloroethylene (TCE) as a prototypical contaminant.
@@
 The ground surface and the pipe are both assumed to be sources of air to the soil domain.
-Both are assumed to be at reference atmospheric pressure.
+Both are assumed to be at reference atmospheric pressure. \par
+
 Vapor transport in the soil is governed by Richard’s equation, a modified version  of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
@@
     \centering
-    \caption{Period before and after the PP was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
+    \caption{Period before and after the preferential pathway was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
     \label{fig:kde-asu-nas}
@@
 % Introduction
-Out of the factors that influence IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
-Three well-studied VI sites are considered, the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
+Of the factors that influence indoor air contaminant concentration IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
+Three well-studied VI sites are considered; the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
 % Univariate IACC description
-The absolute IACC at these sites vary significantly, therefore some means of comparing them all to each other is necessary.
-Additionally, the focus in this section is the driver for variations in IACC, and representation of the variations between the different sites must also be achieved, specifically variations on the scale of order of magnitudes is of the greatest interest, as these are the largest hinderance to proper site IACC characterization.
-To achieve these goals, the log-10 deviation from the mean IACC ($\mu$) within each dataset is calculated, e.g. $10\mu$ indicates the IACC is an order of magnitude above the mean IACC for the dataset.
-On the y-margin in figures \ref{fig:kde-asu-nas} and \ref{fig:kde-indianapolis}, the univariate KDE distribution of deviation from the mean may be seen. \par
+The absolute IACCs at these sites vary significantly, therefore some means of comparing them to each other is necessary.
+Additionally, the focus in this section is identifying the driver for variations in IACC, and consistent representation of the variations between the different sites must also be achieved.
+Here variations on the scale of orders of magnitude is of interest, as these are of greatest concern in proper site IACC characterization.
+To achieve these goals, log-10 deviation from the mean IACC (represented as $\mu$) within each dataset is calculated, e.g.
+$10\mu$ indicates an IACC that is an order of magnitude above the mean IACC for that dataset.
+On the y-margins in Figures \ref{fig:kde-asu-nas} and \ref{fig:kde-indianapolis}, the univariate KDE distribution of deviation from the mean may be seen. \par
+
+What these curves on the y-margins characterize are the distributions of sampled IACC at each site.
+Naturally, the distributions are centered around the mean ($\mu = 1$).
+In same caes, they look close to simple Gaussians in form (e.g. PCE, Tetrachloroethene at Indianapolis).
+In most of the other cases, the distributions are curved and not of a simple nature, strongly suggesting that there might be several factors determinig the IACC. \par
+
 % Univariate pressure description
-The indoor/outdoor pressure differences are simply taken as it, and their univariate KDE distributions are shown on the x-margin.
-A negative pressure difference indicate that the building is underpressurized relative to outdoor.\par
+The indoor/outdoor pressure differences are simply shown as reported, and their univariate KDE distributions are shown on the x-margins.
+A negative pressure difference indicates that the building is underpressurized relative to ambient.\par
 % Bivariate description
 The relationship between the deviation from mean IACC, and indoor/outdoor pressure difference may be seen in the central portion of each figure, in the form of a bivariate KDE distribution.
-The p-value and Pearson's r-value for each bivariate distribution is shown in the legend.\par
+The p-value and Pearson's r-value for each bivariate distribution are shown in the legend.\par
 % Sub-figure description (details about sites)
 In Figure \ref{fig:kde-asu-nas} the North Island and ASU house datasets are plotted, with blue representing the North Island site, and with red and green representing the ASU house.
-The ASU house dataset is split into two different parts, the first (red) is data from the time period before the PP was discovered, the second (green) is from after the PP was sealed off.
-At both of these sites, TCE was the only contaminant considered.
-The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex are considered in Figure \ref{fig:kde-indianapolis}.\par
+The ASU house dataset is split into two different parts, the first (red) are data from the time period before the preferential pathway was discovered, the second (green) is from after the preferential pathway was sealed off.
+At both the ASU house and North Island sites, TCE was the only contaminant considered.
+The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex were considered in Figure \ref{fig:kde-indianapolis}.\par
 % North Island discussion
@@
 The two ASU datasets are not only significantly different from the North Island dataset, but also different from each other.
-Consider the dataset before the PP was discovered (red) and one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit rarely.)
+Consider the dataset before the preferential pathway was discovered (red) and one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit rarely.)
 This is more variation than was observed at North Island, and yet the distribution of pressure differences is narrow, mostly varying by a few Pa around zero (~$\pm 2$ Pa), and with $r = -0.50$, suggests that pressure difference is not insignificant but still a weak predictor for the variations in IACC.\par
 
-The picture changes completely after the PP was closed off, where a similar distribution of pressure differences is observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
-This clearly shows just how much variance the PP at this site contributed with, and how critical it is to assess whether a PP is present at a particular site in general, lest potentially face unacceptable uncertainty in determining the relevant contaminant exposure.
-It is also clear that in the absence of a PP, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are here insignficant in determining variation in IACC.
-This also suggests that the PP at the ASU house acted not only as a PP for contaminant vapors, but for air in general - allowing the impermeable soil surrounding the house to be circumvented.
-PPs, and the factors that influnce them will be explored later in this article.\par
+The picture changes completely after the preferential pathway was closed off, where a similar distribution of pressure differences is observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
+This clearly shows just how much variance the preferential pathway at this site contributed with, and how critical it is to assess whether a preferential pathway is present at a particular site in general, lest potentially face unacceptable uncertainty in determining the relevant contaminant exposure.
+It is also clear that in the absence of a preferential pathway, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are here insignficant in determining variation in IACC.
+This also suggests that the preferential pathway at the ASU house acted not only as a preferential pathway for contaminant vapors, but for air in general - allowing the impermeable soil surrounding the house to be circumvented.
+preferential pathways, and the factors that influnce them will be explored later in this article.\par
 % Indianapolis discussion
-The Indianapolis duplex also featured a PP, but despite this, there is significantly less variance in IACC of both Chloroform and PCE.
-The exact reason for this is not known, but considering that the PP at the Indianapolis site was a sanitary sewer, and at ASU the PP was a foundation drain system, it is not inconceivable that their dynamics are quite different.
+The Indianapolis duplex also featured a preferential pathway, but despite this, there is significantly less variance in IACC of both Chloroform and PCE.
+The exact reason for this is not known, but considering that the preferential pathway at the Indianapolis site was a sanitary sewer, and at ASU the preferential pathway was a foundation drain system, it is not inconceivable that their dynamics are quite different.
 Nevertheless, both PCE and Chloroform are more or less log-normally distributed, with (mostly) less than an order of magnitude variance around the mean, indicating that the varation in IACC are species independent.\par
@@
 The pressure difference distribution exhibit some bimodal characteristics, with a larger variation than was observed at the ASU site.
-This may explain why the r-values for both PCE and Chloroform, $r=-0.51$ and $r=-0.53$ respectively, are so similar to the ASU site before the PP there was closed; the larger pressure differences, which occured more often than ASU, increase contaminant entry rates to the extent that a PP was needed to "achieve" at the ASU site.\par
+This may explain why the r-values for both PCE and Chloroform, $r=-0.51$ and $r=-0.53$ respectively, are so similar to the ASU site before the preferential pathway there was closed; the larger pressure differences, which occured more often than ASU, increase contaminant entry rates to the extent that a preferential pathway was needed to "achieve" at the ASU site.\par
 % Summary discussion
-All of these data clearly suggest that there is significant variation in IACC that cannot be easily explained by variations in indoor/outdoor pressure difference - even with a PP.
-It is also clear that PP may contribute with unacceptable levels of variations, and screening for PPs should be part of any site investiation.
-It may also be prudent, even in the absence of PP, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air.
+All of these data clearly suggest that there is significant variation in IACC that cannot be easily explained by variations in indoor/outdoor pressure difference - even with a preferential pathway.
+It is also clear that preferential pathway may contribute with unacceptable levels of variations, and screening for preferential pathways should be part of any site investiation.
+It may also be prudent, even in the absence of preferential pathway, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air.
 E.g. if a single sample is less than an order of magnitude from some target limit, it may be justified to perform subsequent samples to better assess the relevant exposure.
@@
 \begin{figure}[htb!]
-  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for a model featuring a PP (\ref{fig:ss-sensitivity-analysis-pp}) and without a PP (\ref{fig:ss-sensitivity-analysis-no-pp}).
+  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for cases featuring a preferential pathway (\ref{fig:ss-sensitivity-analysis-pp}) and without a preferential pathway (\ref{fig:ss-sensitivity-analysis-no-pp}).
 Results compared with field data from ASU house.}
@@
   \begin{subfigure}{0.55\textwidth}
-    \caption{PP present. Sensitivity to the presence of a gravel sub-base and contamination in the PP considered.}
+    \caption{preferential pathway present. Sensitivity to the presence of a gravel sub-base and contamination in the preferential pathway considered.}
     \label{fig:ss-sensitivity-analysis-pp}
@@
   \begin{subfigure}{0.55\textwidth}
-    \caption{No PP preseent. Sensitivitiy to the presence of a gravel sub-base considered.}
+    \caption{No preferential pathway present. Sensitivitiy to the presence of a gravel sub-base considered.}
     \label{fig:ss-sensitivity-analysis-no-pp}
@@
 \end{figure}
+
+Clearly, preferential pathways pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover; even at a well studied site like the ASU house it took years to discover the preferential pathway.
+Therefore, there is a need to consider other indications that a preferential pathway is present or a potential issue.\par
+
+To investigate this, inspiration is taken from the ASU house, where there clearly was a preferential pathway containing contaminant vapor, which exited into a gravel (permeable) sub-base.
+The impact of these features is simulated by examining the IACC at different indoor/outdoor pressure differences for each combination of cases.
+Instead of representing IACC as absolute concentration, we now non-dimensionalize with respect to the vapor in equilibrium with the groundwater contaminant concentration.
+This non-dimensionalized property is commonly called the attenuation factor and is denoted by $\alpha_\mathrm{gw}$.
+The result of a steady-state sensitivity analysis relative to several factors is seen in Figure \ref{fig:ss-sensitivity-analysis}.\par
+
+Figure \ref{fig:ss-sensitivity-analysis-pp} deal with cases where a preferential pathway is present and, the impact of having a gravel vs.
+soil sub-base and vapor contaminant vs.
+clean air in the preferential pathway ($\chi=1$ and $\chi=0$ respectively) are considered.
+Figure \ref{fig:ss-sensitivity-analysis-no-pp} considers the cases absent a preferential pathway, for reference, but still considers the impact of a gravel vs. soil sub-base.
+To validate these results, the IACC at the ASU house (as $\alpha_\mathrm{gw}$) vs. indoor/outdoor pressure difference are also plotted.
+Only data from period before the preferential pathway was discovered, and after the preferential pathway was closed are considered in each respective figure (\ref{fig:ss-sensitivity-analysis-pp} and \ref{fig:ss-sensitivity-analysis-no-pp}).
+To improve visibility of the ASU data, it's average $\alpha_\mathrm{gw}$ are aggregated into up to 40 evenly spaced bins on the x-axis.\par
+
+Begining with Figure \ref{fig:ss-sensitivity-analysis-pp} it immediately clear that absent a gravel sub-base, the impact of a preferential pathway is minimal, as the orange and blue lines show.
+Since in the model, it is assumed that the soil type is sandy clay, which is relatively impermeable, there is simply too much resistance to advection for changes in pressurization to matter.
+Additionally, whether the preferential pathway is filled with contaminant vapors or not does matter; without a permeable sub-base, the transport of contaminant vapor in the sub-base is too inhibited to be effectively transported into the building.\par
+
+Moving on to the case of when a gravel sub-base is present, and the preferential pathway is filled with contaminant vapors ($\chi=1$).
+In this scenario the preferential pathway has a massive impact on $\alpha_\mathrm{gw}$, spanning more than four orders of magnitude.
+This is explained by the increased role of advective transport under these conditions.
+The gravel sub-base allows vapor to flow more easily, and the preferential pathway is able to circumvent the large resistance to transport due to the surrounding soil, realizing the advective potential offered by the gravel sub-base.
+When the building is underpressurized under these conditions, a significant amount of contaminant vapor is able to be pulled from the preferential pathway, leading to the large increase of $\alpha_\mathrm{gw}$.\par
+
+On the other hand, due to the increased advective potential, overpressurization more effectively prevents contaminant vapors from entering the structure, and most of the contaminant entering the building does so through diffusion.
+This also explains why there is no difference between the uncontaminted and contaminted preferential pathway cases (red and green) when the building is overpressurized.
+But these two cases (red and green) differ signficant when the building is underpressurized, obviously because there simply is no additional contaminant being supplied by the preferential pathway, $\alpha_\mathrm{gw}$ thus limits to whatever diffuses from the contaminted groundwater source.\par
+
+It is interesting to note that $\alpha_\mathrm{gw}$ is higher with an uncontaminted preferential pathway (green) than the cases without a gravel sub-base.
+One could expect that pulling so much clean air into the sub-base region would significantly dilute the contaminant vapors and if anything cause a decrease in $\alpha_\mathrm{gw}$ relative to blue and orange.
+However, the preferential pathway is small relative to the sub-base (10 cm diameter pipe), thus impacting the contaminant concentration in a limited region of the sub-base, but still allows for higher advection in the entire sub-base region, leading to a $\alpha_\mathrm{gw}$ plateau that is limited by contaminant transport from the groundwater source.\par
+
+These case most resembling the ASU site (with an open preferential pathway) is the red case, and fits the data fairly well, especially for small under-/overpressurization values.
+The most significant deviations occur as the under-/overpressurization becomes larger.
+However, the model is not intended to replicate the ASU data perfectly, but rather investigate the factors make a preferential pathway the most impactful in a general sense, and perfect simulation is not to be expected.
+This is especially reflected in the choice of vapor contaminant concentration in the preferential pathway, with it either being equal to the groundwater contaminant vapor concentration ($\chi=1$) or clean ($\chi=0$), neither of which is likely to be true at the ASU site.
+The idea with picking $\chi=1$ and $\chi=0$ is to give a span of $\alpha_\mathrm{gw}$ values that one can may expect a preferential pathway to fall between.
+Thus, this helps explain why $\alpha_\mathrm{gw}$ is overpredicted when the structure is increasingly underpressurized - the contaminant vapor in the preferential pathway at the ASU house was lower than the groundwater contaminant vapor concentration.
+This does not explain why underprediction occurs at increasing overpressurization, but it is likely due to the steady-state nature of the simulations.
+Regardless, the simulation are able to capture the general trends well, with not too significant deviation, giving validity to the cases simulated.\par
+
+In Figure \ref{fig:ss-sensitivity-analysis-no-pp} cases where a preferential pathway is absent are considered, i.e. a "normal" VI scenario, which may be considered reference scenarios.
+Here the only cases considered is whether there is a gravel or soil sub-base.
+It is apparent that without a preferential pathway, the indoor/outdoor pressure difference changes do not cause any signficant change in $\alpha_\mathrm{gw}$, due to the resistance to advective transport in the surrounding soil - showing the importance of the preferential pathway for increasing advection.
+Furthermore, the gravel sub-base does increase the overall $\alpha_\mathrm{gw}$, partly due to slightly higher advective potential but also due to the contaminant vapors more easily diffuse in the more permeable sub-base.
+The data from the ASU house after the preferential pathway there had been close agrees well with these model predictions, clearly again demonstrating the role of preferential pathway in increasing advective potential at a site.
+It should also be stated that if the surrounding soil was more permeable, e.g. sand, there would be more of a relationship between $\alpha_\mathrm{gw}$ and pressurization.\par
+
+Based on the simulated cases in Figure \ref{fig:ss-sensitivity-analysis} it can be concluded that a preferential pathway can be a signficant contributor to VI, but only when some conditions are met.
+\begin{enumerate}
+  \item The preferential pathway has to supply additional vapor contaminant.
+  \item There must be a source of increased advective potential, which may be due to the preferential pathway itself, or other site features or simply due to more permeable surrounding soil.
+  \item A permeable sub-base must exist to realize this advective potential.
+\end{enumerate}
+In the absense of one or all of these conditions, a preferential pathway is unlikely to be a signficant contributor to VI.
+In lieu of actually uncovering an actual preferential pathway at a site, it may be easier for a site investigator to look for the above criteria.
+Strategies for screening for preferential pathway is also discussed by \citeauthor{nielsen_remediation_2017}\cite{nielsen_remediation_2017}. \par
 

### 2018-10-21 tex/main.tex
Added Eric's changes

@@
     \centering
-    \caption{Period before and after the preferential pathway was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
+    \caption{Period before and after the preferential pathway (PP) was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
     \label{fig:kde-asu-nas}
@@
     \centering
-    \caption{Chloroform and PCE considered separeately at Indianapolis duplex 422.}
+    \caption{Chloroform and PCE considered at the Indianapolis duplex 422.}
     \label{fig:kde-indianapolis}
@@
 At both the ASU house and North Island sites, TCE was the only contaminant considered.
-The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex were considered in Figure \ref{fig:kde-indianapolis}.\par
+The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex were considered in Figure \ref{fig:kde-indianapolis}, since they were particulary tracked.\par
 % North Island discussion
 From Figure \ref{fig:kde-asu-nas}, it is apparent that the three datasets differ significantly from each other.
-The North Island site exhibit the greatest variation in both pressure difference and IACC, in fact, looking at the univariate pressure distribution - it is almost entirely flat and spanning from extremly high and low values.
-A likely explation for this is the poor condition of the structure, rendering it highly susceptible to envivornmental influnces.\par
+The North Island site exhibits the greatest variation in both pressure difference and IACC, in fact, looking at the univariate pressure distribution - it is almost entirely flat and spanning from extremly high and low values.
+A likely explation for this is the reportedly poor condition of the structure, rendering it highly susceptible to envivornmental influnces.\par
 
-To accompany this, there is significant variation in IACC, that has a relativly strong link with pressure difference ($r = -0.71$), although there is clearly a distinct peak (around $0.5\mu$) that is primally associated with the smaller negative and positive pressure differences ($~-4 \leq \Delta p \leq 8$).
-This demonstrates that at this particular site, the very large variations (an order of magnitude or more) are primarily influenced by the unusually large pressure difference fluctations.
-Likely, this relationship is amplified by the fact that the soil surrounding the structure is sandy and therefore relatively permeable - increasing the influence of pressure differences.
-However, despite all this, pressure differences can only partly explain the variations observed at this site.\par
+There is also a significant variation in IACC at the North Island site and this has a relativly strong link to pressure difference ($r = -0.71$), although there is clearly a distinct peak (around $0.5\mu$) that is associated with the smaller negative and positive pressure differences ($~-4 \leq \Delta p \leq 8$).
+This demonstrates that at this particular site, the very large IACC variations (an order of magnitude or more) are primarily influenced by the unusually large pressure difference fluctations.
+Likely, this relationship is created by the fact that the soil surrounding the structure is sandy and therefore relatively permeable - increasing the influence of pressure differences on inflow of contaminant from the soil.
+However, despite this, pressure differences can only partly explain the variations observed at this site.\par
 % ASU discussion
 The two ASU datasets are not only significantly different from the North Island dataset, but also different from each other.
-Consider the dataset before the preferential pathway was discovered (red) and one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit rarely.)
-This is more variation than was observed at North Island, and yet the distribution of pressure differences is narrow, mostly varying by a few Pa around zero (~$\pm 2$ Pa), and with $r = -0.50$, suggests that pressure difference is not insignificant but still a weak predictor for the variations in IACC.\par
+Considering the dataset before the preferential pathway was discovered (red), one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit the latter only rarely.)
+This is more variation than was observed at North Island, and yet the observed distribution of pressure differences is narrow, mostly varying by a few Pa around zero (~$\pm 2$ Pa).
+With $r = -0.50$, for this dataset, this suggests that pressure difference is not insignificant but still only a weak predictor for the variations in IACC.\par
 
-The picture changes completely after the preferential pathway was closed off, where a similar distribution of pressure differences is observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
-This clearly shows just how much variance the preferential pathway at this site contributed with, and how critical it is to assess whether a preferential pathway is present at a particular site in general, lest potentially face unacceptable uncertainty in determining the relevant contaminant exposure.
-It is also clear that in the absence of a preferential pathway, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are here insignficant in determining variation in IACC.
-This also suggests that the preferential pathway at the ASU house acted not only as a preferential pathway for contaminant vapors, but for air in general - allowing the impermeable soil surrounding the house to be circumvented.
-preferential pathways, and the factors that influnce them will be explored later in this article.\par
+The picture changes completely after the preferential pathway was closed off, where a similar distribution of pressure differences were observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
+This clearly shows just how much variance the preferential pathway at this site contributed, and how critical it is to assess whether a preferential pathway is present at a particular site since this may lead to large uncertainties in determining relevant contaminant exposures.
+It is also clear that in the absence of a preferential pathway, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are insignficant in determining variation in IACC.\par
 % Indianapolis discussion
-The Indianapolis duplex also featured a preferential pathway, but despite this, there is significantly less variance in IACC of both Chloroform and PCE.
+The Indianapolis duplex also featured a preferential pathway, but despite this, there is significantly lower variance in IACC for both Chloroform and PCE\cite{mchugh_evidence_2017}.
 The exact reason for this is not known, but considering that the preferential pathway at the Indianapolis site was a sanitary sewer, and at ASU the preferential pathway was a foundation drain system, it is not inconceivable that their dynamics are quite different.
-Nevertheless, both PCE and Chloroform are more or less log-normally distributed, with (mostly) less than an order of magnitude variance around the mean, indicating that the varation in IACC are species independent.\par
+Nevertheless, both PCE and Chloroform are more or less log-normally distributed, with (mostly) less than an order of magnitude variance around the mean, indicating that the varations in IACC are species independent.\par
 
-The pressure difference distribution exhibit some bimodal characteristics, with a larger variation than was observed at the ASU site.
+The pressure difference distributions for the Indianapolis site exhibit some bimodal characteristics, with a larger variation than was observed at the ASU site.
 This may explain why the r-values for both PCE and Chloroform, $r=-0.51$ and $r=-0.53$ respectively, are so similar to the ASU site before the preferential pathway there was closed; the larger pressure differences, which occured more often than ASU, increase contaminant entry rates to the extent that a preferential pathway was needed to "achieve" at the ASU site.\par
 % Summary discussion
-All of these data clearly suggest that there is significant variation in IACC that cannot be easily explained by variations in indoor/outdoor pressure difference - even with a preferential pathway.
-It is also clear that preferential pathway may contribute with unacceptable levels of variations, and screening for preferential pathways should be part of any site investiation.
-It may also be prudent, even in the absence of preferential pathway, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air.
-E.g. if a single sample is less than an order of magnitude from some target limit, it may be justified to perform subsequent samples to better assess the relevant exposure.
-These data also call the role of advection in VI into question, as "normal" pressure difference ranges are such a poor predictor of IACC variation, which is the opposite of what one would expect if advection was a dominant transport mechanism.\par
-
-\begin{comment}
-
-% 2D KDE plot of deviation from mean TCE in indoor air and pressure difference from the ASU house and North Island NAS
-\begin{figure}[htb!]
-  \caption{Probability density plots of the deviation from the mean concentration along with the indoor-outdoor pressure difference at the ASU house both before and after the preferential pathway was closed and at North Island NAS.}
-  \label{fig:kde-asu-nas}
-  \includegraphics[width=\textwidth]{asu-nas.pdf}
-\end{figure}
-
-Figure \ref{fig:kde-asu-nas} show the result of the statistical analysis of the field data from the ASU and NAS North Island sites.  On the axes, the univariate KDE’s are shown.
-The y-margin shows the distribution of the log-10 deviation from the mean TCE concentration in indoor air and the x-margin shows the distribution of indoor-outdoor pressure differences.
-A negative pressure difference indicates that the structure is underpressurized, and thus air flows into the building, and a positive pressure difference indicates the opposite.
-Thus the curves on the axis margins are the distribution functions of the respective measured variable values, considered independently.
-They show the distributions of the results from the many different sampling events.
-The distribution functions of pressure fluctuations are relatively simpler than those for the indoor air concentration measurements, and the pressure distribution functions for the pressure difference measurements at the ASU house are much simpler and narrower than that for comparable measurements at NAS North Island (see the distribution functions on the top margin of the figure).
-Comparing the narrow red and green distribution functions for the ASU house results shows that whether the valve to the preferential pathway (PP) is open or not, roughly the same indoor-outdoor pressure differentials were seen.
-This is not at all surprising, given that the existence of a preferential pathway buried beneath the foundation slab would not be expected to influence the indoor-outdoor pressure differential.
-But when comparing the ASU house pressure differentials with those reported for the NAS North Island site (the blue distribution function), it is immediately apparent how much larger the pressure swings are at the latter site.
-On this basis, if one associates contaminant entry rate with pressure differential alone, the NAS North Island site would be expected to show a much larger fluctuation in indoor air contaminant concentration than would the ASU house (but of course the preferential pathway changes the situation, as described below).
-
-Turning to the univariate distributions for indoor air contaminant concentrations, the results show that even with the preferential pathway closed off, the results from the ASU house clearly show variations of an order of magnitude.
-The construction of the KDE is naturally such that the overall distribution gives a mean of what is characterized as "$1.0 \mu$" in the figure, but around that mean the distribution of all measured concentrations falls within an order of magnitude of the mean.
-The implication is that the most probable result from a single concentration sampling event will be a value around the mean, but the possibility that a single sampling event might give a result that is an order of magnitude higher or lower than a second independent sampling event cannot be ruled out, even under non-preferential pathway circumstances.
-Thus, to the extent that the ASU house serves as a prototype for “normal” VI scenarios (when the preferential pathway was closed), it is only when any particular sampling event provides a result that is an order of magnitude below some action threshold that one can be highly confident that further sampling will not change this outcome.
-The more sampling events that are involved, such that a population mean can confidently be estimated, the less likely that any future sampling event would give a value that is more than about a half an order of magnitude beyond that mean.
-These are of course very rough rules of thumb, based upon this analysis of this data set.
-
-When the preferential pathway beneath the ASU house is open, a very different picture emerges.
-The calculated univariate distribution function takes on a much more complicated, somewhat bimodal character.
-It should be recalled that the mean of all of the results remains, as before, at $"1.0 \mu"$.
-The red distribution curve shows the origin of what has been reported as the “three orders of magnitude” variability in the indoor air contaminant concentrations.
-On the one hand, there is a significant probability that any one measured concentration can be lower, up to even an order of magnitude below, the mean of the whole data set.
-On the other hand, there is a spread to the upside of the mean by two orders of magnitude.
-So in other words, the existence of an open preferential pathway in this case offers little guidance on what might be expected from any single sampling event.
-This could easily have been inferred from the already raw published data from this site, but this analysis offers a statistical picture that suggests that there are distinctly different operational situations that give rise to different peaks in the observed distributions.
-This, in turn, points to different operational parameters that have not yet been fully captured.
-
-There is inherently greater spread in the ASU house data when the preferential pathway is open.
-The variations in a direction greater than the mean show much greater spread, and arguably the choice of a Gaussian kernel is not necessarily the best.
-The bivariate analysis below will shed more light on this point.
-
-The univariate KDE distributions form the NAS North Island site show a distinctly bimodal distribution (the blue curve).
-There is a population of sampling results above the mean and another below the mean.
-Given that there was no influence of preferential pathways at this site, these results immediately flag the possibility that one is dealing with sample concentration populations that are strongly influenced by the other variable displayed on this plot - the indoor-outdoor pressure difference.
-When the interior of the building is depressurized relative to ambient, there is a tendency towards indoor air contaminant concentrations higher than the mean, and when the building is pressurized, the values tend to fall below the mean.
-This immediately suggests the examination of bivariate distributions, which are also shown on Figure \ref{fig:kde-asu-nas}, in the center of the plot.
-
-The center of Figure \ref{fig:kde-asu-nas} shows the bivariate KDEs constructed based on the relationship between the two aforementioned variables.
-The color-coding indicates the source of the dataset, with green again representing the ASU house dataset after the preferential pathway was closed, red is for before the pathway was closed, and blue the NAS North Island dataset.
-The Pearson’s r-values and p-value for the relation between TCE in indoor air and indoor-outdoor pressure difference for each dataset are also shown.
-The bivariate distributions in Figure \ref{fig:kde-asu-nas} show even more starkly the differences between the ASU house datasets and the NAS North Island dataset.
-We believe that this is primarily due to the significant different indoor-outdoor pressure differences existing at these two sites.
-Again, NAS North Island showed a much larger spread of indoor-outdoor pressure differential values, seen in the distributions on the top x-margin; the ASU house pressure difference distributions were quite Gaussian with a variation of just a few Pascals, whereas the NAS North Island distribution is rather more flat and includes a much larger span of values.
-
-The bivariate distributions make clear one aspect of the ASU house results.
-The “outlier” points at almost two orders of magnitude beyond the mean are associated with the existence of a preferential pathway and an unusually high indoor-outdoor pressure differential.
-But for that particular combination of circumstances, the variability in indoor air contaminant concentrations, given existence of the preferential pathway, would have been “only” two orders of magnitude.
-Comparing the green and red bivariate distributions, it is clear that the existence of the preferential pathway increases the potential spread of measured concentration values.
-But beyond this, the correlation statistics show fairly convincingly that the pressure driving force is connected with the observed indoor air contaminant concentrations (even leaving out the few points at high depressurization), but that the correlation between indoor air contaminant concentration and pressure driving force is weak under “normal” VI conditions, where the preferential pathway was closed off.
-This strongly suggests that under the latter conditions, advective entry was not a significant driver, and that ordinary diffusive processes were playing a role in bringing contaminant into the house from the area beneath the slab.
-The above results reaffirm the problem that the potential existence of preferential pathways can cause for VI site investigations, i.e.
-they can greatly reduce the reliability of any single sampling event, requiring a larger number of samples to properly assess the site.
-The existence of preferential pathways to the vicinity of a structure can enhance the effect of other variables, such as small pressure differences.
-In this case, any single sampling event under “natural” conditions might give a value that is two orders of magnitude different from another sampling event.
-Conservatism in that instance would suggest that any sampling results within two orders of magnitude of an action threshold would require re-sampling.
-While the present results are taken from a single site, there is every reason to believe that this site can be representative of a large number of real sites.
-
-Turning to the bivariate results for the NAS North Island site, the bivariate distributions leave little doubt that indoor-outdoor pressure differential is a key driver for measured indoor air contaminant concentrations.
-The correlation statistics confirm the visual conclusion.
-There is little surprise in the results, in that higher extents of depressurization lead to more contaminant entry.
-Note that but for the portions of the blue contours at high and low pressure differences, the distribution of indoor air concentration values, relative to a mean, would be close to those from the ASU house.
-In other words, at least some of the high temporal variability at NAS North Island comes from the rather extreme range of pressure differentials that exist there.
-This supports a conclusion that under “normal” VI conditions, absent a preferential pathway, with the normally encountered pressure differentials of a few Pa, temporal variations in measured indoor air contaminant concentrations of an order of magnitude must be expected.
-Clearly it would be useful to have a few more datasets from different sites to support such a general conclusion.
-
-\begin{figure}[htb!]
-  \caption{Probability density plots of the deviation from the mean Chloroform and Tetrachloroethene concentrations along with the indoor-outdoor pressure difference at the Indianapolis site}
-  \label{fig:kde-indianapolis}
-  \includegraphics[width=\textwidth]{indianapolis-422.pdf}
-\end{figure}
-
-Figure \ref{fig:kde-indianapolis} shows the same as Figure \ref{fig:kde-asu-nas} but at the Indianapolis site, and the variations of Chloroform and PCE are considered instead of TCE.
-The concentration of both Chloroform and PCE are still roughly equally log-normally distributed within half an order of magnitude above and below the mean, similar to the variation observed at the ASU house after the land drain was closed, as well as to the North Island site, if one would consider comparable indoor-outdoor pressure differences.
-The distribution of pressure difference at the Indianapolis site is between that of ASU and North Island, with the structure there typically being more depressurized than ASU but less than North Island.
-The relationship between pressure difference and IACC is much stronger at the Indianapolis site than it was at the ASU house, but less so than at North Island.
-Given what is known about the soil around Indianapolis site, this is somewhat surprising, as it is not significantly different from the ASU site (which North Island is).
-This may be evidence of the interaction with the PP that exists at the site, although the extreme variations in IACC cannot be observed here.
-\end{comment}
+All of these data clearly suggest that there is significant variation in IACC that cannot easily be explained by variations in indoor/outdoor pressure difference - even with a preferential pathway.
+It is also clear that a preferential pathway may contribute to unacceptable levels of IACC variations, and screening for preferential pathways should be part of any site investiation.
+It may also be prudent, even in the absence of a preferential pathway, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air, e.g.
+if a single sample is less than an order of magnitude below some target limit, it may be justified to perform subsequent sampling to better assess the relevant exposure.
+These data also call the role of advection in VI into question, since "normal" pressure difference ranges are such a poor predictor of IACC variation.
+This is the opposite of what one would expect if advection were the dominant transport mechanism for entry of contaminant into a structure, under steady-state conditions.\par
 
@@
 \end{figure}
-Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion, and another highly dynamic process.
+Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion, and part of another highly dynamic process.
 Exploring the impact of fluctuations in air exchange rate is thus also important for the development of sampling strategies.
@@
 This allows for exploration of the impact that fluctuating air exchange rate has on the final IACC by solving the CST equation, using the recorded contaminant entry rate as one input, and assuming a constant air exchange rate.
-The rationale is to show the extent of variability in IACC due to factors other than fluctuations in air exchange rates.
-We have taken this factor out of consideration by assuming a single constant exchange rate, characteristic of the whole period.
+The rationale is to show the extent of variability in IACCs due to factors other than fluctuations in air exchange rates.
+We have taken this factor out of consideration by below assuming a single constant exchange rate, characteristic of the whole period.
 
-In Figure \ref{fig:asu-iacc-ae-overview}, the measured IACC and air exchange rate across time are shown, with the predicted IACC assuming a constant air exchange rate.
+In Figure \ref{fig:asu-iacc-ae-overview}, the measured IACC and air exchange rate across time are shown, compared to the predicted IACC assuming instead a constant air exchange rate.
 The median air exchange rate value for the pre-CPM period was chosen as this seemed to be the most representative value for this time period.
-From Figure \ref{fig:asu-iacc-ae-overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC assuming constant and actual fluctuating exchange rates are quite similar.
+From Figure \ref{fig:asu-iacc-ae-overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC values assuming constant vs.
+actual fluctuating exchange rates are quite similar.
 In other words, in this instance it is unlikely that the variations in IACC were driven by fluctuations in air exchange rate.
@@
 This is more clearly shown in Figure \ref{fig:violin-iacc-aet}, where KDEs of the IACC are constructed assuming the constant and fluctuating air exchange rate cases.
-This figure shows the relative probability of being at a particular TCE in indoor air concentration at the ASU house.
+This figure shows the relative probability of being at a particular TCE level in indoor air concentration at the ASU house.
 These are compared to each other for 1. the full measurement period, 2. the non-CPM periods and 3. the CPM period.
@@
 This is because the actual air exchange rate for this CPM period is significantly higher than the assumed median value, and much more contaminant is actually exchanged with the exterior.
-The shape of the IACC distribution is not very different from that which takes the air exchange fluctuations into account however, indicating that fluctuating exchange rate is not a major contributor to observed IACC variation for this period (the distinction being drawn between variation and absolute levels of IACC).
+The shape of the IACC distribution is not very different from that which takes the air exchange fluctuations into account, indicating that fluctuating exchange rate is not a major contributor to observed IACC variation for this period (the distinction being drawn between variation and absolute levels of IACC).
 
@@
 It is clear here that there is only a very minor difference made in assuming a constant vs.
-real fluctating air exchange rate, again indicating that account for the minor fluctuations of air exchange rate's impact on variations in IACC is not important.
+real fluctating air exchange rate, again indicating that accounting for the minor fluctuations of air exchange rate's impact on variations in IACC is not important.
 \citeauthor{rackes_time-averaged_2016}\cite{rackes_time-averaged_2016} drew similar conclusions regarding the relationship between IACC and air exchange rate.
@@
 
-Clearly, preferential pathways pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover; even at a well studied site like the ASU house it took years to discover the preferential pathway.
+Clearly, preferential pathways pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover.
+Even at a well studied site like the ASU house it took years to discover the preferential pathway.
 Therefore, there is a need to consider other indications that a preferential pathway is present or a potential issue.\par
@@
 Figure \ref{fig:ss-sensitivity-analysis-no-pp} considers the cases absent a preferential pathway, for reference, but still considers the impact of a gravel vs. soil sub-base.
-To validate these results, the IACC at the ASU house (as $\alpha_\mathrm{gw}$) vs. indoor/outdoor pressure difference are also plotted.
+To give perspective to these results, the IACC at the ASU house (as $\alpha_\mathrm{gw}$) vs. indoor/outdoor pressure difference are also plotted.
+But the reader is reminded that the model was examined for a structure of the same scale as the ASU house, but which is not an exact representation of it.
 Only data from period before the preferential pathway was discovered, and after the preferential pathway was closed are considered in each respective figure (\ref{fig:ss-sensitivity-analysis-pp} and \ref{fig:ss-sensitivity-analysis-no-pp}).
-To improve visibility of the ASU data, it's average $\alpha_\mathrm{gw}$ are aggregated into up to 40 evenly spaced bins on the x-axis.\par
+To improve visibility of the ASU data, it's average $\alpha_\mathrm{gw}$ values were aggregated into up to 40 evenly spaced points.\par
 
-Begining with Figure \ref{fig:ss-sensitivity-analysis-pp} it immediately clear that absent a gravel sub-base, the impact of a preferential pathway is minimal, as the orange and blue lines show.
-Since in the model, it is assumed that the soil type is sandy clay, which is relatively impermeable, there is simply too much resistance to advection for changes in pressurization to matter.
-Additionally, whether the preferential pathway is filled with contaminant vapors or not does matter; without a permeable sub-base, the transport of contaminant vapor in the sub-base is too inhibited to be effectively transported into the building.\par
+Begining with Figure \ref{fig:ss-sensitivity-analysis-pp} it is immediately clear that absent a gravel sub-base, the impact of a preferential pathway is minimal, as comparing the orange and blue lines shows.
+Since in the model, it is assumed that the soil type is sandy clay, which is relatively impermeable, there is simply too much resistance to advection in the soil for changes in pressurization to matter.
+Additionally, whether the preferential pathway is filled with contaminant vapors or not does not matter; without a permeable sub-base, the transport of contaminant vapor in the sub-base is too slow to allow an advective entry path to contribute with.\par
 
-Moving on to the case of when a gravel sub-base is present, and the preferential pathway is filled with contaminant vapors ($\chi=1$).
-In this scenario the preferential pathway has a massive impact on $\alpha_\mathrm{gw}$, spanning more than four orders of magnitude.
+When a gravel sub-base is present, and the preferential pathway is filled with contaminant vapors ($\chi=1$).
+Tthe preferential pathway has a major impact on $\alpha_\mathrm{gw}$, spanning more than four orders of magnitude.
 This is explained by the increased role of advective transport under these conditions.
-The gravel sub-base allows vapor to flow more easily, and the preferential pathway is able to circumvent the large resistance to transport due to the surrounding soil, realizing the advective potential offered by the gravel sub-base.
-When the building is underpressurized under these conditions, a significant amount of contaminant vapor is able to be pulled from the preferential pathway, leading to the large increase of $\alpha_\mathrm{gw}$.\par
+The gravel sub-base allows vapor to readily flow.
+When the building is underpressurized under these conditions, a significant amount of contaminant vapor is pulled in from the preferential pathway, leading to the large increase of $\alpha_\mathrm{gw}$.\par
 
-On the other hand, due to the increased advective potential, overpressurization more effectively prevents contaminant vapors from entering the structure, and most of the contaminant entering the building does so through diffusion.
-This also explains why there is no difference between the uncontaminted and contaminted preferential pathway cases (red and green) when the building is overpressurized.
-But these two cases (red and green) differ signficant when the building is underpressurized, obviously because there simply is no additional contaminant being supplied by the preferential pathway, $\alpha_\mathrm{gw}$ thus limits to whatever diffuses from the contaminted groundwater source.\par
+On the other hand, when advection is stopped by overpressurization, most of the contaminant entering the building does so only through diffusion.
+This also explains why there is no difference whether the preferential pathway brings only clean air or contaminated air (red and green) when the building is overpressurized.
+These two cases differ signficantly from one another when the building is underpressurized, obviously because if there is no additional contaminant being supplied by the preferential pathway, the contaminant comes only from what diffuses from the contaminted groundwater source.\par
 
-It is interesting to note that $\alpha_\mathrm{gw}$ is higher with an uncontaminted preferential pathway (green) than the cases without a gravel sub-base.
-One could expect that pulling so much clean air into the sub-base region would significantly dilute the contaminant vapors and if anything cause a decrease in $\alpha_\mathrm{gw}$ relative to blue and orange.
-However, the preferential pathway is small relative to the sub-base (10 cm diameter pipe), thus impacting the contaminant concentration in a limited region of the sub-base, but still allows for higher advection in the entire sub-base region, leading to a $\alpha_\mathrm{gw}$ plateau that is limited by contaminant transport from the groundwater source.\par
+It is interesting to note that $\alpha_\mathrm{gw}$ is higher with an uncontaminted preferential pathway (green) entering a gravel sub-base than the cases without a gravel sub-base.
+One could expect that pulling so much clean air into the sub-base region might significantly dilute the contaminant vapors and if anything cause a decrease in $\alpha_\mathrm{gw}$ (relative to blue and orange).
+However, the preferential pathway is small relative to the dimensions of the sub-base (10 cm diameter pipe), and its impact is only on the contaminant concentration in a limited region of the sub-base.
+This gravel sub-base still allows for higher advection throughout the entire sub-base region, leading to a $\alpha_\mathrm{gw}$ plateau that is limited by contaminant transport from the groundwater source, and any small amount of dilution from the preferential pathway is negligible.\par
 
-These case most resembling the ASU site (with an open preferential pathway) is the red case, and fits the data fairly well, especially for small under-/overpressurization values.
+These case most resembling the ASU site (with an open preferential pathway) is the red case, and the trend agrees with the field data fairly well, especially for small under-/overpressurization values.
 The most significant deviations occur as the under-/overpressurization becomes larger.
-However, the model is not intended to replicate the ASU data perfectly, but rather investigate the factors make a preferential pathway the most impactful in a general sense, and perfect simulation is not to be expected.
-This is especially reflected in the choice of vapor contaminant concentration in the preferential pathway, with it either being equal to the groundwater contaminant vapor concentration ($\chi=1$) or clean ($\chi=0$), neither of which is likely to be true at the ASU site.
-The idea with picking $\chi=1$ and $\chi=0$ is to give a span of $\alpha_\mathrm{gw}$ values that one can may expect a preferential pathway to fall between.
-Thus, this helps explain why $\alpha_\mathrm{gw}$ is overpredicted when the structure is increasingly underpressurized - the contaminant vapor in the preferential pathway at the ASU house was lower than the groundwater contaminant vapor concentration.
-This does not explain why underprediction occurs at increasing overpressurization, but it is likely due to the steady-state nature of the simulations.
-Regardless, the simulation are able to capture the general trends well, with not too significant deviation, giving validity to the cases simulated.\par
+Again, the model is not intended to replicate the ASU data perfectly, but rather investigate the factors make a preferential pathway the most impactful in a general sense, and perfect simulation is not to be expected.
+This is especially reflected in the choice of vapor contaminant concentration in the preferential pathway, with it either being equal to the groundwater contaminant vapor concentration ($\chi=1$) or clean ($\chi=0$), neither of which is likely to be perfectly true at the ASU site.
+The idea with picking $\chi=1$ and $\chi=0$ is to give a span of $\alpha_\mathrm{gw}$ values within which a preferential pathway may fall.
+This helps explain why $\alpha_\mathrm{gw}$ is overpredicted when the structure is increasingly underpressurized - the contaminant vapor in the preferential pathway at the ASU house was lower than the groundwater contaminant vapor concentration.
+The simulation are able to capture the general trends well.\par
 
-In Figure \ref{fig:ss-sensitivity-analysis-no-pp} cases where a preferential pathway is absent are considered, i.e. a "normal" VI scenario, which may be considered reference scenarios.
-Here the only cases considered is whether there is a gravel or soil sub-base.
+In Figure \ref{fig:ss-sensitivity-analysis-no-pp} cases where a preferential pathway is absent are considered, i.e. "normal" VI scenarios, which may be considered reference scenarios.
+Here the only issues considered is whether there is a gravel or soil sub-base.
 It is apparent that without a preferential pathway, the indoor/outdoor pressure difference changes do not cause any signficant change in $\alpha_\mathrm{gw}$, due to the resistance to advective transport in the surrounding soil - showing the importance of the preferential pathway for increasing advection.
-Furthermore, the gravel sub-base does increase the overall $\alpha_\mathrm{gw}$, partly due to slightly higher advective potential but also due to the contaminant vapors more easily diffuse in the more permeable sub-base.
-The data from the ASU house after the preferential pathway there had been close agrees well with these model predictions, clearly again demonstrating the role of preferential pathway in increasing advective potential at a site.
-It should also be stated that if the surrounding soil was more permeable, e.g. sand, there would be more of a relationship between $\alpha_\mathrm{gw}$ and pressurization.\par
+Furthermore, the gravel sub-base does increase the overall $\alpha_\mathrm{gw}$, partly due to slightly higher advective potential but also due to the contaminant vapors being able to more easily diffuse in the more permeable sub-base.
+The trend in and even the value sof the data from the ASU house after the preferential pathway had been closed agree well with these model predictions, clearly again demonstrating the role of preferential pathway in increasing advective potential at a site.
+It should also be stated that if the surrounding soil was more permeable, e.g. sand, there would be more of a dependence of $\alpha_\mathrm{gw}$ on pressurization, as seen in North Island.\par
 
-Based on the simulated cases in Figure \ref{fig:ss-sensitivity-analysis} it can be concluded that a preferential pathway can be a signficant contributor to VI, but only when some conditions are met.
+Based on the simulated cases in Figure \ref{fig:ss-sensitivity-analysis} it can be concluded that a preferential pathway can be a signficant contributor to VI, but only when particular conditions are met.
 \begin{enumerate}
   \item The preferential pathway has to supply additional vapor contaminant.
-  \item There must be a source of increased advective potential, which may be due to the preferential pathway itself, or other site features or simply due to more permeable surrounding soil.
-  \item A permeable sub-base must exist to realize this advective potential.
+  \item There must be a source of increased advective potential, which may be due to the presence of the preferential pathway itself, or other site features, or simply due to more permeable surrounding soil.
+  \item A permeable sub-base must exist to realize increased advective potential.
 \end{enumerate}
-In the absense of one or all of these conditions, a preferential pathway is unlikely to be a signficant contributor to VI.
-In lieu of actually uncovering an actual preferential pathway at a site, it may be easier for a site investigator to look for the above criteria.
-Strategies for screening for preferential pathway is also discussed by \citeauthor{nielsen_remediation_2017}\cite{nielsen_remediation_2017}. \par
+In the absense of any of these conditions, a preferential pathway is unlikely to be a signficant contributor to VI.
+Strategies for screening for preferential pathway are discussed by \citeauthor{nielsen_remediation_2017}\cite{nielsen_remediation_2017}. \par
 

### 2018-10-22 python/transient-response/cstr-reponse.py
Cleaned up code and generated figure

@@ -12,161 +12,142 @@ import plotly.io as pio
 
 data_dir = './data/preferential-pathway-sensitivity/'
 figures_dir = './figures/transient-response/'
+
+# parameters
+V = 200.0
+
+# reads data file
 data = pd.read_csv(
     data_dir + 'param-no-preferential-pathway.csv',
     header=4,
 )
-
-
-V = 200.0
+# only interested in the case with a gravel sub-base and no contaminant in the
+# pp
 data = data[(data.SB == 1) & (data.chi == 0)]
-n_func = interp1d(data.p.values, data.n.values)
 
 
+def process_data(data):
+    data.n *= 3600.0 # converts mol/s to mol/hr
+    n_func = interp1d(data.p.values, data.n.values) # interpolation function
+    # all possible pressures and air exchanges rates
+    ps = np.arange(-10,11)
+    Aes = np.arange(0.1, 1.6, 0.1)
+    # calculates the indoor air concentration and puts all the input variables
+    # and concentration into lists
+    p, Ae, c, n = [], [], [], []
+    for Ae_now in Aes:
+        for p_now in ps:
+                Ae.append(Ae_now)
+                p.append(p_now)
+                c.append(n_func(p_now)/V/Ae_now)
+                n.append(n_func(p_now))
+    # puts data into dataframe
+    df = pd.DataFrame(
+        {
+        'p': p,
+        'Ae': Ae,
+        'c': c,
+        'n': n,
+        }
+    )
+    # reference state
+    ref_state = ((df.p == -2) & (df.Ae == 0.1))
+    ref = df[ref_state].copy()
+    # potential target states
+    target_state = ((ref.c.values/df.c >= 10.0) | (ref.c.values/df.c <= 0.1)) & (df.p % 5 == 0) & (df.p != 0)
+    target = df[target_state].copy()
+    return df, ref, target
 
-ps = np.arange(-10,11)
-Aes = np.arange(0.1, 1.6, 0.1)
-Aes /= 3600.0
+def plot_contour(df, ref, target):
+    # data to plot
+    data = [
+        go.Contour(
+            z=df.c.apply(np.log10),
+            x=df.p,
+            y=df.Ae,
+        ),
+        go.Scatter(
+            x = ref.p,
+            y = ref.Ae
+        ),
+        go.Scatter(
+            x = target.p,
+            y = target.Ae,
[…]

### 2018-12-12 python/kde-iacc-pressure/get_asu_data.py
Junk

@@ -1,39 +1,0 @@
-import sqlite3
-import pandas as pd
-import matplotlib.pyplot as plt
-
-db_dir = 'C://Users/jstroem/lib/vapor-intrusion-dbs/'
-
-db_asu = sqlite3.connect(db_dir + 'hill-afb.db')
-
-"""asu = pd.read_sql_query(
-    "SELECT \
-        date(C.time) as t, \
-        AVG(C.concentration) AS c, \
-        P.pressure AS p \
-    FROM \
-        \"td-basement\" AS C, \
-        daily_average_pressure AS P \
-    GROUP BY \
-        date(C.time);",
-    db_asu,
-)"""
-
-
-
-asu = pd.read_sql_query(
-    "SELECT \
-        date(C.time) as t, \
-        AVG(C.concentration) AS c, \
-        AVG(P.pressure) AS p \
-    FROM \
-        \"td-basement\" AS C \
-    INNER JOIN daily_average_pressure on date(C.time) == date(daily_average_pressure.time);",
-    db_asu,
-)
-
-print(asu)
-
-
-asu.plot(x='t',y=['c','p'])
-plt.show()

### 2018-12-12 python/transient-response/cstr-reponse.py
Changed pressure limits and csv column names

@@ -13,10 +13,14 @@ figures_dir = './figures/transient-response/'
 
 # parameters
 V = 200.0
-df = pd.read_csv('./data/soils/all-data.csv')
+#df = pd.read_csv('./data/soils/all-data.csv') # no PP data
+
+# pp data stuff
+df = pd.read_csv('./data/preferential-pathway-sensitivity/param-preferential-pathway.csv', header=4) # PP data (only 1 soil)
+df['soil_type'] = np.repeat('sandy-clay',len(df))
+df = df[ (df['SB']==1) & (df['chi']==1) ]
 
 Ae = 0.5
-
 t1, t2 = [], []
 dc1, dc2 = [], []
 Aes = []
@@ -27,8 +31,8 @@ c_max, c_min = [], []
 df.n *= 3600.0
 for Ae in [0.5, 1.0, 1.5]:
     for soil in df['soil_type'].unique():
-        ref = df[ (df['soil_type'] == soil) & ( df.p==-5 ) ]
-        target = df[ (df['soil_type'] == soil) & ( df.p==5 ) ]
+        ref = df[ (df['soil_type'] == soil) & ( df.p==-0.5 ) ]
+        target = df[ (df['soil_type'] == soil) & ( df.p==0.5 ) ]
         Aes.append(Ae)
         soils.append(soil)
         # unsteady cstr method
@@ -41,6 +45,7 @@ for Ae in [0.5, 1.0, 1.5]:
         tau = 240 # max allowed time
         # initial/reference concentration
         y0 = ref.n/V/Ae
+        print(y0)
         c_max.append(y0.values[0])
         for n, p in zip(target.n.values, target.p.values):
             # solving for target state change in variable values
@@ -67,7 +72,7 @@ for Ae in [0.5, 1.0, 1.5]:
             t1.append(t_eq)
             dc1.append(abs(float(c-y0)))
             dcdt1.append(float((c-y0)/t_eq))
-            #print('Min. reached after %2.1f hours' % t_eq)
+            print('Min. reached after %2.1f hours' % t_eq)
             # going back to reference state variables
             n = ref.n
             while solver.y < 0.99*y0.values:
@@ -82,7 +87,7 @@ for Ae in [0.5, 1.0, 1.5]:
             t2.append(t_org)
             dc2.append(abs(float(y0-c)))
[…]

### 2018-12-12 python/transient-response/rate-of-change.py
Changed db dir

@@ -8,7 +8,7 @@ import seaborn as sns
 
 data_dir = './data/preferential-pathway-sensitivity/'
 #db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-db_dir = '/home/jonathan/Dropbox/vapor-intrusion-dbs/'
+db_dir = '/home/jonathan/Dropbox/var/'
 
 #db_dir = 'C://Users/jstroem/lib/vapor-intrusion-dbs/'
 
@@ -176,7 +176,7 @@ custom_x_ticks(ax)
 ax.set_title('PP Closed')
 
 # resampling plots
-# TODO: repeat this analysis but within each season 
+# TODO: repeat this analysis but within each season
 time_scales = ('h', 'D', 'W', 'M')
 
 fig, ax = plt.subplots()

### 2018-12-13 python/transient-response/asu_change_in_concentration.py
Created a bunch of figures

@@ -1,0 +1,231 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.integrate import LSODA
+from scipy.interpolate import interp1d
+from scipy.stats import gaussian_kde
+import sqlite3
+import seaborn as sns
+
+# TODO: add indianapolis data and add to present analysis
+
+data_dir = './data/preferential-pathway-sensitivity/'
+db_dir = '/home/jonathan/Dropbox/var/'
+fig_dir = './figures/rate_of_change/'
+db = sqlite3.connect(db_dir + 'HillAFB.db')
+db_indianapolis = sqlite3.connect(db_dir + 'Indianapolis.db')
+
+def get_season(x):
+    seasons = {
+        'Winter': (12, 2),
+        'Spring': (3, 5),
+        'Summer': (6, 8),
+        'Fall': (9, 11),
+    }
+    if (x == 12) or (x == 1) or (x == 2):
+        return 'Winter'
+    elif (x == 3) or (x == 4) or (x == 5):
+        return 'Spring'
+    elif (x == 6) or (x == 7) or (x == 8):
+        return 'Summer'
+    elif (x == 9) or (x == 10) or (x == 11):
+        return 'Fall'
+    else:
+        return 'Error'
+
+class ProcessData:
+    def __init__(self,db):
+
+        return
+
+    def get_rate_of_change(self, df, time_scale='1h'):
+        df['TimeChange'] = (df['StopTime'].diff()/np.timedelta64(time_scale[0], time_scale[1])).apply(float)
+        df['RateOfChange'] = df['Concentration'].diff()/df['TimeChange']
+        return df
+
+    def get_season(self, df):
+        # finds the month and season of each timestamp
+        asu['Month'] = asu['StopTime'].map(lambda x: x.month)
+        asu['Season'] = asu['Month'].apply(lambda x: get_season(x))
+        return df
+
+def process_asu(resample_time=False):
+    asu = pd.read_sql_query("SELECT * FROM TDBasement;",
+        db,
+    ).sort_values(by='StopTime')
+
+    asu['StopTime'] = asu['StopTime'].apply(pd.to_datetime)
+
+    if resample_time != False:
+        asu = asu.resample(resample_time, on='StopTime', kind='timestamp').mean().reset_index()
+
[…]

### 2018-12-13 python/multivariate_analysis/asu_season.py
Generated figures

@@ -77,4 +77,6 @@ post_cpm.drop(columns=cols_to_drop,inplace=True)
 sns.pairplot(post_cpm, hue="Season", hue_order=['Winter','Fall','Spring','Summer'])
 
 # TODO: change axis labels (adding units) and adjust ticks for the indoor concentration
+plt.savefig('./figures/multivariate_analysis/asu_season.pdf',dpi=300)
+
 plt.show()

### 2018-12-13 python/multivariate_analysis/indianapolis_season.py

@@ -7,7 +7,6 @@ import sqlite3
 import seaborn as sns
 
 data_dir = './data/preferential-pathway-sensitivity/'
-#db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
 db_dir = '/home/jonathan/Dropbox/var/'
 
 db = sqlite3.connect(db_dir + 'Indianapolis.db')
@@ -98,4 +97,5 @@ indianapolis = indianapolis.drop(columns=['StopTime','Month'])
 # TODO: change axis labels (adding units) and adjust ticks for the indoor concentration
 
 sns.pairplot(indianapolis, hue="Season", hue_order=['Winter','Fall','Spring','Summer'])
+plt.savefig('./figures/multivariate_analysis/indianapolis_season.pdf',dpi=300)
 plt.show()

### 2018-12-13 tex/main.tex
Added some new figures

@@
   {./../figures/preferential-pathway-sensitivity/},
-  {./../figures/transient-response/}
+  {./../figures/transient-response/},
+  {./../figures/multivariate_analysis/},
+  {./../figures/rate_of_change/},
 }
@@
   % "Overview" plot
-  \begin{subfigure}{0.55\textwidth}
+  \begin{subfigure}{0.45\textwidth}
     \caption{preferential pathway present. Sensitivity to the presence of a gravel sub-base and contamination in the preferential pathway considered.}
@@
   % Violinplot figure
-  \begin{subfigure}{0.55\textwidth}
+  \begin{subfigure}{0.45\textwidth}
     \caption{No preferential pathway present. Sensitivitiy to the presence of a gravel sub-base considered.}
@@
 
-\begin{comment}
+\subsubsection{Multivariate Analysis}
 
-Clearly, PPs pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover; even at a well studied site like the ASU house it took years to discover the PP.
-Therefore, there is a need to consider other indications that a PP is present.
-
-% Sensitivity analysis plot showing when a PP is relevant
 \begin{figure}[htb!]
-  \caption{Attenuation factor relative to contaminant vapor in equilibrium with groundwater as a function of indoor-outdoor pressure difference.
-The effects of a preferential pathway, that is either contaminated or uncontaminated as well as that of having a gravel sub-base vs.
-uniform sandy clay soil are considered.}
-  \label{fig:model_results_sensitivity_analysis}
-  % Upper left
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_0}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_0.png}
-  \end{subfigure}
-  % Upper right
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_1}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_1.png}
-  \end{subfigure}
-  % Middle left
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_2}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_2.png}
-  \end{subfigure}
-  % Middle right
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_3}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_3.png}
-  \end{subfigure}
-  % Lower left
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_4}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_4.png}
-  \end{subfigure}
-  % Lower right
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:model_results_sensitivity_analysis_5}
-    \includegraphics[width=\textwidth]{ss_pp_sensitivity_analysis_5.png}
-  \end{subfigure}
+  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the ASU site. Only period after land drain was closed considered.}
+  \label{fig:asu_season}
+  \includegraphics[width=\textwidth]{asu_season.pdf}
 \end{figure}
 
-One way has been suggested by the situation at the ASU house, where there clearly was a PP containing contaminant vapor, and the PP exited into a gravel (permeable) sub-base.
-This situation can be examined using the VI model introduced earlier by examining how IACC is impacted by various factors over a range of indoor-outdoor pressure difference values.
-Instead of representing IACC as absolute concentration, we now non-dimensionalize with respect to the vapor in equilibrium with the groundwater contaminant concentration.
-This non-dimensionalized property is commonly called the attenuation factor and is denoted by $\alpha_\mathrm{gw}$.
-The result of a sensitivity analysis relative to several factors is seen in Figure \ref{fig:model_results_sensitivity_analysis}.
-These examine scenarios in which the indoor-outdoor pressure difference is a driver for VI.
-
-In this figure, all the graphs in the left column show results for a scenario, in which the house has a gravel sub-base, with the remaining surrounding soil assumed to be sandy clay.
-All the right column graphs features only sandy clay soil, directly in contact with the slab.
-Figure \ref{fig:model_results_sensitivity_analysis_0} considers an uncontaminated PP ($\chi = 0$), i.e. the PP delivers clean air to beneath the slab.
-In \ref{fig:model_results_sensitivity_analysis_1} there is no PP present at all, giving a reference state for a "normal" VI scenario.
-\ref{fig:model_results_sensitivity_analysis_2} features a contaminated PP ($\chi = 1$), i.e. the PP delivers contaminant vapor at a concentration characteristic of underlying groundwater.
-
-In terms of conditions that may lead to significant temporal transients in $\alpha_\mathrm{gw}$, driven by indoor-outdoor pressure difference, most of the cases in this analysis show negative results.
-There are only two combinations of factors where significant changes in $\alpha_\mathrm{gw}$ accompany the changes in indoor-outdoor pressure difference, both involving the permeable (gravel) sub-base into which a PP enters.
-In \ref{fig:model_results_sensitivity_analysis_2}, the PP is filled with contaminant vapor whereas figure \ref{fig:model_results_sensitivity_analysis} involves the PP only delivering "clean" air to the gravel sub-base.
-When the building is overpressurized (positive indoor-outdoor pressure difference), the two cases exhibit similar behavior, i.e. $\alpha_\mathrm{gw}$ is low as expected.
-
-When the building is underpressurized, the uncontaminated PP case involves barely any increase in $\alpha_\mathrm{gw}$ while there is a very significant increase in $\alpha_\mathrm{gw}$ in the contaminated PP case.
-For the contaminated PP scenario, this is quite easy to understand, as the extra air flow provided by the PP into the gravel sub-base very effectively disperses contaminant vapor across the sub-base and the contaminant entry rate is increased.
-
-When the PP is provides clean air, the increase in air flow into the sub-base with depressurization is the same.
-However, the resulting increase in air entry rate does not bring in more contaminant as clean air is being drawn in.
-This means that most of the contaminant entry rate is still only controlled by the usual process of diffusion through the soil, giving a result that is actually similar to that for the "no PP and with gravel sub-base" scenario.
-Thus, for a PP to be a significant contributor to contaminant entry at a VI site the conditions such as featured in the bottom left case are required, i.e.
-a permeable sub-base and the PP must be a source of contaminant.
-
-Various techniques for determining the potential for sewer gas entry into a house are described by \citeauthor{nielsen_remediation_2017-1}\cite{nielsen_remediation_2017-1}.
-Collecting contaminant vapor samples from nearby manholes is also recommended, however one should be aware that contaminant vapors can potentially travel very long distances in a sewer system\cite{mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-
-% Models specific to ASU & North Island
 \begin{figure}[htb!]
-  \caption{Attenuation factor vs.
-indoor-outdoor pressure difference when considering VI scenarios similar to the ASU house and the North Island NAS site (\ref{fig:ss_asu} and \ref{fig:ss_north_island} respectively).}
-  \label{fig:model_results_asu_north_island}
-  % ASU
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:ss_asu}
-    \includegraphics[width=\textwidth]{ss_asu.png}
-  \end{subfigure}
-  % North Island
-  \begin{subfigure}[t]{0.45\textwidth}
-    \caption{ }
-    \label{fig:ss_north_island}
-    \includegraphics[width=\textwidth]{ss_north_island.png}
-  \end{subfigure}
+  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the Indianapolis site. Only PCE considered.}
+  \label{fig:indianapolis_season}
+  \includegraphics[width=\textwidth]{indianapolis_season.pdf}
 \end{figure}
 
-At the ASU house, after the PP was closed, there was still roughly an order of magnitude variation in IACC, the reasons for which Figure \ref{fig:model_results_sensitivity_analysis} does not address.
-Likewise, there was more than two orders of magnitude observed at the North Island site over the indoor-outdoor pressure difference ranges simulated despite there not being a PP present.
-Therefore, it is of interest to examine these sites a bit further using the model.
-
-The foundation crack at the ASU house seems by all accounts to be quite small, roughly 180x1 cm and located close to the PP\cite{guo_vapor_2015}.
-This crack is smaller than the one assumed in the model used in Figure \ref{fig:model_results_sensitivity_analysis}.
-The influence of indoor-outdoor pressure difference is explored for a case in which a smaller crack area is assumed.
-We also consider the effects of having a significantly lower air exchange rate, 0.1 vs. the regular 0.5 per hour.
-
-The results of the case shown in Figure \ref{fig:model_results_sensitivity_analysis_2} is shown in Figure \ref{fig:ss_asu}.
-Even though there is an increase in $\alpha_\mathrm{gw}$ with increased underpressurization, it is no where near the roughly one order of magnitude recorded at the ASU house after the PP was closed.
-So while crack dimensions can clearly influence $\alpha_\mathrm{gw}$, they cannot be used to infer a bigger role for pressure difference in temporal variations in $\alpha_\mathrm{gw}$ or IACC.
-The influence of air exchange rate on $\alpha_\mathrm{gw}$ predictions is also shown in Figure \ref{fig:model_results_asu_north_island}.
-Clearly, the value of air exchange rate at steady-state conditions will only influence absolute concentrations in the structure.
-Ultimately, the reason for the transient variability in IACC during the post-CPM period at the ASU house remains elusive, as a Pearson's r analysis between IACC and various factors do not yield any explanation.
-
-In the case of the North Island site, variations in IACC are clearly not related to the existence of a PP.
-The soil underlying the North Island site is more sand-like\cite{hosangadi_high-frequency_2017}.
-The significance of this is that sand soil is much more permeable than sandy clay soil, which means that $\alpha_\mathrm{gw}$ will be much more sensitive to changes in indoor-outdoor pressure difference.
-Thus, another "normal" VI scenario was run but this time with sand soil instead of sandy clay, the results of which may be seen in Figure \ref{fig:ss_north_island}.
-Under these conditions, $\alpha_\mathrm{gw}$ changes significantly with pressurization, spanning roughly two orders of magnitude.
-This is consistent with what was observed at North Island, as there the span of IACC values was roughly the same across this range of indoor-outdoor pressure difference values.
-This simulation, together with the Pearson $r = -0.64$ for North Island in Figure \ref{fig:kde-asu-nas} suggest that under conditions where the soil around a structure is relatively permeable, indoor-outdoor pressure difference can be a significant driving forces for transient changes in IACC.
-Therefore, recording the indoor-outdoor pressure difference at a site (in particular if the soil is permeable) may offer insight into the potential for transient changes in IACC (some good methods for doing this are suggested by \citeauthor{nielsen_remediation_2017-1}\cite{nielsen_remediation_2017-1}).
-
-% Transient part
-\subsubsection{Transient Modeling}
-
-Analyzing the impact of a PP using steady-state simulations such as those in Figures \ref{fig:model_results_sensitivity_analysis} and does not revealing about how quickly IACC can change in the presence of a PP.
-A transient simulation of a VI scenario that is characterized by a PP has been performed, where only the indoor-outdoor pressure difference was temporally changed.
-The pressure difference is changed in a sinusoidal fashion, defined by \eqref{eq:p_transient},
-\begin{equation}
-  \label{eq:p_transient}
-  p_\mathrm{in/out} = 10 \sin{(2 \pi t)} - 0.5
-\end{equation}
-where $t$ is given in days and $p_\mathrm{in/out}$ is the indoor-outdoor pressure difference in Pa.
-Figure \ref{fig:transient_simulation} shows the results of this simulation.
-
-\begin{figure}
-  \caption{Transient response of TCE in indoor air (as attenuation factor) and in the surrounding soil \& gravel sub-base for a PP VI scenario subject to sinusoidal indoor-outdoor pressure fluctuations.
-  \ref{fig:alpha_response} shows the attenuation factor and indoor-outdoor pressure across time.
-  Figures \ref{fig:soil_30_hr} \& \ref{fig:soil_42_hr} show the contaminant concentration (normalized to the contaminant vapor in equilibrium with groundwater) in the surrounding soil \& gravel sub-base at different depths and times.}
-  \label{fig:transient_simulation}
-  % Transient response figure
-  \begin{subfigure}[t]{0.75\textwidth}
-    \caption{ }
-    \label{fig:alpha_response}
-    \includegraphics[width=\textwidth]{alpha_transient.png}
-  \end{subfigure}
-  % Soil t = 30 hr ...
-  \begin{subfigure}[c]{0.45\textwidth}
-    \caption{t = 30 hr}
-    \label{fig:soil_30_hr}
-    \includegraphics[width=\textwidth]{tri_layer_soil_concs1.png}
-  \end{subfigure}
-  % Soil t = 42 hr ...
-  \begin{subfigure}[c]{0.45\textwidth}
-    \caption{t = 42 hr}
-    \label{fig:soil_42_hr}
-    \includegraphics[width=\textwidth]{tri_layer_soil_concs2.png}
-  \end{subfigure}
-  % Color map
-  \begin{subfigure}[c]{0.05\textwidth}
-    \phantomcaption{ }
-    %\label{fig:cmap}
-    \includegraphics[width=\textwidth]{cmap.png}
-  \end{subfigure}
-\end{figure}
-
-Figure \ref{fig:alpha_response} shows the IACC response, given as attenuation factor relative to contaminant vapor in equilibrium with groundwater, and the indoor-outdoor pressure difference with time.
-What is apparent is that $\alpha_\mathrm{gw}$ closely follows the indoor-outdoor pressure difference, with the max./min.
-$\alpha_\mathrm{gw}$ being reached roughly 0.5 hours after the indoor-outdoor pressure difference reaches its max./min..
-The reader should be aware that the assumption that the indoor air space can be modeled as a CST overestimates how dynamic this process is as a real house probably would not be perfectly mixed on as a short time scale as this.
-Regardless, this suggests that VI sites characterized by PPs are highly dynamic and IACC's may change quickly in response to changing indoor-outdoor pressure differences, consistent with what was observed at the ASU house\cite{holton_temporal_2013}.
-Again, the dynamics of the indoor-outdoor pressure difference at a VI site can offer insight into the potential for transient behavior of IACC.
-
-Figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} show the contaminant concentration (normalized to the contaminant vapor in equilibrium with groundwater) in the soil and gravel sub-base at three different depths at times of 30 and 42 hr respectively corresponding to roughly the maximum in overpressurization and depressurization respectively.
-The top layer is 3 m above groundwater and immediately under the house foundation (indicated by the black square).
-The middle and bottom layers are 2 and 1 m above groundwater respectively.
-The black streamlines show the contaminant transport flow paths.
-The region closest to the PP is expanded in the circular insets.
-
-The impact of the PP on the contaminant vapor concentration in the subsurface is apparent in these figures, as one easily sees how (especially) the gravel sub-base is filled up.
-The contaminant vapor concentration is the highest closest to the PP (which is located at the edge of the gravel sub-base) and decreases with increased distance from the PP.
-Note that a very small amount of contaminant vapor leaves the gravel sub-base, as the sandy clay has a relatively large resistance to transport compared with gravel.
-This makes it difficult to detect the contribution of the PP outside the foundation footprint.
-However, this may not always be true, as vapor contaminant from the PP would disperse more easily into more permeable soils.
-
-It should also be noted that the vapor contaminant emanating from the PP in this model is relatively high concentration; the vapor contaminant concentration is equal to that in equilibrium with the groundwater source.
-This does not always seem to be the case, as VI sites that have uncovered PPs find that the vapor contaminant emanating from the PP can be lower than that\cite{guo_vapor_2015}.
-Therefore, it may be much more difficult than it may appear in Figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} to localize a vapor contaminant hotspot as a way to determining the presence of a PP at a particular VI site.
-
-The insert circles in figures \ref{fig:soil_30_hr} and \ref{fig:soil_42_hr} also show that for the simulated time period, only a small amount of the contaminant enters the structure (as evident by the blue low concentration zone around the perimeter), yet a very large increase in $\alpha_\mathrm{gw}$ occurs.
-The concentration in the rest of the sub-base is unchanged.
-This suggests that once a PP has deposited contaminant underneath a structure, it may persist for a long time period, as pointed out by \citeauthor{guo_vapor_2015}\cite{guo_vapor_2015}.
-
-\end{comment}
+Pressure alone does not seem adequately be able to predict the variability observed at the ASU and Indianapolis VI sites, necessitating a multivariate analysis to better understand the distribution of indoor air contaminant concentration.
+% Add references motivating the choice of parameters
+This is acheived by examining the distribution various parameters recorded at these sites, and how they relate to eachother.
+It is known that VI potential differs across the seasons, thus the seasonal distribution of each parameter is considered.
 
@@
 \end{acknowledgement}
+
+\subsubsection{Rates of Change in Indoor Air Contaminant Concentration}
+
+\begin{figure}[htb!]
+  \caption{}
+  \label{fig:}
+  % "Overview" plot
+  \begin{subfigure}{0.45\textwidth}
+    \caption{}
+    \label{fig:}
+    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_open.pdf}
+  \end{subfigure}
+  % Violinplot figure
+  \begin{subfigure}{0.45\textwidth}
+    \caption{}
+    \label{fig:}
+    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_closed.pdf}
+  \end{subfigure}
+  % Indianapolis
+  \begin{subfigure}{0.45\textwidth}
+    \caption{}
+    \label{fig:}
+    \includegraphics[width=\textwidth]{rate_of_change_season_indianapolis.pdf}
+  \end{subfigure}
+  % Indianapolis
+  \begin{subfigure}{0.45\textwidth}
+    \caption{}
+    \label{fig:}
+    \includegraphics[width=\textwidth]{rate_of_change_simulations.pdf}
+  \end{subfigure}
+\end{figure}
 

### 2018-12-13 tex/main.tex
Finished writing multivariate draft

@@
 \begin{figure}[htb!]
-  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the ASU site. Only period after land drain was closed considered.}
+  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the ASU site. Only period after the land drain was closed considered.}
   \label{fig:asu_season}
@@
 Pressure alone does not seem adequately be able to predict the variability observed at the ASU and Indianapolis VI sites, necessitating a multivariate analysis to better understand the distribution of indoor air contaminant concentration.
-% Add references motivating the choice of parameters
+% TODO: Add references motivating the choice of parameters
 This is acheived by examining the distribution various parameters recorded at these sites, and how they relate to eachother.
 It is known that VI potential differs across the seasons, thus the seasonal distribution of each parameter is considered.
+Winter is defined as from December to Feburary, spring from March to April, summer from June to August and fall from September to November.
+For the ASU site dataset, only the period after the land drain was closed is considered, and the Indianapolis site data only considers the period before the SSD system was installed, and only considers the concentration of PCE.
+The distributions of parameters that these two sites share are:
+\begin{itemize}
+  \item Daily average of the logarithmic indoor air contaminant concentration (TCE at ASU and PCE at Indianapolis).
+  \item Daily average groundwater concentration.
+  \item Daily average indoor/outdoor pressure difference.
+  \item Daily average vapor concentration in the immediate sub-foundation.
+\end{itemize}
+At the ASU site, the daily average air exchange rate is also given, while at the Indianapolis the daily average outdoor temperature is given.
+All of this is shown in Figures (\ref{fig:asu_season}) and (\ref{fig:indianapolis_season}). \par
+
+On the diagonal plots in both of these figures, the KDE distributions of each parameter is plotted, with the y-axis denotes the probability density and the x-axis the numeric value associated with that probability density.
+The off-diagonals plots are symmetrical, and show the relationship between two parameters.
+E.g. in the bottom left plot in Figure (\ref{fig:asu_season}), the relationship between the subfoundation TCE concentration and air exchange rate is shown.
+A tight cluster resembling a straight line would indicate a linear relationship between the two factors, and the opposite would indicate no real relationship between the two factors exists. \par
+
+What is immediately apparent from these two figures is that there is a very clear seasonal discrepancy between most of the parameters plotted.
+Most significant of which is that the indoor air contaminant concentration is the lowest and exhibit the least amount of variation during summer.
+During winter the opposite may be seen, where indoor air contaminant concentations are generally higher, and exhibit higher variabilty.
+There is roughly an order of magnitude difference in indoor air concentration between summer and winter, with a few factors (2-3) of variabilty during summer and close to an order of magnitude during winter (albeit this is quite rare).
+Indoor air contaminant concentrations fall somewhere inbetween this for the other seasons.
+This suggests that maximum VI potential is achieved during winter or colder periods, indicating this may be the best period for VI site investgations, although one be aware of the larger variability during this period too.
+\par
+
+The reasons for these seasonal trends is not clear by a simple univariate analysis, as has been previously discussed.
+This is supported by looking at the relationships between the TCE in indoor air and other parameters in Figure (\ref{fig:asu_season}), where no one clear linear relationship exists, but there are some general correlations that can be made, e.g.
+one can clearly see that higher depressurziation is associated with higher TCE in indoor air.
+The TCE in indoor air during summer is a good example of how multiple factors play a role here to determine VI.
+For instance, the subfoundation concentration is the highest during this period, and the air exchange rate is the lowest, which would indicate that the highest concentrations of TCE in indoor air would be found during summer.
+But, by looking at the pressure distributions, the house is practically always overpressurized, preventing contaminant entry and reducing variability. \par
+
+At both of these sites, summer is characterized by low contaminant concentrations and low variability as well as a tendency for the structures to be overpressurized, as can be seen in Figure (\ref{fig:indianapolis_season}).
+This is likely due to less heating required during warmer periods, removing the stack effect.
+It is also interesting to note that at the ASU site, the air exchange rate is the lowest during warmer periods, which is the opposite of what would be expected, as windows and doors are more likely to be open during these periods, increasing air exchange.
+However, increasing air exchange rate would simply further decrease indoor air contaminant concentrations during these periods, which is consistent with the previous conclusions.
+% TODO: Add reference to the Ae found at different sites during different seasons
+
+\subsubsection{Rates of Change in Indoor Air Contaminant Concentration}
+
+\begin{figure}[htb!]
+  \caption{Hourly changes in (logarithmic) indoor air contaminant concentration at the ASU and Indianapolis sites}
+  \label{fig:rate_of_change}
+  % "Overview" plot
+  \begin{subfigure}{0.45\textwidth}
+    \caption{ASU site, PP open. Seasonal effect considered.}
+    \label{fig:rate_of_change_season_asu_pp_open}
+    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_open.pdf}
+  \end{subfigure}
+  % Violinplot figure
+  \begin{subfigure}{0.45\textwidth}
+    \caption{ASU site, PP closed. Seasonal effect considered.}
+    \label{fig:rate_of_change_season_asu_pp_closed}
+    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_closed.pdf}
+  \end{subfigure}
+  % Indianapolis
+  \begin{subfigure}{0.45\textwidth}
+    \caption{Indianapolis site, only PCE considered. Seasonal effect considered.}
+    \label{fig:rate_of_change_season_indianapolis}
+    \includegraphics[width=\textwidth]{rate_of_change_season_indianapolis.pdf}
+  \end{subfigure}
+  % Indianapolis
+  \begin{subfigure}{0.45\textwidth}
+    \caption{ASU and Indianapolis sites, compared to predicted rates of changes.}
+    \label{fig:rate_of_change_simulations}
+    \includegraphics[width=\textwidth]{rate_of_change_simulations.pdf}
+  \end{subfigure}
+  % "Overview" plot
+  \begin{subfigure}{0.45\textwidth}
+    \caption{Different resampling frequencies of the Indianapolis site data of the entire ASU site data.}
+    \label{fig:resampling_asu}
+    \includegraphics[width=\textwidth]{resampling_asu.pdf}
+  \end{subfigure}
+  % Violinplot figure
+  \begin{subfigure}{0.45\textwidth}
+    \caption{Different resampling frequencies of the Indianapolis site data, only PCE considered.}
+    \label{fig:resampling_indianapolis}
+    \includegraphics[width=\textwidth]{resampling_indianapolis.pdf}
+  \end{subfigure}
+\end{figure}
+
+Most of the analysis so far has investigated how various factors impact indoor air contaminant concentrations, but this does says nothing about how quickly this can change across time.
+Knowing how much indoor air contaminant concentrations may change over a given time period is important to understand reduce the number of times indoor air samples have to be collected to estabilish some certainty that the relevant indoor air contaminant concentration has been captured.
+\par
+
+To examine this, the ASU and Indianapolis datasets are used again, and the
+
+\begin{equation}
+  \frac{\Delta\log{(c)}}{\Delta t} = \frac{\log{(c(t_2))} - \log{(c(t_1))}}{t_2 - t_1}
+\end{equation}
 
@@
 \end{acknowledgement}
-
-\subsubsection{Rates of Change in Indoor Air Contaminant Concentration}
-
-\begin{figure}[htb!]
-  \caption{}
-  \label{fig:}
-  % "Overview" plot
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_open.pdf}
-  \end{subfigure}
-  % Violinplot figure
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_closed.pdf}
-  \end{subfigure}
-  % Indianapolis
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_indianapolis.pdf}
-  \end{subfigure}
-  % Indianapolis
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_simulations.pdf}
-  \end{subfigure}
-\end{figure}
 

### 2018-12-13 python/kde-iacc-pressure/ase-seasons.py
Referencing dataset average instead of seasonal average and changed name

@@ -11,9 +11,9 @@ import matplotlib.patches as mpatches
 
 data_dir = './data/preferential-pathway-sensitivity/'
 #db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-db_dir = '/home/jonathan/Dropbox/vapor-intrusion-dbs/'
+db_dir = '/home/jonathan/Dropbox/var/'
 
-db = sqlite3.connect(db_dir + 'hill-afb.db')
+db = sqlite3.connect(db_dir + 'HillAFB.db')
 
 indoor = pd.read_sql_query( "SELECT * FROM DailyAverages;", db, )
 soil_gas = pd.read_sql_query( "SELECT * FROM AverageSubSurfaceSoilGasConcentration;", db, )
@@ -93,7 +93,7 @@ for dataset in (pre_cpm,post_cpm,):
         df = dataset[dataset['Season']==season]
         print("Dataset: %s, Season: %s" % (dataset.name,season))
         # TODO: use dataset mean instead of seasonal mean
-        df['IndoorConcentration'] -= df['IndoorConcentration'].mean()
+        df['IndoorConcentration'] -= dataset['IndoorConcentration'].mean()
 
         if dp is True:
             df['p'] -= df['p'].mean()
@@ -156,7 +156,7 @@ for dataset in (pre_cpm,post_cpm,):
     g.ax_joint.set_yticklabels(my_ytick_labels)
     g.ax_joint.legend(handles=label_patches, loc='upper left')
     plt.tight_layout()
-    plt.savefig('./figures/kde-iacc-pressure/%s-seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
+    plt.savefig('./figures/kde-iacc-pressure/%s_seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
     # TODO: dynamically change axis limits?
     #plt.show()
     plt.clf()

### 2018-12-13 python/kde-iacc-pressure/indianapolis-seasons.py

@@ -11,9 +11,9 @@ import matplotlib.patches as mpatches
 
 data_dir = './data/preferential-pathway-sensitivity/'
 #db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-db_dir = '/home/jonathan/Dropbox/vapor-intrusion-dbs/'
+db_dir = '/home/jonathan/Dropbox/var/'
 
-db = sqlite3.connect(db_dir + 'indianapolis.db')
+db = sqlite3.connect(db_dir + 'Indianapolis.db')
 
 observation = pd.read_sql_query(
     "SELECT StopDate AS StopTime FROM Observation_Status_Data WHERE Variable='Mitigation' AND Value='not yet installed';", db, )
@@ -109,8 +109,8 @@ for dataset in (indianapolis,):
         #dataset = dataset.name
         df = dataset[dataset['Season']==season]
         print("Dataset: %s, Season: %s" % (dataset.name,season))
-        # TODO: use dataset mean instead of seasonal mean 
-        df['IndoorConcentration'] -= df['IndoorConcentration'].mean()
+        # TODO: use dataset mean instead of seasonal mean
+        df['IndoorConcentration'] -= dataset['IndoorConcentration'].mean()
 
         if dp is True:
             df['p'] -= df['p'].mean()
@@ -173,7 +173,7 @@ for dataset in (indianapolis,):
     g.ax_joint.set_yticklabels(my_ytick_labels)
     g.ax_joint.legend(handles=label_patches, loc='upper left')
     plt.tight_layout()
-    plt.savefig('./figures/kde-iacc-pressure/%s-seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
+    plt.savefig('./figures/kde-iacc-pressure/%s_seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
     # TODO: dynamically change axis limits?
     #plt.show()
     plt.clf()

### 2018-12-24 tex/main_old.tex
Added backup

--- 2018-12-13 tex/main.tex
@@
 
-Pressure alone does not seem adequately be able to predict the variability observed at the ASU and Indianapolis VI sites, necessitating a multivariate analysis to better understand the distribution of indoor air contaminant concentration.
+Indoor/outdoor pressure differences or air exchange rate alone do not seem adequately be able to predict the variability observed at the ASU and Indianapolis VI sites, necessitating a multivariate analysis to better understand the distribution of indoor air contaminant concentration.
 % TODO: Add references motivating the choice of parameters
 This is acheived by examining the distribution various parameters recorded at these sites, and how they relate to eachother.
-It is known that VI potential differs across the seasons, thus the seasonal distribution of each parameter is considered.
-Winter is defined as from December to Feburary, spring from March to April, summer from June to August and fall from September to November.
-For the ASU site dataset, only the period after the land drain was closed is considered, and the Indianapolis site data only considers the period before the SSD system was installed, and only considers the concentration of PCE.
 The distributions of parameters that these two sites share are:
@@
 At the ASU site, the daily average air exchange rate is also given, while at the Indianapolis the daily average outdoor temperature is given.
-All of this is shown in Figures (\ref{fig:asu_season}) and (\ref{fig:indianapolis_season}). \par
+All of this is shown in Figures (\ref{fig:asu_season}) and (\ref{fig:indianapolis_season}).
+It is known that VI potential differs across the seasons, thus the seasonal distribution of each parameter is considered.
+Winter is defined as from December to Feburary, spring from March to April, summer from June to August and fall from September to November.
+For the ASU site dataset, only the period after the land drain was closed is considered, and the Indianapolis site data only considers the period before the SSD system was installed, and only considers the concentration of PCE.\par
 
@@
 The off-diagonals plots are symmetrical, and show the relationship between two parameters.
-E.g. in the bottom left plot in Figure (\ref{fig:asu_season}), the relationship between the subfoundation TCE concentration and air exchange rate is shown.
+E.g. in the bottom left plot in Figure \ref{fig:asu_season}, the relationship between the subfoundation TCE concentration and air exchange rate is shown.
 A tight cluster resembling a straight line would indicate a linear relationship between the two factors, and the opposite would indicate no real relationship between the two factors exists. \par
@@
 What is immediately apparent from these two figures is that there is a very clear seasonal discrepancy between most of the parameters plotted.
-Most significant of which is that the indoor air contaminant concentration is the lowest and exhibit the least amount of variation during summer.
+Most significant of which is that the indoor air contaminant concentration is the lowest, and exhibit the least amount of variation during summer.
 During winter the opposite may be seen, where indoor air contaminant concentations are generally higher, and exhibit higher variabilty.
@@
 Indoor air contaminant concentrations fall somewhere inbetween this for the other seasons.
-This suggests that maximum VI potential is achieved during winter or colder periods, indicating this may be the best period for VI site investgations, although one be aware of the larger variability during this period too.
+This suggests that maximum VI potential is achieved during winter or colder periods, indicating this may be the best period for VI site investgations, although one should be aware of the larger variability during this period.
 \par
@@
 The reasons for these seasonal trends is not clear by a simple univariate analysis, as has been previously discussed.
-This is supported by looking at the relationships between the TCE in indoor air and other parameters in Figure (\ref{fig:asu_season}), where no one clear linear relationship exists, but there are some general correlations that can be made, e.g.
+This is supported by looking at the relationships between the TCE in indoor air and other parameters in Figure \ref{fig:asu_season}, where no one clear linear relationship exists, but there are some general correlations that can be made, e.g.
 one can clearly see that higher depressurziation is associated with higher TCE in indoor air.
-The TCE in indoor air during summer is a good example of how multiple factors play a role here to determine VI.
+The TCE in indoor air during summer is a good example of how multiple factors can play a complex role in determining VI potential.
 For instance, the subfoundation concentration is the highest during this period, and the air exchange rate is the lowest, which would indicate that the highest concentrations of TCE in indoor air would be found during summer.
@@
 
-Most of the analysis so far has investigated how various factors impact indoor air contaminant concentrations, but this does says nothing about how quickly this can change across time.
-Knowing how much indoor air contaminant concentrations may change over a given time period is important to understand reduce the number of times indoor air samples have to be collected to estabilish some certainty that the relevant indoor air contaminant concentration has been captured.
+Most of the analysis so far has investigated how various factors impact indoor air contaminant concentrations, but this revleas little regarding how quickly this can change over time.
+Knowing how much indoor air contaminant concentration may change over a given time period is important to reduce the number of times indoor air samples have to be collected to establish the relevant indoor air contaminant concentration.
 \par
 
-To examine this, the ASU and Indianapolis datasets are used again, and the
+To examine this, the ASU and Indianapolis datasets are used again, where the rate of change in indoor air contaminant concentration is calculated, and probability distributions for these valuese given using the KDE method.
+The rate of change in indoor air contaminant concentration is given by \eqref{eq:rate_of_change}
+\begin{equation}
+  \frac{\Delta\log{(c)}}{\Delta t} = \frac{\log{(c(t_2))} - \log{(c(t_1))}}{t_2 - t_1} \label{eq:rate_of_change}
+\end{equation}
+where $c$ is the indoor air contaminant concentration, $t$ is some timeperiod where the concentration is known and $\log$ is the base-10 logarithm.
+In the ASU dataset, indoor air contaminant concentrations were (for the most part) taken at 4-hour intervals\cite{holton_long-term_2015}, while at the Indianapolis site samples were not collected with the same regularity, but several samples were taken everyday for any given sampling location\cite{u.s._environmental_protection_agency_assessment_2015}, giving high time-resolution datasets to work with.
 
-\begin{equation}
-  \frac{\Delta\log{(c)}}{\Delta t} = \frac{\log{(c(t_2))} - \log{(c(t_1))}}{t_2 - t_1}
-\end{equation}
+In Figure \ref{fig:rate_of_change_season_asu_pp_open} and \ref{fig:rate_of_change_season_asu_pp_closed} the distributions of the rate of change on an hourly basis, i.e.
+how much the contaminant concentration changed over one hour ($t_2 - t_1 = 1 \; \mathrm{hour}$), is shown divided up into each season.
+Both of the figures consider non-CPM periods, but Figure \ref{fig:rate_of_change_season_asu_pp_open} only contains data from the period when the preferential pathway still open and \ref{fig:rate_of_change_season_asu_pp_closed} data from after the the preferential pathway was closed.
+
+These distributions show that on an hourly basis, only minor changes in indoor air contaminant concentration may be expected; roughly one could at most expect at 50\% increase and a 25\% decrease in indoor air contaminant concentration over a hour at this site.
+No significant seasonal difference in the rate of changes seems to exist within either period, except for during summer after the land drain was closed, where greater rates of change indoor air contaminant concentration were more prevalent, albeit still minor.
+The preferential pathway had a minor impact on the rate of change, increasing the probability of relatievly greater changes, but again only minor differences.
+
+In Figure \ref{fig:rate_of_change_season_indianapolis} the distribution of hourly rate of change in PCE concentration at the Indianapolis site is shown.
+The conclusions from the ASU house hold here as well, with only minor rates of change observed and no seasonal difference.
+
+This suggests that no significant changes in indoor air contaminant concentration may be expected on a hourly basis regardless of the presence of a preferential pathway as well as the hourly rate of change is seasonally independent.
+
+Indoor air contaminant concentrations may however change much more significantly after a longer time period.
+This may be investigated by resampling the two datasets over different time periods and aggregating by the mean concentration within each time period.
+The rate of change is again calculated using \eqref{eq:rate_of_change}, but unlike in the previous figures, $t_2 - t_1 = 1$ day, week or month.
+
+Since there was no major difference in rate of change in indoor air concentratin between the preferential pathway open/closed periods, the entire non-CPM period of the ASU house dataset is used for this analysis.
+Figure \ref{fig:resampling_asu} and \ref{fig:resampling_indianapolis} shows the distributions of rate of change in indoor air contaminant concentration resampled over a daily, weekly and monthly basis at the ASU house and Indianapolis house respectively.
+
+As one might expect, the probability of a larger change in indoor air contaminant increases, as the resampling timeperiod increases.
+Only minor changes in indoor air concentration may be expected on a daily basis, albeit larger changes are more likely to occur than on an hourly basis.
+At the ASU site, as the resampling timeperiod increases, larger changes become more probable, where on a monthly resampling period half an order magnitude change start to become probable.
+At the Indianapolis site, the rate of change over the resampling periods are smaller, and less probable than at the ASU site.
+But similarily, after a monthly period, larger changes begin to become more prevalent.
+
+One should keep in mind that by using the mean to aggregate the indoor air concentration, there is a tendency for data to be "smoothed out", missing some of the outliers, especially as the resampling period increases.
+Regardless, this analysis suggests that sampling indoor air on a daily basis may be unnecessary as little new information may be gained.
+And sampling on a monthly basis may be too long of a period, and important information may be missed.
 

### 2018-12-26 tex/main.tex
Created new version of manuscript

--- 2018-12-13 tex/main.tex
@@
 \abbreviations{VI}
-\keywords{VI, Preferential pathways, Temporal transient variability}
+\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Factor analysis, Modeling}
 
@@
 \begin{abstract}
-Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the variability at VI sites located near Hill Air Force Base in Utah, an EPA experimental house in Indianapolis, IN and Naval Air Station North Island in California using statistical analysis methods and three-dimensional subsurface computational fluid dynamics modeling.
-The results suggest that an order of magnitude variation in indoor air contaminant concentrations may be expected at ”normal” VI sites where preferential pathways do not play a role, whereas three or more orders of magnitude can be observed at sites characterized by preferential pathways.
-A computational fluid dynamics modeling sensitivity analysis reveals that it is not only the presence of contaminant vapor in a preferential pathway that is required to see large observed variations, but there must also exist a permeable region beneath the structure.
-Large temporal fluctuations in indoor air contaminant concentrations may be observed where no preferential pathways exist, but this requires a particular combination of large fluctuations in pressure driving force and high soil permeability.
+
 \end{abstract}
@@
 
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-VI has been a recognized problem for some time, but many aspects remain poorly understood, particularly with respect to the causes of large temporal transients in indoor air concentrations.
-There is uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-\par
-
-Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
-All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations varied significantly with time - orders of magnitude on the timescale of a day or days.
-\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}. \par
-
-In one instance the source of the variation was clearly established during the study; at the ASU house a field drain pipe (or “land drain”), which connected to a sewer system, was discovered beneath the house, and careful isolation of this source led to a clear conclusion that this preferential pathway significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway to an area beneath a structure would always be expected to lead to large fluctuations in indoor air contaminant concentrations.
-\par
-
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
-That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been generally implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study estimate that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
-In some of these cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant via the soil beneath a structure. \par
-
-It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is a site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site. \par
-
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as ”normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at  such normal  VI  sites vs. those characterized by preferential pathways.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.
-
 \section{Methods}
 
-\subsection{Statistical Analysis Of Field Data}
+\subsection{Statistical Analysis}
 
-To begin to characterize transient behavior in indoor air contaminant concentrations, actual datasets are analyzed to establish common levels of variability at VI sites.
-For this purpose, the datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-This paper relies on statistical analysis of published field data, and readers are referred to the original works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.
-\par
+\subsubsection{Indianapolis Dataset}
 
-The ASU house data were obtained over a period of a few years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” operation.
-This caused greater than normal advective flow from the subsurface into the house, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-This period of CPM testing is considered separately from the otherwise ”natural” VI conditions in the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining the dataset, noting that during some of the testing, this pathway was deliberately cut off, resulting in what we have termed “normal” VI conditions in which the main source of contaminant was believed to be groundwater.
-\par
+\subsubsection{"ASU House" Dataset}
 
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to large internal pressure fluctuations, much more extensive than those typically recorded at the ASU house during normal operations.
-Additionally, the underlying soil at NAS North Island is sandy and more permeable than that at the ASU site, which, as will be shown, contributes to the indoor air contaminant concentrations being more sensitive to pressure fluctuations\cite{hosangadi_high-frequency_2017}.
-\par
-
-Likewise, the Indianapolis site investigation spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
-The goal of the SSD testing was to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above.
-Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this preferential pathway removed, making it difficult to assess how significant the role of the preferential pathway was at this site.
-Regardless of this it is of interest to consider the data from this site due to how extensive and complete the data collection was. \par
-
-The typical variation in indoor air contaminant concentrations with time will first be considered below in the case of the ASU house during ”natural”, (i.e.
-non-CPM conditions), in the case of the NAS North Island site over the entire available dataset, and for the Indianapolis case we consider the variations before the installation of the SSD system.
-The deviations in indoor air concentration from the mean TCE (and Chloroform and PCE at the Indianapolis site) values, as well as the indoor-outdoor pressure differentials associated with these concentrations were examined.
-Both univariate and bivariate kernel density estimations (KDE) were constructed.
-KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (and that there are statistical fluctuations associated with each sampling event).
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
-The distributions of the individual parameters and the relationship between them will be examined using the KDE method.
-
-\subsection{Modeling Work}
-In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
-This model was implemented in a finite element solver package, COMSOL Multiphysics.
-In the present work, there has been an addition of a preferential pathway to the ”standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-
-The modeled structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is assumed to consist of only of the basement, assumed as having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of certain other features can be.
-\par
-
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy clay (except as noted).
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
-\par
-
-Where relevant The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
-The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, allowing the contaminant concentration in the pipe to be parameterized.
-\par
-
-The groundwater beneath the structure is assumed to be homogeneously contaminated with trichloroethylene (TCE) as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-The ground surface and the pipe are both assumed to be sources of air to the soil domain.
-Both are assumed to be at reference atmospheric pressure. \par
-
-Vapor transport in the soil is governed by Richard’s equation, a modified version  of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
-The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
-The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of vapor contaminant in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The equations and the boundary conditions are given in Table \ref{tbl:eqns-bc-parameters}.
-
-\begin{table}[htb!]
-  \centering
-  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}
-  \label{tbl:eqns-bc-parameters}
-  \bigskip
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  % Governing equations
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subcaption{Governing equations}
-  \begin{tabular}{l l}
-    \toprule
-    % Indoor air space equation
-    Unsteady-CSTR                 & $V\frac{d u}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - u A_e V$ \\
-    % Richard's equation
-    Richard's equation            & $\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$ \\
-    % Millington-Quirk equation
-    Millington-Quirk              & $D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$ \\
-    % Advection-diffusion equation
-    Advection-diffusion equation  & $\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$ \\
-    % van Genuchten's equations
-    \multirow{3}{*}{van Genuchten equations}     & $\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$ \\
-                                  & $\theta_g = \theta_t - \theta_w$ \\
-                                  & $k_r = (1 - \mathrm{Se})^{l} [1 - (\mathrm{Se}^{-m})^m]^2$ \\
-                                  & $m = 1 - 1/n$ \\
-    \bottomrule
-  \end{tabular}
-  \bigskip
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  % Boundary conditions
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subcaption{Boundary conditions}
-  \begin{tabular}{l l l}
-    \toprule
-    \textbf{Boundary}          & \textbf{Richard's equation}      &   \textbf{Advection-diffusion equation} \\
-    % Foundation crack
-    At foundation crack  & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$                            & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
-    % Groundwater source
-    At groundwater source &  N/A & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$ \\
-    % Ground surface
-    At ground surface      & $p = 0 \; \mathrm{(Pa)}$  & $c = 0 \; \mathrm{(\mu g/m^3)}$ \\
-    % Preferential pathway
-    Exit of preferential pathway  & $p = 0 \; \mathrm{(Pa)}$  & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$ \\
-    \bottomrule
-  \end{tabular}
-  \bigskip
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  % Soil input parameters
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subcaption{Soil \& gravel properties\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
-  \begin{tabular}{l l l l l l l}
-    \toprule
-    % Descriptions
-    Soil & $\text{Permeability} \; \mathrm{(m^2)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$  & $\theta_s$  & $\theta_r$  & $\alpha \; \mathrm{(1/m)}$  & $n$ \\
-    % Gravel
-    Gravel     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
-    % Sand
-    Sand     & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053        & 3.5       & 3.2 \\
-    % Sandy clay
-    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
-    \bottomrule
-  \end{tabular}
-  \bigskip
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  % TCE input parameters
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subcaption{Trichloroethylene (diluted in air) properties\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
-  \begin{tabular}{l l l l l l}
-    \toprule
-    $D_\mathrm{air} \; \mathrm{(m^2/h)}$  & $D_\mathrm{water} \; \mathrm{(m^2/h)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$ & $\mathrm{Viscosity} \; \mathrm{(Pa \cdot s)}$  & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
-    $2.47 \cdot 10^{-2}$  & $3.67 \cdot 10^{-6}$  & 1.614 & $1.86 \cdot 10^{-5}$  & 0.403 & 131.39 \\
-    \bottomrule
-  \end{tabular}
-  \bigskip
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  % Building input parameters
-  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \subcaption{Building properties}
-  \begin{tabular}{l l l}
-    \toprule
-    $V_\mathrm{base} \; \mathrm{(m^3)}$  & $L_\mathrm{slab} \; \mathrm{(cm)}$  & $A_e \; \mathrm{(1/hr)}$ \\
-    %
-    300  &  15  & 0.5 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\subsection{Drivers For Indoor Air Contaminant Variability}
-\subsubsection{The Role Of Building Depressurization/Pressurization}
-% KDE figures
-\begin{figure}[htb!]
-  \caption{KDE analysis of IACC dependence on indoor/outdoor pressure difference at the ASU house and North Island site (\ref{fig:kde-asu-nas}) and the Indianapolis site (\ref{fig:kde-indianapolis}).
-p-values and Pearson's r-values shown for each dataset.}
-  \label{fig:kde-analysis}
-  % ASU/North Island plot
-  \begin{subfigure}{\textwidth}
-    \centering
-    \caption{Period before and after the preferential pathway (PP) was shut at the ASU house considered separately; North Island dataset considered in its entirety.}
-    \label{fig:kde-asu-nas}
-    \includegraphics[height=0.4\textheight,keepaspectratio]{nas-asu-house-iacc-deviating.pdf}
-  \end{subfigure}
-  % Indianapolis plot
-  \begin{subfigure}{\textwidth}
-    \centering
-    \caption{Chloroform and PCE considered at the Indianapolis duplex 422.}
-    \label{fig:kde-indianapolis}
-    \includegraphics[height=0.4\textheight,keepaspectratio]{indianapolis-422.pdf}
-  \end{subfigure}
-\end{figure}
-% Introduction
-Of the factors that influence indoor air contaminant concentration IACC in VI, pressure is one of the most dynamic ones, and the relationship between changes in indoor/outdoor pressure difference and IACC are examined in Figure \ref{fig:kde-analysis}.
-Three well-studied VI sites are considered; the "ASU house" near Hill AFB in Utah, a building at North Island NAS in California and a duplex in Indianapolis.\par
-% Univariate IACC description
-The absolute IACCs at these sites vary significantly, therefore some means of comparing them to each other is necessary.
-Additionally, the focus in this section is identifying the driver for variations in IACC, and consistent representation of the variations between the different sites must also be achieved.
-Here variations on the scale of orders of magnitude is of interest, as these are of greatest concern in proper site IACC characterization.
-To achieve these goals, log-10 deviation from the mean IACC (represented as $\mu$) within each dataset is calculated, e.g.
-$10\mu$ indicates an IACC that is an order of magnitude above the mean IACC for that dataset.
-On the y-margins in Figures \ref{fig:kde-asu-nas} and \ref{fig:kde-indianapolis}, the univariate KDE distribution of deviation from the mean may be seen. \par
-
-What these curves on the y-margins characterize are the distributions of sampled IACC at each site.
-Naturally, the distributions are centered around the mean ($\mu = 1$).
-In same caes, they look close to simple Gaussians in form (e.g. PCE, Tetrachloroethene at Indianapolis).
-In most of the other cases, the distributions are curved and not of a simple nature, strongly suggesting that there might be several factors determinig the IACC. \par
-
-% Univariate pressure description
-The indoor/outdoor pressure differences are simply shown as reported, and their univariate KDE distributions are shown on the x-margins.
-A negative pressure difference indicates that the building is underpressurized relative to ambient.\par
-% Bivariate description
-The relationship between the deviation from mean IACC, and indoor/outdoor pressure difference may be seen in the central portion of each figure, in the form of a bivariate KDE distribution.
-The p-value and Pearson's r-value for each bivariate distribution are shown in the legend.\par
-% Sub-figure description (details about sites)
-In Figure \ref{fig:kde-asu-nas} the North Island and ASU house datasets are plotted, with blue representing the North Island site, and with red and green representing the ASU house.
-The ASU house dataset is split into two different parts, the first (red) are data from the time period before the preferential pathway was discovered, the second (green) is from after the preferential pathway was sealed off.
-At both the ASU house and North Island sites, TCE was the only contaminant considered.
-The IACC of Chloroform and PCE in the 422 part of the Indianapolis duplex were considered in Figure \ref{fig:kde-indianapolis}, since they were particulary tracked.\par
-% North Island discussion
-From Figure \ref{fig:kde-asu-nas}, it is apparent that the three datasets differ significantly from each other.
-The North Island site exhibits the greatest variation in both pressure difference and IACC, in fact, looking at the univariate pressure distribution - it is almost entirely flat and spanning from extremly high and low values.
-A likely explation for this is the reportedly poor condition of the structure, rendering it highly susceptible to envivornmental influnces.\par
-
-There is also a significant variation in IACC at the North Island site and this has a relativly strong link to pressure difference ($r = -0.71$), although there is clearly a distinct peak (around $0.5\mu$) that is associated with the smaller negative and positive pressure differences ($~-4 \leq \Delta p \leq 8$).
-This demonstrates that at this particular site, the very large IACC variations (an order of magnitude or more) are primarily influenced by the unusually large pressure difference fluctations.
-Likely, this relationship is created by the fact that the soil surrounding the structure is sandy and therefore relatively permeable - increasing the influence of pressure differences on inflow of contaminant from the soil.
-However, despite this, pressure differences can only partly explain the variations observed at this site.\par
-% ASU discussion
-The two ASU datasets are not only significantly different from the North Island dataset, but also different from each other.
-Considering the dataset before the preferential pathway was discovered (red), one can see that the IACC varies significantly (from $0.05 \mu$ to more than $50 \mu$, albeit the latter only rarely.)
-This is more variation than was observed at North Island, and yet the observed distribution of pressure differences is narrow, mostly varying by a few Pa around zero (~$\pm 2$ Pa).
-With $r = -0.50$, for this dataset, this suggests that pressure difference is not insignificant but still only a weak predictor for the variations in IACC.\par
-
-The picture changes completely after the preferential pathway was closed off, where a similar distribution of pressure differences were observed, but with a signficant reduction in IACC variance to around $\pm 0.5 \mu$, resembling something of a log-normal distribution.
-This clearly shows just how much variance the preferential pathway at this site contributed, and how critical it is to assess whether a preferential pathway is present at a particular site since this may lead to large uncertainties in determining relevant contaminant exposures.
-It is also clear that in the absence of a preferential pathway, the pressure difference becomes even less important, with the p- and r-value indicating that pressure differences are insignficant in determining variation in IACC.\par
-% Indianapolis discussion
-The Indianapolis duplex also featured a preferential pathway, but despite this, there is significantly lower variance in IACC for both Chloroform and PCE\cite{mchugh_evidence_2017}.
-The exact reason for this is not known, but considering that the preferential pathway at the Indianapolis site was a sanitary sewer, and at ASU the preferential pathway was a foundation drain system, it is not inconceivable that their dynamics are quite different.
-Nevertheless, both PCE and Chloroform are more or less log-normally distributed, with (mostly) less than an order of magnitude variance around the mean, indicating that the varations in IACC are species independent.\par
-
-The pressure difference distributions for the Indianapolis site exhibit some bimodal characteristics, with a larger variation than was observed at the ASU site.
-This may explain why the r-values for both PCE and Chloroform, $r=-0.51$ and $r=-0.53$ respectively, are so similar to the ASU site before the preferential pathway there was closed; the larger pressure differences, which occured more often than ASU, increase contaminant entry rates to the extent that a preferential pathway was needed to "achieve" at the ASU site.\par
-% Summary discussion
-All of these data clearly suggest that there is significant variation in IACC that cannot easily be explained by variations in indoor/outdoor pressure difference - even with a preferential pathway.
-It is also clear that a preferential pathway may contribute to unacceptable levels of IACC variations, and screening for preferential pathways should be part of any site investiation.
-It may also be prudent, even in the absence of a preferential pathway, to add around a half to one order of magnitude margin of error to early screening measurements of indoor air, e.g.
-if a single sample is less than an order of magnitude below some target limit, it may be justified to perform subsequent sampling to better assess the relevant exposure.
-These data also call the role of advection in VI into question, since "normal" pressure difference ranges are such a poor predictor of IACC variation.
-This is the opposite of what one would expect if advection were the dominant transport mechanism for entry of contaminant into a structure, under steady-state conditions.\par
-
-\subsubsection{The Role Of Air Exchange Rate}
-% Air exchange rate plots
-\begin{figure}[htb!]
-  \caption{Comparison between the recorded  and the calculated TCE in indoor air at the ASU house, assuming constant air exchange rate.
-\ref{fig:asu-iacc-ae-overview} shows the TCE in indoor air across time as well as the exchange rate.
-\ref{fig:violin-iacc-aet} shows the distribution of these values for three periods.}
-  \label{fig:ae-analysis}
-  % "Overview" plot
-  \begin{subfigure}{\textwidth}
-    \caption{ }
-    \label{fig:asu-iacc-ae-overview}
-    \includegraphics[width=\textwidth]{tce_indoor_air_exchange_rate_impact.pdf}
-  \end{subfigure}
-  % Violinplot figure
-  \begin{subfigure}{0.75\textwidth}
-    \caption{ }
-    \label{fig:violin-iacc-aet}
-    \includegraphics[width=\textwidth]{violin-asu-iacc.pdf}
-  \end{subfigure}
-\end{figure}
-Air exchange rate, which measures how often the interior air in a given building is exchanged with exterior air, is the main parameter characterizing contaminant expulsion, and part of another highly dynamic process.
-Exploring the impact of fluctuations in air exchange rate is thus also important for the development of sampling strategies.
-At the ASU house, the contaminant entry rate into the structure (called emission rate in that study) as well as the building air exchange rate were recorded across time.
-This allows for exploration of the impact that fluctuating air exchange rate has on the final IACC by solving the CST equation, using the recorded contaminant entry rate as one input, and assuming a constant air exchange rate.
-The rationale is to show the extent of variability in IACCs due to factors other than fluctuations in air exchange rates.
-We have taken this factor out of consideration by below assuming a single constant exchange rate, characteristic of the whole period.
-
-In Figure \ref{fig:asu-iacc-ae-overview}, the measured IACC and air exchange rate across time are shown, compared to the predicted IACC assuming instead a constant air exchange rate.
-The median air exchange rate value for the pre-CPM period was chosen as this seemed to be the most representative value for this time period.
-From Figure \ref{fig:asu-iacc-ae-overview} it may be seen quite clearly that for the non-CPM periods, the calculated IACC values assuming constant vs.
-actual fluctuating exchange rates are quite similar.
-In other words, in this instance it is unlikely that the variations in IACC were driven by fluctuations in air exchange rate.
-
-This is more clearly shown in Figure \ref{fig:violin-iacc-aet}, where KDEs of the IACC are constructed assuming the constant and fluctuating air exchange rate cases.
-This figure shows the relative probability of being at a particular TCE level in indoor air concentration at the ASU house.
-These are compared to each other for 1. the full measurement period, 2. the non-CPM periods and 3. the CPM period.
-Looking at the full period, it does not appear as though the two cases of steady and fluctuating air exchange rate are comparable, the distribution functions for IACC are different and offset.
-The curves to the different sides of the vertical line for the "full" measurement period represent the probability distributions for the indicated values of IACC.
-It appears as though assuming a constant air exchange rate shifts the IACC probability distribution to higher values.
-The reason for this are more apparent when considering the CPM and non-CPM periods separately.
-
-Considering only the CPM period, assuming a constant air exchange rate does not influence the shape of the IACC distribution, but assuming constant exchange rate shifts the IACC to a higher values.
-This is because the actual air exchange rate for this CPM period is significantly higher than the assumed median value, and much more contaminant is actually exchanged with the exterior.
-The shape of the IACC distribution is not very different from that which takes the air exchange fluctuations into account, indicating that fluctuating exchange rate is not a major contributor to observed IACC variation for this period (the distinction being drawn between variation and absolute levels of IACC).
-
-The non-CPM period is the condition under which field practitioners would normally collect samples at a VI site.
-It is clear here that there is only a very minor difference made in assuming a constant vs.
-real fluctating air exchange rate, again indicating that accounting for the minor fluctuations of air exchange rate's impact on variations in IACC is not important.
-\citeauthor{rackes_time-averaged_2016}\cite{rackes_time-averaged_2016} drew similar conclusions regarding the relationship between IACC and air exchange rate.
-
-\subsection{The Influence of Preferential Pathways}
-
-% Point is:
-% If preferential pathways exist, then roughly 1-3 orders of magnitude variability are possible.
-
-% Definition: A preferential pathway is a pathway that allows for contaminant vapor to enter a structure directly or close to the building - circumventing much of the resistance to transport in the soil.
-
-% How do I deal with this? How do I know if preferential pathways may be an issue?
-
-% Scenario: Preferential pathway directly connected to the building.
-% - Identify sewer & plumbing fixtures.
-% - Drains and sumps may be sources.
-
-% Scenario: Preferential pathway in the near slab area.
-% - Permeable sub-base necessary.
-% - May have a limited “sphere of influence”.
-% -- Limit likely not reached for a normal house.
-% -- Limit more likely to be reached for larger structures.
-
-% Relevant for both: Check local manhole/sewer for presence of contaminant - likely source for the preferential pathway.
-
-\subsubsection{Conclusions From Steady-State Modeling}
-
-\begin{figure}[htb!]
-  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for cases featuring a preferential pathway (\ref{fig:ss-sensitivity-analysis-pp}) and without a preferential pathway (\ref{fig:ss-sensitivity-analysis-no-pp}).
-Results compared with field data from ASU house.}
-  \label{fig:ss-sensitivity-analysis}
-  % "Overview" plot
-  \begin{subfigure}{0.45\textwidth}
-    \caption{preferential pathway present. Sensitivity to the presence of a gravel sub-base and contamination in the preferential pathway considered.}
-    \label{fig:ss-sensitivity-analysis-pp}
-    \includegraphics[width=\textwidth]{yes-pp.pdf}
-  \end{subfigure}
-  % Violinplot figure
-  \begin{subfigure}{0.45\textwidth}
-    \caption{No preferential pathway present. Sensitivitiy to the presence of a gravel sub-base considered.}
-    \label{fig:ss-sensitivity-analysis-no-pp}
-    \includegraphics[width=\textwidth]{no-pp.pdf}
-  \end{subfigure}
-\end{figure}
-
-Clearly, preferential pathways pose a major problem for site investigations, in particular because they may be difficult or even impossible to anticipate or uncover.
-Even at a well studied site like the ASU house it took years to discover the preferential pathway.
-Therefore, there is a need to consider other indications that a preferential pathway is present or a potential issue.\par
-
-To investigate this, inspiration is taken from the ASU house, where there clearly was a preferential pathway containing contaminant vapor, which exited into a gravel (permeable) sub-base.
-The impact of these features is simulated by examining the IACC at different indoor/outdoor pressure differences for each combination of cases.
-Instead of representing IACC as absolute concentration, we now non-dimensionalize with respect to the vapor in equilibrium with the groundwater contaminant concentration.
-This non-dimensionalized property is commonly called the attenuation factor and is denoted by $\alpha_\mathrm{gw}$.
-The result of a steady-state sensitivity analysis relative to several factors is seen in Figure \ref{fig:ss-sensitivity-analysis}.\par
-
-Figure \ref{fig:ss-sensitivity-analysis-pp} deal with cases where a preferential pathway is present and, the impact of having a gravel vs.
-soil sub-base and vapor contaminant vs.
-clean air in the preferential pathway ($\chi=1$ and $\chi=0$ respectively) are considered.
-Figure \ref{fig:ss-sensitivity-analysis-no-pp} considers the cases absent a preferential pathway, for reference, but still considers the impact of a gravel vs. soil sub-base.
-To give perspective to these results, the IACC at the ASU house (as $\alpha_\mathrm{gw}$) vs. indoor/outdoor pressure difference are also plotted.
-But the reader is reminded that the model was examined for a structure of the same scale as the ASU house, but which is not an exact representation of it.
-Only data from period before the preferential pathway was discovered, and after the preferential pathway was closed are considered in each respective figure (\ref{fig:ss-sensitivity-analysis-pp} and \ref{fig:ss-sensitivity-analysis-no-pp}).
-To improve visibility of the ASU data, it's average $\alpha_\mathrm{gw}$ values were aggregated into up to 40 evenly spaced points.\par
-
-Begining with Figure \ref{fig:ss-sensitivity-analysis-pp} it is immediately clear that absent a gravel sub-base, the impact of a preferential pathway is minimal, as comparing the orange and blue lines shows.
-Since in the model, it is assumed that the soil type is sandy clay, which is relatively impermeable, there is simply too much resistance to advection in the soil for changes in pressurization to matter.
-Additionally, whether the preferential pathway is filled with contaminant vapors or not does not matter; without a permeable sub-base, the transport of contaminant vapor in the sub-base is too slow to allow an advective entry path to contribute with.\par
-
-When a gravel sub-base is present, and the preferential pathway is filled with contaminant vapors ($\chi=1$).
-Tthe preferential pathway has a major impact on $\alpha_\mathrm{gw}$, spanning more than four orders of magnitude.
-This is explained by the increased role of advective transport under these conditions.
-The gravel sub-base allows vapor to readily flow.
-When the building is underpressurized under these conditions, a significant amount of contaminant vapor is pulled in from the preferential pathway, leading to the large increase of $\alpha_\mathrm{gw}$.\par
-
-On the other hand, when advection is stopped by overpressurization, most of the contaminant entering the building does so only through diffusion.
-This also explains why there is no difference whether the preferential pathway brings only clean air or contaminated air (red and green) when the building is overpressurized.
-These two cases differ signficantly from one another when the building is underpressurized, obviously because if there is no additional contaminant being supplied by the preferential pathway, the contaminant comes only from what diffuses from the contaminted groundwater source.\par
-
-It is interesting to note that $\alpha_\mathrm{gw}$ is higher with an uncontaminted preferential pathway (green) entering a gravel sub-base than the cases without a gravel sub-base.
-One could expect that pulling so much clean air into the sub-base region might significantly dilute the contaminant vapors and if anything cause a decrease in $\alpha_\mathrm{gw}$ (relative to blue and orange).
-However, the preferential pathway is small relative to the dimensions of the sub-base (10 cm diameter pipe), and its impact is only on the contaminant concentration in a limited region of the sub-base.
-This gravel sub-base still allows for higher advection throughout the entire sub-base region, leading to a $\alpha_\mathrm{gw}$ plateau that is limited by contaminant transport from the groundwater source, and any small amount of dilution from the preferential pathway is negligible.\par
-
-These case most resembling the ASU site (with an open preferential pathway) is the red case, and the trend agrees with the field data fairly well, especially for small under-/overpressurization values.
-The most significant deviations occur as the under-/overpressurization becomes larger.
-Again, the model is not intended to replicate the ASU data perfectly, but rather investigate the factors make a preferential pathway the most impactful in a general sense, and perfect simulation is not to be expected.
-This is especially reflected in the choice of vapor contaminant concentration in the preferential pathway, with it either being equal to the groundwater contaminant vapor concentration ($\chi=1$) or clean ($\chi=0$), neither of which is likely to be perfectly true at the ASU site.
-The idea with picking $\chi=1$ and $\chi=0$ is to give a span of $\alpha_\mathrm{gw}$ values within which a preferential pathway may fall.
-This helps explain why $\alpha_\mathrm{gw}$ is overpredicted when the structure is increasingly underpressurized - the contaminant vapor in the preferential pathway at the ASU house was lower than the groundwater contaminant vapor concentration.
-The simulation are able to capture the general trends well.\par
-
-In Figure \ref{fig:ss-sensitivity-analysis-no-pp} cases where a preferential pathway is absent are considered, i.e. "normal" VI scenarios, which may be considered reference scenarios.
-Here the only issues considered is whether there is a gravel or soil sub-base.
-It is apparent that without a preferential pathway, the indoor/outdoor pressure difference changes do not cause any signficant change in $\alpha_\mathrm{gw}$, due to the resistance to advective transport in the surrounding soil - showing the importance of the preferential pathway for increasing advection.
-Furthermore, the gravel sub-base does increase the overall $\alpha_\mathrm{gw}$, partly due to slightly higher advective potential but also due to the contaminant vapors being able to more easily diffuse in the more permeable sub-base.
-The trend in and even the value sof the data from the ASU house after the preferential pathway had been closed agree well with these model predictions, clearly again demonstrating the role of preferential pathway in increasing advective potential at a site.
-It should also be stated that if the surrounding soil was more permeable, e.g. sand, there would be more of a dependence of $\alpha_\mathrm{gw}$ on pressurization, as seen in North Island.\par
-
-Based on the simulated cases in Figure \ref{fig:ss-sensitivity-analysis} it can be concluded that a preferential pathway can be a signficant contributor to VI, but only when particular conditions are met.
-\begin{enumerate}
-  \item The preferential pathway has to supply additional vapor contaminant.
-  \item There must be a source of increased advective potential, which may be due to the presence of the preferential pathway itself, or other site features, or simply due to more permeable surrounding soil.
-  \item A permeable sub-base must exist to realize increased advective potential.
-\end{enumerate}
-In the absense of any of these conditions, a preferential pathway is unlikely to be a signficant contributor to VI.
-Strategies for screening for preferential pathway are discussed by \citeauthor{nielsen_remediation_2017}\cite{nielsen_remediation_2017}. \par
-
-\subsubsection{Multivariate Analysis}
-
-\begin{figure}[htb!]
-  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the ASU site. Only period after land drain was closed considered.}
-  \label{fig:asu_season}
-  \includegraphics[width=\textwidth]{asu_season.pdf}
-\end{figure}
-
-\begin{figure}[htb!]
-  \caption{Seasonal distribution of relevant VI parameters and indoor air contaminant concentration at the Indianapolis site. Only PCE considered.}
-  \label{fig:indianapolis_season}
-  \includegraphics[width=\textwidth]{indianapolis_season.pdf}
-\end{figure}
-
-Pressure alone does not seem adequately be able to predict the variability observed at the ASU and Indianapolis VI sites, necessitating a multivariate analysis to better understand the distribution of indoor air contaminant concentration.
-% Add references motivating the choice of parameters
-This is acheived by examining the distribution various parameters recorded at these sites, and how they relate to eachother.
-It is known that VI potential differs across the seasons, thus the seasonal distribution of each parameter is considered.
+\subsection{Temporal Variability of Indoor Air Contaminant Concentration}
 
@@
 
-\subsubsection{Rates of Change in Indoor Air Contaminant Concentration}
-
-\begin{figure}[htb!]
-  \caption{}
-  \label{fig:}
-  % "Overview" plot
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_open.pdf}
-  \end{subfigure}
-  % Violinplot figure
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_asu_pp_closed.pdf}
-  \end{subfigure}
-  % Indianapolis
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_season_indianapolis.pdf}
-  \end{subfigure}
-  % Indianapolis
-  \begin{subfigure}{0.45\textwidth}
-    \caption{}
-    \label{fig:}
-    \includegraphics[width=\textwidth]{rate_of_change_simulations.pdf}
-  \end{subfigure}
-\end{figure}
-
-\begin{table}[htb!]
-  \caption{Abbreviations \& symbols}
-  \begin{tabular}{l l}
-  \toprule
-  \textbf{Abbreviation or symbol}     & \textbf{Explanation} \\
-  % General abbreviations
-  VI                                  & Vapor intrusion \\
-  PP                                  & Preferential pathway \\
-  $L_\mathrm{slab}$                   & Thickness of the foundation concrete slab \\
-  % Richard's equation
-  $p$                                 & Pressure \\
-  % Unsteady-cstr
-  $u$                                 & TCE in indoor air concentration \\
-  $c$                                 & TCE in soil gas concentration \\
-  % Diffusion stuff
-  $D_\mathrm{air}$                    & Diffusion of TCE in air \\
-  $D_\mathrm{water}$                  & Diffusion of TCE in water \\
-  $D_\mathrm{eff}$                    & Effective Diffusion of TCE in soil \\
-  $K_H$                               & Henry's Law constant \\
-  % van Genuchten stuff
-  $\theta_g, \theta_w\;\&\;\theta_s$  & Vapor, water filled \& saturated soil porosity \\
-  $\theta_r$                          & Residual moisture porosity \\
-  $\alpha,\; l,\; n\;\&\; m$          & van Genuchten parameters \\
-  $\mathrm{Se}$                       & Soil moisture saturation \\
-  $k_r$                               & Relative permeability \\
-  $\alpha_\mathrm{gw}$                & Attenuation factor \\
-  \bottomrule
-  \end{tabular}
-\end{table}
-
 \bibliography{library}

### 2018-12-26 python/temporal_variability/indoor_air_concentration.py
Analysis of variability in IACC

@@ -1,0 +1,230 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.integrate import LSODA
+from scipy.interpolate import interp1d
+from scipy.stats import gaussian_kde
+import sqlite3
+import seaborn as sns
+
+# TODO: add indianapolis data and add to present analysis
+
+data_dir = './data/preferential-pathway-sensitivity/'
+db_dir = '/home/jonathan/Dropbox/var/'
+fig_dir = './figures/rate_of_change/'
+db = sqlite3.connect(db_dir + 'HillAFB.db')
+db_indianapolis = sqlite3.connect(db_dir + 'Indianapolis.db')
+
+def get_season(x):
+    seasons = {
+        'Winter': (12, 2),
+        'Spring': (3, 5),
+        'Summer': (6, 8),
+        'Fall': (9, 11),
+    }
+    if (x == 12) or (x == 1) or (x == 2):
+        return 'Winter'
+    elif (x == 3) or (x == 4) or (x == 5):
+        return 'Spring'
+    elif (x == 6) or (x == 7) or (x == 8):
+        return 'Summer'
+    elif (x == 9) or (x == 10) or (x == 11):
+        return 'Fall'
+    else:
+        return 'Error'
+
+class ProcessData:
+    def __init__(self,db):
+
+        return
+
+    def get_rate_of_change(self, df, time_scale='1h'):
+        df['TimeChange'] = (df['StopTime'].diff()/np.timedelta64(time_scale[0], time_scale[1])).apply(float)
+        df['RateOfChange'] = df['Concentration'].diff()/df['TimeChange']
+        return df
+
+    def get_season(self, df):
+        # finds the month and season of each timestamp
+        asu['Month'] = asu['StopTime'].map(lambda x: x.month)
+        asu['Season'] = asu['Month'].apply(lambda x: get_season(x))
+        return df
+
+def process_asu(resample_time=False):
+    asu = pd.read_sql_query("SELECT * FROM TDBasement;",
+        db,
+    ).sort_values(by='StopTime')
+
+    asu['StopTime'] = asu['StopTime'].apply(pd.to_datetime)
+
+    if resample_time != False:
+        asu = asu.resample(resample_time, on='StopTime', kind='timestamp').mean().reset_index()
+
[…]

### 2018-12-26 python/temporal_variability/non_dim_concentration.py
Starting analysis of change in non-dimensional concentration

@@ -1,0 +1,75 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.integrate import LSODA
+from scipy.interpolate import interp1d
+import sqlite3
+import seaborn as sns
+
+data_dir = './data/preferential-pathway-sensitivity/'
+db_dir = '/home/jonathan/Dropbox/var/'
+db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+
+db = sqlite3.connect(db_dir + 'HillAFB.db')
+
+pressure = pd.read_sql_query( "SELECT StopTime, Pressure FROM PressureDifference;", db, )
+indoor = pd.read_sql_query( "SELECT StopTime, Concentration AS IndoorConcentration FROM TDBasement;", db, )
+air_exchange_rate = pd.read_sql_query( "SELECT StopTime, AirExchangeRate FROM Tracer;", db, )
+soil_gas = pd.read_sql_query( "SELECT StopTime, MAX(Concentration) AS SubSurfaceConcentration FROM SoilGasConcentration WHERE Depth=0.0 AND (Location='1' OR Location='2' OR Location='3' OR Location='4' OR Location='5' OR Location='6' OR Location='7') GROUP BY StopTime;", db, )
+groundwater = pd.read_sql_query( "SELECT StopTime, AVG(Concentration) AS GroundwaterConcentration FROM GroundwaterConcentration WHERE Depth=2.7 GROUP BY StopTime;", db, )
+
+
+
+for df in (indoor,soil_gas,groundwater):
+    df['StopTime'] = df['StopTime'].apply(pd.to_datetime)
+
+df = pressure.set_index('StopTime').combine_first(indoor.set_index('StopTime')).combine_first(air_exchange_rate.set_index('StopTime')).combine_first(soil_gas.set_index('StopTime')).combine_first(groundwater.set_index('StopTime')).reset_index()
+df.interpolate(inplace=True,type='linear')
+df.dropna(inplace=True)
+
+
+df['Month'] = df['StopTime'].map(lambda x: x.month)
+
+
+def get_season(x):
+    seasons = {
+        'Winter': (12, 2),
+        'Spring': (3, 5),
+        'Summer': (6, 8),
+        'Fall': (9, 11),
+    }
+    if (x == 12) or (x == 1) or (x == 2):
+        return 'Winter'
[…]

### 2018-12-26 python/temporal_variability/indoor_air_concentration.py
Cosmetic changes

@@ -11,6 +11,8 @@ import seaborn as sns
 
 data_dir = './data/preferential-pathway-sensitivity/'
 db_dir = '/home/jonathan/Dropbox/var/'
+db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+
 fig_dir = './figures/rate_of_change/'
 db = sqlite3.connect(db_dir + 'HillAFB.db')
 db_indianapolis = sqlite3.connect(db_dir + 'Indianapolis.db')
@@ -142,7 +144,7 @@ def process_simulation(path):
     return simulation
 
 def custom_x_ticks(ax, my_xticks=np.log10([0.5, 1.0, 5.0])):
-    my_xtick_labels = ["%1.1e" % x_tick for x_tick in 10.0**ax.get_xticks()]
+    my_xtick_labels = ["%1.3f" % x_tick for x_tick in 10.0**ax.get_xticks()]
     #ax.set_xlim((my_xticks[0], my_xticks[-1]))
     #ax.set_xticks(my_xticks)
     ax.set_xticklabels(my_xtick_labels)
@@ -170,7 +172,7 @@ def plot_facetgrid(df):
     g.despine(bottom=True, left=True)
     g.axes[-1][0].set_xlabel('Contaminant in Indoor Air $\\mathrm{(\\mu g/m^3)}$')
     custom_x_ticks(g.axes[-1][0],my_xticks=np.log10([0.001,0.01,0.1,1,10,100]))
-    return
+    return g
 #plt.show()
 
 
@@ -178,13 +180,14 @@ asu, pre_cpm, post_cpm = process_asu()
 asu = asu[asu['PP'] != 'CPM'].reset_index() # removes cpm period
 asu['Category'] = 'PP ' + asu['PP'] + ', ' + asu['Season']
 sns.set_palette(('Blue','Green','Red','Orange','Orange','Blue','Green','Red',))
-plot_facetgrid(asu)
+g = plot_facetgrid(asu)
+g.axes[-1][0].set_xlabel('TCE in Indoor Air $\\mathrm{(\\mu g/m^3)}$')
 
 
 indianapolis = process_indianapolis()
 indianapolis['Category'] = indianapolis['Specie'] + ', ' + indianapolis['Season']
 sns.set_palette(('Red','Red','Red','Orange','Orange','Orange','Blue','Blue','Blue',))
-plot_facetgrid(indianapolis)
[…]

### 2018-12-27 python/temporal_variability/indoor_air_concentration.py
Generated new figures

@@ -11,7 +11,7 @@ import seaborn as sns
 
 data_dir = './data/preferential-pathway-sensitivity/'
 db_dir = '/home/jonathan/Dropbox/var/'
-db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+#db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
 
 fig_dir = './figures/rate_of_change/'
 db = sqlite3.connect(db_dir + 'HillAFB.db')
@@ -173,8 +173,6 @@ def plot_facetgrid(df):
     g.axes[-1][0].set_xlabel('Contaminant in Indoor Air $\\mathrm{(\\mu g/m^3)}$')
     custom_x_ticks(g.axes[-1][0],my_xticks=np.log10([0.001,0.01,0.1,1,10,100]))
     return g
-#plt.show()
-
 
 asu, pre_cpm, post_cpm = process_asu()
 asu = asu[asu['PP'] != 'CPM'].reset_index() # removes cpm period
@@ -182,12 +180,14 @@ asu['Category'] = 'PP ' + asu['PP'] + ', ' + asu['Season']
 sns.set_palette(('Blue','Green','Red','Orange','Orange','Blue','Green','Red',))
 g = plot_facetgrid(asu)
 g.axes[-1][0].set_xlabel('TCE in Indoor Air $\\mathrm{(\\mu g/m^3)}$')
-
+plt.savefig('./figures/temporal_variability/asu.pdf',dpi=300)
 
 indianapolis = process_indianapolis()
 indianapolis['Category'] = indianapolis['Specie'] + ', ' + indianapolis['Season']
 sns.set_palette(('Red','Red','Red','Orange','Orange','Orange','Blue','Blue','Blue',))
 g = plot_facetgrid(indianapolis)
+plt.savefig('./figures/temporal_variability/indianapolis.pdf',dpi=300)
+
 #plt.show()
 
 resampled = pd.DataFrame({'Resampling': [], 'Delta': [], 'Dataset': []})
@@ -230,4 +230,4 @@ my_ytick_labels = ["%1.1f" % y_tick for y_tick in 10.0**g.ax.get_yticks()]
 g.ax.set_yticklabels(my_ytick_labels)
 g.ax.set_xlabel('Resampling Period')
 g.ax.set_ylabel('$c_\\mathrm{max}/c_\\mathrm{min}$')
-plt.show()
+plt.savefig('./figures/temporal_variability/resampling.pdf',dpi=300)

### 2018-12-27 tex/main.tex
Started working on new manuscript

@@
 
-\subsection{Statistical Analysis}
+\subsection{Statistical Analysis of Field Data}
 
-\subsubsection{Indianapolis Dataset}
+This paper heavily relies on statistical analysis of high resolution datasets from two well-studied VI sites, one near Hill AFB in Utah (called the ASU house) and another in Indianapolis, IN (simply called as such.)
+Analysis is performed using the SciPy, NumPy, Pandas, and Seaborn Python packages.
 
-\subsubsection{"ASU House" Dataset}
+Probability distributions of various parameters are constructed using the kernel density estimation (KDE) method\cite{altman_introduction_1992}, which is implemented in the SciPy package.
 

### 2019-01-21 python/transient-response/24h-analysis.py
Deprecated

@@ -1,44 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy.integrate import LSODA
-from scipy.interpolate import interp1d
-import sqlite3
-import seaborn as sns
-
-db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-#db_dir = 'C://Users/jstroem/lib/vapor-intrusion-dbs/'
-
-db = sqlite3.connect(db_dir + 'hill-afb.db')
-
-asu = pd.read_sql_query("SELECT * from daily_averages;", db,
-                        ).sort_values('time')
-asu.time = asu.time.apply(pd.to_datetime)
-
-asu['pressure'] *= -1
-asu['dc'] = asu['concentration'].apply(np.log10).diff()
-asu['dp'] = asu['pressure'].diff()
-asu = asu.dropna()
-
-asu['csign'] = asu['dc'].apply(np.sign)
-asu['psign'] = asu['dp'].apply(np.sign)
-
-asu['csigncumsum'] = asu['csign'].cumsum()
-asu['psigncumsum'] = asu['psign'].cumsum()
-asu['ccumsum'] = asu['dc'].cumsum()
-asu['pcumsum'] = asu['dp'].cumsum()
-
-
-#asu['concentration'] = asu['concentration'].apply(np.log10)
-#asu = asu.replace([np.inf, -np.inf], np.nan).dropna()
-
-fig, ax1 = plt.subplots()
-ax2 = ax1.twinx()
-
-asu.plot(x='time', y='concentration', color='red', logy=True, ax=ax1)
-asu.plot(x='time', y=['csigncumsum', 'psigncumsum'], ax=ax2)
-
-
-sns.jointplot(asu['psigncumsum'], asu['concentration'], kind='reg')
-
-plt.show()

### 2019-01-21 python/temporal_variability/indoor_air_concentration.py
Added more resampling times

@@ -10,8 +10,8 @@ import seaborn as sns
 # TODO: add indianapolis data and add to present analysis
 
 data_dir = './data/preferential-pathway-sensitivity/'
-db_dir = '/home/jonathan/Dropbox/var/'
-#db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+#db_dir = '/home/jonathan/Dropbox/var/'
+db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
 
 fig_dir = './figures/rate_of_change/'
 db = sqlite3.connect(db_dir + 'HillAFB.db')
@@ -208,7 +208,7 @@ names = (
 )
 
 for dataset, name in zip(datasets,names):
-    for resampling_time in ('1D','1W','3W','1M',):
+    for resampling_time in ('1D','2D','3D','1W','2W','3W','1M','2M','3M','6M','1Y'):
         r = dataset.resample(resampling_time ,on='StopTime', kind='timestamp')
         r = r['Concentration'].agg([np.mean, np.max, np.min, np.std])
         to_be_appended = pd.DataFrame({

### 2019-01-21 python/temporal_variability/non_dim_concentration.py
Tmp

@@ -10,7 +10,7 @@ from scipy import stats
 
 data_dir = './data/preferential-pathway-sensitivity/'
 db_dir = '/home/jonathan/Dropbox/var/'
-#db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
 
 db = sqlite3.connect(db_dir + 'HillAFB.db')
 
@@ -74,21 +74,31 @@ filter3 = (asu['StopTime'] > phase3['StartTime'].values[0])
 
 pre_cpm = asu.loc[filter1].copy()
 post_cpm = asu.loc[filter3].copy()
+non_cpm = pd.concat([pre_cpm, post_cpm])
+
+
+fig, ax = plt.subplots()
+r, p = stats.pearsonr( non_cpm['Pressure'], non_cpm['AirExchangeRate'],)
+ax.set_title('r = %1.2f, p = %1.2f' % (r, p))
+
+sns.kdeplot(non_cpm['Pressure'], non_cpm['AirExchangeRate'],clip=((-3,3),(0,2)), shade_lowest=False,shade=True,  ax=ax)
+
+
 
 # Attenunation from subsurface concentration correlation with pressure and air exchange plots
 fig, ((ax1, ax2),(ax3,ax4)) = plt.subplots(2,2,sharey=True)
 
 r, p = stats.pearsonr( pre_cpm['Pressure'], pre_cpm['AttenuationSubSurface'],)
-ax1.set_title('r = %1.2f' % r)
+ax1.set_title('r = %1.2f, p = %1.2f' % (r, p))
 
 r, p = stats.pearsonr( pre_cpm['AirExchangeRate'], pre_cpm['AttenuationSubSurface'],)
-ax2.set_title('r = %1.2f' % r)
+ax2.set_title('r = %1.2f, p = %1.2f' % (r, p))
 r, p = stats.pearsonr( post_cpm['Pressure'], post_cpm['AttenuationSubSurface'],)
 
-ax3.set_title('r = %1.2f' % r)
+ax3.set_title('r = %1.2f, p = %1.2f' % (r, p))
 r, p = stats.pearsonr( post_cpm['AirExchangeRate'], post_cpm['AttenuationSubSurface'],)
 
-ax4.set_title('r = %1.2f' % r)
+ax4.set_title('r = %1.2f, p = %1.2f' % (r, p))
 
[…]

### 2019-01-23 python/temporal_variability/indoor_air_concentration.py
Removed 1 year resampling time

@@ -208,7 +208,7 @@ names = (
 )
 
 for dataset, name in zip(datasets,names):
-    for resampling_time in ('1D','2D','3D','1W','2W','3W','1M','2M','3M','6M','1Y'):
+    for resampling_time in ('1D','2D','3D','1W','2W','3W','1M','2M','3M','6M',):
         r = dataset.resample(resampling_time ,on='StopTime', kind='timestamp')
         r = r['Concentration'].agg([np.mean, np.max, np.min, np.std])
         to_be_appended = pd.DataFrame({

### 2019-01-23 python/temporal_variability/non_dim_concentration.py
Changed plotting order

@@ -6,7 +6,7 @@ from scipy.interpolate import interp1d
 import sqlite3
 import seaborn as sns
 from scipy import stats
-
+from scipy.optimize import curve_fit
 
 data_dir = './data/preferential-pathway-sensitivity/'
 db_dir = '/home/jonathan/Dropbox/var/'
@@ -75,10 +75,28 @@ filter3 = (asu['StopTime'] > phase3['StartTime'].values[0])
 pp_status = lambda x: 'Open' if x < phase1['StopTime'].values[0] else ('Closed' if x > phase3['StartTime'].values[0] else 'CPM')
 asu['PP'] = asu['StopTime'].apply(pp_status)
 
-g = sns.PairGrid(asu.loc[(asu['PP']!='CPM') & (asu['Pressure'] > -5)][['AttenuationSubSurface','Pressure','AirExchangeRate','PP']], hue='PP')
+g = sns.PairGrid(asu.loc[(asu['PP']!='CPM') & (asu['Pressure'] > -5)][['Pressure','AirExchangeRate','AttenuationSubSurface','PP']], hue='PP')
 #g = g.map_lower(sns.kdeplot, shade=True, shade_lowest=False)
-g = g.map_upper(sns.regplot, x_bins=20, truncate=True )
-g = g.map_lower(sns.regplot, x_bins=20, truncate=True )
+g = g.map_upper(plt.scatter)
+g = g.map_lower(sns.regplot, x_bins=20, truncate=True, lowess=True )
 g = g.map_diag(sns.kdeplot, shade=True)
 g = g.add_legend()
+
+
+# non-linear fit
+
+def func(p, a, b):
+    Ae = a*np.exp(b*p)
+    return Ae
+
+xdata = asu.loc[(asu['PP']!='CPM') & (asu['Pressure'] > -5)]['Pressure'].values
+ydata = asu.loc[(asu['PP']!='CPM') & (asu['Pressure'] > -5)]['AirExchangeRate'].values
+
+popt, pcov = curve_fit(func, xdata, ydata)
+
+
+fig, ax = plt.subplots()
+plt.plot(xdata,ydata,'o',label='Data')
+plt.plot(xdata, func(xdata, *popt), 'o', label='fit: a=%5.3f, b=%5.3f' % tuple(popt))
+plt.legend()
 plt.show()

### 2019-01-23 tex/main.tex
Tmp
Started writing on new paper, adding plots

@@
   {./../figures/transient-response/},
-  {./../figures/multivariate_analysis/},
-  {./../figures/rate_of_change/},
+  {./../figures/multivariate_analysis/}
 }
@@
 
+The significant temporal varibility in indoor air contaminant concentrations at vapor intrusion (VI) sites pose a major impedent for assertaining the relevant human exposure to vapor contaminant.
+Exactly how significantly the indoor air concentration may vary, and the causes of the variability is poorly understood.
+Improving our understanding of these two factors are crucial to reduce uncertainty in determining indoor air contaminant exposures, and reducing the length of these investigations.
+
+Two well-documentated VI sites both showed significant temporal variability in indoor air contaminant concentrations.
+One is a two-story house near Hill AFB in Utah (called the ASU house in this paper), and the other a duplex in Indianapolis, IN.
+
+The discovery of preferential pathways for contaminant entry at VI sites has further
+
 \section{Methods}
@@
 
+\begin{figure}[htb!]
+  \caption{ }
+  \label{fig:temporal_variability}
+  % ASU/North Island plot
+  \begin{subfigure}{0.49\textwidth}
+    \centering
+    \caption{ }
+    \label{fig:temporal_variability_asu}
+    \includegraphics[width=\textwidth]{temporal_variability/asu.pdf}
+  \end{subfigure}
+  % Indianapolis plot
+  \begin{subfigure}{0.49\textwidth}
+    \centering
+    \caption{ }
+    \label{fig:temporal_variability_indianapolis}
+    \includegraphics[width=\textwidth]{temporal_variability/indianapolis.pdf}
+  \end{subfigure}
+\end{figure}
+
+\subsection{Maximum Change in Indoor Air Contaminant Concentration}
+
+\begin{figure}[htb!]
+  \caption{ }
+  \label{fig:resampling}
+  \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
+\end{figure}
+
+\subsubsection{Conclusions From Steady-State Modeling}
+
+\begin{figure}[htb!]
+  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for cases featuring a preferential pathway (\ref{fig:ss-sensitivity-analysis-pp}) and without a preferential pathway (\ref{fig:ss-sensitivity-analysis-no-pp}).
+Results compared with field data from ASU house.}
+  \label{fig:ss-sensitivity-analysis}
+  % "Overview" plot
+  \begin{subfigure}{0.45\textwidth}
+    \caption{preferential pathway present. Sensitivity to the presence of a gravel sub-base and contamination in the preferential pathway considered.}
+    \label{fig:ss-sensitivity-analysis-pp}
+    \includegraphics[width=\textwidth]{yes-pp.pdf}
+  \end{subfigure}
+  % Violinplot figure
+  \begin{subfigure}{0.45\textwidth}
+    \caption{No preferential pathway present. Sensitivitiy to the presence of a gravel sub-base considered.}
+    \label{fig:ss-sensitivity-analysis-no-pp}
+    \includegraphics[width=\textwidth]{no-pp.pdf}
+  \end{subfigure}
+\end{figure}
+
 \begin{acknowledgement}

### 2019-02-21 code/resampling.py
Moved legend inside

@@ -1,0 +1,67 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.integrate import LSODA
+from scipy.interpolate import interp1d
+from scipy.stats import gaussian_kde
+import sqlite3
+import seaborn as sns
+
+
+# global settings
+num_bins = 10
+fig_path = './figures/pair_grids/'
+ext = '.png'
+dpi = 300
+
+
+asu = pd.read_csv('./data/asu_house.csv')
+indianapolis = pd.read_csv('./data/indianapolis.csv')
+
+
+resampled = pd.DataFrame({'Resampling': [], 'MaxDelta': [], 'Dataset': []})
+
+datasets = (
+    asu[asu['Phase']=='Open'],
+    asu[asu['Phase']=='Closed'],
+    indianapolis[indianapolis['Specie']=='Trichloroethene'],
+    indianapolis[indianapolis['Specie']=='Tetrachloroethene'],
+    indianapolis[indianapolis['Specie']=='Chloroform'],
+)
+names = (
+    'ASU House, PP Open',
+    'ASU House, PP Closed',
+    'Indianapolis, TCE',
+    'Indianapolis, PCE',
+    'Indianapolis, Chloroform',
+)
+
+for dataset, name in zip(datasets,names):
+    dataset['Time'] = dataset['Time'].apply(pd.to_datetime)
+    print('Using dataset: %s' % name)
+    for resampling_time in ('1D','2D','3D','1W','2W','3W','1M','2M','3M','6M',):
+        r = dataset.resample(resampling_time ,on='Time', kind='timestamp')
+        r = r['logIndoorConcentration'].agg([np.mean, np.max, np.min, np.std])
+        to_be_appended = pd.DataFrame({
+            'Resampling': np.repeat(resampling_time,len(r)),
+            'MaxDelta': r['amax'].values-r['amin'].values,
+            'Dataset': np.repeat(name,len(r)),
+        })
+        resampled = resampled.append(to_be_appended,ignore_index=True)
+
+resampled.dropna(inplace=True)
+
+#print(resampled)
+sns.set_palette(sns.color_palette("muted"))
+
+
+g = sns.catplot(x="Resampling", y="MaxDelta", hue='Dataset', kind="point", data=resampled, legend_out=False, aspect=1.5)
+
[…]

### 2019-02-21 code/land_drain_scenarios.py
Minor fixes

@@ -53,7 +53,7 @@ g = sns.lmplot(
     fit_reg=False,
     legend_out=False,
     legend=False,
-    aspect=2,
+    aspect=1.5,
 )
 
 ax = g.axes[0][0]
@@ -97,5 +97,5 @@ ax.set_title('Modeling PP scenarios, assuming constant $A_e$, and comparing to \
 
 plt.tight_layout()
 
-plt.savefig('./figures/simulation_predictions/land_drain_scenarios.png',dpi=300)
+plt.savefig('./figures/simulation_predictions/land_drain_scenarios.pdf',dpi=300)
 #plt.show()

### 2019-02-21 code/resampling.py

@@ -62,6 +62,6 @@ g.ax.set_yticklabels(my_ytick_labels)
 g.ax.set_xlabel('Period')
 g.ax.set_ylabel('$c_\\mathrm{max}/c_\\mathrm{min}$')
 g.ax.set_title('Maximum variability within a given period')
-
+plt.tight_layout()
 #plt.show()
-plt.savefig('./figures/temporal_variability/resampling.png',dpi=300)
+plt.savefig('./figures/temporal_variability/resampling.pdf',dpi=300)

### 2019-02-22 tex/main.tex
Working on paper

@@
   {./../figures/kde-iacc-pressure/},
-  {./../figures/preferential-pathway-sensitivity/},
-  {./../figures/transient-response/},
+  {./../figures/simulation_predictions/},
+  {./../figures/pair_grids/},
   {./../figures/multivariate_analysis/}
@@
 
-\subsection{Statistical Analysis of Field Data}
+\subsection{Statistical Analysis}
 
@@
 Analysis is performed using the SciPy, NumPy, Pandas, and Seaborn Python packages.
-
 Probability distributions of various parameters are constructed using the kernel density estimation (KDE) method\cite{altman_introduction_1992}, which is implemented in the SciPy package.
 
-\subsection{Temporal Variability of Indoor Air Contaminant Concentration}
+At the Indianapolis site, we only consider the 422 side of the duplex, which was heated.
+Indoor air concentration in this case, specifically refer to the indoor air contaminant concentration in the 422 basement.
+Soilgas concentration refers only to samples taken from SGP2.
+The indoor/outdoor pressure difference is the 422 basement and exterior pressure difference.
+
+The indoor/outdoor pressure difference were recorded every two minutes.
+Indoor air samples were taken every four hours.
+
+\subsection{Vapor Intrusion Model}
+
+\section{Results & Discussion}
+
+\subsection{Statistical Analysis of Field Data}
+
+\begin{figure}[!h]
+	\centering
+	\begin{minipage}[c]{0.49\textwidth}
+		\centering
+    \caption{ }
+    \label{fig:asu_phase}
+    \includegraphics[width=\textwidth]{asu_phase.pdf}
+	\end{minipage}
+	%\hspace{3cm}
+	\begin{minipage}[c]{0.49\textwidth}
+		\centering
+    \caption{ }
+    \label{fig:indianapolis_species}
+    \includegraphics[width=\textwidth]{indianapolis_species.pdf}
+	\end{minipage}
+\end{figure}
+
+\subsection{Seasonal Variability}
 
@@
   \caption{ }
-  \label{fig:temporal_variability}
+  \label{fig:seasonal_analysis}
   % ASU/North Island plot
@@
     \caption{ }
-    \label{fig:temporal_variability_asu}
-    \includegraphics[width=\textwidth]{temporal_variability/asu.pdf}
+    \label{fig:seasonal_analysis_asu}
+    \includegraphics[width=\textwidth]{asu_closed_season.pdf}
   \end{subfigure}
@@
     \caption{ }
-    \label{fig:temporal_variability_indianapolis}
-    \includegraphics[width=\textwidth]{temporal_variability/indianapolis.pdf}
+    \label{fig:seasonal_analysis_indianapolis}
+    \includegraphics[width=\textwidth]{indianapolis_pce_season.pdf}
   \end{subfigure}
@@
 
-\begin{figure}[htb!]
+\begin{figure}[!h]
+	\centering
+	\begin{minipage}[c]{0.49\textwidth}
+		\centering
     \caption{ }
@@
     \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
+	\end{minipage}
+	%\hspace{3cm}
+	\begin{minipage}[c]{0.49\textwidth}
+		\centering
+    \caption{Simulated cases}
+    \label{fig:land_drain_scenarios}
+    \includegraphics[width=\textwidth]{land_drain_scenarios.pdf}
+	\end{minipage}
 \end{figure}
 
-\subsubsection{Conclusions From Steady-State Modeling}
+\subsubsection{Modeling Indoor Air Variability}
 
 \begin{figure}[htb!]
-  \caption{Sensitivity analysis of IACC dependence on indoor/outdoor pressure difference for cases featuring a preferential pathway (\ref{fig:ss-sensitivity-analysis-pp}) and without a preferential pathway (\ref{fig:ss-sensitivity-analysis-no-pp}).
-Results compared with field data from ASU house.}
-  \label{fig:ss-sensitivity-analysis}
-  % "Overview" plot
-  \begin{subfigure}{0.45\textwidth}
-    \caption{preferential pathway present. Sensitivity to the presence of a gravel sub-base and contamination in the preferential pathway considered.}
-    \label{fig:ss-sensitivity-analysis-pp}
-    \includegraphics[width=\textwidth]{yes-pp.pdf}
+  \caption{ }
+  \label{fig:simulation_sampling}
+  % ASU/North Island plot
+  \begin{subfigure}{0.49\textwidth}
+    \centering
+    \caption{ }
+    \label{fig:simulation_sampling_pp_open}
+    \includegraphics[width=\textwidth]{sampling_simulation_pp_open.pdf}
   \end{subfigure}
-  % Violinplot figure
-  \begin{subfigure}{0.45\textwidth}
-    \caption{No preferential pathway present. Sensitivitiy to the presence of a gravel sub-base considered.}
-    \label{fig:ss-sensitivity-analysis-no-pp}
-    \includegraphics[width=\textwidth]{no-pp.pdf}
+  % Indianapolis plot
+  \begin{subfigure}{0.49\textwidth}
+    \centering
+    \caption{ }
+    \label{fig:simulation_sampling_pp_closed}
+    \includegraphics[width=\textwidth]{sampling_simulation_pp_closed.pdf}
   \end{subfigure}

### 2019-02-22 code/land_drain_scenarios.py

@@ -53,7 +53,7 @@ g = sns.lmplot(
     fit_reg=False,
     legend_out=False,
     legend=False,
-    aspect=1.5,
+    #aspect=1.5,
 )
 
 ax = g.axes[0][0]
@@ -74,7 +74,7 @@ handles = handles[1:]
 labels = (
     'PP present',
     'PP absent',
-    'PP present, gravel sub-base absent',
+    'PP present, \ngravel sub-base\nabsent',
     'Data, PP open',
     'Data, PP closed',
 )
@@ -82,7 +82,8 @@ labels = (
 ax.legend(
     handles,
     labels,
-    ncol=2,
+    #ncol=2,
+    loc='best',
 )
 
 ax.set_xlim([-6,6])
@@ -92,7 +93,7 @@ ax.set_xlabel('$\\Delta p_\\mathrm{in/out}$ (Pa)')
 ax.set_ylabel('$\\alpha_\\mathrm{gw}$')
 my_ytick_labels = ["%1.0e" % y_tick for y_tick in 10.0**ax.get_yticks()]
 ax.set_yticklabels(my_ytick_labels)
-ax.set_title('Modeling PP scenarios, assuming constant $A_e$, and comparing to \"ASU house\" field data')
+ax.set_title('Modeling PP scenarios, assuming constant $A_e$,\nand comparing to \"ASU house\" field data')
 
 
 plt.tight_layout()

### 2019-02-22 code/resampling.py

@@ -22,11 +22,11 @@ indianapolis = pd.read_csv('./data/indianapolis.csv')
 resampled = pd.DataFrame({'Resampling': [], 'MaxDelta': [], 'Dataset': []})
 
 datasets = (
-    asu[asu['Phase']=='Open'],
-    asu[asu['Phase']=='Closed'],
-    indianapolis[indianapolis['Specie']=='Trichloroethene'],
-    indianapolis[indianapolis['Specie']=='Tetrachloroethene'],
-    indianapolis[indianapolis['Specie']=='Chloroform'],
+    asu[asu['Phase'] == 'Open'],
+    asu[asu['Phase'] == 'Closed'],
+    indianapolis[indianapolis['Specie'] == 'Trichloroethene'],
+    indianapolis[indianapolis['Specie'] == 'Tetrachloroethene'],
+    indianapolis[indianapolis['Specie'] == 'Chloroform'],
 )
 names = (
     'ASU House, PP Open',
@@ -36,26 +36,34 @@ names = (
     'Indianapolis, Chloroform',
 )
 
-for dataset, name in zip(datasets,names):
+for dataset, name in zip(datasets, names):
     dataset['Time'] = dataset['Time'].apply(pd.to_datetime)
     print('Using dataset: %s' % name)
-    for resampling_time in ('1D','2D','3D','1W','2W','3W','1M','2M','3M','6M',):
-        r = dataset.resample(resampling_time ,on='Time', kind='timestamp')
+    for resampling_time in ('1D', '2D', '3D', '1W', '2W', '3W', '1M', '2M', '3M', '6M',):
+        r = dataset.resample(resampling_time, on='Time', kind='timestamp')
         r = r['logIndoorConcentration'].agg([np.mean, np.max, np.min, np.std])
         to_be_appended = pd.DataFrame({
-            'Resampling': np.repeat(resampling_time,len(r)),
-            'MaxDelta': r['amax'].values-r['amin'].values,
-            'Dataset': np.repeat(name,len(r)),
+            'Resampling': np.repeat(resampling_time, len(r)),
+            'MaxDelta': r['amax'].values - r['amin'].values,
+            'Dataset': np.repeat(name, len(r)),
         })
-        resampled = resampled.append(to_be_appended,ignore_index=True)
+        resampled = resampled.append(to_be_appended, ignore_index=True)
 
[…]

### 2019-02-22 code/simulation_predictions.py

@@ -131,4 +131,8 @@ for sim_case, phase in zip(sim_cases, phases):
     make_plots(asu_sort)
     plt.savefig(fig_dir+'sampling_simulation_pp_'+phase.lower()+ext,dpi=dpi)
 
+
+
+
+# TODO: how many samples are needed to capture like 99% of the distribution?
 #plt.show()

### 2019-02-22 tex/main.tex
Added introduction & method

@@
 
-The significant temporal varibility in indoor air contaminant concentrations at vapor intrusion (VI) sites pose a major impedent for assertaining the relevant human exposure to vapor contaminant.
-Exactly how significantly the indoor air concentration may vary, and the causes of the variability is poorly understood.
-Improving our understanding of these two factors are crucial to reduce uncertainty in determining indoor air contaminant exposures, and reducing the length of these investigations.
+Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
+VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
+VI has been a recognized problem for some time, but many aspects remain poorly understood, particularly with respect to the causes of large temporal transients in indoor air concentrations.
+There is uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
+\par
 
-Two well-documentated VI sites both showed significant temporal variability in indoor air contaminant concentrations.
-One is a two-story house near Hill AFB in Utah (called the ASU house in this paper), and the other a duplex in Indianapolis, IN.
+Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
+All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
+All have shown that these concentrations varied significantly with time - orders of magnitude on the timescale of a day or days.
+\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}. \par
 
-The discovery of preferential pathways for contaminant entry at VI sites has further
+In one instance the source of the variation was clearly established during the study; at the ASU house a field drain pipe (or “land drain”), which connected to a sewer system, was discovered beneath the house, and careful isolation of this source led to a clear conclusion that this preferential pathway significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
+While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway to an area beneath a structure would always be expected to lead to large fluctuations in indoor air contaminant concentrations.
+\par
+
+Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
+That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
+Sewer lines have been generally implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
+A Danish study estimate that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
+Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
+In some of these cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
+While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant via the soil beneath a structure. \par
+
+It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
+One example is a site at NAS North Island at which no preferential pathways have been identified.
+Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
+It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site. \par
+
+This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
+In this work, the latter scenarios are referred to as ”normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
+Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at  such normal  VI  sites vs. those characterized by preferential pathways.
+The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.
 
@@
 
-At the Indianapolis site, we only consider the 422 side of the duplex, which was heated.
-Indoor air concentration in this case, specifically refer to the indoor air contaminant concentration in the 422 basement.
-Soilgas concentration refers only to samples taken from SGP2.
-The indoor/outdoor pressure difference is the 422 basement and exterior pressure difference.
+To begin to characterize transient behavior in indoor air contaminant concentrations, actual datasets are analyzed to establish common levels of variability at VI sites.
+For this purpose, the datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
+This paper relies on statistical analysis of published field data, and readers are referred to the original works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.
+\par
 
-The indoor/outdoor pressure difference were recorded every two minutes.
-Indoor air samples were taken every four hours.
+The ASU house data were obtained over a period of a few years.
+During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” operation.
+This caused greater than normal advective flow from the subsurface into the house, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
+This period of CPM testing is considered separately from the otherwise ”natural” VI conditions in the analysis.
+Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining the dataset, noting that during some of the testing, this pathway was deliberately cut off, resulting in what we have termed “normal” VI conditions in which the main source of contaminant was believed to be groundwater.
+\par
+
+The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to large internal pressure fluctuations, much more extensive than those typically recorded at the ASU house during normal operations.
+Additionally, the underlying soil at NAS North Island is sandy and more permeable than that at the ASU site, which, as will be shown, contributes to the indoor air contaminant concentrations being more sensitive to pressure fluctuations\cite{hosangadi_high-frequency_2017}.
+\par
+
+Likewise, the Indianapolis site investigation spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
+The goal of the SSD testing was to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above.
+Only the period before the installation of this system was considered in the present analysis.
+It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this preferential pathway removed, making it difficult to assess how significant the role of the preferential pathway was at this site.
+Regardless of this it is of interest to consider the data from this site due to how extensive and complete the data collection was. \par
+
+The typical variation in indoor air contaminant concentrations with time will first be considered below in the case of the ASU house during ”natural”, (i.e.
+non-CPM conditions), in the case of the NAS North Island site over the entire available dataset, and for the Indianapolis case we consider the variations before the installation of the SSD system.
+The deviations in indoor air concentration from the mean TCE (and Chloroform and PCE at the Indianapolis site) values, as well as the indoor-outdoor pressure differentials associated with these concentrations were examined.
+Both univariate and bivariate kernel density estimations (KDE) were constructed.
+KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
+This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (and that there are statistical fluctuations associated with each sampling event).
+In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
+The distributions of the individual parameters and the relationship between them will be examined using the KDE method.
 
@@
 
-\section{Results & Discussion}
+\section{Results \& Discussion}
 

### 2019-02-28 code/simulation_predictions.py
Fill in figure

@@ -163,36 +163,61 @@ for season in seasons:
 
 
 class SimPrediction:
-    def __init__(self):
-        asu_span = asu.loc[(asu['Phase']=='Closed') & ( (asu['IndoorOutdoorPressure']>=-5) & (asu['IndoorOutdoorPressure']<=5) )]
-        sim_span = df[df['Simulation']=='No Pp']
+    def __init__(self,status='Closed'):
 
-        grp = asu_span[['IndoorOutdoorPressure','AirExchangeRate']].groupby(pd.cut(asu_span['IndoorOutdoorPressure'], np.arange(-5, 5.5, 0.5))).describe()
+        options = {
+            'Open': ['Open', 'Pp'],
+            'Closed': ['Closed', 'No Pp'],
+        }
+        asu_span = asu.loc[(asu['Phase']==options[status][0]) & ( (asu['IndoorOutdoorPressure']>=-5) & (asu['IndoorOutdoorPressure']<=5) )]
+        sim_span = df[df['Simulation']==options[status][1]]
 
+        grp = asu_span[['IndoorOutdoorPressure','AirExchangeRate']].groupby(pd.cut(asu_span['IndoorOutdoorPressure'], np.arange(-5, 5.5, 0.5))).describe(percentiles=[0.05, 0.95])
+        print(grp['IndoorOutdoorPressure'])
         p_in = grp['IndoorOutdoorPressure'].index
         p_in = np.arange(-5, 5, 0.5)
         #print(p_in.categories.astype('float'))
 
-        Ae_min = grp[('AirExchangeRate','min')].values
-        Ae_max = grp[('AirExchangeRate','max')].values
+        Ae_min = grp[('AirExchangeRate','5%')].values
+        Ae_max = grp[('AirExchangeRate','95%')].values
+        Ae_mean = grp[('AirExchangeRate','mean')].values
 
         alpha_min = []
         alpha_max = []
+        alpha_mean = []
 
         func = self.get_interp_func(sim_span)
-        for p, min, max in zip(p_in, Ae_min, Ae_max):
+        for p, min, max, mean in zip(p_in, Ae_min, Ae_max, Ae_mean):
 
             alpha_min.append(func(p,max)[0])
             alpha_max.append(func(p,min)[0])
+            alpha_mean.append(func(p,mean)[0])
 
-
+        alpha_min = np.array(alpha_min)
+        alpha_max = np.array(alpha_max)
+        alpha_mean = np.array(alpha_mean)
 
         fig, ax = plt.subplots()
[…]

### 2019-02-28 code/pair_grids.py
Making simplified pair grid plot

@@ -2,6 +2,7 @@ import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
+import matplotlib.gridspec as gridspec
 from scipy import stats
 
 
@@ -38,7 +39,7 @@ def custom_axis(g):
                 g.axes[i][j].set_ylabel(replacements[ylabel])
     return g
 
-
+"""
 # demonstrate the variability
 
 # asu
@@ -100,4 +101,125 @@ for specie in indianapolis['Specie'].unique():
     g.fig.suptitle(specie)
     plt.savefig(fig_path+'indianapolis_season_'+specie.lower()+ext,dpi=dpi)
 
-#plt.show()
+"""
+
+class SimpleGrid:
+    def __init__(self):
+        # loads data
+        asu = pd.read_csv('./data/asu_house.csv')
+        indianapolis = pd.read_csv('./data/indianapolis.csv')
+        nas = pd.read_csv('./data/north_island.csv').dropna()
+
+        # processes data
+        asu = asu.loc[(asu['Phase']!='CPM')] # & ((asu['IndoorOutdoorPressure'] >= -5) & (asu['IndoorOutdoorPressure'] <= 5))
+        indianapolis = indianapolis.loc[indianapolis['Specie']=='Tetrachloroethene']
+        # simpler figure for publication
+        gs = gridspec.GridSpec(3, 3)
+
+
+
+        plt.figure()
+        ax1 = plt.subplot(gs[0, 0]) # p distributions ASU
+        ax2 = plt.subplot(gs[0, 1]) # - " - Indianapolis
+        ax3 = plt.subplot(gs[0, 2]) # - " - nas
+        ax4 = plt.subplot(gs[1, 0])
+        ax5 = plt.subplot(gs[1, 1])
+        ax6 = plt.subplot(gs[1, 2])
+        ax7 = plt.subplot(gs[2, 0])
+
+        for phase in asu['Phase'].unique():
+            sns.kdeplot(
+                data=asu.loc[asu['Phase']==phase]['IndoorOutdoorPressure'],
+                shade=True,
+                ax=ax1,
+                label=phase,
+            )
+            sns.regplot(
+                data=asu.loc[asu['Phase']==phase][['IndoorOutdoorPressure','logIndoorConcentration']],
+                x='IndoorOutdoorPressure',
+                y='logIndoorConcentration',
+                x_bins=np.linspace(-5,5,10),
+                ax=ax4,
+            )
+            sns.regplot(
[…]

### 2019-02-28 code/process_datasets.py

@@ -84,7 +84,7 @@ class Indianapolis:
             },
             inplace=True,
         )
-        pressure['IndoorOutdoorPressure'] *= -1 # changing sign to my convention
+        #pressure['IndoorOutdoorPressure'] *= -1 # changing sign to my convention (wasn't it needed?)
         pressure.drop(columns=['Variable','Location'],inplace=True)
         return pressure
 
@@ -257,5 +257,40 @@ class ASUHouse:
         subslab.drop(columns=['Depth','Location'],inplace=True)
         return subslab
 
-ind = Indianapolis()
-asu = ASUHouse()
+
+class NorthIsland:
+    def __init__(self):
+
+        self.db = sqlite3.connect(get_dropbox_path() + '/var/NorthIslandNAS.db')
+        df = self.get_data()
+        df.to_csv('./data/north_island.csv')
+        return
+
+    def get_data(self):
+        indoor = self.get_indoor_air()
+        pressure = self.get_pressure()
+
+        df = pd.merge_asof(indoor, pressure, on='Time')
+
+        return df
+
+    def get_indoor_air(self):
+        indoor = pd.read_sql_query(
+            "SELECT Time, Concentration AS IndoorConcentration FROM IndoorAirConcentration;",
+            self.db,
+        ).sort_values(by='Time')
+
+        indoor['logIndoorConcentration'] = indoor['IndoorConcentration'].apply(np.log10).replace([-np.inf,np.inf], np.nan)
+        return indoor
+
+    def get_pressure(self):
+        pressure = pd.read_sql_query(
+            "SELECT Time, Pressure AS IndoorOutdoorPressure FROM IndoorOutdoorPressureDifference;",
+            self.db,
+        ).sort_values(by='Time')
+        pressure['IndoorOutdoorPressure'] *= -1 
+        return pressure
+
+#ind = Indianapolis()
+#asu = ASUHouse()
+nas = NorthIsland()

### 2019-03-01 code/pair_grids.py
Generated figures

@@ -16,7 +16,7 @@ asu = asu.loc[(asu['IndoorOutdoorPressure'] >= -5) & (asu['IndoorOutdoorPressure
 # global settings
 num_bins = 10
 fig_path = './figures/pair_grids/'
-ext = '.png'
+ext = '.pdf'
 dpi = 300
 
 # method to set custom axis labels
@@ -216,7 +216,7 @@ class SimpleGrid:
         # TODO: fix the plot colors in the ASU column
         # TODO: increase resolution of ASU KDE
         ax1.legend()
-
+        plt.savefig(fig_path+'simple_grid'+ext,dpi=300)
         plt.tight_layout()
         return
 

### 2019-03-01 code/simulation_predictions.py

@@ -11,6 +11,11 @@ from scipy.interpolate import interp2d
 asu = pd.read_csv('./data/asu_house.csv')
 #simulation = pd.read_csv('./data/asu_house.csv')
 
+# figure saving settings
+fig_dir = './figures/simulation_predictions/'
+ext = '.png'
+dpi = 300
+
 dfs = []
 for file in ('pp','pp_uniform','no_pp',):
     df = pd.read_csv('./data/simulation/sweep_'+file+'.csv',header=4)
@@ -114,10 +119,7 @@ df['logIndoorConcentration'] = df['IndoorConcentration'].apply(np.log10)
 df['logAttenuationSubslab'] = df['AttenuationSubslab'].apply(np.log10)
 df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
 
-# figure saving settings
-fig_dir = './figures/simulation_predictions/'
-ext = '.pdf'
-dpi = 300
+
 
 """
 sim_cases = ('Pp','No Pp',)
@@ -173,7 +175,6 @@ class SimPrediction:
         sim_span = df[df['Simulation']==options[status][1]]
 
         grp = asu_span[['IndoorOutdoorPressure','AirExchangeRate']].groupby(pd.cut(asu_span['IndoorOutdoorPressure'], np.arange(-5, 5.5, 0.5))).describe(percentiles=[0.05, 0.95])
-        print(grp['IndoorOutdoorPressure'])
         p_in = grp['IndoorOutdoorPressure'].index
         p_in = np.arange(-5, 5, 0.5)
         #print(p_in.categories.astype('float'))
@@ -217,7 +218,7 @@ class SimPrediction:
         )
 
         sns.scatterplot(
-            data=asu_span.loc[asu_span['IndoorConcentration']>=0.001],
+            data=asu_span.loc[asu_span['IndoorConcentration']>=0.05],
             x='IndoorOutdoorPressure',
             y='logAttenuationAvgGroundwater',
             ax=ax,
@@ -228,6 +229,7 @@ class SimPrediction:
 
 
         plt.legend()
+        plt.savefig(fig_dir+'simulation_prediction_span_'+status.lower()+ext,dpi=300)
         plt.show()
         #print(grp[('AirExchangeRate','min')].values)
 
@@ -251,4 +253,5 @@ class SimPrediction:
 
 gs = gridspec.GridSpec(2, 2)
 
[…]

### 2019-03-01 tex/main.tex
Updated figures

@@
 		\centering
-	\begin{minipage}[c]{0.49\textwidth}
-		\centering
     \caption{ }
-    \label{fig:asu_phase}
-    \includegraphics[width=\textwidth]{asu_phase.pdf}
-	\end{minipage}
-	%\hspace{3cm}
-	\begin{minipage}[c]{0.49\textwidth}
-		\centering
-    \caption{ }
-    \label{fig:indianapolis_species}
-    \includegraphics[width=\textwidth]{indianapolis_species.pdf}
-	\end{minipage}
+    \label{fig:pair_grid}
+    \includegraphics[width=\textwidth]{simple_grid.pdf}
 \end{figure}
-
-\subsection{Seasonal Variability}
-
-\begin{figure}[htb!]
-  \caption{ }
-  \label{fig:seasonal_analysis}
-  % ASU/North Island plot
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \caption{ }
-    \label{fig:seasonal_analysis_asu}
-    \includegraphics[width=\textwidth]{asu_closed_season.pdf}
-  \end{subfigure}
-  % Indianapolis plot
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \caption{ }
-    \label{fig:seasonal_analysis_indianapolis}
-    \includegraphics[width=\textwidth]{indianapolis_pce_season.pdf}
-  \end{subfigure}
-\end{figure}
-
-\subsection{Maximum Change in Indoor Air Contaminant Concentration}
 
@@
     \label{fig:simulation_sampling_pp_open}
-    \includegraphics[width=\textwidth]{sampling_simulation_pp_open.pdf}
+    \includegraphics[width=\textwidth]{simulation_prediction_span_open.pdf}
   \end{subfigure}
@@
     \label{fig:simulation_sampling_pp_closed}
-    \includegraphics[width=\textwidth]{sampling_simulation_pp_closed.pdf}
+    \includegraphics[width=\textwidth]{simulation_prediction_span_closed.pdf}
   \end{subfigure}

### 2019-03-04 code/simulation_predictions.py
Added structure notes

@@ -179,8 +179,8 @@ class SimPrediction:
         p_in = np.arange(-5, 5, 0.5)
         #print(p_in.categories.astype('float'))
 
-        Ae_min = grp[('AirExchangeRate','5%')].values
-        Ae_max = grp[('AirExchangeRate','95%')].values
+        Ae_min = grp[('AirExchangeRate','min')].values
+        Ae_max = grp[('AirExchangeRate','max')].values
         Ae_mean = grp[('AirExchangeRate','mean')].values
 
         alpha_min = []

### 2019-03-04 tex/main.tex
Added structure notes
Writing

@@
 
+\section{Methods}
+
 \subsection{Vapor Intrusion Model}
+
+\begin{equation}
+  V \frac{\partial c_\mathrm{in}}{\partial t} = n - c_\mathrm{in} A_e V \label{eq:indoor_air}
+\end{equation}
 
@@
 		\centering
-    \caption{ }
+    \caption{Distributions of indoor/outdoor pressure differences (top row), and its correlation with indoor air contaminant concentrations (middle rows), and air exchange rate (bottom row), at three different VI sites - ASU house, Indianapolis site, and North Island NAS.}
     \label{fig:pair_grid}
@@
 \end{figure}
+
+Our discussion of examining the temporal variability of indoor air contaminant concentration begins by considering one of the most dynamic factors that influence VI - the indoor/outdoor pressure difference ($p_\mathrm{in/out}$).
+In the top row of Figure \ref{fig:pair_grid}, the distribution of $p_\mathrm{in/out}$ at three difference VI sites, the ASU house, Indianapolis site, and North Island NAS, are plotted.
+The distributions of $p_\mathrm{in/out}$ at these three site resemble normal-like distribution, with a mean of a few negativa Pa, indicating that the most probable state is that these sites are slightly depressized.
+At the ASU house and the Indianapolis site, the range of $p_\mathrm{in/out}$ values are relatively similar, whereas the North Island NAS site has significalty larger span of values.
+This is most likely due to the poor structural condition of the North Island site, making it much more suspectible to barometric pressure fluctations; the ASU house and Indianapolis site regular residential structure in good condition.
+The periods when the preferential pathway at the ASU house was open and closed are considered separately.
+% TODO: Add some sort of interesting conclusion
+
+Now we move on to look at $p_\mathrm{in/out}$ is correlated with the indoor air contaminant concentration ($c_\mathrm{in}$).
+The middle row of Figure \ref{fig:pair_grid} shows the log-10 values of $c_\mathrm{in}$ vs. $p_\mathrm{in/out}$, placed into evenly spaced bins, are plotted.
+Each bin represents the mean $c_\mathrm{in}$ and the errorbars show the 95\% confidence interval. % TODO: Change this description once you've fixed the figure.
+A linear regression curve is also fitted to the data, with the shaded part of the regression again indicating the 95\% confidence interval. % TODO: Make sure this is right.
+At the ASU house, it is readibly apparent that the preferential pathway has a profound effect on how sensitive $c_\mathrm{in}$ is to $p_\mathrm{in/out}$, showing that the closing of the preferential pathway fundamentally changes the site.
+Perhaps surprsingly, when the preferential pathway is closed $p_\mathrm{in/out}$ is not strongly correlated with $c_\mathrm{in}$.
+
+It is clear that $p_\mathrm{in/out}$ sets a trend, i.e.
+increased depressurziation is associated with higher $c_\mathrm{in}$, but there is still considerable variability for a given $p_\mathrm{in/out}$.
+
+$p_\mathrm{in/out}$ is important in determining the contaminant entry rate into a structure, but it also will partly determine the air exchange rate.
+How significantly $p_\mathrm{in/out}$ determines $A_e$ varies significantly with site specific conditions, e.g. HVAC operation, if windows are open/closed, how well insulted the structure is etc.
+$A_e$ is thus highly variate for a given $p_\mathrm{in/out}$ hypothesize that most of this variability may be explained by variation in air exchange rate ($A_e$).
+
+Air exchange
+
+This leads us to examine how $A_e$ is associated with
+
+The difference in correlation of $c_\mathrm{in}$ vs.
+$p_\mathrm{in/out}$ for the "open" and "closed" periods, as well as the $p_\mathrm{in/out}$ is more evenly distributed around 0 Pa support with our choice of model, where we assume that the preferential pathway acts by enhancing the "advective potential" of a VI site through improving communication between the indoor and outdoor.
+
+% ASU:
+% Pressure distributions similar
+% Yet drastically different c
+% => clearly driven by PP
+% Pressure relationship better when open than close
+% => Enhanced advection, motivating our choice of model
+% Poor relationship for closed,
+% => Diffusion dominated transport
+% Yet significant uncertainty and variance in indoor air?
+% Look at the influence of Ae on this
+% Pressure is related to Ae, but not determinant
+% => still large variability in Ae for a given p, what effect does this have?
+
+% Indianapolis & North Island
+% Unforunately no Ae data here like ASU, cannot look at that influence
+% Stronger correlation between p and c here
+% Indianapolis had some type of PP, though less significant than at ASU
+% North Island had very large pressure changes, may explain the stronger relationship
+
+% Constant Ae simulations
+% First consider only pressure influence (constant Ae)
+% PP Open
+% Clearly we can capture the trend when p < 0, but not well when p > 0
+% When p > 0, diffusion limit, but still variability, why?
+% => Indicate fluctuating Ae
+% Uniform soil
+% Removing gravel sub-base removes the "upswing"
+% Permeable subbase region necessary to realize "advective potential"
+% PP closed
+% Kind of capture a "trend", but not variability
+% Pressure not significant and/or advective transport too inhibited to be signifncant
+% => Indicate fluctating Ae driver for variability here
+% Main point:
+% Pressure controls Ae and entry rate, but in closed and uniform cases, it is insufficnet to dramatically increase entry rate
+% When open, this is no longer true
+% Pressure also help control Ae, but there is great uncertainty
+% This leads to the difficult to predict variability
+% => Need to incorporate Ae fluctuations to understand variability
+
+% New PP simulations with extreme Ae values for each pressure considered
+% Now we capture most of the variability for open and closed
+% Strong indicator that fluctating Ae is responsible for a significant portion of the variability
+% Needs to be controlled better (hard)
+% Or the span of values considered to give upper and lower limits on variability
+% Use EPA handbook to estimate variability (pretty much can always be one OOM)
+
+% This says nothing about at what speed or timescale change may occur
+% We look at the difference between max/min across various timespans
+% See that on daily to a week, absent a PP, little variability is expected
+% => 24 hr average samples adequate
+% => no need to sample more than a few days apart
+% Across time, more of the extreme values of Ae and pressure have occured, kind of ergodically, giving rise to this increase in variability
+% Remember p and Ae are more or less normally distributed, (Ae more skewed)
+% => extreme values unlikely but do happen across time
+
+% Problem with attenuation to subslab
+% Spatial variability in subslab may be greater than temporal indoor variability
+% ASU perfect example, here there is greater alpha_sb variability than c
+% => May be more uncertain
+% => False positive for indoor sources
+% => But could also indicate presence of PP in subbase
+% Much larger spatial variability in subbase with PP present than without
+% (Show plot with 2-4 panes)
+% 1./2. show ASU/Indianapolis alpha_sb
+% 3./4. model alpha_sb distributions with and without PP
 

### 2019-03-05 code/kde-iacc-pressure/ase-seasons.py
Removed old todos

@@ -92,7 +92,6 @@ for dataset in (pre_cpm,post_cpm,):
         #dataset = dataset.name
         df = dataset[dataset['Season']==season]
         print("Dataset: %s, Season: %s" % (dataset.name,season))
-        # TODO: use dataset mean instead of seasonal mean
         df['IndoorConcentration'] -= dataset['IndoorConcentration'].mean()
 
         if dp is True:
@@ -157,7 +156,6 @@ for dataset in (pre_cpm,post_cpm,):
     g.ax_joint.legend(handles=label_patches, loc='upper left')
     plt.tight_layout()
     plt.savefig('./figures/kde-iacc-pressure/%s_seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
-    # TODO: dynamically change axis limits?
     #plt.show()
     plt.clf()
     plt.cla()

### 2019-03-05 code/kde-iacc-pressure/indianapolis-seasons.py

@@ -109,7 +109,6 @@ for dataset in (indianapolis,):
         #dataset = dataset.name
         df = dataset[dataset['Season']==season]
         print("Dataset: %s, Season: %s" % (dataset.name,season))
-        # TODO: use dataset mean instead of seasonal mean
         df['IndoorConcentration'] -= dataset['IndoorConcentration'].mean()
 
         if dp is True:
@@ -174,7 +173,6 @@ for dataset in (indianapolis,):
     g.ax_joint.legend(handles=label_patches, loc='upper left')
     plt.tight_layout()
     plt.savefig('./figures/kde-iacc-pressure/%s_seasons.pdf' % dataset.name.replace(',','').replace(' ','_').lower(), dpi=300)
-    # TODO: dynamically change axis limits?
     #plt.show()
     plt.clf()
     plt.cla()

### 2019-03-05 code/land_drain_scenarios.py

@@ -37,7 +37,6 @@ df.rename(
 # data choosing
 df['AirExchangeRate'] *= 3600 # convert from 1/s to 1/hr
 df['logIndoorConcentration'] = df['IndoorConcentration'].apply(np.log10)
-#df['AttenuationSubslab'] *= 2e3 # this seems to fix the problem... any way to get it? TODO: Looking into this more
 df['logAttenuationSubslab'] = df['AttenuationSubslab'].apply(np.log10)
 df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
 
@@ -58,7 +57,7 @@ g = sns.lmplot(
 
 ax = g.axes[0][0]
 sns.lineplot(
-    data=df.loc[(df['AirExchangeRate'] >= 0.4) & (df['AirExchangeRate'] <= 0.6)][['IndoorOutdoorPressure','logAttenuationGroundwater','Simulation']],#TODO: find nicer way, looks like shit
+    data=df.loc[(df['AirExchangeRate'] >= 0.4) & (df['AirExchangeRate'] <= 0.6)][['IndoorOutdoorPressure','logAttenuationGroundwater','Simulation']],
     x='IndoorOutdoorPressure',
     y='logAttenuationGroundwater',
     hue='Simulation',

### 2019-03-05 code/multivariate_analysis/asu_season.py

@@ -76,7 +76,6 @@ pre_cpm.drop(columns=cols_to_drop,inplace=True)
 post_cpm.drop(columns=cols_to_drop,inplace=True)
 sns.pairplot(post_cpm, hue="Season", hue_order=['Winter','Fall','Spring','Summer'])
 
-# TODO: change axis labels (adding units) and adjust ticks for the indoor concentration
 plt.savefig('./figures/multivariate_analysis/asu_season.pdf',dpi=300)
 
 plt.show()

### 2019-03-05 code/multivariate_analysis/indianapolis_season.py

@@ -103,7 +103,6 @@ def get_season(x):
 indianapolis['Season'] = indianapolis['Month'].apply(lambda x: get_season(x))
 #indianapolis = indianapolis.drop(columns=['StopTime','Month'])
 
-# TODO: change axis labels (adding units) and adjust ticks for the indoor concentration
 sns.pairplot(indianapolis.loc[indianapolis['Specie']=='Chloroform'][['Pressure','AttenuationSubSurface','AttenuationGroundwater','Season']], hue="Season", hue_order=['Winter','Fall','Spring','Summer'])
 #plt.savefig('./figures/multivariate_analysis/indianapolis_season.pdf',dpi=300)
 #plt.show()

### 2019-03-05 code/simulation_predictions.py

@@ -115,7 +115,6 @@ def make_plots(asu):
 # data choosing
 df['AirExchangeRate'] *= 3600 # convert from 1/s to 1/hr
 df['logIndoorConcentration'] = df['IndoorConcentration'].apply(np.log10)
-#df['AttenuationSubslab'] *= 2e3 # this seems to fix the problem... any way to get it? TODO: Looking into this more
 df['logAttenuationSubslab'] = df['AttenuationSubslab'].apply(np.log10)
 df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
 

### 2019-03-05 code/temporal_variability/indoor_air_concentration.py

@@ -7,8 +7,6 @@ from scipy.stats import gaussian_kde
 import sqlite3
 import seaborn as sns
 
-# TODO: add indianapolis data and add to present analysis
-
 data_dir = './data/preferential-pathway-sensitivity/'
 #db_dir = '/home/jonathan/Dropbox/var/'
 db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'

### 2019-03-05 code/transient-response/asu_change_in_concentration.py

@@ -7,7 +7,6 @@ from scipy.stats import gaussian_kde
 import sqlite3
 import seaborn as sns
 
-# TODO: add indianapolis data and add to present analysis
 
 data_dir = './data/preferential-pathway-sensitivity/'
 db_dir = '/home/jonathan/Dropbox/var/'
@@ -158,8 +157,6 @@ indianapolis = process_indianapolis()
 
 # figure 1: differences in rate of change for pre and post CPM periods
 # and simulated comparison
-# TODO: add same simulation data but for with a land drain
-# TODO: figure out how to show how well or poorly your simulation fits the data
 fig, ax = plt.subplots()
 sns.kdeplot(pre_cpm['RateOfChange'], gridsize=5e3, ax=ax, label='ASU, PP open')
 sns.kdeplot(post_cpm['RateOfChange'], gridsize=5e3, ax=ax, label='ASU, PP closed')
@@ -182,7 +179,6 @@ custom_x_ticks(ax, my_xticks=np.log10([0.5, 0.75, 1.0, 1.5, 2.0]))
 ax.legend(loc='best')
 plt.savefig(fig_dir+'rate_of_change_simulations.pdf',dpi=300)
 # table 1: simulated rates of change for various Ae and soils
-# TODO: print out all the data you wish to tabulate here
 
 
 # figure 2: rate of change over the seasons

### 2019-03-05 code/transient-response/rate-of-change.py

@@ -176,7 +176,6 @@ custom_x_ticks(ax)
 ax.set_title('PP Closed')
 
 # resampling plots
-# TODO: repeat this analysis but within each season
 time_scales = ('h', 'D', 'W', 'M')
 
 fig, ax = plt.subplots()

### 2019-03-05 tex/main.tex
Writing

@@
 
+% PRESSURE DISTRIBUTION DISCUSSION
 Our discussion of examining the temporal variability of indoor air contaminant concentration begins by considering one of the most dynamic factors that influence VI - the indoor/outdoor pressure difference ($p_\mathrm{in/out}$).
@@
 At the ASU house and the Indianapolis site, the range of $p_\mathrm{in/out}$ values are relatively similar, whereas the North Island NAS site has significalty larger span of values.
-This is most likely due to the poor structural condition of the North Island site, making it much more suspectible to barometric pressure fluctations; the ASU house and Indianapolis site regular residential structure in good condition.
+This is most likely due to the poor structural condition of the North Island site, making it much more suspectible to barometric pressure fluctations; the ASU house and Indianapolis site are regular residential structure in good condition.
 The periods when the preferential pathway at the ASU house was open and closed are considered separately.
@@
 
+% P_IN/OUT VS. C_IN - ASU HOUSE
 Now we move on to look at $p_\mathrm{in/out}$ is correlated with the indoor air contaminant concentration ($c_\mathrm{in}$).
@@
 
+% - " - - INDIANAPOLIS AND NORTH ISLAND NAS
+At the Indianapolis and North Island NAS sites, we again see the association of higher $c_\mathrm{in}$ when the structures are depressurized, but with still significant variability for a given $p_\mathrm{in/out}$.
+
+% MODEL CHOICE DISCUSSION
+The difference in correlation of $c_\mathrm{in}$ vs.
+$p_\mathrm{in/out}$ for the "open" and "closed" periods, as well as the $p_\mathrm{in/out}$ is more evenly distributed around 0 Pa support our choice of model, where we assume that the preferential pathway acts by enhancing the "advective potential" of a VI site through improving communication between the indoor and outdoor.
+
+% AE DISCUSSION
 It is clear that $p_\mathrm{in/out}$ sets a trend, i.e.
 increased depressurziation is associated with higher $c_\mathrm{in}$, but there is still considerable variability for a given $p_\mathrm{in/out}$.
-
 $p_\mathrm{in/out}$ is important in determining the contaminant entry rate into a structure, but it also will partly determine the air exchange rate.
-How significantly $p_\mathrm{in/out}$ determines $A_e$ varies significantly with site specific conditions, e.g. HVAC operation, if windows are open/closed, how well insulted the structure is etc.
-$A_e$ is thus highly variate for a given $p_\mathrm{in/out}$ hypothesize that most of this variability may be explained by variation in air exchange rate ($A_e$).
-
-Air exchange
-
-This leads us to examine how $A_e$ is associated with
-
-The difference in correlation of $c_\mathrm{in}$ vs.
-$p_\mathrm{in/out}$ for the "open" and "closed" periods, as well as the $p_\mathrm{in/out}$ is more evenly distributed around 0 Pa support with our choice of model, where we assume that the preferential pathway acts by enhancing the "advective potential" of a VI site through improving communication between the indoor and outdoor.
+How significantly $p_\mathrm{in/out}$ determines $A_e$ varies with site specific conditions, e.g. HVAC operation, if windows are open/closed, how well insulted the structure is etc.
+$A_e$ is thus highly variate for a given $p_\mathrm{in/out}$, and we hypothesize that most of the variability in $c_\mathrm{in}$ at a specific $p_\mathrm{in/out}$ may be explained by variation in air exchange rate ($A_e$).
+In the bottom left of Figure \ref{fig:pair_grid} we can see that larger depressurziation is associated with higher $A_e$, while overpressurazation is associated with lower $A_e$, but with still significant uncertainty.
 
@@
 % North Island had very large pressure changes, may explain the stronger relationship
+
+\subsubsection{Influence Of Pressure On Indoor Air Variability}
+
+\begin{figure}[!h] % TODO: Adjust aspect ratio
+	\centering
+  \caption{Simulated cases}
+  \label{fig:land_drain_scenarios}
+  \includegraphics[width=0.5\textwidth]{land_drain_scenarios.pdf}
+\end{figure}
+
+To examine the role that $p_\mathrm{in/out}$ and $A_e$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$, and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
+The first of these scenarios is aimed to be similar to the period when the preferential pathway is open at the ASU house, and thus the modeled preferential pathway is active.
+In the second scenario, this preferential pathway feature is deactivated, akin to the period when the preferential pathway was closed at the ASU site.
+And in the third, the preferential pathway is again modeled, but the permeable gravel sub-base is instead assumed consist of the sandy loam, just like the surrounding soil, giving an "uniform soil" scenario.
+% TODO: Make sure that the soil type is correct
+The results of these modeled scearios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios}.
+The data is here binned into evenly spaced bins, with each dot representing the mean $\alpha_\mathrm{gw}$ and the error bars denote the 95\% confidence intervals.
+
+By examining the first scenario (blue) it is readily apparent that the model is able to capture the behavior of the preferential pathway when the structure is depressized ($p_\mathrm{in/out} < 0$) but not very well when the house is overpressurized ($p_\mathrm{in/out} > 0$).
+This shows that the preferential pathways acts by increasing advective potential, allowing even relatively small increases in depressurziation dramatically increases the contaminant entry rate into the house.
+Conversely, when the structure is overpressurized, contaminant entry is more effectively inhibited, leading to diffusion being the more important transport mechanism from the sub-base to the indoor air space, leading to the decrease in $\alpha_\mathrm{gw}$.
+Yet, when $p_\mathrm{in/out} > 0$ there is a slight increase in $\alpha_\mathrm{gw}$ which can be explained by the fact that lower $A_e$ are associated with overpressurazation, leading to an increase in $\alpha_\mathrm{gw}$,
+
+An interesting contrast to this scenario is third, when the preferential pathway is still present but the permeable gravel sub-base is not.
+Under these conditions, there is almost increase in advective transport as the impermeable sub-base effectively inhibits this.
+This leads to the conclusion that a unless there is some way for a preferential pathway to effectively communicate between the sub-base region (or wherever the preferential pathway is), then it is unlikely to find the dramatic effect observed at the ASU house.
+
+Lastly, we consider the second scenario, where there is no preferential pathway present (but there is a permeable gravel sub-base).
+Here the general (weak) trend of increased $\alpha_\mathrm{gw}$ with increase depressurziation is captured, but none of the variability is captured.
+There is inadequate potential for advective transport and thus $p_\mathrm{in/out}$ does not significantly increase the contaminant entry rate.
+Furthermore, the seemingly random variability is not captured.
+Clearly, the $A_e$ fluctations must be incorporated into the modeling methodoly to more accurately predict the variability in $\alpha_\mathrm{gw}$.
 
@@
 
-% New PP simulations with extreme Ae values for each pressure considered
-% Now we capture most of the variability for open and closed
-% Strong indicator that fluctating Ae is responsible for a significant portion of the variability
-% Needs to be controlled better (hard)
-% Or the span of values considered to give upper and lower limits on variability
-% Use EPA handbook to estimate variability (pretty much can always be one OOM)
+\subsubsection{Combined Influence Of Pressure And Air Exchange Rate On Indoor Air Variability}
 
-% This says nothing about at what speed or timescale change may occur
-% We look at the difference between max/min across various timespans
-% See that on daily to a week, absent a PP, little variability is expected
-% => 24 hr average samples adequate
-% => no need to sample more than a few days apart
-% Across time, more of the extreme values of Ae and pressure have occured, kind of ergodically, giving rise to this increase in variability
-% Remember p and Ae are more or less normally distributed, (Ae more skewed)
-% => extreme values unlikely but do happen across time
-
-% Problem with attenuation to subslab
-% Spatial variability in subslab may be greater than temporal indoor variability
-% ASU perfect example, here there is greater alpha_sb variability than c
-% => May be more uncertain
-% => False positive for indoor sources
-% => But could also indicate presence of PP in subbase
-% Much larger spatial variability in subbase with PP present than without
-% (Show plot with 2-4 panes)
-% 1./2. show ASU/Indianapolis alpha_sb
-% 3./4. model alpha_sb distributions with and without PP
-
-\begin{figure}[!h]
-	\centering
-	\begin{minipage}[c]{0.49\textwidth}
-		\centering
-    \caption{ }
-    \label{fig:resampling}
-    \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
-	\end{minipage}
-	%\hspace{3cm}
-	\begin{minipage}[c]{0.49\textwidth}
-		\centering
-    \caption{Simulated cases}
-    \label{fig:land_drain_scenarios}
-    \includegraphics[width=\textwidth]{land_drain_scenarios.pdf}
-	\end{minipage}
-\end{figure}
-
-\subsubsection{Modeling Indoor Air Variability}
-
-\begin{figure}[htb!]
+\begin{figure}[htb!] % TODO: Only plot 5th and 95th percentile
   \caption{ }
@@
 
+To model how much variability that may be expected for a given indoor/outdoor pressyre difference, $p_\mathrm{in/out}$ is binned from 5 to -5 Pa in 0.5 Pa intervals, and the maximum and minimum $A_e$ in that interval are used as additional inputs to the model.
+% TODO: Update this description if different interval is used
+
+An analysis similar to that in Figure \ref{fig:land_drain_scenarios}
+
+% New PP simulations with extreme Ae values for each pressure considered
+% Now we capture most of the variability for open and closed
+% Strong indicator that fluctating Ae is responsible for a significant portion of the variability
+% Needs to be controlled better (hard)
+% Or the span of values considered to give upper and lower limits on variability
+% Use EPA handbook to estimate variability (pretty much can always be one OOM)
+
+\subsection{Variation Timeperiod} % TODO: Improve title
+
+\begin{figure}[htb!] % TODO: Adjust aspect ratio
+  \centering
+  \caption{ (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
+  \label{fig:resampling}
+  \includegraphics[width=0.5\textwidth]{temporal_variability/resampling.pdf}
+\end{figure}
+
+% RESAMPLING TIME ANALYSIS
+So far our discussion has focused on how much variability in $c_\mathrm{in}$ can be expected, and some drivers for this, at the three different sites.
+However, this tells just little about how much $c_\mathrm{in}$ may vary over a given timeperiod.
+To examine this, the data is resampled over a chosen timeperiod, and the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are plotted for this period.
+I.e.
+if $c_\mathrm{in}$ samples were collected every four hours over a period of a year, and the data is resampled on a daily basis, then $c_\mathrm{max}/c_\mathrm{min}$ is returned for within each day, giving 365 data points.
+For this analysis, we consider the period when the preferential pathway was open/closed at the ASU house, and the three indoor air contaminant found at the Indianapolis site (TCE, PCE, and Chloroform).
+The North Island NAS dataset is excluded here, as the data that we have available only spans a few days versus year(s) for the ASU house and Indianapolis sites.
+
+Figure \ref{fig:resampling} shows the the expected maximum variability over different resampling periods (from 1 day to 6 months).
+As one might expect, the longer the resampling period, the larger the maximum variability is, spanning from less than a threefold difference, to two to three orders of magnitude.
+That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
+What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may may take a few weeks before an order of magnitude maximum variability is reached.
+The most significant result of this analysis is that the maximum variability is quite small across a few days, suggesting that e.g.
+24-hour passive samplers will resolve the daily temporal variabilty in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
+
+% Ergodicity argument here
+
+% This says nothing about at what speed or timescale change may occur
+% We look at the difference between max/min across various timespans
+% See that on daily to a week, absent a PP, little variability is expected
+% => 24 hr average samples adequate
+% => no need to sample more than a few days apart
+% Across time, more of the extreme values of Ae and pressure have occured, kind of ergodically, giving rise to this increase in variability
+% Remember p and Ae are more or less normally distributed, (Ae more skewed)
+% => extreme values unlikely but do happen across time
+
+\subsection{Attenuation To Subslab} % TODO: Improve title
+
+% Problem with attenuation to subslab
+% Spatial variability in subslab may be greater than temporal indoor variability
+% ASU perfect example, here there is greater alpha_sb variability than c
+% => May be more uncertain
+% => False positive for indoor sources
+% => But could also indicate presence of PP in subbase
+% Much larger spatial variability in subbase with PP present than without
+% (Show plot with 2-4 panes)
+% 1./2. show ASU/Indianapolis alpha_sb
+% 3./4. model alpha_sb distributions with and without PP
+
 \begin{acknowledgement}

### 2019-03-06 code/land_drain_scenarios.py
New flucating Ae figures

@@ -43,59 +43,142 @@ df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
 #fig, ax = plt.subplots()
 
 
-g = sns.lmplot(
-    data=asu.loc[asu['Phase']!='CPM'][['IndoorOutdoorPressure','logAttenuationAvgGroundwater','Phase']],
-    x='IndoorOutdoorPressure',
-    y='logAttenuationAvgGroundwater',
-    hue='Phase',
-    x_bins=np.linspace(-5,5,20),
-    fit_reg=False,
-    legend_out=False,
-    legend=False,
-    #aspect=1.5,
-)
+class ConstantAe:
+    def __init__(self):
 
-ax = g.axes[0][0]
-sns.lineplot(
-    data=df.loc[(df['AirExchangeRate'] >= 0.4) & (df['AirExchangeRate'] <= 0.6)][['IndoorOutdoorPressure','logAttenuationGroundwater','Simulation']],
-    x='IndoorOutdoorPressure',
-    y='logAttenuationGroundwater',
-    hue='Simulation',
-    hue_order=['Pp', 'No Pp', 'Pp Uniform',],
-    ax=ax,
-    #legend=False,
-)
+        g = sns.lmplot(
+            data=asu.loc[asu['Phase']!='CPM'][['IndoorOutdoorPressure','logAttenuationAvgGroundwater','Phase']],
+            x='IndoorOutdoorPressure',
+            y='logAttenuationAvgGroundwater',
+            hue='Phase',
+            x_bins=np.linspace(-5,5,20),
+            fit_reg=False,
+            legend_out=False,
+            legend=False,
+            #aspect=1.5,
+        )
 
-
-handles, labels = ax.get_legend_handles_labels()
-handles = handles[1:]
-
-labels = (
-    'PP present',
-    'PP absent',
-    'PP present, \ngravel sub-base\nabsent',
-    'Data, PP open',
-    'Data, PP closed',
-)
-
-ax.legend(
-    handles,
-    labels,
-    #ncol=2,
-    loc='best',
-)
-
-ax.set_xlim([-6,6])
-ax.set_ylim([-7,-3.7])
-
-ax.set_xlabel('$\\Delta p_\\mathrm{in/out}$ (Pa)')
-ax.set_ylabel('$\\log{(\\alpha_\\mathrm{gw})}$')
[…]

### 2019-03-06 code/simulation_predictions.py

@@ -170,59 +170,47 @@ class SimPrediction:
             'Open': ['Open', 'Pp'],
             'Closed': ['Closed', 'No Pp'],
         }
-        asu_span = asu.loc[(asu['Phase']==options[status][0]) & ( (asu['IndoorOutdoorPressure']>=-5) & (asu['IndoorOutdoorPressure']<=5) )]
-        sim_span = df[df['Simulation']==options[status][1]]
+        quantiles = list(asu.loc[asu['Phase']==options[status][0]]['logAttenuationAvgGroundwater'].quantile([0.05,0.95]))
+        ae_describe = asu['AirExchangeRate'].describe(percentiles=[0.05, 0.95])
 
-        grp = asu_span[['IndoorOutdoorPressure','AirExchangeRate']].groupby(pd.cut(asu_span['IndoorOutdoorPressure'], np.arange(-5, 5.5, 0.5))).describe(percentiles=[0.05, 0.95])
-        p_in = grp['IndoorOutdoorPressure'].index
-        p_in = np.arange(-5, 5, 0.5)
-        #print(p_in.categories.astype('float'))
 
-        Ae_min = grp[('AirExchangeRate','min')].values
-        Ae_max = grp[('AirExchangeRate','max')].values
-        Ae_mean = grp[('AirExchangeRate','mean')].values
+        data = asu.loc[
+            (asu['Phase']==options[status][0]) &
+            ((asu['IndoorOutdoorPressure']>=-5) & (asu['IndoorOutdoorPressure']<=5)) #&
+            #(
+            #    (asu['logAttenuationAvgGroundwater']>=quantiles[0]) &
+            #    (asu['logAttenuationAvgGroundwater']<=quantiles[1])
+            #)
+        ]
 
-        alpha_min = []
-        alpha_max = []
-        alpha_mean = []
 
-        func = self.get_interp_func(sim_span)
-        for p, min, max, mean in zip(p_in, Ae_min, Ae_max, Ae_mean):
-
-            alpha_min.append(func(p,max)[0])
-            alpha_max.append(func(p,min)[0])
-            alpha_mean.append(func(p,mean)[0])
-
-        alpha_min = np.array(alpha_min)
-        alpha_max = np.array(alpha_max)
-        alpha_mean = np.array(alpha_mean)
+        sim = df.loc[
+            (df['Simulation']==options[status][1]) &
+            (
+                (df['AirExchangeRate'] >= ae_describe['5%']) &
[…]

### 2019-03-06 tex/main.tex
Working on new paper

@@
 
-\subsection{Statistical Analysis}
-
-This paper heavily relies on statistical analysis of high resolution datasets from two well-studied VI sites, one near Hill AFB in Utah (called the ASU house) and another in Indianapolis, IN (simply called as such.)
-Analysis is performed using the SciPy, NumPy, Pandas, and Seaborn Python packages.
-Probability distributions of various parameters are constructed using the kernel density estimation (KDE) method\cite{altman_introduction_1992}, which is implemented in the SciPy package.
+\subsection{Statistical Analysis Of Field Data}
 
@@
 
-\section{Methods}
+\subsection{Modeling Work}
+In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
+This model was implemented in a finite element solver package, COMSOL Multiphysics.
+In the present work, there has been an addition of a preferential pathway to the ”standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
+As in the earlier studies, only the vadose zone soil domain is directly modeled.
 
-\subsection{Vapor Intrusion Model}
+The modeled structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
+The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the perimeter of the house.
+All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
+The indoor control volume is assumed to consist of only of the basement, assumed as having a total volume of $300 \; \mathrm{m^3}$.
+Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of certain other features can be.
+\par
 
-\begin{equation}
-  V \frac{\partial c_\mathrm{in}}{\partial t} = n - c_\mathrm{in} A_e V \label{eq:indoor_air}
-\end{equation}
+The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy clay (except as noted).
+Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
+\par
+
+Where relevant The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
+The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, allowing the contaminant concentration in the pipe to be parameterized.
+\par
+
+The groundwater beneath the structure is assumed to be homogeneously contaminated with trichloroethylene (TCE) as a prototypical contaminant.
+The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
+The ground surface and the pipe are both assumed to be sources of air to the soil domain.
+Both are assumed to be at reference atmospheric pressure. \par
+
+Vapor transport in the soil is governed by Richard’s equation, a modified version  of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
+The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
+The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
+The transport of vapor contaminant in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
+The equations and the boundary conditions are given in Table \ref{tbl:eqns-bc-parameters}.
+
+\begin{table}[htb!]
+  \centering
+  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}
+  \label{tbl:eqns-bc-parameters}
+  \bigskip
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % Governing equations
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subcaption{Governing equations}
+  \begin{tabular}{l l}
+    \toprule
+    % Indoor air space equation
+    Unsteady-CSTR                 & $V\frac{d u}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - u A_e V$ \\
+    % Richard's equation
+    Richard's equation            & $\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$ \\
+    % Millington-Quirk equation
+    Millington-Quirk              & $D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$ \\
+    % Advection-diffusion equation
+    Advection-diffusion equation  & $\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$ \\
+    % van Genuchten's equations
+    \multirow{3}{*}{van Genuchten equations}     & $\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$ \\
+                                  & $\theta_g = \theta_t - \theta_w$ \\
+                                  & $k_r = (1 - \mathrm{Se})^{l} [1 - (\mathrm{Se}^{-m})^m]^2$ \\
+                                  & $m = 1 - 1/n$ \\
+    \bottomrule
+  \end{tabular}
+  \bigskip
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % Boundary conditions
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subcaption{Boundary conditions}
+  \begin{tabular}{l l l}
+    \toprule
+    \textbf{Boundary}          & \textbf{Richard's equation}      &   \textbf{Advection-diffusion equation} \\
+    % Foundation crack
+    At foundation crack  & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$                            & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
+    % Groundwater source
+    At groundwater source &  N/A & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$ \\
+    % Ground surface
+    At ground surface      & $p = 0 \; \mathrm{(Pa)}$  & $c = 0 \; \mathrm{(\mu g/m^3)}$ \\
+    % Preferential pathway
+    Exit of preferential pathway  & $p = 0 \; \mathrm{(Pa)}$  & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$ \\
+    \bottomrule
+  \end{tabular}
+  \bigskip
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % Soil input parameters
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subcaption{Soil \& gravel properties\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
+  \begin{tabular}{l l l l l l l}
+    \toprule
+    % Descriptions
+    Soil & $\text{Permeability} \; \mathrm{(m^2)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$  & $\theta_s$  & $\theta_r$  & $\alpha \; \mathrm{(1/m)}$  & $n$ \\
+    % Gravel
+    Gravel     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
+    % Sand
+    Sand     & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053        & 3.5       & 3.2 \\
+    % Sandy clay
+    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
+    \bottomrule
+  \end{tabular}
+  \bigskip
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % TCE input parameters
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subcaption{Trichloroethylene (diluted in air) properties\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
+  \begin{tabular}{l l l l l l}
+    \toprule
+    $D_\mathrm{air} \; \mathrm{(m^2/h)}$  & $D_\mathrm{water} \; \mathrm{(m^2/h)}$  & $\mathrm{Density} \; \mathrm{(kg/m^3)}$ & $\mathrm{Viscosity} \; \mathrm{(Pa \cdot s)}$  & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
+    $2.47 \cdot 10^{-2}$  & $3.67 \cdot 10^{-6}$  & 1.614 & $1.86 \cdot 10^{-5}$  & 0.403 & 131.39 \\
+    \bottomrule
+  \end{tabular}
+  \bigskip
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % Building input parameters
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \subcaption{Building properties}
+  \begin{tabular}{l l l}
+    \toprule
+    $V_\mathrm{base} \; \mathrm{(m^3)}$  & $L_\mathrm{slab} \; \mathrm{(cm)}$  & $A_e \; \mathrm{(1/hr)}$ \\
+    %
+    300  &  15  & 0.5 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   \caption{Simulated cases}
-  \label{fig:land_drain_scenarios}
-  \includegraphics[width=0.5\textwidth]{land_drain_scenarios.pdf}
+  \label{fig:land_drain_scenarios_constant_ae}
+  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_constant_ae.pdf}
 \end{figure}
 
-To examine the role that $p_\mathrm{in/out}$ and $A_e$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$, and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
+To examine the role that $p_\mathrm{in/out}$ and $A_e$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$ (the mean $A_e$ at the ASU house), and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
 The first of these scenarios is aimed to be similar to the period when the preferential pathway is open at the ASU house, and thus the modeled preferential pathway is active.
@@
 % TODO: Make sure that the soil type is correct
-The results of these modeled scearios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios}.
-The data is here binned into evenly spaced bins, with each dot representing the mean $\alpha_\mathrm{gw}$ and the error bars denote the 95\% confidence intervals.
+The results of these modeled scearios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios_constant_ae}.
+The data is here binned into evenly spaced bins, with each dot representing the mean $\alpha_\mathrm{gw}$ and the error bars denote the one standard deviation.
 
@@
 
-\begin{figure}[htb!] % TODO: Only plot 5th and 95th percentile
-  \caption{ }
-  \label{fig:simulation_sampling}
-  % ASU/North Island plot
-  \begin{subfigure}{0.49\textwidth}
+\begin{figure}[!h] % TODO: Adjust aspect ratio
 	\centering
-    \caption{ }
-    \label{fig:simulation_sampling_pp_open}
-    \includegraphics[width=\textwidth]{simulation_prediction_span_open.pdf}
-  \end{subfigure}
-  % Indianapolis plot
-  \begin{subfigure}{0.49\textwidth}
-    \centering
-    \caption{ }
-    \label{fig:simulation_sampling_pp_closed}
-    \includegraphics[width=\textwidth]{simulation_prediction_span_closed.pdf}
-  \end{subfigure}
+  \caption{Simulated cases}
+  \label{fig:land_drain_scenarios_fluctuating_ae}
+  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
 \end{figure}
 
-To model how much variability that may be expected for a given indoor/outdoor pressyre difference, $p_\mathrm{in/out}$ is binned from 5 to -5 Pa in 0.5 Pa intervals, and the maximum and minimum $A_e$ in that interval are used as additional inputs to the model.
-% TODO: Update this description if different interval is used
+% INTRODUCE THE FIGURE
+To predict the variability of $\alpha_\mathrm{gw}$ for a given pressure, we use the minimum and maximum $A_e$ values observed at the ASU site (0.06 and 3.38 per hour respectively), and calculate $\alpha_\mathrm{gw}$ over the same range of $p_\mathrm{in/out}$ values.
+This gives the range of values that $\alpha_\mathrm{gw}$ may fluctuate over for a given $p_\mathrm{in/out}$.
+Using this method, the analysis in Figure \ref{fig:land_drain_scenarios_constant_ae} is repeated, i.e. that the cases when the preferential pathways is open and closed respectively are considered.
+The result of this analysis can be seen in Figure \ref{fig:land_drain_scenarios_fluctuating_ae}, where this range of $\alpha_\mathrm{gw}$ is represented by the shaded areas, and the solid line is the predicted $\alpha_\mathrm{gw}$ when $A_e$ assumes the mean value of 0.50 per hour.
+Data from ASU is binned in 40 evenly spaced bins, where the dot represents the mean value, and the error bars represent one standard deviation.
 
-An analysis similar to that in Figure \ref{fig:land_drain_scenarios}
+The trends in Figure \ref{fig:land_drain_scenarios_fluctuating_ae} are the same as in Figure \ref{fig:land_drain_scenarios_constant_ae}, but it is clear that much more variability is accounted for, especially considering the majority of datapoints were recorded between -3 and 3 Pa (the bins are evenly spaced, not evenly sized).
+More e
 

### 2019-03-06 code/land_drain_scenarios.py

@@ -51,11 +51,11 @@ class ConstantAe:
             x='IndoorOutdoorPressure',
             y='logAttenuationAvgGroundwater',
             hue='Phase',
-            x_bins=np.linspace(-5,5,20),
+            x_bins=np.linspace(-5,5,40),
             fit_reg=False,
             legend_out=False,
             legend=False,
-            #aspect=1.5,
+            aspect=1.5,
         )
 
         ax = g.axes[0][0]
@@ -101,13 +101,16 @@ class ConstantAe:
         plt.tight_layout()
 
         plt.savefig('./figures/simulation_predictions/land_drain_scenarios_constant_ae.png',dpi=300)
+        plt.savefig('./figures/simulation_predictions/land_drain_scenarios_constant_ae.pdf',dpi=300)
+
         plt.show()
         return
 
 
 class FluctuatingAe:
     def __init__(self):
-        ae_describe = asu['AirExchangeRate'].describe(percentiles=[0.05, 0.95])
+        ae_describe = asu.loc[asu['Phase']!='CPM']['AirExchangeRate'].describe(percentiles=[0.05, 0.95])
+        print(ae_describe)
         data = asu.loc[
             ((asu['IndoorOutdoorPressure']>=-5) & (asu['IndoorOutdoorPressure']<=5)) &
             (asu['Phase']!='CPM')
@@ -120,8 +123,8 @@ class FluctuatingAe:
 
         sim = df.loc[
             (
-                (df['AirExchangeRate'] >= ae_describe['5%']) &
-                (df['AirExchangeRate'] <= ae_describe['95%'])
+                (df['AirExchangeRate'] >= ae_describe['min']) &
+                (df['AirExchangeRate'] <= ae_describe['max'])
             )
         ]
         g = sns.lmplot(
@@ -129,23 +132,36 @@ class FluctuatingAe:
             x='IndoorOutdoorPressure',
             y='logAttenuationAvgGroundwater',
             hue='Phase',
-            x_bins=np.linspace(-5,5,20),
+            x_bins=np.linspace(-5,5,40),
+            #x_bins=20,
             fit_reg=False,
             legend_out=False,
             legend=False,
-            #aspect=1.5,
+            aspect=1.5,
         )
 
         ax = g.axes[0][0]
-        sns.lineplot(
+        ax = sns.lineplot(
             data=sim,
             x='IndoorOutdoorPressure',
             y='logAttenuationGroundwater',
             hue='Simulation',
[…]

### 2019-03-07 code/pair_grids/large_grids.py
Splitting and organising

@@ -1,0 +1,102 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import matplotlib.gridspec as gridspec
+from scipy import stats
+
+
+
+asu = pd.read_csv('./data/asu_house.csv')
+indianapolis = pd.read_csv('./data/indianapolis.csv')
+
+# data processing
+asu = asu.loc[(asu['IndoorOutdoorPressure'] >= -5) & (asu['IndoorOutdoorPressure'] <= 5)] # TODO: kind of ugly "hack", should instead adjust axes
+
+# global settings
+num_bins = 10
+fig_path = './figures/pair_grids/'
+ext = '.pdf'
+dpi = 300
+
+# method to set custom axis labels
+def custom_axis(g):
+
+    replacements = {
+        'IndoorOutdoorPressure': '$\\Delta p_\\mathrm{in/out}$ (Pa)',
+        'AirExchangeRate': '$A_e$ (1/h)',
+        'logIndoorConcentration': '$\\log{(c_\\mathrm{in})}$ ($\\mathrm{\\mu g/m^3}$)',
+        'logAttenuationSubslab': '$\\log{(\\alpha_\\mathrm{subslab})}$'
+    }
+
+    for i in range(len(g.axes)):
+        for j in range(len(g.axes)):
+            xlabel = g.axes[i][j].get_xlabel()
+            ylabel = g.axes[i][j].get_ylabel()
+            if xlabel in replacements.keys():
+                g.axes[i][j].set_xlabel(replacements[xlabel])
+            if ylabel in replacements.keys():
+                g.axes[i][j].set_ylabel(replacements[ylabel])
+    return g
+
+
+# demonstrate the variability
+
+# asu
+# PP hue
+g = sns.PairGrid(
+    asu.loc[asu['Phase']!='CPM'][['IndoorOutdoorPressure','AirExchangeRate','logIndoorConcentration','logAttenuationSubslab','Phase']],
+    hue='Phase',
+)
+g = g.map_upper(plt.scatter)
+g = g.map_lower(sns.regplot, x_bins=num_bins, truncate=True, )
+g = g.map_diag(sns.kdeplot, shade=True)
+g = g.add_legend()
+g = custom_axis(g)
+plt.savefig(fig_path+'asu_phase'+ext,dpi=dpi)
+
+# season hue
+for phase in asu['Phase'].unique():
+    try:
+        g = sns.PairGrid(
[…]

### 2019-03-07 code/pair_grids/simple_grids.py

@@ -1,107 +1,10 @@
+
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 import seaborn as sns
 import matplotlib.gridspec as gridspec
 from scipy import stats
-
-
-
-asu = pd.read_csv('./data/asu_house.csv')
-indianapolis = pd.read_csv('./data/indianapolis.csv')
-
-# data processing
-asu = asu.loc[(asu['IndoorOutdoorPressure'] >= -5) & (asu['IndoorOutdoorPressure'] <= 5)] # TODO: kind of ugly "hack", should instead adjust axes
-
-# global settings
-num_bins = 10
-fig_path = './figures/pair_grids/'
-ext = '.pdf'
-dpi = 300
-
-# method to set custom axis labels
-def custom_axis(g):
-
-    replacements = {
-        'IndoorOutdoorPressure': '$\\Delta p_\\mathrm{in/out}$ (Pa)',
-        'AirExchangeRate': '$A_e$ (1/h)',
-        'logIndoorConcentration': '$\\log{(c_\\mathrm{in})}$ ($\\mathrm{\\mu g/m^3}$)',
-        'logAttenuationSubslab': '$\\log{(\\alpha_\\mathrm{subslab})}$'
-    }
-
-    for i in range(len(g.axes)):
-        for j in range(len(g.axes)):
-            xlabel = g.axes[i][j].get_xlabel()
-            ylabel = g.axes[i][j].get_ylabel()
-            if xlabel in replacements.keys():
-                g.axes[i][j].set_xlabel(replacements[xlabel])
-            if ylabel in replacements.keys():
-                g.axes[i][j].set_ylabel(replacements[ylabel])
-    return g
-
-"""
-# demonstrate the variability
-
-# asu
-# PP hue
-g = sns.PairGrid(
-    asu.loc[asu['Phase']!='CPM'][['IndoorOutdoorPressure','AirExchangeRate','logIndoorConcentration','logAttenuationSubslab','Phase']],
-    hue='Phase',
-)
-g = g.map_upper(plt.scatter)
-g = g.map_lower(sns.regplot, x_bins=num_bins, truncate=True, )
-g = g.map_diag(sns.kdeplot, shade=True)
-g = g.add_legend()
-g = custom_axis(g)
-plt.savefig(fig_path+'asu_phase'+ext,dpi=dpi)
-
-# season hue
-for phase in asu['Phase'].unique():
-    try:
-        g = sns.PairGrid(
[…]

### 2019-03-11 tmp.py
Not needed

@@ -1,6 +1,0 @@
-import sys
-print(sys.path)
-
-from .code.utils import get_dropbox_path
-
-#print(get_dropbox_path())

### 2019-03-11 code/figures.py
Generated new figures

@@ -29,7 +29,7 @@ class Figure1:
 
 
         datasets = (nas, asu_open, asu_closed,)
-        labels = ('North Island NAS', 'ASU house, land drain open', 'ASU house, land drain closed',)
+        labels = ('North Island NAS', 'ASU house, PP open', 'ASU house, PP closed',)
 
         fig, ax = plt.subplots()
 
@@ -63,13 +63,16 @@ class Figure1:
             xlim=[-30,15],
             ylim=[-1,1.5],
             xlabel='$p_\\mathrm{in/out} \\; \\mathrm{(Pa)}$',
-            ylabel='$\\log{(c_\\mathrm{in})} \\; \\mathrm{(\\mu g/m^3)}$',
+            ylabel='$c_\\mathrm{in}/c_\\mathrm{in,mean}$',
             title='Relationship between indoor/outdoor pressure difference and\nTCE in indoor air (normalized to mean concentration)',
             #yscale='log',
             yticks=yticks,
             yticklabels=yticklabels,
         )
         plt.legend(loc='upper left')
+        plt.savefig('./figures/2d_kde/nas_asu_pp.pdf', dpi=300)
+        plt.savefig('./figures/2d_kde/nas_asu_pp.png', dpi=300)
+
         plt.show()
 
         return
@@ -86,4 +89,6 @@ def get_log_ticks(start, stop):
 
 
     return ticks, labels
+
+
 Figure1(y_data_log=True,norm_conc=True)

### 2019-03-11 code/temporal_variability/pressure.py
Analysis for Yijun

@@ -1,0 +1,44 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import matplotlib.gridspec as gridspec
+from scipy import stats
+import sqlite3
+db_dir = 'C:\\Users\\jstroem\\Dropbox\\var\\'
+
+db = sqlite3.connect(db_dir + 'HillAFB.db')
+
+asu = pd.read_sql_query("SELECT StopTime AS Time, Pressure FROM PressureDifference;",
+        db,
+    ).sort_values(by='Time')
+asu['Time'] = asu['Time'].apply(pd.to_datetime)
+
+
+phases = pd.read_sql_query("SELECT * from phases;", db)
+phases['StartTime'] = phases['StartTime'] .apply(pd.to_datetime)
+phases['StopTime'] = phases['StopTime'] .apply(pd.to_datetime)
+
+phase1 = phases[(phases['CPM'] == 'off') & (phases['LandDrain'] == 'open')]
+phase3 = phases[(phases['CPM'] == 'off') & (phases['LandDrain'] == 'closed')]
+
+pp_status = lambda x: 'Open' if x < phase1['StopTime'].values[0] else ('Closed' if x > phase3['StartTime'].values[0] else 'CPM')
+asu['Phase'] = asu['Time'].apply(pp_status)
+
+asu = asu.loc[asu['Phase']!='CPM']
+
+asu['dp'] = asu['Pressure'].diff().apply(np.abs)
+asu['dt'] = asu['Time'].diff()/np.timedelta64(1, 'm')
+asu['dpdt'] = asu['dp']/asu['dt']
+asu['dpdt'] = asu['dpdt'].replace([-np.inf,np.inf], np.nan)
+asu.dropna(inplace=True)
+
+print(asu['dpdt'].describe(percentiles=[0.01, 0.05, 0.95, 0.99]))
+
+
+sns.boxplot(
+    x="Phase",
+    y="dpdt",
+    data=asu,
+)
+plt.show()

### 2019-03-11 code/figures.py
Adjusted plots

@@ -64,7 +64,7 @@ class Figure1:
             ylim=[-1,1.5],
             xlabel='$p_\\mathrm{in/out} \\; \\mathrm{(Pa)}$',
             ylabel='$c_\\mathrm{in}/c_\\mathrm{in,mean}$',
-            title='Relationship between indoor/outdoor pressure difference and\nTCE in indoor air (normalized to mean concentration)',
+            title='Relationship between indoor/outdoor pressure difference and\nTCE in indoor air (normalized to dataset mean concentration)',
             #yscale='log',
             yticks=yticks,
             yticklabels=yticklabels,
@@ -90,5 +90,20 @@ def get_log_ticks(start, stop):
 
     return ticks, labels
 
+
+class AttenuationSubslab:
+    def __init__(self):
+        asu = pd.read_csv('./data/asu_house.csv')
+
+
+        fig, (ax1, ax2) = plt.subplots()
+
+
+        for phase in asu['Phase'].unique()
+        sns.kdeplot(
+            data=asu.loc[asu['Phase']==phase]
+        )
+
+        return
 
 Figure1(y_data_log=True,norm_conc=True)

### 2019-03-11 code/resampling.py

@@ -62,7 +62,7 @@ g = sns.catplot(
     kind="point",
     data=resampled,
     legend_out=False,
-    # aspect=1.5,
+    aspect=1.5,
 )
 
 my_ytick_labels = ["%1.1f" % y_tick for y_tick in 10.0**g.ax.get_yticks()]
@@ -71,5 +71,7 @@ g.ax.set_xlabel('Period')
 g.ax.set_ylabel('$c_\\mathrm{max}/c_\\mathrm{min}$')
 g.ax.set_title('Maximum variability within a given period')
 plt.tight_layout()
-# plt.show()
+
 plt.savefig('./figures/temporal_variability/resampling.pdf', dpi=300)
+plt.savefig('./figures/temporal_variability/resampling.png', dpi=300)
+plt.show()

### 2019-03-11 tex/main.tex
Reordered figures

@@
   {./../figures/},
-  {./../figures/air-exchange-rate/},
-  {./../figures/kde-iacc-pressure/},
+  {./../figures/2d_kde/},
   {./../figures/simulation_predictions/},
   {./../figures/pair_grids/},
-  {./../figures/multivariate_analysis/}
+  {./../figures/temporal_variability/},
 }
@@
     \caption{Distributions of indoor/outdoor pressure differences (top row), and its correlation with indoor air contaminant concentrations (middle rows), and air exchange rate (bottom row), at three different VI sites - ASU house, Indianapolis site, and North Island NAS.}
-    \label{fig:pair_grid}
-    \includegraphics[width=\textwidth]{simple_grid.pdf}
+    \label{fig:kde}
+    \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
 \end{figure}
@@
 
+\subsection{Variation Timeperiod} % TODO: Improve title
+
+\begin{figure}[htb!] % TODO: Adjust aspect ratio
+  \centering
+  \caption{ (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
+  \label{fig:resampling}
+  \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
+\end{figure}
+
+% RESAMPLING TIME ANALYSIS
+So far our discussion has focused on how much variability in $c_\mathrm{in}$ can be expected, and some drivers for this, at the three different sites.
+However, this tells just little about how much $c_\mathrm{in}$ may vary over a given timeperiod.
+To examine this, the data is resampled over a chosen timeperiod, and the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are plotted for this period.
+I.e.
+if $c_\mathrm{in}$ samples were collected every four hours over a period of a year, and the data is resampled on a daily basis, then $c_\mathrm{max}/c_\mathrm{min}$ is returned for within each day, giving 365 data points.
+For this analysis, we consider the period when the preferential pathway was open/closed at the ASU house, and the three indoor air contaminant found at the Indianapolis site (TCE, PCE, and Chloroform).
+The North Island NAS dataset is excluded here, as the data that we have available only spans a few days versus year(s) for the ASU house and Indianapolis sites.
+
+Figure \ref{fig:resampling} shows the the expected maximum variability over different resampling periods (from 1 day to 6 months).
+As one might expect, the longer the resampling period, the larger the maximum variability is, spanning from less than a threefold difference, to two to three orders of magnitude.
+That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
+What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may may take a few weeks before an order of magnitude maximum variability is reached.
+The most significant result of this analysis is that the maximum variability is quite small across a few days, suggesting that e.g.
+24-hour passive samplers will resolve the daily temporal variabilty in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
+
 % ASU:
@@
 
+\begin{figure}[!h] % TODO: Adjust aspect ratio
+	\centering
+  \caption{Simulated cases}
+  \label{fig:land_drain_scenarios_fluctuating_ae}
+  \includegraphics[width=\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
+\end{figure}
+
+\begin{comment}
 \begin{figure}[!h] % TODO: Adjust aspect ratio
@@
   \label{fig:land_drain_scenarios_fluctuating_ae}
-  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
+  \includegraphics[width=\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
 \end{figure}
@@
 % Use EPA handbook to estimate variability (pretty much can always be one OOM)
-
-\subsection{Variation Timeperiod} % TODO: Improve title
-
-\begin{figure}[htb!] % TODO: Adjust aspect ratio
-  \centering
-  \caption{ (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
-  \label{fig:resampling}
-  \includegraphics[width=0.5\textwidth]{temporal_variability/resampling.pdf}
-\end{figure}
-
-% RESAMPLING TIME ANALYSIS
-So far our discussion has focused on how much variability in $c_\mathrm{in}$ can be expected, and some drivers for this, at the three different sites.
-However, this tells just little about how much $c_\mathrm{in}$ may vary over a given timeperiod.
-To examine this, the data is resampled over a chosen timeperiod, and the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are plotted for this period.
-I.e.
-if $c_\mathrm{in}$ samples were collected every four hours over a period of a year, and the data is resampled on a daily basis, then $c_\mathrm{max}/c_\mathrm{min}$ is returned for within each day, giving 365 data points.
-For this analysis, we consider the period when the preferential pathway was open/closed at the ASU house, and the three indoor air contaminant found at the Indianapolis site (TCE, PCE, and Chloroform).
-The North Island NAS dataset is excluded here, as the data that we have available only spans a few days versus year(s) for the ASU house and Indianapolis sites.
-
-Figure \ref{fig:resampling} shows the the expected maximum variability over different resampling periods (from 1 day to 6 months).
-As one might expect, the longer the resampling period, the larger the maximum variability is, spanning from less than a threefold difference, to two to three orders of magnitude.
-That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
-What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may may take a few weeks before an order of magnitude maximum variability is reached.
-The most significant result of this analysis is that the maximum variability is quite small across a few days, suggesting that e.g.
-24-hour passive samplers will resolve the daily temporal variabilty in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
+\end{comment}
 

### 2019-03-12 code/figures.py
Added statistical analysis

@@ -45,6 +45,9 @@ class Figure1:
 
 
             r, p = stats.pearsonr(data[x], data2)
+
+            print(label)
+            print(data[x].describe(percentiles=[0.05,0.95]), data['IndoorConcentration'].describe(percentiles=[0.05,0.95]), data2.describe(percentiles=[0.05,0.95]))
             sns.kdeplot(
                 data=data[x],
                 data2=data2,
@@ -70,8 +73,8 @@ class Figure1:
             yticklabels=yticklabels,
         )
         plt.legend(loc='upper left')
-        plt.savefig('./figures/2d_kde/nas_asu_pp.pdf', dpi=300)
-        plt.savefig('./figures/2d_kde/nas_asu_pp.png', dpi=300)
+        #plt.savefig('./figures/2d_kde/nas_asu_pp.pdf', dpi=300)
+        #plt.savefig('./figures/2d_kde/nas_asu_pp.png', dpi=300)
 
         plt.show()
 
@@ -99,10 +102,10 @@ class AttenuationSubslab:
         fig, (ax1, ax2) = plt.subplots()
 
 
-        for phase in asu['Phase'].unique()
-        sns.kdeplot(
-            data=asu.loc[asu['Phase']==phase]
-        )
+        for phase in asu['Phase'].unique():
+            sns.kdeplot(
+                data=asu.loc[asu['Phase']==phase]
+            )
 
         return
 

### 2019-03-12 tex/main.tex
Writing

@@
 
-\begin{figure}[!h]
+\begin{figure}[htb!]
 		\centering
@@
 
-% PRESSURE DISTRIBUTION DISCUSSION
-Our discussion of examining the temporal variability of indoor air contaminant concentration begins by considering one of the most dynamic factors that influence VI - the indoor/outdoor pressure difference ($p_\mathrm{in/out}$).
-In the top row of Figure \ref{fig:pair_grid}, the distribution of $p_\mathrm{in/out}$ at three difference VI sites, the ASU house, Indianapolis site, and North Island NAS, are plotted.
-The distributions of $p_\mathrm{in/out}$ at these three site resemble normal-like distribution, with a mean of a few negativa Pa, indicating that the most probable state is that these sites are slightly depressized.
-At the ASU house and the Indianapolis site, the range of $p_\mathrm{in/out}$ values are relatively similar, whereas the North Island NAS site has significalty larger span of values.
-This is most likely due to the poor structural condition of the North Island site, making it much more suspectible to barometric pressure fluctations; the ASU house and Indianapolis site are regular residential structure in good condition.
-The periods when the preferential pathway at the ASU house was open and closed are considered separately.
-% TODO: Add some sort of interesting conclusion
+\begin{table}[htb!]
+  \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
+  \begin{tabular}{c c c c c c c}
+    \toprule
+    & \multicolumn{2}{c|}{\textbf{North Island NAS}} & \multicolumn{2}{c|}{\textbf{ASU house PP Open}} & \multicolumn{2}{c}{\textbf{ASU House PP Closed}} \\
+    Percentile & 5th & 95th & 5th & 95th & 5th & 95th \\
+    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1 & -2.1 & 2.27 \\
+    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2 & 3.3 & 0.4 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
-% P_IN/OUT VS. C_IN - ASU HOUSE
-Now we move on to look at $p_\mathrm{in/out}$ is correlated with the indoor air contaminant concentration ($c_\mathrm{in}$).
-The middle row of Figure \ref{fig:pair_grid} shows the log-10 values of $c_\mathrm{in}$ vs. $p_\mathrm{in/out}$, placed into evenly spaced bins, are plotted.
-Each bin represents the mean $c_\mathrm{in}$ and the errorbars show the 95\% confidence interval. % TODO: Change this description once you've fixed the figure.
-A linear regression curve is also fitted to the data, with the shaded part of the regression again indicating the 95\% confidence interval. % TODO: Make sure this is right.
-At the ASU house, it is readibly apparent that the preferential pathway has a profound effect on how sensitive $c_\mathrm{in}$ is to $p_\mathrm{in/out}$, showing that the closing of the preferential pathway fundamentally changes the site.
-Perhaps surprsingly, when the preferential pathway is closed $p_\mathrm{in/out}$ is not strongly correlated with $c_\mathrm{in}$.
+The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) is an important driving force in VI, drawing in (or preventing) contaminants from entering a structure.
+% TODO: Add references
+Changes in $p_\mathrm{in/out}$ is also a dynamic and fast process, impacting VI more rapidly than e.g.
+fluctuations in groundwater depth or contaminant concentration does; these processes may take weeks or even months to impact the overlying structure.
+Therefore it is reasonable to assume that the temporal variability in indoor air contaminant concentration $c_\mathrm{in}$ may be driven by changes in $p_\mathrm{in/out}$.
 
-% - " - - INDIANAPOLIS AND NORTH ISLAND NAS
-At the Indianapolis and North Island NAS sites, we again see the association of higher $c_\mathrm{in}$ when the structures are depressurized, but with still significant variability for a given $p_\mathrm{in/out}$.
+We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$  by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
+The KDE plot allows us to view the distribution of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and how well these correlate.
+For this analysis we consider two VI sites, North Island NAS, and the ASU house, with the ASU house dataset divided up into two periods, before and after the land drain (called preferential pathway (PP) from here on) had been closed.
+The preferential pathway significantly impacted the ASU house to such an extent that these two periods can essentially be considered as two different VI sites, and allowing us to examine the effect it had.
 
-% MODEL CHOICE DISCUSSION
-The difference in correlation of $c_\mathrm{in}$ vs.
-$p_\mathrm{in/out}$ for the "open" and "closed" periods, as well as the $p_\mathrm{in/out}$ is more evenly distributed around 0 Pa support our choice of model, where we assume that the preferential pathway acts by enhancing the "advective potential" of a VI site through improving communication between the indoor and outdoor.
+In Figure \ref{fig:kde} $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing us to compare the impact $p_\mathrm{in/out}$ had on $c_\mathrm{in}$.
+A value of 10 on the y-axis indicate that $c_\mathrm{in}$ is 10 times greater than the mean here, and 0.1 indicate that it is one tenth of the mean.
+The Pearson's r between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend.
 
-% AE DISCUSSION
-It is clear that $p_\mathrm{in/out}$ sets a trend, i.e.
-increased depressurziation is associated with higher $c_\mathrm{in}$, but there is still considerable variability for a given $p_\mathrm{in/out}$.
-$p_\mathrm{in/out}$ is important in determining the contaminant entry rate into a structure, but it also will partly determine the air exchange rate.
-How significantly $p_\mathrm{in/out}$ determines $A_e$ varies with site specific conditions, e.g. HVAC operation, if windows are open/closed, how well insulted the structure is etc.
-$A_e$ is thus highly variate for a given $p_\mathrm{in/out}$, and we hypothesize that most of the variability in $c_\mathrm{in}$ at a specific $p_\mathrm{in/out}$ may be explained by variation in air exchange rate ($A_e$).
-In the bottom left of Figure \ref{fig:pair_grid} we can see that larger depressurziation is associated with higher $A_e$, while overpressurazation is associated with lower $A_e$, but with still significant uncertainty.
+An interesting aspect of the North Island NAS site is that $p_\mathrm{in/out}$ varies so significantly here, with the 5th and 9th percentile $p_\mathrm{in/out}$ -19.9 and 7.4 Pa respectively.
+This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (PP open), and -2.1 and 2.27 Pa (PP closed).
 
-\subsection{Variation Timeperiod} % TODO: Improve title
+The large (roughly one order of magnitude) under- and overpressurization of the North Island NAS site leads to roughly a one order magnitude increase and decrease from the mean $c_\mathrm{in}$.
+Combined with $r=-0.73$, it is quite clear that at this site $p_\mathrm{in/out}$ largely determines $c_\mathrm{in}$ (there is still variability in $c_\mathrm{in}$ for a given $p_\mathrm{in/out}$, which we address later).
+This is the same principle that governs the controlled pressure method (CPM) concept.
+
+Turning to the ASU house datasets, we see a quite different situation.
+Here the variability of $c_\mathrm{in}$ is just as large, or even larger than at North Island NAS, yet the $p_\mathrm{in/out}$ varies far less.
+At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed respectively are relatively comparable, but the 5th and 95th percentiles values of differ significantly $c_\mathrm{in}/c_\mathrm{in,mean}$ as may be seen in Table \ref{tbl:percentiles}.
+
+It is clear that the PP dramatically increases the house's sensitivity to $p_\mathrm{in/out}$, and the site is fundamentally different during the period when the PP was open.
+Yet, the magnitude of change in $c_\mathrm{in}$ is far greater than the change in $p_\mathrm{in/out}$, leading to the conclusion that it is not just increased flow rates into the structure that causes this, but there must also be an increased amount of contaminant as well as a much larger spatial variability in the sub-base.
+This is a topic that we will return to later.
+
+The magnitude of change in $c_\mathrm{in}$ when the PP is closed is more in line with the magnitude of change in $p_\mathrm{in/out}$, but there is still more variability than one would expect.
+We hypothesize that this is largely due to fluctuations in air exchange rate, which we will examine later.
+
+\subsection{Change In Indoor Air Contaminant Concentration Over Time} % TODO: Improve title
 
@@
 % RESAMPLING TIME ANALYSIS
-So far our discussion has focused on how much variability in $c_\mathrm{in}$ can be expected, and some drivers for this, at the three different sites.
-However, this tells just little about how much $c_\mathrm{in}$ may vary over a given timeperiod.
-To examine this, the data is resampled over a chosen timeperiod, and the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are plotted for this period.
+So far we have discussed how the relationship between $c_\mathrm{in}$ and $p_\mathrm{in/out}$, and how significantly $c_\mathrm{in}$ may vary at a site.
+But this tells us little about how much variability in $c_\mathrm{in}$ may be expected over time.
+For this analysis we turn to the ASU house data (PP open/closed considered separately again), and another well-studied VI site, the Indianapolis site in Indiana.
+Both of these sites collected high frequency $c_\mathrm{in}$ samples over a significant periods and are therefore suitable. % TODO: Reference to the time periods here
+Furthermore, at Indianapolis site the $c_\mathrm{in}$ of three different contaminants, Chloroform, Trichloroethylene (TCE), and Tetrachloroethylene (PCE) were collected, allowing us to see if there is any difference in the variability of each.
+The North Island NAS dataset only spans a few days, and is therefore excluded from this analysis.
+
+To demonstrate how significantly $c_\mathrm{in}$ can vary across time, we calculate at the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) over a given time period.
 I.e.
 if $c_\mathrm{in}$ samples were collected every four hours over a period of a year, and the data is resampled on a daily basis, then $c_\mathrm{max}/c_\mathrm{min}$ is returned for within each day, giving 365 data points.
-For this analysis, we consider the period when the preferential pathway was open/closed at the ASU house, and the three indoor air contaminant found at the Indianapolis site (TCE, PCE, and Chloroform).
-The North Island NAS dataset is excluded here, as the data that we have available only spans a few days versus year(s) for the ASU house and Indianapolis sites.
+Resampling periods of one, two, three, days, weeks, and months are chosen and Figure \ref{fig:resampling} shows the result of this analysis.
 
-Figure \ref{fig:resampling} shows the the expected maximum variability over different resampling periods (from 1 day to 6 months).
 As one might expect, the longer the resampling period, the larger the maximum variability is, spanning from less than a threefold difference, to two to three orders of magnitude.
 That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
-What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may may take a few weeks before an order of magnitude maximum variability is reached.
+What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may take a few weeks before an order of magnitude maximum variability is reached.
 The most significant result of this analysis is that the maximum variability is quite small across a few days, suggesting that e.g.
-24-hour passive samplers will resolve the daily temporal variabilty in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
+24-hour passive samplers will resolve the daily temporal variability in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
+There is also little difference between maximum variability of $c_\mathrm{in}$ the ASU house (when the PP is closed), and the three contaminants at the Indianapolis site.
 

### 2019-03-13 code/figures.py
Attenuation from subslab plot

@@ -81,13 +81,17 @@ class Figure1:
         return
 
 
-def get_log_ticks(start, stop):
+def get_log_ticks(start, stop, style='f'): # TODO: Remove the unnecessary ticklabels
 
     ticks = np.array([])
     for int_now in np.arange(np.floor(start),np.ceil(stop)+1):
         ticks = np.append(ticks, np.arange(0.1,1.1,0.1)*10.0**int_now)
 
-    labels = ['%1.1f' % tick for tick in ticks]
+
+    if style == 'f':
+        labels = ['%1.1f' % tick for tick in ticks]
+    elif style=='e':
+        labels = ['%1.1e' % tick for tick in ticks]
     ticks = np.log10(ticks)
 
 
@@ -97,16 +101,24 @@ def get_log_ticks(start, stop):
 class AttenuationSubslab:
     def __init__(self):
         asu = pd.read_csv('./data/asu_house.csv')
+        asu = asu.loc[asu['Phase']!='CPM']
 
+        ax = sns.boxplot(x="Phase", y="logAttenuationSubslab", data=asu)
 
-        fig, (ax1, ax2) = plt.subplots()
+        ticks, labels = get_log_ticks(-4,2, style='e')
+        ax.set(
+            xlabel='Preferential pathway status',
+            ylabel='Attenuation from subslab',
+            #yticks=ticks,
+            #yticklabels=labels,
+        )
 
+        plt.tight_layout()
+        plt.savefig('./figures/temporal_variability/asu_attenuation_subslab.pdf', dpi=300)
+        plt.savefig('./figures/temporal_variability/asu_attenuation_subslab.png', dpi=300)
 
-        for phase in asu['Phase'].unique():
-            sns.kdeplot(
-                data=asu.loc[asu['Phase']==phase]
-            )
-
+        plt.show()
         return
 
-Figure1(y_data_log=True,norm_conc=True)
+#Figure1(y_data_log=True,norm_conc=True)
+AttenuationSubslab()

### 2019-03-13 code/figures.py
New combined simulation figure

@@ -120,5 +120,92 @@ class AttenuationSubslab:
         plt.show()
         return
 
+
+class Modeling:
+    # TODO: Add log ticks
+    # TODO: Fix plot colors
+    # TODO: Remove fill for PP uniform case
+    # TODO: Fix legend labels
+    def __init__(self):
+        sim = PreferentialPathway().data
+        asu = pd.read_csv('./data/asu_house.csv')
+        asu = asu.loc[
+            (asu['Phase']!='CPM') &
+            (
+                (asu['IndoorOutdoorPressure'] >=-5) &
+                (asu['IndoorOutdoorPressure'] <= 5)
+            ) &
+            (
+                (asu['AirExchangeRate'] <= 1.5) &
+                (asu['AirExchangeRate'] >= 0.1)
+            )
+        ]
+
+        print(list(sim.loc[sim['Simulation']=='Pp Uncontaminated']))
+        fig, (ax1,ax2) = plt.subplots(2,1, sharex=True, sharey=True)
+
+        # ax1
+        sns.regplot(
+            data=asu.loc[asu['Phase']=='Open'],
+            x='IndoorOutdoorPressure',
+            y='logAttenuationAvgGroundwater',
+            ax=ax1,
+            fit_reg=False,
+            x_bins=np.linspace(-5,5,40),
+            ci='sd',
+            label='Data',
+        )
+
+        sns.lineplot(
+            data=sim,
+            x='IndoorOutdoorPressure',
+            y='logAttenuationGroundwater',
+            hue='Simulation',
+            hue_order=['Pp','Pp Uniform','Pp Uncontaminated',],
+            ax=ax1,
+        )
+
+
+        # ax2
+        sns.regplot(
+            data=asu.loc[asu['Phase']=='Closed'],
+            x='IndoorOutdoorPressure',
+            y='logAttenuationAvgGroundwater',
+            ax=ax2,
+            fit_reg=False,
+            x_bins=np.linspace(-5,5,40),
+            ci='sd',
+            label='Data'
+        )
+
+        sns.lineplot(
+            data=sim,
+            x='IndoorOutdoorPressure',
+            y='logAttenuationGroundwater',
+            hue='Simulation',
+            hue_order=['No Pp'],
+            ax=ax2,
+        )
+
+        ax1.set(
+            ylabel='$\\alpha_\\mathrm{gw}$',
+            title='Preferential pathway open',
+        )
[…]

### 2019-03-13 code/get_simulation_data.py

@@ -6,12 +6,15 @@ import pandas as pd
 class PreferentialPathway:
     def __init__(self):
         dfs = []
-        for file in ('pp','pp_uniform','no_pp',):
+        for file in ('pp','pp_uniform','no_pp','pp_uncontaminated',):
             df = pd.read_csv('./data/simulation/sweep_'+file+'.csv',header=4)
             df['Simulation'] = np.repeat(file.replace('_',' ').title(),len(df))
+            if file == 'pp_uncontaminated':
+                df['% Ae'] = np.repeat(0.5/3600, len(df))
+                df.rename(columns={'% p_in': 'p_in'},inplace=True)
             dfs.append(df)
 
-        df = pd.concat(dfs)
+        df = pd.concat(dfs,sort=True)
         df.rename(
             columns={
                 '% Ae': 'AirExchangeRate',

### 2019-03-13 tex/main.tex
Writing

@@
 
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy clay (except as noted).
+The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy loam (except as noted).
 Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
@@
     Gravel     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
-    % Sand
-    Sand     & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053        & 3.5       & 3.2 \\
-    % Sandy clay
-    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
+    % sandy loam
+    Sandy Loam    & $5.9 \cdot 10^{-13}$  & 1460    & 0.39        & 0.039        & 2.7       & 1.4 \\
     \bottomrule
@@
 
+\subsection{Variability Of Attenuation From Subslab}
+
+\begin{figure}[htb!]
+  \caption{Boxplot of attenuation from subslab at the ASU house site.
+The box shows the quartiles of the distribution, the whiskers the extent of the distribution, and diamonds are points that are considered outlier.}\label{fig:attenuation_subslab}
+  \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
+\end{figure}
+
+To examine how $p_\mathrm{in/out}$ affects the contaminant entry rate it would be most convenient to analyze how $p_\mathrm{in/out}$ is correlated with the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) at a site - in particularly for understanding how the PP at the ASU house affected the site dynamics.
+By normalizing the $c_\mathrm{in}$ to $c_\mathrm{subslab}$ the $p_\mathrm{in/out}$ influence on the contaminant availability in the subslab area is mitigated.
+The $c_\mathrm{subslab}$ are taken from the soilgas probe labeled as "6" at the ASU house, which is the probe located closest to the preferential pathway, and to a reported breach in the foundation\cite{guo_identification_2015}.
+However, as Figure \ref{fig:attenuation_subslab} shows, there are some issues with doing so.
+
+First, we can state that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ does not vary that significantly, and is mostly around the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03.
+% TODO: Reference the EPA report that recommends this
+During the period when the preferential pathway was open, there is considerably more variability, and it is not uncommon for the $\alpha_\mathrm{subslab}$ to exceed unity.
+
+Even though probe "6" is located in such close proximity to the preferential pathway, and the breach in the foundation, the contaminant concentration 'hotspot' is still missed.
+This suggests that a preferential pathway may have an extremely localized influence in the subslab region, causing significant spatial variability in $c_\mathrm{subslab}$.
+This shows how difficult it may be to locate a preferential pathway through subslab soilgas sampling.
+
+The large $\alpha_\mathrm{subslab}$ values (compared to the recommended EPA value of 0.03) can often indicate that there may be indoor sources present a site, which there were none of at the ASU house.
+Thus, in addition to potentially indicating indoor contaminant sources, large $\alpha_\mathrm{subslab}$ may also indicate the presence of a preferential pathway.
+
 \subsubsection{Influence Of Pressure On Indoor Air Variability}
 
-\begin{figure}[!h] % TODO: Adjust aspect ratio
+\begin{figure}[htb!]
 	\centering
   \caption{Simulated cases}
-  \label{fig:land_drain_scenarios_fluctuating_ae}
-  \includegraphics[width=\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
+  \label{fig:land_drain_scenarios}
+  \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
 \end{figure}
 
-\begin{comment}
-\begin{figure}[!h] % TODO: Adjust aspect ratio
-	\centering
-  \caption{Simulated cases}
-  \label{fig:land_drain_scenarios_constant_ae}
-  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_constant_ae.pdf}
-\end{figure}
-
-To examine the role that $p_\mathrm{in/out}$ and $A_e$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$ (the mean $A_e$ at the ASU house), and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
+To examine the role that $p_\mathrm{in/out}$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$ (the mean $A_e$ at the ASU house), and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
 The first of these scenarios is aimed to be similar to the period when the preferential pathway is open at the ASU house, and thus the modeled preferential pathway is active.
@@
 Clearly, the $A_e$ fluctations must be incorporated into the modeling methodoly to more accurately predict the variability in $\alpha_\mathrm{gw}$.
+
+\begin{comment}
+\begin{figure}[!h] % TODO: Adjust aspect ratio
+	\centering
+  \caption{Simulated cases}
+  \label{fig:land_drain_scenarios_constant_ae}
+  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_constant_ae.pdf}
+\end{figure}
 

### 2019-03-14 tex/main.tex
Working on the last part

@@
 \author{Yijun Yao}
-\affiliation[Zhejiang University]{Zhejiang University, Hangzhou, China}
+\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
 \author{Eric M. Suuberg}
@@
 
-\subsubsection{Influence Of Pressure On Indoor Air Variability}
+\subsubsection{Modeling Results} % TODO: Improve title
 
@@
 	\centering
-  \caption{Simulated cases}
+  \caption{Simulated preferential pathway scenarios compared to ASU house field data (top panel).
+No preferential pathway scenarios in bottom panel.
+Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars one standard deviation.}
   \label{fig:land_drain_scenarios}
@@
 \end{figure}
+
+To begin a more thorough discussion regarding the role and impact a preferential pathway may have at a particular site, we turn to our model described in the methods section.
+% TODO: Add reference to model figure
+Here, $A_e$ is assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ is varied from -5 to 5 Pa.
+The predicted $c_\mathrm{in}$ is then normalized to the groundwater concentration $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
+
+The predicted $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ is given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios} (ignore the shaded blue area for now).
+These predicted values are compared to $\alpha_\mathrm{gw}$ from the ASU house data from the period the preferential pathway was open, given by the blue dots.
+
+The model captures the trend
+
+Note that this modeled scenario is
 
@@
 And in the third, the preferential pathway is again modeled, but the permeable gravel sub-base is instead assumed consist of the sandy loam, just like the surrounding soil, giving an "uniform soil" scenario.
-% TODO: Make sure that the soil type is correct
-The results of these modeled scearios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios_constant_ae}.
+The results of these modeled scenarios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios_constant_ae}.
 The data is here binned into evenly spaced bins, with each dot representing the mean $\alpha_\mathrm{gw}$ and the error bars denote the one standard deviation.

### 2019-03-14 code/figures.py

@@ -163,6 +163,7 @@ class Modeling:
             hue='Simulation',
             hue_order=['Pp','Pp Uniform','Pp Uncontaminated',],
             ax=ax1,
+            ci='sd',
         )
 
 
@@ -185,6 +186,7 @@ class Modeling:
             hue='Simulation',
             hue_order=['No Pp'],
             ax=ax2,
+            ci='sd'
         )
 
         ax1.set(

### 2019-03-14 code/figures.py
Finished figure

@@ -129,6 +129,7 @@ class Modeling:
     def __init__(self):
         sim = PreferentialPathway().data
         asu = pd.read_csv('./data/asu_house.csv')
+        #print(asu['AirExchangeRate'].describe(percentiles=[0.1, 0.9]))
         asu = asu.loc[
             (asu['Phase']!='CPM') &
             (
@@ -141,8 +142,33 @@ class Modeling:
             )
         ]
 
-        print(list(sim.loc[sim['Simulation']=='Pp Uncontaminated']))
-        fig, (ax1,ax2) = plt.subplots(2,1, sharex=True, sharey=True)
+
+        # isolates the Ae = 0.5 case for the uniform soil modeling
+        sim_data_to_remove = [
+            (sim['Simulation']=='Pp Uniform') &
+            (
+                (sim['AirExchangeRate'] < 0.45) |
+                (sim['AirExchangeRate'] > 0.55)
+            )
+        ]
+
+        sim_data_to_remove = np.invert(sim_data_to_remove)
+        sim = sim[sim_data_to_remove[0]]
+
+
+        fig, (ax1,ax2) = plt.subplots(2,1, sharex=True, sharey=True, figsize=[6.4, 6.4], dpi=300)
+
+        pp_max = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==0.1)]
+        pp = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==0.5)]
+        pp_min = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==1.5)]
+
+
+        ax1.plot(pp['IndoorOutdoorPressure'], pp['logAttenuationGroundwater'],label='PP')
+        ax1.fill_between(pp['IndoorOutdoorPressure'], pp_min['logAttenuationGroundwater'], pp_max['logAttenuationGroundwater'],alpha=0.5)
+
+
+
+
 
         # ax1
         sns.regplot(
@@ -153,19 +179,31 @@ class Modeling:
             fit_reg=False,
             x_bins=np.linspace(-5,5,40),
             ci='sd',
-            label='Data',
+            label='ASU house',
+            color=sns.color_palette()[0]
         )
-
         sns.lineplot(
-            data=sim,
+            data=sim.loc[sim['Simulation']=='Pp Uniform'],
             x='IndoorOutdoorPressure',
             y='logAttenuationGroundwater',
-            hue='Simulation',
[…]

### 2019-03-14 code/get_simulation_data.py

@@ -33,10 +33,10 @@ class PreferentialPathway:
 
         # data choosing
         df['AirExchangeRate'] *= 3600 # convert from 1/s to 1/hr
+        df['AirExchangeRate'] = np.around(df['AirExchangeRate'], 1)
         df['logIndoorConcentration'] = df['IndoorConcentration'].apply(np.log10)
         df['logAttenuationSubslab'] = df['AttenuationSubslab'].apply(np.log10)
         df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
-
 
         self.data = df
         return

### 2019-03-14 tex/main.tex
Finished draft

--- 2019-03-13 tex/main.tex
@@
 This caused greater than normal advective flow from the subsurface into the house, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-This period of CPM testing is considered separately from the otherwise ”natural” VI conditions in the analysis.
+This period of CPM testing is thus excluded from the otherwise ”natural” VI conditions in the analysis.
 Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining the dataset, noting that during some of the testing, this pathway was deliberately cut off, resulting in what we have termed “normal” VI conditions in which the main source of contaminant was believed to be groundwater.
@@
 
-The typical variation in indoor air contaminant concentrations with time will first be considered below in the case of the ASU house during ”natural”, (i.e.
-non-CPM conditions), in the case of the NAS North Island site over the entire available dataset, and for the Indianapolis case we consider the variations before the installation of the SSD system.
-The deviations in indoor air concentration from the mean TCE (and Chloroform and PCE at the Indianapolis site) values, as well as the indoor-outdoor pressure differentials associated with these concentrations were examined.
-Both univariate and bivariate kernel density estimations (KDE) were constructed.
+Much of the analysis relies on an probability density estimation technique called "kernel density estimation",
 KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
@@
 In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
-The distributions of the individual parameters and the relationship between them will be examined using the KDE method.
+The numpy, scipy, and pandas python packages were also used to conduct all statistical analysis and processing, with plotting done by the matplotlib, seaborn packages. % TODO: Add references to the
 
 \subsection{Modeling Work}
+
+\begin{figure}[htb!]
+  \caption{Graphic of the preferential pathway mode.
+Note that the gravel sub-base material may be switched to the same as the surrounding soil, and that the preferential pathway may be turned off.}\label{fig:model}
+  %\includegraphics{model.png}
+\end{figure}
+
 In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
@@
 		\centering
-    \caption{Distributions of indoor/outdoor pressure differences (top row), and its correlation with indoor air contaminant concentrations (middle rows), and air exchange rate (bottom row), at three different VI sites - ASU house, Indianapolis site, and North Island NAS.}
+    \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}
     \label{fig:kde}
@@
 fluctuations in groundwater depth or contaminant concentration does; these processes may take weeks or even months to impact the overlying structure.
-Therefore it is reasonable to assume that the temporal variability in indoor air contaminant concentration $c_\mathrm{in}$ may be driven by changes in $p_\mathrm{in/out}$.
+Therefore, it is reasonable to assume that the temporal variability in indoor air contaminant concentration $c_\mathrm{in}$ may be driven by changes in $p_\mathrm{in/out}$.
 
 We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plot allows us to view the distribution of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and how well these correlate.
+The KDE plots allow us to view the distribution of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and how well these correlate.
 For this analysis we consider two VI sites, North Island NAS, and the ASU house, with the ASU house dataset divided up into two periods, before and after the land drain (called preferential pathway (PP) from here on) had been closed.
-The preferential pathway significantly impacted the ASU house to such an extent that these two periods can essentially be considered as two different VI sites, and allowing us to examine the effect it had.
+The preferential pathway significantly impacted the ASU house to such an extent that these two periods can essentially be considered as two different VI sites, and allowing us to examine the impact of the preferential pathway.
 
-In Figure \ref{fig:kde} $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing us to compare the impact $p_\mathrm{in/out}$ had on $c_\mathrm{in}$.
+In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing us to compare the impact $p_\mathrm{in/out}$ had on $c_\mathrm{in}$.
 A value of 10 on the y-axis indicate that $c_\mathrm{in}$ is 10 times greater than the mean here, and 0.1 indicate that it is one tenth of the mean.
@@
 
-An interesting aspect of the North Island NAS site is that $p_\mathrm{in/out}$ varies so significantly here, with the 5th and 9th percentile $p_\mathrm{in/out}$ -19.9 and 7.4 Pa respectively.
+An interesting aspect of the North Island NAS site is that $p_\mathrm{in/out}$ varies so significantly, and the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
 This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (PP open), and -2.1 and 2.27 Pa (PP closed).
@@
 Turning to the ASU house datasets, we see a quite different situation.
-Here the variability of $c_\mathrm{in}$ is just as large, or even larger than at North Island NAS, yet the $p_\mathrm{in/out}$ varies far less.
+The variability of $c_\mathrm{in}$ is just as large, or even larger than at North Island NAS, yet the $p_\mathrm{in/out}$ varies far less.
 At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed respectively are relatively comparable, but the 5th and 95th percentiles values of differ significantly $c_\mathrm{in}/c_\mathrm{in,mean}$ as may be seen in Table \ref{tbl:percentiles}.
 
-It is clear that the PP dramatically increases the house's sensitivity to $p_\mathrm{in/out}$, and the site is fundamentally different during the period when the PP was open.
+It is clear that the preferential pathway dramatically increases the ASU house's sensitivity to $p_\mathrm{in/out}$, and the site is fundamentally different during the period when the preferential pathway was open.
 Yet, the magnitude of change in $c_\mathrm{in}$ is far greater than the change in $p_\mathrm{in/out}$, leading to the conclusion that it is not just increased flow rates into the structure that causes this, but there must also be an increased amount of contaminant as well as a much larger spatial variability in the sub-base.
@@
 
-\begin{figure}[htb!] % TODO: Adjust aspect ratio
+\begin{figure}[htb!]
   \centering
-  \caption{ (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
+  \caption{Maximum change in indoor air contaminant concentration that may be expected over a given time period. (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
   \label{fig:resampling}
@@
 For this analysis we turn to the ASU house data (PP open/closed considered separately again), and another well-studied VI site, the Indianapolis site in Indiana.
-Both of these sites collected high frequency $c_\mathrm{in}$ samples over a significant periods and are therefore suitable. % TODO: Reference to the time periods here
+Both of these sites collected high frequency $c_\mathrm{in}$ samples over a significant periods and are therefore suitable for this analysis\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
 Furthermore, at Indianapolis site the $c_\mathrm{in}$ of three different contaminants, Chloroform, Trichloroethylene (TCE), and Tetrachloroethylene (PCE) were collected, allowing us to see if there is any difference in the variability of each.
@@
 
-To demonstrate how significantly $c_\mathrm{in}$ can vary across time, we calculate at the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) over a given time period.
+To demonstrate how significantly $c_\mathrm{in}$ can vary across time, we calculate the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) over a given time period.
 I.e.
@@
 
-As one might expect, the longer the resampling period, the larger the maximum variability is, spanning from less than a threefold difference, to two to three orders of magnitude.
+As one might expect, the longer the resampling period, the larger the maximum variability is: spanning from less than a threefold difference, to two to three orders of magnitude.
 That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
@@
 24-hour passive samplers will resolve the daily temporal variability in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
-There is also little difference between maximum variability of $c_\mathrm{in}$ the ASU house (when the PP is closed), and the three contaminants at the Indianapolis site.
-
-% ASU:
-% Pressure distributions similar
-% Yet drastically different c
-% => clearly driven by PP
-% Pressure relationship better when open than close
-% => Enhanced advection, motivating our choice of model
-% Poor relationship for closed,
-% => Diffusion dominated transport
-% Yet significant uncertainty and variance in indoor air?
-% Look at the influence of Ae on this
-% Pressure is related to Ae, but not determinant
-% => still large variability in Ae for a given p, what effect does this have?
-
-% Indianapolis & North Island
-% Unforunately no Ae data here like ASU, cannot look at that influence
-% Stronger correlation between p and c here
-% Indianapolis had some type of PP, though less significant than at ASU
-% North Island had very large pressure changes, may explain the stronger relationship
+There is also little difference between maximum variability of $c_\mathrm{in}$ the ASU house (when the PP is closed), and the three contaminants at the Indianapolis site, suggesting that this may be an "usual" amount of variability over these time periods.
 
@@
 
-To examine how $p_\mathrm{in/out}$ affects the contaminant entry rate it would be most convenient to analyze how $p_\mathrm{in/out}$ is correlated with the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) at a site - in particularly for understanding how the PP at the ASU house affected the site dynamics.
+To examine how $p_\mathrm{in/out}$ affects the contaminant entry rate (and subsequently the indoor air contaminant concentration) it would be convenient to analyze how $p_\mathrm{in/out}$ is correlated with the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) at a site - in particularly for understanding how the PP at the ASU house affected the site dynamics.
 By normalizing the $c_\mathrm{in}$ to $c_\mathrm{subslab}$ the $p_\mathrm{in/out}$ influence on the contaminant availability in the subslab area is mitigated.
+
 The $c_\mathrm{subslab}$ are taken from the soilgas probe labeled as "6" at the ASU house, which is the probe located closest to the preferential pathway, and to a reported breach in the foundation\cite{guo_identification_2015}.
-However, as Figure \ref{fig:attenuation_subslab} shows, there are some issues with doing so.
+However, as Figure \ref{fig:attenuation_subslab} shows, there are some issues with using $\alpha_\mathrm{subslab}$ for subsequent analysis.
 
-First, we can state that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ does not vary that significantly, and is mostly around the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03.
-% TODO: Reference the EPA report that recommends this
+First, we can state that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ does not vary that significantly, and is mostly around the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
 During the period when the preferential pathway was open, there is considerably more variability, and it is not uncommon for the $\alpha_\mathrm{subslab}$ to exceed unity.
@@
 	\centering
-  \caption{Simulated cases}
+  \caption{Simulated preferential pathway scenarios compared to ASU house field data (top panel).
+No preferential pathway scenarios in bottom panel.
+Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars one standard deviation for that binned mean.}
   \label{fig:land_drain_scenarios}
@@
 
-To examine the role that $p_\mathrm{in/out}$ play in determining $c_\mathrm{in}$ variability (given as attenuation from groundwater vapor source $\alpha_\mathrm{gw}$ here), we first consider modeling VI scenarios with a constant $A_e = 0.5 \; \mathrm{(1/hr)}$ (the mean $A_e$ at the ASU house), and varying $p_\mathrm{in/out}$ from 5 to -5 Pa.
-The first of these scenarios is aimed to be similar to the period when the preferential pathway is open at the ASU house, and thus the modeled preferential pathway is active.
-In the second scenario, this preferential pathway feature is deactivated, akin to the period when the preferential pathway was closed at the ASU site.
-And in the third, the preferential pathway is again modeled, but the permeable gravel sub-base is instead assumed consist of the sandy loam, just like the surrounding soil, giving an "uniform soil" scenario.
-% TODO: Make sure that the soil type is correct
-The results of these modeled scearios are then compared to the preferential pathways "open" and "closed" periods from the ASU data, and may be seen in Figure \ref{fig:land_drain_scenarios_constant_ae}.
-The data is here binned into evenly spaced bins, with each dot representing the mean $\alpha_\mathrm{gw}$ and the error bars denote the one standard deviation.
+To begin a more thorough discussion regarding the role and impact a preferential pathway may have at a particular site, we turn to our model presented in the methods section.
+Here, $A_e$ is assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ is varied from -5 to 5 Pa.
+The predicted $c_\mathrm{in}$ is then normalized to the groundwater concentration $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
 
-By examining the first scenario (blue) it is readily apparent that the model is able to capture the behavior of the preferential pathway when the structure is depressized ($p_\mathrm{in/out} < 0$) but not very well when the house is overpressurized ($p_\mathrm{in/out} > 0$).
-This shows that the preferential pathways acts by increasing advective potential, allowing even relatively small increases in depressurziation dramatically increases the contaminant entry rate into the house.
-Conversely, when the structure is overpressurized, contaminant entry is more effectively inhibited, leading to diffusion being the more important transport mechanism from the sub-base to the indoor air space, leading to the decrease in $\alpha_\mathrm{gw}$.
-Yet, when $p_\mathrm{in/out} > 0$ there is a slight increase in $\alpha_\mathrm{gw}$ which can be explained by the fact that lower $A_e$ are associated with overpressurazation, leading to an increase in $\alpha_\mathrm{gw}$,
+The predicted $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ is given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios} (ignore the shaded blue areas for now).
+These predicted values are compared to $\alpha_\mathrm{gw}$ from the ASU house data from the period the preferential pathway was open, given by the blue dots.
 
-An interesting contrast to this scenario is third, when the preferential pathway is still present but the permeable gravel sub-base is not.
-Under these conditions, there is almost increase in advective transport as the impermeable sub-base effectively inhibits this.
-This leads to the conclusion that a unless there is some way for a preferential pathway to effectively communicate between the sub-base region (or wherever the preferential pathway is), then it is unlikely to find the dramatic effect observed at the ASU house.
+The model successfully predicts the increase in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
+Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) occurs.
+This asymmetry is due to two factors.
 
-Lastly, we consider the second scenario, where there is no preferential pathway present (but there is a permeable gravel sub-base).
-Here the general (weak) trend of increased $\alpha_\mathrm{gw}$ with increase depressurziation is captured, but none of the variability is captured.
-There is inadequate potential for advective transport and thus $p_\mathrm{in/out}$ does not significantly increase the contaminant entry rate.
-Furthermore, the seemingly random variability is not captured.
-Clearly, the $A_e$ fluctations must be incorporated into the modeling methodoly to more accurately predict the variability in $\alpha_\mathrm{gw}$.
+First, the preferential pathway acts as preferential air source, circumventing the large resistance to flow in the surrounding soil, and allowing a small change in $p_\mathrm{in/out}$ to much more dramatically increase the advective flow into the structure from the subslab region.
+A second simulation, where the model is rerun with the preferential pathway present, but the permeable (gravel) layer in the subslab is removed and instead is assumed to be of the same material as the surrounding soil (sandy loam), giving a "uniform soil" scenario (orange line in the top panel of Figure \ref{fig:land_drain_scenarios}).
+This simulation demonstrates that without a permeable subslab to effectively allow the "advective potential" to be realized, a preferential pathway may not impact a VI site; a preferential pathway requires good communication between the indoor environment and itself to be as a significant contributor to VI as was seen at the ASU house site.
 
-\begin{comment}
-\begin{figure}[!h] % TODO: Adjust aspect ratio
-	\centering
-  \caption{Simulated cases}
-  \label{fig:land_drain_scenarios_constant_ae}
-  \includegraphics[width=0.5\textwidth]{land_drain_scenarios_constant_ae.pdf}
-\end{figure}
+The second factor is that the preferential pathway must contain contaminant vapors to be impactful.
+In a third simulation, the permeable subslab region is included, but the preferential pathway is filled with clean air, i.e.
+just allows the "advective potential" to be realized, but without any additional contaminant introduced.
+The result of this simulation is given by the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
+This shows that while there was a slightly larger increase in $\alpha_\mathrm{gw}$ compared to the "uniform soil" scenario as the structure is increasingly depressurized, it is nowhere near as significant as when the preferential pathway was filled with contaminant vapors.
+The contaminated and uncontaminated preferential pathway scenarios (blue and green lines respectively) thus give the range of $\alpha_\mathrm{gw}$ that one may see for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.
 
-% Constant Ae simulations
-% First consider only pressure influence (constant Ae)
-% PP Open
-% Clearly we can capture the trend when p < 0, but not well when p > 0
-% When p > 0, diffusion limit, but still variability, why?
-% => Indicate fluctuating Ae
-% Uniform soil
-% Removing gravel sub-base removes the "upswing"
-% Permeable subbase region necessary to realize "advective potential"
-% PP closed
-% Kind of capture a "trend", but not variability
-% Pressure not significant and/or advective transport too inhibited to be signifncant
-% => Indicate fluctating Ae driver for variability here
-% Main point:
-% Pressure controls Ae and entry rate, but in closed and uniform cases, it is insufficnet to dramatically increase entry rate
-% When open, this is no longer true
-% Pressure also help control Ae, but there is great uncertainty
-% This leads to the difficult to predict variability
-% => Need to incorporate Ae fluctuations to understand variability
+The model is also able to capture $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ when a preferential pathway is absent (but there is a permeable subslab region), as the bottom panel in Figure \ref{fig:land_drain_scenarios} shows.
+However, what these simulations fail to predict is the considerable variability of $\alpha_\mathrm{gw}$ for a given $p_\mathrm{in/out}$, as shown by the error bars in the two panels (denoting one standard deviation).
+A significant portion of this variability may be explained by variation in air exchange rate, $A_e$.
 
-\subsubsection{Combined Influence Of Pressure And Air Exchange Rate On Indoor Air Variability}
+Air exchange rate determines which portion of the indoor is exchanged with outdoor air, reducing the indoor air contaminant concentration.
+A high air exchange is associated with lower $c_\mathrm{in}$ and low with higher $c_\mathrm{in}$.
+$A_e$ is correlated with $p_\mathrm{in/out}$, i.e.
+that a depressurized structure has higher $A_e$ than an overpressurized structure as more air from the outside may be brought in, but with significant variability.
+Determining any determinant relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena has a huge effect on air exchange, i.e.
+for a given $p_\mathrm{in/out}$.
+Fortunately, the range of $A_e$ values are typically bounded, with the U.S.
+EPA reporting that (depending on region) the 10th percentile values range from 0.16 to 0.2 per hour, and the 50th and 90th percentiles range from 0.35 to 0.49 and 1.21 to 1.49 per hour respectively\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
+These values are in line with what as found at the both the ASU house and the Indianapolis site as Table \ref{tbl:air_exchange_rate} shows.
 
-\begin{figure}[!h] % TODO: Adjust aspect ratio
-	\centering
-  \caption{Simulated cases}
-  \label{fig:land_drain_scenarios_fluctuating_ae}
-  \includegraphics[width=\textwidth]{land_drain_scenarios_fluctuating_ae.pdf}
-\end{figure}
+To more accurately predict the range of $\alpha_\mathrm{gw}$ values for a given $p_\mathrm{in/out}$, the preferential pathway (top panel, blue line) and the no preferential pathway (bottom panel, blue line) scenarios are rerun with $A_e$ equal to 0.1, 0.5, and 1.5 per hour, giving us a predicted upper and lower bound for $\alpha_\mathrm{gw}$.
+These bounds are given by the blue region above and below the central line, with the upper and lower bound associated with 0.1 and 1.5 air exchange rate per hour respectively.
+Doing this allows us to capture most of the variability in $\alpha_\mathrm{gw}$, indicating that a significant portion of the variability in $\alpha_\mathrm{gw}$ for a given $p_\mathrm{in/out}$ may be attributed to fluctuations in air exchange rate.
+This could be used during VI site investigations to estimate how significantly the indoor air contaminant concentration may fluctuate.
 
-% INTRODUCE THE FIGURE
-To predict the variability of $\alpha_\mathrm{gw}$ for a given pressure, we use the minimum and maximum $A_e$ values observed at the ASU site (0.06 and 3.38 per hour respectively), and calculate $\alpha_\mathrm{gw}$ over the same range of $p_\mathrm{in/out}$ values.
-This gives the range of values that $\alpha_\mathrm{gw}$ may fluctuate over for a given $p_\mathrm{in/out}$.
-Using this method, the analysis in Figure \ref{fig:land_drain_scenarios_constant_ae} is repeated, i.e. that the cases when the preferential pathways is open and closed respectively are considered.
-The result of this analysis can be seen in Figure \ref{fig:land_drain_scenarios_fluctuating_ae}, where this range of $\alpha_\mathrm{gw}$ is represented by the shaded areas, and the solid line is the predicted $\alpha_\mathrm{gw}$ when $A_e$ assumes the mean value of 0.50 per hour.
-Data from ASU is binned in 40 evenly spaced bins, where the dot represents the mean value, and the error bars represent one standard deviation.
-
-The trends in Figure \ref{fig:land_drain_scenarios_fluctuating_ae} are the same as in Figure \ref{fig:land_drain_scenarios_constant_ae}, but it is clear that much more variability is accounted for, especially considering the majority of datapoints were recorded between -3 and 3 Pa (the bins are evenly spaced, not evenly sized).
-More e
-
-% New PP simulations with extreme Ae values for each pressure considered
-% Now we capture most of the variability for open and closed
-% Strong indicator that fluctating Ae is responsible for a significant portion of the variability
-% Needs to be controlled better (hard)
-% Or the span of values considered to give upper and lower limits on variability
-% Use EPA handbook to estimate variability (pretty much can always be one OOM)
-\end{comment}
-
-% Ergodicity argument here
-
-% This says nothing about at what speed or timescale change may occur
-% We look at the difference between max/min across various timespans
-% See that on daily to a week, absent a PP, little variability is expected
-% => 24 hr average samples adequate
-% => no need to sample more than a few days apart
-% Across time, more of the extreme values of Ae and pressure have occured, kind of ergodically, giving rise to this increase in variability
-% Remember p and Ae are more or less normally distributed, (Ae more skewed)
-% => extreme values unlikely but do happen across time
-
-\subsection{Attenuation To Subslab} % TODO: Improve title
-
-% Problem with attenuation to subslab
-% Spatial variability in subslab may be greater than temporal indoor variability
-% ASU perfect example, here there is greater alpha_sb variability than c
-% => May be more uncertain
-% => False positive for indoor sources
-% => But could also indicate presence of PP in subbase
-% Much larger spatial variability in subbase with PP present than without
-% (Show plot with 2-4 panes)
-% 1./2. show ASU/Indianapolis alpha_sb
-% 3./4. model alpha_sb distributions with and without PP
+\begin{table}[htb!]
+  \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
+  \begin{tabular}{l c c c}
+    \toprule
+    \textbf{Percentile} & \textbf{10th} & \textbf{50th} & \textbf{90th} \\
+    EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995} & 0.16-0.2 & 0.35-0.49 & 1.21-1.49 \\
+    ASU house\cite{holton_temporal_2013,guo_identification_2015} & 0.25 & 0.5 & 1.15 \\
+    Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34 & 0.74 & 1.27 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 

### 2019-03-20 tex/main.tex
Added Eric's changed (intro and first part of methods)

@@
 % title
-\title{Transient Variability In Vapor Intrusion And The Factors That Influence It}
+\title{Temporal Variations In Vapor Intrusion Induced Indoor Air Contaminant Concentrations }
 
@@
 Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
+Such variations make it difficult for those charged with protecting human health to formulate a response should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
+There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
+What represents a reasonable sampling strategy for a particular site a single 8 hr sample?
+Repeated 8 hr samples?
+Month-long samples?
+Continuous monitoring?
+
 VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-VI has been a recognized problem for some time, but many aspects remain poorly understood, particularly with respect to the causes of large temporal transients in indoor air concentrations.
-There is uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-\par
-
+The basic nature of VI has been understood for some time and it has been the subject of much study, but some aspects remain poorly understood, such as the causes of the sometimes observed large temporal transients in indoor air concentrations.
 Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
 All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations varied significantly with time - orders of magnitude on the timescale of a day or days.
-\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}. \par
+All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.
 
-In one instance the source of the variation was clearly established during the study; at the ASU house a field drain pipe (or “land drain”), which connected to a sewer system, was discovered beneath the house, and careful isolation of this source led to a clear conclusion that this preferential pathway significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway to an area beneath a structure would always be expected to lead to large fluctuations in indoor air contaminant concentrations.
-\par
+In one instance the source of the variation was clearly established during the study of the site.
+At the ASU house a drain pipe (or “land drain”) connected to a sewer system was discovered beneath the house.
+Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
+While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.
 
@@
 That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been generally implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study estimate that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while the consideration of a role of possible sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
-In some of these cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many cases, this scenario is not further considered here, where the focus is on pathways that deliver contaminant via the soil beneath a structure. \par
+Sewer lines have been previously implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
+A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
+Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
 
+In some of the above cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
+While potentially important in many instances, this scenario is not further considered here where the focus is on pathways that deliver contaminant via the soil beneath a structure.
 It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is a site at NAS North Island at which no preferential pathways have been identified.
+One example is the site at NAS North Island at which no preferential pathways have been identified.
 Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site. \par
+It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.
 
 This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as ”normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at  such normal  VI  sites vs. those characterized by preferential pathways.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are also discussed.
+In this work, the latter scenarios are referred to as “normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
+Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at such normal VI sites vs.
+those characterized by preferential pathways within the soil beneath the site.
+The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.
 
@@
 
-To begin to characterize transient behavior in indoor air contaminant concentrations, actual datasets are analyzed to establish common levels of variability at VI sites.
-For this purpose, the datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-This paper relies on statistical analysis of published field data, and readers are referred to the original works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.
-\par
+To frame the question of just how much variability in indoor air contaminant concentrations is actually seen, field datasets have been analyzed.
+For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
+Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.
 
-The ASU house data were obtained over a period of a few years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” operation.
-This caused greater than normal advective flow from the subsurface into the house, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-This period of CPM testing is thus excluded from the otherwise ”natural” VI conditions in the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining the dataset, noting that during some of the testing, this pathway was deliberately cut off, resulting in what we have termed “normal” VI conditions in which the main source of contaminant was believed to be groundwater.
-\par
+The ASU house data were obtained over a period of several years.
+During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
+The period of CPM testing is thus excluded from the analysis.
+Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.
 
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to large internal pressure fluctuations, much more extensive than those typically recorded at the ASU house during normal operations.
-Additionally, the underlying soil at NAS North Island is sandy and more permeable than that at the ASU site, which, as will be shown, contributes to the indoor air contaminant concentrations being more sensitive to pressure fluctuations\cite{hosangadi_high-frequency_2017}.
-\par
+The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to “large” internal pressure fluctuations.
+By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
+The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.
 
-Likewise, the Indianapolis site investigation spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD).
-The goal of the SSD testing was to mitigate the VI risk by drastically depressurizing the sub-slab area underneath the house, preventing the contaminants from entering the structure above.
+The Indianapolis site investigation also spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD) for VI mitigation.
 Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}, however at no point was this preferential pathway removed, making it difficult to assess how significant the role of the preferential pathway was at this site.
-Regardless of this it is of interest to consider the data from this site due to how extensive and complete the data collection was. \par
+It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}.
+Unlike at the ASU house, this preferential pathway was never removed or blocked, making it impossible to isolate the role of the preferential pathway at this site.
+It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
 
-Much of the analysis relies on an probability density estimation technique called "kernel density estimation",
-KDE is a technique that estimates the probability distribution of a random variable(s) by using multiple kernels, or weighting functions, and in this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values (and that there are statistical fluctuations associated with each sampling event).
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott's rule.
-The numpy, scipy, and pandas python packages were also used to conduct all statistical analysis and processing, with plotting done by the matplotlib, seaborn packages. % TODO: Add references to the
+Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
+KDE is a technique used for estimating the probability distribution of a random variable(s) by using multiple kernels, or weighting functions to characterize the data sets.
+In this case, Gaussian kernels are used to create the KDEs.
+This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values and that there are statistical fluctuations associated with each sampling event.
+In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
+The numpy, scipy, and pandas python packages were also used to conduct all statistical analysis and processing, with plotting done by the matplotlib, seaborn packages.
+% TODO: Add references to the packages
 

### 2019-03-20 code/figures.py
Made new figures

@@ -243,6 +243,68 @@ class Modeling:
 
         return
 
+class IndianapolisTime:
+    def __init__(self):
+        data = pd.read_csv('./data/indianapolis.csv')
+        data['Time'] = data['Time'].apply(pd.to_datetime)
+        data.sort_values(by='Time')
+
+        fig, ax = plt.subplots(dpi=300)
+
+        sns.lineplot(
+            data=data,
+            x='Time',
+            y='IndoorConcentration',
+            hue='Specie',
+            ax=ax,
+        )
+
+        ax.set_yscale('log')
+        ax.set(
+            title='Indoor air contaminant concentration at the Indianapolis site',
+            ylabel='$c_\\mathrm{in} \; \\mathrm{(\\mu g/m^3)}$',
+        )
+
+        plt.tight_layout()
+        plt.savefig('./figures/temporal_variability/time_indianapolis.png')
+        plt.savefig('./figures/temporal_variability/time_indianapolis.pdf')
+
+        plt.show()
+        return
+
+class AirExchangeRateKDE:
+    def __init__(self):
+
+        data = pd.read_csv('./data/asu_house.csv').dropna()
+        data = data.loc[data['Phase']!='CPM']
+
+
+        fig, ax = plt.subplots(dpi=300)
+        sns.kdeplot(
+            data=data['IndoorOutdoorPressure'],
+            data2=data['AirExchangeRate'],
+            shade=True,
+            shade_lowest=False,
+            gridsize=200,
+        )
+
+        ax.set(
+            title='2D KDE showing distributions and relationship between\nindoor/outdoor pressure and air exchange rate ',
+            ylabel='$A_e \; \\mathrm{(1/hr)}$',
+            xlabel='$p_\\mathrm{in/out} \; \\mathrm{(Pa)}$',
+            ylim=[0,1.5],
+            xlim=[-5,5],
+        )
+
+        plt.tight_layout()
+        plt.savefig('./figures/2d_kde/pressure_air_exchange_rate.png')
+        plt.savefig('./figures/2d_kde/pressure_air_exchange_rate.pdf')
+
+        plt.show()
+        return
+
 #Figure1(y_data_log=True,norm_conc=True)
 #AttenuationSubslab()
-Modeling()
+#Modeling()
+#IndianapolisTime()
+AirExchangeRateKDE()

### 2019-03-20 code/preferential-pathway-sensitivity/sensitivity-analysis.py
No longer needed

@@ -1,124 +1,0 @@
-import numpy as np
-import pandas as pd
-import pandas.tools.plotting as pdplt
-import matplotlib.pyplot as plt
-from matplotlib import cm
-import seaborn as sns
-import sqlite3
-data_dir = './data/preferential-pathway-sensitivity/'
-db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-
-"""
-Prepares model data
-"""
-# no PP data
-no_pp = pd.read_csv(
-    data_dir + 'param-no-preferential-pathway.csv',
-    header=4,
-)
-no_pp['PP'] = pd.Series(np.repeat('No', len(no_pp['p'])))
-# PP data
-pp = pd.read_csv(
-    data_dir + 'param-preferential-pathway.csv',
-    header=4,
-)
-pp['PP'] = pd.Series(np.repeat('Yes', len(no_pp['p'])))
-df = pd.concat([pp, no_pp],axis=0,sort=True)
-# adds friendlier tags for gravel sub-base present (SB) and if the PP is
-# contaminated (chi)
-df['Gravel sub-base'] = pd.Series(df['SB'].apply(lambda x: 'Yes' if x == 1 else 'No'))
-df['Contaminated PP'] = pd.Series(df['chi'].apply(lambda x: 'Yes' if x == 1 else 'No'))
-# data from permeable model, unknown if I want to add it,
-# might be able to concat at it and assign SB = Sand or whatever
-permamble = pd.read_csv(
-    data_dir + 'param-permeable-soil.csv',
-    header=4,
-)
-"""
-Prepares ASU data for comparison
-"""
-db_asu = sqlite3.connect(db_dir + 'asu_house.db')
-asu = pd.read_sql_query(
-    "SELECT \
-        day, \
-        pressure_difference AS Pressure, \
-        tce_emission_rate, \
-        building_flow_rate, tce_groundwater \
-    FROM \
-        parameters; \
-    ", db_asu,
-).interpolate(limit_area='inside',limit=250).dropna()
-# removes data that is outside model data range
-asu['Pressure'] = -1.0*asu['Pressure']
-asu = asu[(asu['Pressure'] < 10) & (asu['Pressure'] > -10)]
-# removes cpm period from dataset
-cpm_start, cpm_end = 780.0, 1157.0
-asu = asu[(asu['day'] < cpm_start) | (asu['day'] > cpm_end)]
-# assigns a PP open/closed column (for sorting purposes)
[…]

### 2019-03-20 code/preferential-pathway-sensitivity/ss_asu_north_island.py

@@ -1,83 +1,0 @@
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-import sqlite3
-from scipy import stats
-from natu import units
-
-# path where the datafiles are stored
-file_path = '/Users/duckerfeller/Dropbox/Research/Data/'
-
-
-asu = pd.read_csv(
-    file_path+'data_asu2.csv', # full path
-    header=4, # dataframe headers are in row 4
-)
-
-nas = pd.read_csv(
-    file_path+'data_north_island.csv', # full path
-    header=4, # dataframe headers are in row 4
-)
-
-fig, (ax1, ax2) = plt.subplots(1,2,figsize=(7,4),sharey=True,sharex=True)
-
-asu = asu.pivot_table(
-    index=['p'], # x-axis
-    values='alpha', # y-axis
-    columns=['Ae'], # levels
-)
-
-titles = (
-    'No PP\nGravel sub-base\nSmaller foundation crack area',
-    'No PP\nSand soil',
-)
-
-fig, ax = plt.subplots(figsize=(5,3))
-
-# plots the pivoted data
-asu.plot(
-    ax=ax,
-    logy=True, # sets y-axis to log scale
-    color='k', # sets the line color to black
-    #title=titles[0], # assigns the custom titles
-    style=['-','--'],
-    legend=False, # removes the default legend (replaced by titeles)
-)
-
-ax.set_xlabel('Indoor/outdoor pressure\ndifference (Pa)')
-ax.set_ylabel('Attenuation factor')
-ax.set_title(titles[0])
-ax.legend(['Ae = 0.1 (1/h)', 'Ae = 0.5 (1/h)'],loc='best')
-plt.tight_layout()
-plt.savefig(
-    '/Users/duckerfeller/Documents/Research/Transient Variability/Figures/ss_asu.png',
-    dpi=300,
-)
-
-
-
-
-fig, ax = plt.subplots(figsize=(5,3))
-# plots the pivoted data
-nas.plot(
-    x='p',
-    y='alpha',
-    ax=ax,
-    logy=True, # sets y-axis to log scale
-    subplots=True, # indicates we want to plot each panda level as a subplot
-    color='k', # sets the line color to black
-    #title=titles[1], # assigns the custom titles
-    legend=False, # removes the default legend (replaced by titeles)
-)
-
[…]

### 2019-03-20 code/preferential-pathway-sensitivity/ss_pp_analysis.py

@@ -1,115 +1,0 @@
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-import sqlite3
-from scipy import stats
-from natu import units
-
-# path where the datafiles are stored
-file_path = '/Users/duckerfeller/Dropbox/Research/Data/'
-
-# names of the data files we wish to load
-file_names = (
-    'data_pp.csv',
-    'data_no_pp.csv',
-)
-
-
-data_list = [] # list to store dataframes
-for file_name in file_names: # loops through files
-    # loads dataframe based on file
-    data = pd.read_csv(
-        file_path+file_name, # full path
-        header=4, # dataframe headers are in row 4
-    )
-    # assigns if there is a PP in the dataset based on the naming of the file
-    data = data.assign(
-        PP = np.repeat(
-            'No' if 'no_pp' in file_name else 'Yes',
-            data.shape[0],
-        )
-    )
-    data_list.append(data) # appends dataframe to list
-
-# concatenates the dataframes
-df = pd.concat(
-    data_list, # dataframes
-    axis=0, # should be added row-wise (on top of each other)
-    ignore_index=True, # do not concat according to index
-)
-
-# removes the case where there is no PP and the PP isn't contaminated
-# (it is superfluous to have two no PP cases...)
-df = df[(df.chi != 0.0) | (df.PP != 'No')]
-
-# pivots the dataframe to 3 "levels" which are chi, PP and SB
-df = df.pivot_table(
-    index='p', # x-axis
-    values='alpha', # y-axis
-    columns=['chi', 'PP', 'SB'], # levels
-)
-
-# all of the titles to use
-titles = (
-    'Uncontamianted PP\nGravel sub-base',
-    'Uncontaminated PP\nSandy clay sub-base',
-    'No PP\nGravel sub-base',
-    'No PP\nSandy clay sub-base',
-    'Contaminated PP\nGravel sub-base',
-    'Contaminated PP\nSandy clay sub-base',
-)
-
-cols = [series for colname, series in df.iteritems()]
-
-
-
-
-for i, col in enumerate(cols):
-    fig, ax = plt.subplots(figsize=(5,3))
-    col.plot(
-        ax=ax,
-        logy=True,
-        legend=False,
-        color='k',
-    )
-    ax.set_ylim([1e-07, 5e-03])
-    ax.set_title(titles[i])
-
-    ax.set_ylabel('Attenuation factor')
[…]

### 2019-03-20 code/kde-iacc-pressure/ase-seasons.py
No longer needed

@@ -1,161 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy.integrate import LSODA
-from scipy.interpolate import interp1d
-import sqlite3
-import seaborn as sns
-from scipy import stats
-import matplotlib.patches as mpatches
-
-
-data_dir = './data/preferential-pathway-sensitivity/'
-#db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-db_dir = '/home/jonathan/Dropbox/var/'
-
-db = sqlite3.connect(db_dir + 'HillAFB.db')
-
-indoor = pd.read_sql_query( "SELECT * FROM DailyAverages;", db, )
-soil_gas = pd.read_sql_query( "SELECT * FROM AverageSubSurfaceSoilGasConcentration;", db, )
-groundwater = pd.read_sql_query( "SELECT StopTime, Concentration FROM AverageGroundWaterConcentration;", db, )
-
-indoor.rename(columns={'Concentration': 'IndoorConcentration',}, inplace=True)
-soil_gas.rename(columns={'Concentration': 'SubSurfaceConcentration',}, inplace=True)
-groundwater.rename(columns={'Concentration': 'GroundwaterConcentration',}, inplace=True)
-
-
-for df in (indoor,soil_gas,groundwater):
-    df['StopTime'] = df['StopTime'].apply(pd.to_datetime)
-
-df = indoor.set_index('StopTime').combine_first(soil_gas.set_index('StopTime')).reset_index().combine_first(groundwater.set_index('StopTime')).reset_index()
-
-df['Month'] = df['StopTime'].map(lambda x: x.month)
-def get_season(x):
-    seasons = {
-        'Winter': (12, 2),
-        'Spring': (3, 5),
-        'Summer': (6, 8),
-        'Fall': (9, 11),
-    }
-    if (x == 12) or (x == 1) or (x == 2):
-        return 'Winter'
-    elif (x == 3) or (x == 4) or (x == 5):
-        return 'Spring'
-    elif (x == 6) or (x == 7) or (x == 8):
-        return 'Summer'
-    elif (x == 9) or (x == 10) or (x == 11):
-        return 'Fall'
-    else:
-        return 'Error'
-df['Season'] = df['Month'].apply(lambda x: get_season(x))
-
-
-df[['IndoorConcentration','EmissionRate']] = df[['IndoorConcentration','EmissionRate']].apply(np.log10)
-df.interpolate(inplace=True)
[…]

### 2019-03-20 code/kde-iacc-pressure/indianapolis-seasons.py

@@ -1,178 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy.integrate import LSODA
-from scipy.interpolate import interp1d
-import sqlite3
-import seaborn as sns
-from scipy import stats
-import matplotlib.patches as mpatches
-
-
-data_dir = './data/preferential-pathway-sensitivity/'
-#db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-db_dir = '/home/jonathan/Dropbox/var/'
-
-db = sqlite3.connect(db_dir + 'Indianapolis.db')
-
-observation = pd.read_sql_query(
-    "SELECT StopDate AS StopTime FROM Observation_Status_Data WHERE Variable='Mitigation' AND Value='not yet installed';", db, )
-observation['SSDStatus'] = np.repeat('SSD not installed', len(observation))
-
-pressure = pd.read_sql_query(
-    "SELECT StopDate AS StopTime, AVG(Value) AS Pressure FROM Differential_Pressure_Data WHERE Variable='Basement.Vs.Exterior' GROUP BY StopDate;", db, )
-
-radon = pd.read_sql_query(
-    "SELECT Date(StopDate) AS StopTime, AVG(Value) AS RadonConcentration FROM Radon_Data_SIRAD WHERE Variable='Radon' GROUP BY StopDate;", db, )
-
-groundwater = pd.read_sql_query(
-    "SELECT Date(StopDate) AS StopTime, AVG(Value) AS GroundwaterConcentration, Variable AS Specie FROM VOC_Data_Groundwater GROUP BY DATE(StopDate);", db, )
-
-sub_surface = pd.read_sql_query(
-    "SELECT StopDate AS StopTime, AVG(Value) AS SubSurfaceConcentration FROM VOC_Data_SoilGas_and_Air WHERE Depth_ft='Basement' GROUP BY StopDate;", db, )
-
-weather = pd.read_sql_query(
-    "SELECT StopDate AS StopTime, AVG(Value) AS Temperature FROM Meteorological_Data WHERE Variable='Temp.Out' AND Units='F' GROUP BY StopDate;", db, )
-
-indoor = pd.read_sql_query(
-    "SELECT Date(StopDate) AS StopTime, AVG(Value) AS IndoorConcentration, Variable AS Specie FROM VOC_Data_SRI_8610_Onsite_GC GROUP BY DATE(StopDate);", db, )
-
[…]

### 2019-03-20 code/kde-iacc-pressure/indianapolis.py

@@ -1,203 +1,0 @@
-import numpy as np
-import sqlite3
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from scipy import stats
-import matplotlib.patches as mpatches
-
-
-
-class Observation:
-    def __init__(self,select,where,grouby):
-        return
-    def opt_select(self,value):
-        return
-class IndoorAir:
-    def __init__(self,method,compound,location,groupby):
-        return
-
-class Indianapolis:
-    def __init__(self):
-        # loads database
-        self.db = sqlite3.connect('/home/jonathan/lib/vapor-intrusion-dbs/indianapolis.db')
-        #self.db = sqlite3.connect('D:\Dropbox\Research\indianapolis.db')
-
-        return
-
-    def ssd(self,status='not yet installed'):
-        select = "O.StopDate AS StopDate"
-        table = "Observation_Status_Data AS O"
-        where = "O.Value = \"%s\"" % status
-        return select, table, where
-
-    def pressure(self,location='422',type='Basement.Vs.Exterior',avg=True):
-        if avg is True:
-            select = "AVG(P.Value) AS Pressure"
-        else:
-            select = "P.Value AS Pressure"
-        table = "Differential_Pressure_Data AS P"
-        where = "P.Location = \"%s\"" % location + " AND " + "P.Variable = \"%s\"" % type
-
-        return select, table, where
-
-    def concentration(self,location='422BaseS',compound='Chloroform',avg=True):
-        if avg is True:
-            select = "AVG(C.Value) AS Concentration"
-        else:
-            select = "C.Value AS Concentration"
-        table = "VOC_Data_SRI_8610_Onsite_GC AS C"
-        where = "C.Location = \"%s\"" % location + " AND " + "C.Variable = \"%s\"" % compound
-
-        return select, table, where
-
-
-
-    def sql_query(self,methods):
-        select = "SELECT "
-        tables = "FROM "
-        wheres = "WHERE "
-
-
-        for method in methods:
-            select += method[0] + ", "
-            tables += method[1] + ", "
-            wheres += method[2] + " AND "
-
-        select = select.rstrip(", ")
-        tables = tables.rstrip(", ")
[…]

### 2019-03-20 code/kde-iacc-pressure/nas-asu-house.py

@@ -1,129 +1,0 @@
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-import sqlite3
-from scipy import stats
-import matplotlib.patches as mpatches
-db_dir = '/home/jonathan/lib/vapor-intrusion-dbs/'
-
-
-# databases
-db_asu = sqlite3.connect(db_dir+'hill-afb.db')
-db_nas = sqlite3.connect(db_dir+'north_island_nas.db')
-
-# loads asu data, interpolates and drops NAs
-asu = pd.read_sql_query("SELECT * from daily_averages;", db_asu)
-phases = pd.read_sql_query("SELECT * from phases;", db_asu)
-
-asu.t = asu.t.apply(pd.to_datetime)
-phases.start = phases.start.apply(pd.to_datetime)
-phases.stop = phases.stop.apply(pd.to_datetime)
-
-phase1 = phases[(phases.cpm == 'off') & (phases.land_drain == 'open')]
-#phase2 = phases[(phases.cpm == 'off') & (phases.land_drain == 'closed')]
-phase3 = phases[(phases.cpm == 'off') & (phases.land_drain == 'closed')]
-
-
-# loads nas data, interpolates and drops NAs
-nas = pd.read_sql_query("SELECT tce_indoor_air AS c, pressure_difference AS p FROM north_island_nas;", db_nas).interpolate(limit_area='inside',limit=250).dropna()
-
-cpm_start, cpm_end = 780.0, 1157.0
-asu_pre_cpm = asu[asu['t'] < phase1['stop'].values[0]].copy()
-asu_post_cpm = asu[asu['t'] > phase3['start'].values[0]].copy()
-
-asu_pre_cpm.name = "ASU House, PP open"
-asu_post_cpm.name = "ASU House, PP closed"
-nas.name = "North Island"
-
-asu_pre_cpm['c'] = asu_pre_cpm['c'].replace(0.0, np.nan).dropna()
-
-dfs = []
-alpha = 0.6
-cmaps = ["Blues", "Reds", "Greens"]
-
-
-label_patches = []
-for dp in [False]:
-    df = pd.DataFrame({'x': [], 'y': []})
-    g = sns.JointGrid(x="x", y="y", data=df) # empty joint grid
-    for i, df in enumerate([nas, asu_pre_cpm, asu_post_cpm]):
-        dataset = df.name
-        print("Dataset: %s" % dataset)
-
[…]

### 2019-03-20 code/kde-iacc-pressure/preferential-pathways.py

@@ -1,141 +1,0 @@
-import numpy as np
-import sqlite3
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from scipy import stats
-import matplotlib.patches as mpatches
-
-
-db_asu = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/asu_house.db')
-db_ind = sqlite3.connect('/Users/duckerfeller/Dropbox/Research/indianapolis.db')
-
-asu = pd.read_sql_query("SELECT day, pressure_difference AS Pressure, tce_emission_rate, building_flow_rate FROM parameters;", db_asu).interpolate(limit_area='inside',limit=250).dropna()
-asu['Concentration'] = asu['tce_emission_rate']/asu['building_flow_rate']
-cpm_start, cpm_end = 780.0, 1157.0
-asu_pre_cpm = asu[(asu['day'] < cpm_start) & (asu['Pressure'] < 7.5)].copy().dropna()
-asu_pre_cpm['Pressure'] = -1.0*asu_pre_cpm['Pressure'].dropna()
-
-
-ind_chloro = pd.read_sql_query(
-    "SELECT \
-        O.StopDate AS StopDate, \
-        AVG(C.Value) AS Concentration, \
-        AVG(P.Value) AS Pressure \
-    FROM \
-        Observation_Status_Data AS O, \
-        VOC_Data_SRI_8610_Onsite_GC AS C, \
-        Differential_Pressure_Data AS P \
-    WHERE \
-        O.Value = 'not yet installed' AND \
-        C.StopDate = O.StopDate AND \
-        C.Location = '422BaseS' AND \
-        P.StopDate = O.StopDate AND \
-        C.Variable = 'Chloroform' AND \
-        P.Variable = 'Basement.Vs.Exterior' AND \
-        P.Location = '422' \
-    GROUP BY \
-        O.StopDate \
-    ;", db_ind)
-
-ind_pce = pd.read_sql_query(
-    "SELECT \
-        O.StopDate AS StopDate, \
-        AVG(C.Value) AS Concentration, \
-        AVG(P.Value) AS Pressure \
-    FROM \
-        Observation_Status_Data AS O, \
-        VOC_Data_SRI_8610_Onsite_GC AS C, \
-        Differential_Pressure_Data AS P \
-    WHERE \
-        O.Value = 'not yet installed' AND \
[…]

### 2019-03-20 code/air-exchange-rate/asu_exchange_rate_analysis.py
No longer needed

@@ -1,191 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import sqlite3
-import seaborn as sns
-
-# returns the error between the predicted and measured TCE in indoor air
-def get_error(data):
-    error = np.array([])
-    n = len(data['day'].values)
-    for i in range(n):
-        try:
-            error = np.append(error,(data['estimate'].values[i]-data['ug/m3'].values[i]))
-        except:
-            error = np.append(error,np.nan)
-    data = data.assign(error=error)
-    return data
-
-# generates a data frame that contains the mean and median absolute error
-# between the predicted and measured TCE indoor air concentrations
-def get_error_table(data):
-    max_error = data['error'].abs().max()
-    min_error = data['error'].abs().min()
-    var = data['error'].var()
-    # mean absolute error
-    mae = data['error'].abs().mean()
-    # medan absolute error
-    mdae = data['error'].abs().median()
-    # creates a new dataframe containing the error estimates
-    error_table = pd.DataFrame(data={'Mean absolute difference': [mae],
-                                'Median absolute difference': [mdae],
-                                'Max. absolute difference': [max_error],
-                                'Min. absolute difference': [min_error],
-                                'Variance': [var]})
-
-    return error_table
-
-# estimates the TCE in indoor air by assuming that entry rate is the same as
-# measured/calculated but keeping air exchange rate*volume to its median value
-def predict_tce_indoor_air(data):
-    cpm_start = 780.0
-    Q_median_no_cpm = data['building_flow_rate'].loc[data['day'] < cpm_start].median()
-    # estimates the TCE in indoor iar
-    #tce_indoor_air_prediction = data['tce_emission_rate']/data['building_flow_rate'].median()
-    tce_indoor_air_prediction = data['tce_emission_rate']/Q_median_no_cpm
-    # assigns predicted value to data dataframe
-    data = data.assign(tce_indoor_air_prediction=tce_indoor_air_prediction)
-
-    return data
-
[…]

### 2019-03-20 code/diurnal.py
Diurnal pressure and air exchange rate data

@@ -1,0 +1,25 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import matplotlib.gridspec as gridspec
+from scipy import stats
+
+
+class Diurnal:
+    def __init__(self):
+        data = pd.read_csv('./data/asu_house.csv')
+        data = data.loc[data['Phase']!='CPM']
+        data['Time'] = data['Time'].apply(pd.to_datetime)
+
+
+        pressure = data[['Time','IndoorOutdoorPressure']].groupby(data['Time'].dt.hour).median()
+        air_exchange_rate = data[['Time','AirExchangeRate']].groupby(data['Time'].dt.hour).median()
+
+        pressure.to_csv('./data/diurnal/pressure.csv')
+        air_exchange_rate.to_csv('./data/diurnal/air_exchange_rate.csv')
+
+        return
+
+
+Diurnal()

### 2019-03-21 code/figures.py
Updated figures

@@ -5,7 +5,7 @@ import seaborn as sns
 import matplotlib.gridspec as gridspec
 from scipy import stats
 from get_simulation_data import Soil, PreferentialPathway
-
+import datetime
 
 """
 This figure shows a case (North Island NAS) where indoor/outdoor pressure
@@ -252,19 +252,19 @@ class IndianapolisTime:
         fig, ax = plt.subplots(dpi=300)
 
         sns.lineplot(
-            data=data,
+            data=data[data['Specie']=='Trichloroethene'],
             x='Time',
             y='IndoorConcentration',
-            hue='Specie',
             ax=ax,
         )
 
         ax.set_yscale('log')
         ax.set(
-            title='Indoor air contaminant concentration at the Indianapolis site',
+            title='Indoor TCE concentration at the Indianapolis site',
             ylabel='$c_\\mathrm{in} \; \\mathrm{(\\mu g/m^3)}$',
+            xlim=([datetime.date(2011, 8, 11), datetime.date(2011, 10, 15)]),
         )
-
+        plt.xticks(rotation=45)
         plt.tight_layout()
         plt.savefig('./figures/temporal_variability/time_indianapolis.png')
         plt.savefig('./figures/temporal_variability/time_indianapolis.pdf')
@@ -276,23 +276,21 @@ class AirExchangeRateKDE:
     def __init__(self):
 
         data = pd.read_csv('./data/asu_house.csv').dropna()
-        data = data.loc[data['Phase']!='CPM']
+        data = data.loc[data['Phase']!='Closed']
 
 
         fig, ax = plt.subplots(dpi=300)
-        sns.kdeplot(
-            data=data['IndoorOutdoorPressure'],
-            data2=data['AirExchangeRate'],
-            shade=True,
-            shade_lowest=False,
-            gridsize=200,
+        sns.regplot(
+            data=data,
+            x='IndoorOutdoorPressure',
+            y='AirExchangeRate'
         )
 
         ax.set(
-            title='2D KDE showing distributions and relationship between\nindoor/outdoor pressure and air exchange rate ',
+            title='2D KDE showing distributions and relationship between\nindoor/outdoor pressure and air exchange rate\nafter the land drain was closed',
             ylabel='$A_e \; \\mathrm{(1/hr)}$',
[…]

### 2019-03-21 tex/main.tex
So many things

@@
     EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995} & 0.16-0.2 & 0.35-0.49 & 1.21-1.49 \\
-    ASU house\cite{holton_temporal_2013,guo_identification_2015} & 0.25 & 0.5 & 1.15 \\
+    ASU house\cite{holton_temporal_2013,guo_identification_2015} & 0.21 & 0.43 & 0.78 \\
     Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34 & 0.74 & 1.27 \\

### 2019-03-21 code/figures.py

@@ -130,17 +130,7 @@ class Modeling:
         sim = PreferentialPathway().data
         asu = pd.read_csv('./data/asu_house.csv')
         #print(asu['AirExchangeRate'].describe(percentiles=[0.1, 0.9]))
-        asu = asu.loc[
-            (asu['Phase']!='CPM') &
-            (
-                (asu['IndoorOutdoorPressure'] >=-5) &
-                (asu['IndoorOutdoorPressure'] <= 5)
-            ) &
-            (
-                (asu['AirExchangeRate'] <= 1.5) &
-                (asu['AirExchangeRate'] >= 0.1)
-            )
-        ]
+        asu = asu.loc[ (asu['Phase']!='CPM') ]
 
 
         # isolates the Ae = 0.5 case for the uniform soil modeling
@@ -156,19 +146,18 @@ class Modeling:
         sim = sim[sim_data_to_remove[0]]
 
 
+        # options
+        min_Ae, max_Ae = 0.3, 0.9 # the min and max air exchange used for the "fill"
+
         fig, (ax1,ax2) = plt.subplots(2,1, sharex=True, sharey=True, figsize=[6.4, 6.4], dpi=300)
 
-        pp_max = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==0.1)]
+        pp_max = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==min_Ae)]
         pp = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==0.5)]
-        pp_min = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==1.5)]
+        pp_min = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==max_Ae)]
 
 
         ax1.plot(pp['IndoorOutdoorPressure'], pp['logAttenuationGroundwater'],label='PP')
         ax1.fill_between(pp['IndoorOutdoorPressure'], pp_min['logAttenuationGroundwater'], pp_max['logAttenuationGroundwater'],alpha=0.5)
-
-
-
-
 
         # ax1
         sns.regplot(
@@ -198,9 +187,9 @@ class Modeling:
         )
 
 
-        no_pp_max = sim.loc[(sim['Simulation']=='No Pp')&(sim['AirExchangeRate']==0.1)]
+        no_pp_max = sim.loc[(sim['Simulation']=='No Pp')&(sim['AirExchangeRate']==min_Ae)]
[…]

### 2019-03-26 code/figures.py
New "error" spans with interpolated Ae values

@@ -156,25 +156,36 @@ class Modeling:
         sim = sim[sim_data_to_remove[0]]
 
         p_vals = sim['IndoorOutdoorPressure'].unique()
-        ae = asu['AirExchangeRate'].describe(percentiles=[0.1,0.9])
+        Ae = asu['AirExchangeRate'].describe(percentiles=[0.05,0.1,0.9,0.95])
 
 
-        new_sim_vals = pd.DataFrame({
-            'IndoorOutdoorPressure': np.append(p_vals, p_vals),
-            'AirExchangeRate': np.append(),
-        })
-        #Ae_vals =
-        print(new_sim_vals)
-        print(analysis[['10%','90%']].values)
+        Ae_low = Ae['5%']
+        Ae_high = Ae['95%']
 
-        # options
-        min_Ae, max_Ae = 0.3, 0.9 # the min and max air exchange used for the "fill"
+        print(Ae_low, Ae_high)
+
+        from scipy.interpolate import interp2d
+
+        for simulation in ['Pp', 'No Pp']:
+            interp_func = interp2d(
+                sim.loc[sim['Simulation']==simulation]['IndoorOutdoorPressure'],
+                sim.loc[sim['Simulation']==simulation]['AirExchangeRate'],
+                sim.loc[sim['Simulation']==simulation]['logAttenuationGroundwater'],
+            )
+            for Ae_now in [Ae_low, Ae_high]:
+                new_sim_vals = pd.DataFrame({
+                    'IndoorOutdoorPressure': p_vals,
+                    'AirExchangeRate': np.repeat(Ae_now, len(p_vals)),
+                    'Simulation': np.repeat(simulation, len(p_vals)),
+                    'logAttenuationGroundwater': interp_func(p_vals, Ae_now),
+                })
+                sim = sim.append(new_sim_vals, ignore_index=True, sort=False)
 
         fig, (ax1,ax2) = plt.subplots(2,1, sharex=True, sharey=True, figsize=[6.4, 6.4], dpi=300)
 
-        pp_max = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==min_Ae)]
+        pp_max = sim.loc[(sim['Simulation']=='Pp')&(sim['AirExchangeRate']==Ae_low)]
[…]

### 2019-03-27 code/resampling.py
Using Yijun's fitting idea

@@ -55,6 +55,29 @@ resampled.dropna(inplace=True)
 sns.set_palette(sns.color_palette("muted"))
 
 
+from scipy.optimize import curve_fit
+
+def func(t, a, b):
+    return a*np.log(b*t)
+
+fit_data = resampled.loc[resampled['Dataset']!='ASU House, PP Open']
+
+def convert_to_days(df):
+
+    days = []
+    for resamp in df['Resampling']:
+
+        days.append(pd.Timedelta(int(resamp[0]),resamp[1]).days)
+
+    df['Days'] = pd.Series(days, index=df.index)
+    df.sort_values(by='Days',inplace=True)
+    return df
+
+fit_data = convert_to_days(fit_data)
+
+popt, pcov = curve_fit(func, fit_data['Days'], fit_data['MaxDelta'])
+
+
 g = sns.catplot(
     x="Resampling",
     y="MaxDelta",
@@ -63,6 +86,14 @@ g = sns.catplot(
     data=resampled,
     legend_out=False,
     aspect=1.5,
+)
+
+ax = g.axes[0][0]
+
+ax.plot(
+    fit_data['Resampling'],
+    func(fit_data['Days'], popt[0], popt[1]),
+    'k-'
 )
 
 my_ytick_labels = ["%1.1f" % y_tick for y_tick in 10.0**g.ax.get_yticks()]

### 2019-03-27 tex/main.tex
What
Added Eric's changes

@@
 % macros
+
+% TODO: Fix all the citation issues
 
@@
 
+\begin{figure}
+ \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
+ \includegraphics[width=\textwidth]{time_indianapolis.pdf}
+\end{figure}
+
 \section{Methods}
@@
 It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
+Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
+There is almost a two order of magnitude variation in the concentration data.
 
@@
 In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The numpy, scipy, and pandas python packages were also used to conduct all statistical analysis and processing, with plotting done by the matplotlib, seaborn packages.
-% TODO: Add references to the packages
+The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.
 
@@
 \begin{figure}[htb!]
-  \caption{Graphic of the preferential pathway mode.
-Note that the gravel sub-base material may be switched to the same as the surrounding soil, and that the preferential pathway may be turned off.}\label{fig:model}
-  %\includegraphics{model.png}
+ \caption{Foundation and vadose zone soil represented in the modeling.
+Note that here a gravel sub-base material is shown, but in certain simulations, that material is absent and the surrounding soil directly contacts the foundation slab.
+Different assumptions are made regarding the preferential pathway, here shown as a pipe entering the gravel sub-base.
+In some cases, the preferential pathway has been "turned off".}\label{fig:model}
+ \includegraphics[width=\textwidth]{model.png}
 \end{figure}
 
-In addition to examining the actual field data, a previously described three -dimensional computational fluid dynamics model of a generic VI impacted house was used to elucidate certain aspects of  the processes.
-This model was implemented in a finite element solver package, COMSOL Multiphysics.
-In the present work, there has been an addition of a preferential pathway to the ”standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
+A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
+In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
 As in the earlier studies, only the vadose zone soil domain is directly modeled.
+Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
 
-The modeled structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
+The modeled VI impacted structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
 The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the  perimeter of the house.
 All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is assumed to consist of only of the basement, assumed as having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of certain other features can be.
-\par
+The indoor control volume is here assumed to consist of only of the basement, having a total volume of $300 \; \mathrm{m^3}$.
+Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of some critical parameters is.\par
 
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house, and is assumed to consist of sandy loam (except as noted).
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases where this sub-base material is assumed to be the same as the surrounding soil (termed  a ”uniform” soil scenario).
-\par
+The modeled surrounding soil domain extends 5 meters from the perimeter of the house and is assumed to consist of sandy loam, except as noted otherwise.
+Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases here this sub-base material is assumed to be the same as the surrounding soil (termed a "uniform" soil scenario).
+The groundwater beneath the structure is assumed to be homogeneously contaminated with TCE selected as a prototypical contaminant.
+The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
+Where relevant, the preferential pathway is modeled as a 10 cm (4”) pipe that opens into the gravel sub-base beneath the structure.
+The air in the pipe is also assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$ (allowing the contaminant concentration in the pipe to be parameterized).
+This model illustrates the concept of a "preferential pathway", as the pipe carries contaminant vapor to the immediate vicinity of the foundation, by a path that circumvents the usual soil diffusion pathway.\par
 
-Where relevant The preferential pathway is modeled as a 10 cm (4”) pipe that exits into the gravel sub-base beneath the structure.
-The air in the pipe is assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$, allowing the contaminant concentration in the pipe to be parameterized.
-\par
-
-The groundwater beneath the structure is assumed to be homogeneously contaminated with trichloroethylene (TCE) as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-The ground surface and the pipe are both assumed to be sources of air to the soil domain.
-Both are assumed to be at reference atmospheric pressure. \par
-
-Vapor transport in the soil is governed by Richard’s equation, a modified version  of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
+The ground surface and the pipe are both sources of air to the soil domain.
+Both are assumed to exist be at reference atmospheric pressure.
+Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
 The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
 The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of vapor contaminant in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The equations and the boundary conditions are given in Table \ref{tbl:eqns-bc-parameters}.
+The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
+The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par
 
@@
  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}
-  \label{tbl:eqns-bc-parameters}
+ \label{tbl:eqns_bc_parameters}
  \bigskip
@@
 
+\subsection{Variation In Indoor Air Contaminant Concentration Over Time}
+
+\begin{figure}[htb!]
+ \centering
+ \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
+  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
+ \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
+\end{figure}
+
+High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$ , such as those in Figure 1, took place at both the ASU House and the Indianapolis House over significant periods\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
+Furthermore, at the Indianapolis site the $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
+The NAS North Island NAS dataset was obtained over a much shorter duration, and is therefore not examined in this portion of the analysis.
+
+Figure \ref{fig:indianapolis} showed a large degree of temporal variation in one of the components, and the data for the other components were quite similar.
+What is apparent upon closer examination of such data is that the actual day-to-day variations are typically not nearly as large as those observed when tracking the data for a longer time.
+To demonstrate this point, the quotient of the maximum and minimum $c_\mathrm{in}$ values (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are shown as a function of time in Figure \ref{fig:resampling}.
+The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
+Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.
+
+For example, if the data are examined in terms of the mean maximum variation observable over the course of 24 hours (one day) the variation is no greater than about a factor of two for any of the contaminants at the Indianapolis house or for TCE at the ASU house (when the preferential pathway was closed).
+The mean variability at the latter was only a bit higher (about a factor of 3) when the preferential pathway was open.
+In other words, a sampling protocol that involves sampling on two consecutive days would typically not uncover the large temporal variations that characterize the site over longer periods of time.
+As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
+So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.
+
+Weeks of temporal separation in sampling events are required to observe the large variations of concern.
+Orders of magnitude differences begin to manifest themselves over the course of months.
+This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
+This would be consistent with requiring months to see the more significant variations.
+
+This analysis also suggests that certain types of preferential pathways contribute to larger variations on shorter timescales (ASU House).
+Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
+This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.
+
+The longer the resampling period, the larger the maximum variability in observed indoor air contaminant concentrations.
+In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
+Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.
+
+If the maximum observed variability is quite small across only a few days as these data suggest, a sampling strategy that involves taking daily samples for several days in a row will likely yield little extra useful information beyond what can be obtained with a single sample.
+
 \subsection{Statistical Analysis of Field Data}
@@
 
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) is an important driving force in VI, drawing in (or preventing) contaminants from entering a structure.
-% TODO: Add references
-Changes in $p_\mathrm{in/out}$ is also a dynamic and fast process, impacting VI more rapidly than e.g.
-fluctuations in groundwater depth or contaminant concentration does; these processes may take weeks or even months to impact the overlying structure.
-Therefore, it is reasonable to assume that the temporal variability in indoor air contaminant concentration $c_\mathrm{in}$ may be driven by changes in $p_\mathrm{in/out}$.
+The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
+The rate of advective entry of soil gas into a structure is frequently cited as playing an important role in determining entry rate of contaminant.
+This advective entry rate is closely linked to the indoor-outdoor pressure difference, as can be caused by the “stack effect”, for example.
+Thus we first consider how much variability there might be in the pressure driving force for advection, and if this can explain the observed variability in observed indoor air contaminant concentrations.\par
 
+The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure.  Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).
 We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plots allow us to view the distribution of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and how well these correlate.
-For this analysis we consider two VI sites, North Island NAS, and the ASU house, with the ASU house dataset divided up into two periods, before and after the land drain (called preferential pathway (PP) from here on) had been closed.
-The preferential pathway significantly impacted the ASU house to such an extent that these two periods can essentially be considered as two different VI sites, and allowing us to examine the impact of the preferential pathway.
+The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
+For this analysis we considered two VI sites, NAS North Island and the ASU House.
+The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  ere on) had been closed.
+By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
 
-In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing us to compare the impact $p_\mathrm{in/out}$ had on $c_\mathrm{in}$.
-A value of 10 on the y-axis indicate that $c_\mathrm{in}$ is 10 times greater than the mean here, and 0.1 indicate that it is one tenth of the mean.
-The Pearson's r between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend.
+In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
+A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.
+Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
+Such issues have of course already been pointed out by the investigators who obtained the data.
 
-An interesting aspect of the North Island NAS site is that $p_\mathrm{in/out}$ varies so significantly, and the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (PP open), and -2.1 and 2.27 Pa (PP closed).
+The power of this KDE representation is that it permits evaluation of the relationship of two independently measured data - the indoor air contaminant concentration and the indoor-outdoor pressure difference.
+Examining the data in this manner immediately points to an important difference between the data from the ASU House and those from NAS North Island.
+At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
+This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).
 
-The large (roughly one order of magnitude) under- and overpressurization of the North Island NAS site leads to roughly a one order magnitude increase and decrease from the mean $c_\mathrm{in}$.
-Combined with $r=-0.73$, it is quite clear that at this site $p_\mathrm{in/out}$ largely determines $c_\mathrm{in}$ (there is still variability in $c_\mathrm{in}$ for a given $p_\mathrm{in/out}$, which we address later).
-This is the same principle that governs the controlled pressure method (CPM) concept.
+The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
+The Pearson’s r value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
+But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.
 
-Turning to the ASU house datasets, we see a quite different situation.
-The variability of $c_\mathrm{in}$ is just as large, or even larger than at North Island NAS, yet the $p_\mathrm{in/out}$ varies far less.
-At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed respectively are relatively comparable, but the 5th and 95th percentiles values of differ significantly $c_\mathrm{in}/c_\mathrm{in,mean}$ as may be seen in Table \ref{tbl:percentiles}.
+The ASU house datasets offer a different picture.
+The variability   of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
+The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r values for the correlations between the variables.
+In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
 
-It is clear that the preferential pathway dramatically increases the ASU house's sensitivity to $p_\mathrm{in/out}$, and the site is fundamentally different during the period when the preferential pathway was open.
-Yet, the magnitude of change in $c_\mathrm{in}$ is far greater than the change in $p_\mathrm{in/out}$, leading to the conclusion that it is not just increased flow rates into the structure that causes this, but there must also be an increased amount of contaminant as well as a much larger spatial variability in the sub-base.
-This is a topic that we will return to later.
+It should be noted that control of indoor depressurization is the principle that underlies the controlled pressure method (CPM) for assessing the potential severity of a VI situation.
+These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.
 
-The magnitude of change in $c_\mathrm{in}$ when the PP is closed is more in line with the magnitude of change in $p_\mathrm{in/out}$, but there is still more variability than one would expect.
-We hypothesize that this is largely due to fluctuations in air exchange rate, which we will examine later.
-
-\subsection{Change In Indoor Air Contaminant Concentration Over Time} % TODO: Improve title
-
-\begin{figure}[htb!]
-  \centering
-  \caption{Maximum change in indoor air contaminant concentration that may be expected over a given time period. (1D is 1 day, 2W is 2 weeks, and 3M is 3 months).}
-  \label{fig:resampling}
-  \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
-\end{figure}
-
-% RESAMPLING TIME ANALYSIS
-So far we have discussed how the relationship between $c_\mathrm{in}$ and $p_\mathrm{in/out}$, and how significantly $c_\mathrm{in}$ may vary at a site.
-But this tells us little about how much variability in $c_\mathrm{in}$ may be expected over time.
-For this analysis we turn to the ASU house data (PP open/closed considered separately again), and another well-studied VI site, the Indianapolis site in Indiana.
-Both of these sites collected high frequency $c_\mathrm{in}$ samples over a significant periods and are therefore suitable for this analysis\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at Indianapolis site the $c_\mathrm{in}$ of three different contaminants, Chloroform, Trichloroethylene (TCE), and Tetrachloroethylene (PCE) were collected, allowing us to see if there is any difference in the variability of each.
-The North Island NAS dataset only spans a few days, and is therefore excluded from this analysis.
-
-To demonstrate how significantly $c_\mathrm{in}$ can vary across time, we calculate the quotient of the maximum and minimum $c_\mathrm{in}$ (denoted as $c_\mathrm{max}/c_\mathrm{min}$) over a given time period.
-I.e.
-if $c_\mathrm{in}$ samples were collected every four hours over a period of a year, and the data is resampled on a daily basis, then $c_\mathrm{max}/c_\mathrm{min}$ is returned for within each day, giving 365 data points.
-Resampling periods of one, two, three, days, weeks, and months are chosen and Figure \ref{fig:resampling} shows the result of this analysis.
-
-As one might expect, the longer the resampling period, the larger the maximum variability is: spanning from less than a threefold difference, to two to three orders of magnitude.
-That such a large variability is observed when the resampling time approaches the length of the entire datasets is hardly surprising, nor is it surprising that the variability is much more significant when the preferential pathway was open than closed.
-What may be surprising is that absent a preferential pathway such as the one at the ASU house, it may take a few weeks before an order of magnitude maximum variability is reached.
-The most significant result of this analysis is that the maximum variability is quite small across a few days, suggesting that e.g.
-24-hour passive samplers will resolve the daily temporal variability in $c_\mathrm{in}$ well ($c_\mathrm{max}/c_\mathrm{min} \approx 1.5$) and that a sampling frequency greater than a few days may yield little extra information.
-There is also little difference between maximum variability of $c_\mathrm{in}$ the ASU house (when the PP is closed), and the three contaminants at the Indianapolis site, suggesting that this may be an "usual" amount of variability over these time periods.
+The data for the ASU House also offer an insight into the role of the preferential pathway.
+At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed are relatively comparable.
+However, the 5th and 95th percentiles values of $c_\mathrm{in}$/$c_\mathrm{in,mean}$ differ significantly as may be seen in Table \ref{tbl:percentiles}.
+It is clear that existence of the preferential pathway dramatically increases the variability in indoor air contaminant concentration.
+This again is entirely consistent with what the investigators of that site have already reported\cite{guo_identification_2015}.
+The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
+These will be explored with the help of a modeling analysis presented below.
 
@@
 \begin{figure}[htb!]
-  \caption{Boxplot of attenuation from subslab at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution, and diamonds are points that are considered outlier.}\label{fig:attenuation_subslab}
+ \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
+The box shows the quartiles of the distribution, the whiskers the extent of the distribution, and diamonds are points that are considered outliers.}\label{fig:attenuation_subslab}
  \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
@@
 
-To examine how $p_\mathrm{in/out}$ affects the contaminant entry rate (and subsequently the indoor air contaminant concentration) it would be convenient to analyze how $p_\mathrm{in/out}$ is correlated with the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) at a site - in particularly for understanding how the PP at the ASU house affected the site dynamics.
-By normalizing the $c_\mathrm{in}$ to $c_\mathrm{subslab}$ the $p_\mathrm{in/out}$ influence on the contaminant availability in the subslab area is mitigated.
+Observed temporal variations in indoor air contaminant concentrations might be explained by temporal variations in subslab contaminant concentrations.
+To examine how variability in subslab contaminant concentration might contribute to variability in indoor air contaminant concentration, data on the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) were examined.
+The dataset utilized for this was that from the ASU House.
+The $c_\mathrm{subslab}$ values were taken from a soil gas probe labeled as "6" at the ASU house.
+This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
+The results are shown in Figure \ref{fig:attenuation_subslab}.
 
-The $c_\mathrm{subslab}$ are taken from the soilgas probe labeled as "6" at the ASU house, which is the probe located closest to the preferential pathway, and to a reported breach in the foundation\cite{guo_identification_2015}.
-However, as Figure \ref{fig:attenuation_subslab} shows, there are some issues with using $\alpha_\mathrm{subslab}$ for subsequent analysis.
+It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
+Thus during the period when the preferential pathway was closed large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.
 
-First, we can state that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ does not vary that significantly, and is mostly around the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-During the period when the preferential pathway was open, there is considerably more variability, and it is not uncommon for the $\alpha_\mathrm{subslab}$ to exceed unity.
+When the PP was open, there was considerably more variability in the subslab concentration values, and the mean value was higher than in the case where the preferential pathway was closed.
+It was also not uncommon for the observed $\alpha_\mathrm{subslab}$ to exceed unity.
+While large $\alpha_\mathrm{subslab}$ values may sometimes indicate indoor sources at a site, there were none at the ASU house.
+A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
+This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.
 
-Even though probe "6" is located in such close proximity to the preferential pathway, and the breach in the foundation, the contaminant concentration 'hotspot' is still missed.
-This suggests that a preferential pathway may have an extremely localized influence in the subslab region, causing significant spatial variability in $c_\mathrm{subslab}$.
-This shows how difficult it may be to locate a preferential pathway through subslab soilgas sampling.
+What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in “normal” VI scenarios.
 
-The large $\alpha_\mathrm{subslab}$ values (compared to the recommended EPA value of 0.03) can often indicate that there may be indoor sources present a site, which there were none of at the ASU house.
-Thus, in addition to potentially indicating indoor contaminant sources, large $\alpha_\mathrm{subslab}$ may also indicate the presence of a preferential pathway.
+\subsection{Modeling Results}
 
-\subsubsection{Modeling Results} % TODO: Improve title
+\subsubsection{Pressure Effects}
 
@@
 No preferential pathway scenarios in bottom panel.
-Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars one standard deviation for that binned mean.}
+Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars represent the 95\% confidence interval.}
  \label{fig:land_drain_scenarios}
@@
 
+Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
+The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
+This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
+The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
+The general modeling conditions are those shown in Table \ref{fig:indianapolis}.
+
+In the calculation whose results are shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
+Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
+Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
+The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater  cgw, giving the attenuation from groundwater $\alpha_\mathrm{gw}$ .
+The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
+These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
+
+The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
+Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.
+
+The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
+First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
+Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.
+
+The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
+This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
+This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway way actually not impact a VI site very much.
+In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
+These requirements were met at the ASU House.
+
+A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
+In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
+The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
+This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
+The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathay.
+
+The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
+These results are shown as the bottom panel in Figure \ref{fig:land_drain_scenarios}.
+These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed.
+
+What the above simulations fail to predict is the actual observed variability of $\alpha_\mathrm{gw}$ at any given $p_\mathrm{in/out}$.
+This variability is captured by the error bars around the data points in the two panels (denoting one standard deviation).
+This variability is roughly an order of magnitude.
+Examining Figure \ref{fig:model} for the case in which the preferential pathway is closed, the “cloud” of data similarly reflect that there is still not captured in the modeling a parameter that determines this variation, independently of the pressure effects.
+
+One other feature of this simulation that warrants noting is that the examined range from $p_\mathrm{in/out}$ from -5 to +5 Pa is larger than the “usual” range from -2 to +2 Pa observed at the ASU House (as may be seen from Figure \ref{fig:model}).
+The data at larger and smaller values of $p_\mathrm{in/out}$ were rather sparse, though there were enough points so that the data for the higher range of pressures could be presented in Figure \ref{fig:land_drain_scenarios}.
+But focusing on the range from -2 to + 2 Pa alone, the pressure dependence of variability is much less visible, just as the data of Figure \ref{fig:model} suggested.
+This again means that there is a need to explore other reasons for the order of magnitude spread seen in concentration data, when the preferential pathway is closed.
+
+\subsubsection{Air Exchange Rate Effects}
+
+Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
+Examination of these distributions point in a clear direction for modifying the model of VI.
+Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
+A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
+Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
+Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
+As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
+
+To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
+These bounds are given by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 hr-1.
+
+It is apparent that assuming variability in air exchange rate allows capturing most of the observed variability in $\alpha_\mathrm{gw}$.
+We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
+Thus we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.
+
+\subsubsection{Results of Transient Simulations}
+
+\begin{figure}
+ \caption{}\label{fig:diurnal}
+ \includegraphics[width=\textwidth]{diurnal.pdf}
+\end{figure}
+
+To see how much variability can be explained due to fluctuations in air exchange rate
+
+The above analyses have been conducted under simulated steady state conditions.
+The final test of the conclusions reached above comes in applying them in an actual transient simulation of a scenario such as that in Figure 2.
+These results are shown in Figure 7.
+
 \begin{acknowledgement}

### 2019-03-27 code/figures.py
Removed todos

@@ -53,7 +53,7 @@ class PressureKDE:
                 data2=data2,
                 shade_lowest=False,
                 shade=True,
-                label=label + ', r = %1.2f' % r, # TODO: Add Pearson's r analysis
+                label=label + ', r = %1.2f' % r,
                 ax=ax,
             )
 
@@ -80,7 +80,7 @@ class PressureKDE:
         return
 
 
-def get_log_ticks(start, stop, style='e'): # TODO: Remove the unnecessary ticklabels
+def get_log_ticks(start, stop, style='e'):
 
     ticks = np.array([])
     ints = np.arange(np.floor(start),np.ceil(stop)+1)
@@ -378,7 +378,7 @@ class Diurnal:
             maxmin(pp_open),
         )
         fig, ((ax1, ax2),(ax3,ax4)) = plt.subplots(2,2, dpi=150)
-        # TODO: Create interpolate function for PP closed cases to make nicer plots
+        
         x_smooth = np.linspace(0,23,100)
         y_smooth = CubicSpline(p_diurnal['Time'], p_diurnal['IndoorOutdoorPressure'])(x_smooth)
         ax1.plot(x_smooth, y_smooth)

### 2019-03-27 code/simulation_predictions.py

@@ -44,7 +44,7 @@ def get_interp_func(df):
     Ae = df['AirExchangeRate']
     alpha = df['logAttenuationGroundwater']
 
-    # TODO: increase pressurization limits in simulation
+    
     # 2d interpolation function
     interp_func = interp2d(
         p_in,

### 2019-03-27 code/resampling.py
Removed curve fit from plot

@@ -88,13 +88,6 @@ g = sns.catplot(
     aspect=1.5,
 )
 
-ax = g.axes[0][0]
-
-ax.plot(
-    fit_data['Resampling'],
-    func(fit_data['Days'], popt[0], popt[1]),
-    'k-'
-)
 
 my_ytick_labels = ["%1.1f" % y_tick for y_tick in 10.0**g.ax.get_yticks()]
 g.ax.set_yticklabels(my_ytick_labels)

### 2019-03-27 code/figures.py
Changed title

@@ -342,7 +342,7 @@ class AirExchangeRateKDE:
             )
 
         ax.set(
-            title='2D KDE showing distributions and relationship between\nindoor/outdoor pressure and air exchange rate\nafter the land drain was closed',
+            title='2D KDE showing distributions and relationship between\nindoor/outdoor pressure and air exchange rate',
             ylabel='$A_e \; \\mathrm{(1/hr)}$',
             xlabel='$p_\\mathrm{in/out} \; \\mathrm{(Pa)}$',
             ylim=[0,1.75],
@@ -378,7 +378,7 @@ class Diurnal:
             maxmin(pp_open),
         )
         fig, ((ax1, ax2),(ax3,ax4)) = plt.subplots(2,2, dpi=150)
-        
+
         x_smooth = np.linspace(0,23,100)
         y_smooth = CubicSpline(p_diurnal['Time'], p_diurnal['IndoorOutdoorPressure'])(x_smooth)
         ax1.plot(x_smooth, y_smooth)
@@ -455,9 +455,9 @@ class Diurnal:
 #Diurnal()
 #PressureKDE(y_data_log=True,norm_conc=True)
 #AttenuationSubslab()
-Modeling()
+#Modeling()
 #IndianapolisTime()
-#AirExchangeRateKDE()
+AirExchangeRateKDE()
 
 """
 df = pd.DataFrame({'x': [0,2], 'y': [1,4]})

### 2019-03-27 tex/supporting-information/esthag_supporting.tex
Created supporting information

@@
 \usepackage{array,multirow,graphicx}
-\graphicspath{{../../Figures/}}
+\graphicspath{{../../figures/2d_kde/}}
 % macros
@@
 \author{Jonathan G. V. Ström}
-\affiliation[Brown University]{Brown University, School of Engineering}
+\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
 \author{Yijun Yao}
-\affiliation[Zhejiang University]{Zhejiang University}
+\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
+\author{Yuanming Guo}
+\affiliation[Arizona State University]{Arizona State University}
 \author{Eric M. Suuberg}
 \email{eric_suuberg@brown.edu}
-\affiliation[Brown University]{Brown University, School of Engineering}
+\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
 
 % title
-\title{Modeling the Impact of Preferential Contaminant Pathways on Vapor Intrusion}
+\title{Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
 
@@
 
-\subsection{Supporting Information Content:}
-
-\begin{itemize}
-  \item x pages: S1 - S6
-  \item 4 figures: Figure S1 - Figure S4
-  \item 0 tables
-\end{itemize}
-
-\section{Method}
-The van Genuchten equation\cite{Genuchten1980a}:
-\begin{equation}
-  \mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m} \label{eq:saturation}
-\end{equation}
-Se is the soil saturation, $\theta_w$, $\theta_r$, $\theta_t$ are the water filled, residual and total porosity of the soil.
-The vapor or gas filled porosity may be calculated by $\theta_g = \theta_t - \theta_w$
-The local soil saturation depends on the parameters $\alpha$, $n$, $m= 1 - 1/n$ and the elevation above groundwater $z$.
-
-Convection-diffusion equation in porous media:
-\begin{equation}
-  \frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c \label{eq:conv-diff}
-\end{equation}
-Where $c_w$ and $c$ are the liquid phase and vapor phase contaminant concentration, $D_\mathrm{eff}$ is
-the effective diffusion coefficient and $\vec{u}$ is the soil gas velocity, as determined by Darcy’s Law.
-It is important to recall that the soil gas flow velocity is not significantly influenced by the contaminant itself, because the contaminant concentrations are normally so low that the soil gas is largely air.
-The above equation is presented in an unsteady form, reflecting the fact that there may be holdup of the contaminant vapor in both the vapor phase as well as in the soil moisture phase. \par
-
-Richards' equation\cite{Richards1931a}:
-\begin{equation}
-  \nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0 \label{eq:richards-equation}
-\end{equation}
-The soil gas flow is driven by a pressure gradient $\nabla p$ and is proportional to the modified permeability of the soil $k_r \kappa_s$ and inversely proportional to the vapor viscosity $\mu$.
-The modified permeability is the saturated fluid permeability $\kappa_s$ corrected using the relative permeability $k_r$ defined by:
-\begin{equation}
-  k_r = (1 - \mathrm{Se})^{0.5} [1 - (\mathrm{Se}^{-m})^m]^2 \label{eq:k_r}
-\end{equation}
-
-Millington-Quirk equation for effective diffusion coefficient in porous media\cite{Millington1961a}.
-\begin{equation}
-  D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2} \label{eq:millington-quirk}
-\end{equation}
-$D_\mathrm{air}$ and $D_\mathrm{water}$ are the diffusion coefficient of the contaminant in air and water respectively and $K_H$ is the dimensionless Henry’s Law constant. \par
-
-The boundary condition for contaminant transport through the foundation crack:
-\begin{align}
-  j_\mathrm{ck} &= u_\mathrm{ck} c_\mathrm{ck} - \frac{u_\mathrm{ck} (c_\mathrm{ck} - c_\mathrm{in})}{1 - \exp{(u_\mathrm{ck}L_\mathrm{slab}/D_\mathrm{air})}} \label{eq:j_ck} \\
-  n_\mathrm{ck} &= \int_{A_\mathrm{ck}} j_\mathrm{ck} dA \label{eq:n_ck}
-\end{align}
-Where $u_\mathrm{ck}$ and $c_\mathrm{ck}$ are the flow velocity and contaminant concentration at the soil side of the crack, respectively, $L_\mathrm{slab}$ is the thickness of the foundation slab (here, 15 cm), $D_\mathrm{air}$ is the diffusion coefficient of the contaminant in air, $j_\mathrm{ck}$ is the flux in the crack and $A$ represents the crack cross-sectional area.
-
-The assumed values for the various parameters mentioned above are shown in Table \ref{tbl:data}.
-\begin{table}[!htb]
-  \centering
-  \caption{Contaminant and soil properties\cite{Abreu2012a,U.S.-Environmental-Protection-Agency2004a}.} \label{tbl:data}
-  \begin{tabular}{l r r r r r r}
-  \toprule
-    Soil & $\text{Permeability} \; [\mathrm{m^2}]$  & $\mathrm{Density} \; [\mathrm{kg/m^3}]$  & $\theta_s$  & $\theta_r$  & $\alpha \; [\mathrm{1/m}]$  & $n$ \\
-    \hline
-    Gravel\cite{Dan2012a}     & $1.3 \cdot 10^{-9}$   & 1680    & 0.42        & 0.005       & 100       & 3.1 \\
-    Sand          & $9.9 \cdot 10^{-12}$  & 1430    & 0.38        & 0.053       & 3.5       & 3.2 \\
-    Loamy Sand    & $1.6 \cdot 10^{-12}$  & 1430    & 0.39        & 0.049       & 3.5       & 1.7 \\
-    Sandy Clay    & $1.7 \cdot 10^{-14}$  & 1470    & 0.39        & 0.12        & 3.3       & 1.2 \\
-                  &                       &         &             &             &           & \\
-    \multicolumn{7}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
-                  &                       &         &             &             &           & \\
-    \multicolumn{7}{c}{Contaminant: Trichloroethylene (Diluted in air)} \\
-    \multicolumn{1}{r}{$D_\mathrm{air} \; [\mathrm{m^2/h}]$}  & $D_\mathrm{water} \; [\mathrm{m^2/h}]$  & $\mathrm{Density} \; [\mathrm{kg/m^3}]$ & $\mathrm{Viscosity} \; [\mathrm{Pa \cdot s}]$  & $K_H$ & $M \; [\mathrm{g/mol}]$ &  \\
-    \hline
-    \multicolumn{1}{r}{$2.47 \cdot 10^{-2}$}  & $3.67 \cdot 10^{-6}$  & 1.614 & $1.86 \cdot 10^{-5}$  & 0.403 & 131.39  &  \\
-    \bottomrule
-    \multicolumn{7}{c}{$\theta_t=$ soil total porosity, $\theta_g=$ gas filled porosity, $\theta_w=$ water filled porosity}
-  \end{tabular}
-\end{table}
-
-\subsection{Transient Behavior Under "Normal" VI Scenarios}
-
-If the air exchange rate $A_e$ is varied sinusoidally across a day between the high and low values as are realistic in a residential setting, one may explore the potential impact of this variable on transient changes in $c_\mathrm{in}$ while keeping $n_\mathrm{ck}$ constant.
-The range of air exchange rate values was taken to be roughly the same as was observed at the ASU house, where air exchange rates varied from 5 to 24 per day with the majority of the time value being between 10 to 20 per day\cite{Holton2013a,Holton2015a}, this is relatively representative of normal residential houses\cite{Murray1995a}.
-As may be seen in Figure \ref{fig:cstr_ae}, $c_\mathrm{in}$ responds quickly with a change in $A_e$, but the change is significantly lower than the few orders of magnitude that has been observed to occur at the ASU house on timescales of a day.
-The modest magnitude of the change is consistent with earlier conclusions on this point\cite{Shen2016a}. \par
-
-Now instead $n_\mathrm{ck}$ may be varied and $A_e$ kept constant.
-It may be seen in Figure \ref{fig:cstr_nck} that $c_\mathrm{in}$ quickly responds to changes in $n_\mathrm{ck}$ and the increase in $c_\mathrm{in}$ is proportional to the increase in $n_\mathrm{ck}$, indicating, not surprisingly, that steep changes in this parameter can be the driving force for increases in $c_\mathrm{in}$.
-\par
-
-\begin{figure}[!htb]
-  \caption[Varying air exchange and contaminant entry rate impact on indoor air concentration]{Varying inputs of the indoor air contaminant concentration calculation.
-(\subref{fig:cstr_ae}) considers the impact of varying air exchange rate with constant contaminant entry rate.
-(\subref{fig:cstr_nck}) considers the impact of varying contaminant entry rate with constant air exchange rate.}
-  \begin{subfigure}[b]{0.9\textwidth}
-    \caption{ } \label{fig:cstr_ae}
-    \includegraphics[width=\textwidth]{cstr_ae.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.9\textwidth}
-    \caption{ } \label{fig:cstr_nck}
-    \includegraphics[width=\textwidth]{cstr_nck.png}
-  \end{subfigure}
+\newpage
+\subsection{Pressure and Air Exchange Rate}
+\begin{figure}
+  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
+The seasonal median $p_\mathrm{in/out}$ and $A_e$ shown.
+Only non-CPM period considered.}
+  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
 \end{figure}
 
-In Figure \ref{fig:stack_effect} the effects of building pressurization/depressurization are considered.
-In order to maximize the impact of pressure swings, the building is assumed to have a gravel sub-base and to be surrounded by sand soil.
-For an already depressurized building, further depressurization only moderately increases $c_\mathrm{in}$.
-Overpressurization of a building however leads to significantly lower $c_\mathrm{in}$, as contaminant entry rates are significantly decreased.
-However, the full extent of this decrease is not shown in the figure; the y-axis was purposely truncated in order to highlight the effects of increased depressurization. \par
-
-\begin{figure}[!htb]
-  \centering
-  \caption[Effects of house depressurization]{A side view of steady-state concentration profiles of "normal" VI scenarios featuring a gravel sub-base with sand soil, with different stack effect magnitudes.
-Color indicate contaminant vapor mole fraction and are logarithmic.
-The attenuation from groundwater source is given for each figure.} \label{fig:stack_effect}
-  % top figure
-  \begin{subfigure}{0.9\textwidth}
-    \centering
-    \caption{ } \label{fig:stack_effect1}
-    \includegraphics[width=\textwidth]{c_in_p_ck.png}
-  \end{subfigure}
-  % left bottom digure
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \caption{-5 Pa, $\alpha = 9.1 \cdot 10^{-6}$} \label{fig:stack_effect2}
-    \includegraphics[width=\textwidth]{soil_concentration_right_side1.png}
-  \end{subfigure}
-  % right bottom figure
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \caption{-20 Pa, $\alpha = 1.6 \cdot 10^{-5}$} \label{fig:stack_effect3}
-    \includegraphics[width=\textwidth]{soil_concentration_right_side2.png}
-  \end{subfigure}
-\end{figure}
-
-Figure \ref{fig:diffusion} illustrates the effective diffusion coefficient for the three different soil types used, ranging from soils allowing for faster diffusion (sand) to soils that show slower vapor diffusion (sandy clay).
-Notice that diffusion is the slowest through the water filled soil, such as just above groundwater, in the capillary fringe, providing a significant resistance to contaminant vapor diffusion.
-The characteristic time scale for transport by diffusion is given by $L^2/D$, where $L$ is the characteristic length scale, which in this case is of order meters (based on the source-slab separation).
-Even with only a 1 m separation and a high $D_\mathrm{eff} = 10^{-6} \; \mathrm{m^2/s}$, response times can only be be as short as weeks.
-Therefore, fluctuations in groundwater contaminant concentrations cannot be the underlying cause of the observed rapid transients in indoor air contaminant concentrations. \par
-
-\begin{figure}[!htb]
-  \caption[Soil moisture \& contaminant transport speed through soil]{In (\subref{fig:soil_properties}) the van Genuchten soil moisture content of various soils as a function of elevation above groundwater along with the corresponding effective diffusion of TCE in those soils.
-The diffusive propagation of TCE through sand soil is shown in (\subref{fig:contaminant_propagation})}\label{fig:diffusion}
-  \centering
-  \begin{subfigure}[t]{0.65\textwidth}
-    \centering
-    \caption{ } \label{fig:soil_properties}
-    \includegraphics[width=\textwidth]{d_eff.png}
-  \end{subfigure}
-  \begin{subfigure}[t]{0.3\textwidth}
-    \centering
-    \caption{ } \label{fig:contaminant_propagation}
-    \includegraphics[width=\textwidth]{c_soil_transient.png}
-  \end{subfigure}
-\end{figure}
-
-\subsection{The Preferential Pathway as a Source of Contaminant Beneath a Structure}
-
-Air flow from the preferential pathway is important for the distribution of vapor contaminant in the sub-base and the soil, and ultimately for the entry rate of contaminant into a building.
-To highlight this, a scenario of a house with a gravel sub-base, surrounded by loamy sand soil with contaminated preferential pathway that exits into the gravel sub-base is considered.
-This scenario is solved at steady-state twice - first with both advection-diffusion enabled (like all other calculations presented) and the second with advection disabled, i.e. diffusion only. \par
-
-The result of this may be seen in Figure \ref{fig:adv-diff_vs_diff}, where (\subref{fig:advection_diffusion}) shows the advection-diffusion result, i.e.
-the regular result and (\subref{fig:diffusion_only}) the diffusion only result.
-Just from observation it is clear that there is a massively different soil contaminant vapor concentration profile, showing that existence of a diffusing preferential pathway will have a very modest impact on VI.
-Contaminant needs to be blown in to have a major effect.
-
-\begin{figure}[!htb]
-  \centering
-  \caption[Importance of advection for the impact of a preferential pathway]{The impact of advection for the distribution of contaminant vapor from a preferential pathway.
-The soil vapor contaminant concentration in the loamy sand soil and gravel sub-base, a short distance away from the preferential pathway shown.
-(\subref{fig:advection_diffusion}) shows the regular advection and diffusion case and (\subref{fig:diffusion_only}) shows the same scenario but advection disabled.} \label{fig:adv-diff_vs_diff}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \caption{Advection enabled} \label{fig:advection_diffusion}
-    \includegraphics[width=\textwidth]{steady-state_preferential_pathway2.png}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.45\textwidth}
-    \centering
-    \caption{Advection disabled} \label{fig:diffusion_only}
-    \includegraphics[width=\textwidth]{steady-state_preferential_pathway_diff_only2.png}
-  \end{subfigure}
-\end{figure}
-
-\newpage
-\bibliography{../library}
-
 \end{document}

### 2019-03-27 tex/main.tex
Draft #3?

@@
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
+\author{Yuanming Guo}
+\affiliation[Arizona State University]{Arizona State University}
 \author{Eric M. Suuberg}
@@
 % title
-\title{Temporal Variations In Vapor Intrusion Induced Indoor Air Contaminant Concentrations }
+\title{Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
 
@@
 No preferential pathway scenarios in bottom panel.
-Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars represent the 95\% confidence interval.}
- \label{fig:land_drain_scenarios}
+Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars represent the 95\% confidence interval.
+Shaded blue represent the predicted variability due to air exchange rate (using 5th and 95th percentile values).}\label{fig:land_drain_scenarios}
  \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
@@
 \begin{figure}
- \caption{}\label{fig:diurnal}
+ \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
  \includegraphics[width=\textwidth]{diurnal.pdf}
@@
 
-To see how much variability can be explained due to fluctuations in air exchange rate
+The above analyses have been conducted under simulated steady state conditions.
+The final test of the conclusions reached above comes in applying them in an actual transient simulations.
+We run the model described in Figure \ref{fig:model} and run 24-hour simulations to see how $c_\mathrm{in}$ fluctuates over a "typical" day, and how much of these fluctuations are attributed to air exchange rate versus indoor/outdoor pressure difference.
+This is achieved by using varying $p_\mathrm{in/out}$ as one model input, and then running a scenario where $A_e$ also fluctuates, and one where $A_e$ is constant throughout the entire simulation period.
+Just like before, we consider this effect for both the preferential pathway open and closed cases.
 
-The above analyses have been conducted under simulated steady state conditions.
-The final test of the conclusions reached above comes in applying them in an actual transient simulation of a scenario such as that in Figure 2.
-These results are shown in Figure 7.
+Using the ASU House dataset, the "typical" daily $p_\mathrm{in/out}$ is found by querying the data what the median, hourly, diurnal $p_\mathrm{in/out}$ was during the non-CPM periods, giving us the median $p_\mathrm{in/out}$ at 00:00, ..., 13:00, ..., 23:00.
+This "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly medians are interpolated using cubic splines).
+The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
+The orange line is the constant air exchange rate value used.
+
+The result of these simulations may be seen in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the preferential pathway open and closed respectively scenarios.
+The "max change" value in the legends is the quotient between the lowest and highest concentrations, i.e. a value of two indicate that the maximum daily concentration is twice as high as the lowest.
+This is the same property that is plotted in Figure \ref{fig:resampling}.
+
+When the preferential pathway is open, we see that there is a maximum change of a factor by around 5, irrespective if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
+The relatively small difference between the diurnal and constant $A_e$ cases indicate that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$.
+This is to be expected as we have already established that the preferential pathway increased the "advective potential", i.e. increasing the influence of $p_\mathrm{in/out}$ on contaminant entry rate.
+Even for the small fluctuations in $p_\mathrm{in/out}$, under the influence of this type of preferential pathway, the contaminant entry rate is a more significant than the contaminant exit rate, which determined by air exchange rate.
+
+When the preferential pathway is closed the story is quite different.
+Here we see that when air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
+This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrate that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small (which is further compounded by the small changes in $p_\mathrm{in/out}$).
+When the air exchange rate is fluctuating the maximum change in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}, showing that for a "typical" day, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.
+
+All of this show the complicated nature of temporal variability in $c_\mathrm{in}$.
+It is important to remember that we have only considered the effect of indoor/outdoor pressure difference and air exchange rate in this artcile, but slower processes, e.g.
+changes in groundwater contaminant concentration or various seasonal effects has a significant impact on VI over time, which we will address in future articles, are too of importance.
+But for shorter time periods, temporal variability in indoor contaminant concentration is dominated by indoor/outdoor pressure difference and/or air exchange rate.
+
+For a site where the communication between the subsurface and the indoor is good i.e.
+advective transport is significant, $p_\mathrm{in/out}$ is or may be the most significant determinant of $c_\mathrm{in}$ and its temporal variability.
+In this paper we show that that such a scenario may arise due to a preferential pathway, such as the one at the ASU House.
+But these conditions may also exist even in the absence of a preferential pathway as the NAS North Island demonstrate.
+At sites where advective transport is not as significant much of the (especially shorter-term) temporal variability in $c_\mathrm{in}$ may be attributed to fluctuations in air exchange rate.
 

### 2019-04-01 tex/main.tex
Minor edits

@@
 Month-long samples?
-Continuous monitoring?
+Continuous monitoring?\par
 
@@
 All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.
+All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.\par
 
@@
 Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.
+While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.\par
 
@@
 A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.
+Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.\par
 
@@
 Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.
+It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.\par
 
@@
 those characterized by preferential pathways within the soil beneath the site.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.
+The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
 
@@
 For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.
+Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.\par
 
@@
 The period of CPM testing is thus excluded from the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.
+Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par
 
@@
 By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
-The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.
+The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.\par
 
@@
 Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
-There is almost a two order of magnitude variation in the concentration data.
+There is almost a two order of magnitude variation in the concentration data.\par
 
@@
 In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.
+The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.\par
 
@@
 Furthermore, at the Indianapolis site the $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration, and is therefore not examined in this portion of the analysis.
+The NAS North Island NAS dataset was obtained over a much shorter duration, and is therefore not examined in this portion of the analysis.\par
 
@@
 The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
-Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.
+Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
 
@@
 As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
-So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.
+So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.\par
 
@@
 This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
-This would be consistent with requiring months to see the more significant variations.
+This would be consistent with requiring months to see the more significant variations.\par
 
@@
 Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
-This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.
+This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.\par
 
@@
 In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
-Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.
+Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.\par
 
-If the maximum observed variability is quite small across only a few days as these data suggest, a sampling strategy that involves taking daily samples for several days in a row will likely yield little extra useful information beyond what can be obtained with a single sample.
+If the maximum observed variability is quite small across only a few days as these data suggest, a sampling strategy that involves taking daily samples for several days in a row will likely yield little extra useful information beyond what can be obtained with a single sample.\par
 
@@
 
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure.  Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).
+The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure. 
+Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).\par
+
 We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
@@
 In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
-A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.
+A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.\par
+
 Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
-Such issues have of course already been pointed out by the investigators who obtained the data.
+Such issues have of course already been pointed out by the investigators who obtained the data.\par
 
@@
 At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).
+This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).\par
 
 The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
-The Pearson’s r value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
-But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.
+The Pearson’s r-value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
+But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.\par
 
@@
 The variability of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
-The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
+The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
+In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.\par
 
 It should be noted that control of indoor depressurization is the principle that underlies the controlled pressure method (CPM) for assessing the potential severity of a VI situation.
-These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.
+These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
 
@@
 The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
-These will be explored with the help of a modeling analysis presented below.
+These will be explored with the help of a modeling analysis presented below.\par
 
@@
 This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}.
+The results are shown in Figure \ref{fig:attenuation_subslab}.\par
 
 It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.
+Thus during the period when the preferential pathway was closed large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
 
@@
 A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
-This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.
+This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.\par
 
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in “normal” VI scenarios.
+What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in “normal” VI scenarios.\par
 
@@
 
-To begin a more thorough discussion regarding the role and impact a preferential pathway may have at a particular site, we turn to our model presented in the methods section.
-Here, $A_e$ is assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ is varied from -5 to 5 Pa.
-The predicted $c_\mathrm{in}$ is then normalized to the groundwater concentration $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
+Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
+The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
+This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
+The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
+The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
 
-The predicted $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ is given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios} (ignore the shaded blue areas for now).
-These predicted values are compared to $\alpha_\mathrm{gw}$ from the ASU house data from the period the preferential pathway was open, given by the blue dots.
+In the calculation whose results are shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
+Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
+Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
+The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater  cgw, giving the attenuation from groundwater $\alpha_\mathrm{gw}$ .
+The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
+These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
 
-The model successfully predicts the increase in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) occurs.
-This asymmetry is due to two factors.
+The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
+Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.\par
 
-First, the preferential pathway acts as preferential air source, circumventing the large resistance to flow in the surrounding soil, and allowing a small change in $p_\mathrm{in/out}$ to much more dramatically increase the advective flow into the structure from the subslab region.
-A second simulation, where the model is rerun with the preferential pathway present, but the permeable (gravel) layer in the subslab is removed and instead is assumed to be of the same material as the surrounding soil (sandy loam), giving a "uniform soil" scenario (orange line in the top panel of Figure \ref{fig:land_drain_scenarios}).
-This simulation demonstrates that without a permeable subslab to effectively allow the "advective potential" to be realized, a preferential pathway may not impact a VI site; a preferential pathway requires good communication between the indoor environment and itself to be as a significant contributor to VI as was seen at the ASU house site.
+The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
+First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
+Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.\par
 
-The second factor is that the preferential pathway must contain contaminant vapors to be impactful.
-In a third simulation, the permeable subslab region is included, but the preferential pathway is filled with clean air, i.e.
-just allows the "advective potential" to be realized, but without any additional contaminant introduced.
-The result of this simulation is given by the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a slightly larger increase in $\alpha_\mathrm{gw}$ compared to the "uniform soil" scenario as the structure is increasingly depressurized, it is nowhere near as significant as when the preferential pathway was filled with contaminant vapors.
-The contaminated and uncontaminated preferential pathway scenarios (blue and green lines respectively) thus give the range of $\alpha_\mathrm{gw}$ that one may see for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.
+The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
+This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
+This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway way actually not impact a VI site very much.
+In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
+These requirements were met at the ASU House.\par
 
-The model is also able to capture $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ when a preferential pathway is absent (but there is a permeable subslab region), as the bottom panel in Figure \ref{fig:land_drain_scenarios} shows.
-However, what these simulations fail to predict is the considerable variability of $\alpha_\mathrm{gw}$ for a given $p_\mathrm{in/out}$, as shown by the error bars in the two panels (denoting one standard deviation).
-A significant portion of this variability may be explained by variation in air exchange rate, $A_e$.
+A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
+In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
+The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
+This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
+The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.\par
 
-Air exchange rate determines which portion of the indoor is exchanged with outdoor air, reducing the indoor air contaminant concentration.
-A high air exchange is associated with lower $c_\mathrm{in}$ and low with higher $c_\mathrm{in}$.
-$A_e$ is correlated with $p_\mathrm{in/out}$, i.e.
-that a depressurized structure has higher $A_e$ than an overpressurized structure as more air from the outside may be brought in, but with significant variability.
-Determining any determinant relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena has a huge effect on air exchange, i.e.
-for a given $p_\mathrm{in/out}$.
-Fortunately, the range of $A_e$ values are typically bounded, with the U.S.
-EPA reporting that (depending on region) the 10th percentile values range from 0.16 to 0.2 per hour, and the 50th and 90th percentiles range from 0.35 to 0.49 and 1.21 to 1.49 per hour respectively\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-These values are in line with what as found at the both the ASU house and the Indianapolis site as Table \ref{tbl:air_exchange_rate} shows.
+The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
+These results are shown as the bottom panel in Figure \ref{fig:land_drain_scenarios}.  
+These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed.\par
 
-To more accurately predict the range of $\alpha_\mathrm{gw}$ values for a given $p_\mathrm{in/out}$, the preferential pathway (top panel, blue line) and the no preferential pathway (bottom panel, blue line) scenarios are rerun with $A_e$ equal to 0.1, 0.5, and 1.5 per hour, giving us a predicted upper and lower bound for $\alpha_\mathrm{gw}$.
-These bounds are given by the blue region above and below the central line, with the upper and lower bound associated with 0.1 and 1.5 air exchange rate per hour respectively.
-Doing this allows us to capture most of the variability in $\alpha_\mathrm{gw}$, indicating that a significant portion of the variability in $\alpha_\mathrm{gw}$ for a given $p_\mathrm{in/out}$ may be attributed to fluctuations in air exchange rate.
-This could be used during VI site investigations to estimate how significantly the indoor air contaminant concentration may fluctuate.
+What the above simulations fail to predict is the actual observed variability of $\alpha_\mathrm{gw}$ at any given $p_\mathrm{in/out}$.
+This variability is captured by the error bars around the data points in the two panels (denoting one standard deviation).
+This variability is roughly an order of magnitude.
+Examining Figure \ref{fig:model} for the case in which the preferential pathway is closed, the “cloud” of data similarly reflect that there is still not captured in the modeling a parameter that determines this variation, independently of the pressure effects.\par
+
+One other feature of this simulation that warrants noting is that the examined range from $p_\mathrm{in/out}$ from -5 to +5 Pa is larger than the “usual” range from -2 to +2 Pa observed at the ASU House (as may be seen from Figure \ref{fig:model}).
+The data at larger and smaller values of $p_\mathrm{in/out}$ were rather sparse, though there were enough points so that the data for the higher range of pressures could be presented in Figure \ref{fig:land_drain_scenarios}.
+But focusing on the range from -2 to + 2 Pa alone, the pressure dependence of variability is much less visible, just as the data of Figure \ref{fig:model} suggested.
+This again means that there is a need to explore other reasons for the order of magnitude spread seen in concentration data, when the preferential pathway is closed.\par
+
+\subsubsection{Air Exchange Rate Effects}
 
@@
 
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.
-
-In the calculation whose results are shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
-Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
-Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater  cgw, giving the attenuation from groundwater $\alpha_\mathrm{gw}$ .
-The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
-These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
-
-The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.
-
-The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
-First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
-Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.
-
-The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
-This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway way actually not impact a VI site very much.
-In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
-These requirements were met at the ASU House.
-
-A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
-In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
-The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
-The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathay.
-
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
-These results are shown as the bottom panel in Figure \ref{fig:land_drain_scenarios}.
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed.
-
-What the above simulations fail to predict is the actual observed variability of $\alpha_\mathrm{gw}$ at any given $p_\mathrm{in/out}$.
-This variability is captured by the error bars around the data points in the two panels (denoting one standard deviation).
-This variability is roughly an order of magnitude.
-Examining Figure \ref{fig:model} for the case in which the preferential pathway is closed, the “cloud” of data similarly reflect that there is still not captured in the modeling a parameter that determines this variation, independently of the pressure effects.
-
-One other feature of this simulation that warrants noting is that the examined range from $p_\mathrm{in/out}$ from -5 to +5 Pa is larger than the “usual” range from -2 to +2 Pa observed at the ASU House (as may be seen from Figure \ref{fig:model}).
-The data at larger and smaller values of $p_\mathrm{in/out}$ were rather sparse, though there were enough points so that the data for the higher range of pressures could be presented in Figure \ref{fig:land_drain_scenarios}.
-But focusing on the range from -2 to + 2 Pa alone, the pressure dependence of variability is much less visible, just as the data of Figure \ref{fig:model} suggested.
-This again means that there is a need to explore other reasons for the order of magnitude spread seen in concentration data, when the preferential pathway is closed.
-
-\subsubsection{Air Exchange Rate Effects}
-
 Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
@@
 Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
+As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.\par
 
 To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are given by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 hr-1.
+These bounds are given by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
 
@@
 We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.
+Thus we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
 
@@
 This is achieved by using varying $p_\mathrm{in/out}$ as one model input, and then running a scenario where $A_e$ also fluctuates, and one where $A_e$ is constant throughout the entire simulation period.
-Just like before, we consider this effect for both the preferential pathway open and closed cases.
+Just like before, we consider this effect for both the preferential pathway open and closed cases.\par
 
@@
 The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the constant air exchange rate value used.
+The orange line is the constant air exchange rate value used.\par
 
@@
 The "max change" value in the legends is the quotient between the lowest and highest concentrations, i.e. a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This is the same property that is plotted in Figure \ref{fig:resampling}.
+This is the same property that is plotted in Figure \ref{fig:resampling}.\par
 
@@
 This is to be expected as we have already established that the preferential pathway increased the "advective potential", i.e. increasing the influence of $p_\mathrm{in/out}$ on contaminant entry rate.
-Even for the small fluctuations in $p_\mathrm{in/out}$, under the influence of this type of preferential pathway, the contaminant entry rate is a more significant than the contaminant exit rate, which determined by air exchange rate.
+Even for the small fluctuations in $p_\mathrm{in/out}$, under the influence of this type of preferential pathway, the contaminant entry rate is a more significant than the contaminant exit rate, which determined by air exchange rate.\par
 
@@
 This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrate that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small (which is further compounded by the small changes in $p_\mathrm{in/out}$).
-When the air exchange rate is fluctuating the maximum change in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}, showing that for a "typical" day, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.
+When the air exchange rate is fluctuating the maximum change in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}, showing that for a "typical" day, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
 
@@
 changes in groundwater contaminant concentration or various seasonal effects has a significant impact on VI over time, which we will address in future articles, are too of importance.
-But for shorter time periods, temporal variability in indoor contaminant concentration is dominated by indoor/outdoor pressure difference and/or air exchange rate.
+But for shorter time periods, temporal variability in indoor contaminant concentration is dominated by indoor/outdoor pressure difference and/or air exchange rate.\par
 
@@
 But these conditions may also exist even in the absence of a preferential pathway as the NAS North Island demonstrate.
-At sites where advective transport is not as significant much of the (especially shorter-term) temporal variability in $c_\mathrm{in}$ may be attributed to fluctuations in air exchange rate.
+At sites where advective transport is not as significant much of the (especially shorter-term) temporal variability in $c_\mathrm{in}$ may be attributed to fluctuations in air exchange rate.\par
 

### 2019-04-01 tex/main.tex
More edits

@@
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
-\author{Yuanming Guo}
-\affiliation[Arizona State University]{Arizona State University}
+%\author{Yuanming Guo}
+%\affiliation[Arizona State University]{Arizona State University}
 \author{Eric M. Suuberg}
@@
   % Indoor air space equation
-  Unsteady-CSTR                            & $V\frac{d u}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - u A_e V$                                                                  \\
+  Unsteady-CST                            & $V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$                                                                  \\
   % Richard's equation
@@
   % Foundation crack
-  At foundation crack          & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$ & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
+  Foundation crack          & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$ & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
   % Groundwater source
-  At groundwater source        & N/A                                      & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$                            \\
+  Groundwater source        & \textit{No flow}                                      & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$                            \\
   % Ground surface
-  At ground surface            & $p = 0 \; \mathrm{(Pa)}$                 & $c = 0 \; \mathrm{(\mu g/m^3)}$                                            \\
+  Groundsurface            & $p = 0 \; \mathrm{(Pa)}$                 & $c = 0 \; \mathrm{(\mu g/m^3)}$                                            \\
   % Preferential pathway
-  Exit of preferential pathway & $p = 0 \; \mathrm{(Pa)}$                 & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$                       \\
+  Preferential pathway & $p = 0 \; \mathrm{(Pa)}$                 & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$                       \\
   \bottomrule
@@
   % Descriptions
-  Soil       & $\text{Permeability} \; \mathrm{(m^2)}$ & $\mathrm{Density} \; \mathrm{(kg/m^3)}$ & $\theta_s$ & $\theta_r$ & $\alpha \; \mathrm{(1/m)}$ & $n$ \\
+  Soil       & $\kappa_s \; \mathrm{(m^2)}$ & $\rho \; \mathrm{(kg/m^3)}$ & $\theta_s$ & $\theta_r$ & $\alpha \; \mathrm{(1/m)}$ & $n$ \\
   % Gravel
@@
   \toprule
-  $D_\mathrm{air} \; \mathrm{(m^2/h)}$ & $D_\mathrm{water} \; \mathrm{(m^2/h)}$ & $\mathrm{Density} \; \mathrm{(kg/m^3)}$ & $\mathrm{Viscosity} \; \mathrm{(Pa \cdot s)}$ & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
+  $D_\mathrm{air} \; \mathrm{(m^2/h)}$ & $D_\mathrm{water} \; \mathrm{(m^2/h)}$ & $\rho \; \mathrm{(kg/m^3)}$ & $\mu \; \mathrm{(Pa \cdot s)}$ & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
   $2.47 \cdot 10^{-2}$                 & $3.67 \cdot 10^{-6}$                   & 1.614                                   & $1.86 \cdot 10^{-5}$                          & 0.403 & 131.39                  \\
@@
 
+\begin{table}[htb!]
+  \caption{List of abbreviations}\label{tbl:abbreviations}
+\begin{tabular}{l l}
+  \toprule
+  % A/ALPHA
+  $A_\mathrm{ck}$ & Crack area \\
+  $A_e$ & Air exchange rate \\
+  $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
+  $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
+  % C
+  $c_\mathrm{in}$ & Indoor air contaminant concentration \\
+  $c$ & Soil-gas contaminant concentration \\
+  $c_w$ & Soil-water contaminant concentration \\
+  $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
+  $\chi$ & Preferential pathway contaminant concentration scaling parameter \\
+  % D
+  $D_\mathrm{eff}$ & Effective diffusion coefficient \\
+  $D_\mathrm{air}$ & Diffusion coefficient in air \\
+  $D_\mathrm{water}$ & Diffusion coefficient in water \\
+  % J
+  $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
+  % K
+  $\kappa_s$ & Saturated soil permeability \\
+  $K_H$ & Dimensionless Henry's law constant \\
+  $k_r$ & Relative permeability \\
+  % L
+  $L_\mathrm{slab}$ & Thickness of the foundation slab \\
+  % M
+  $M$ & Molar mass \\
+  $\mu$ & Contaminant vapor viscosity \\
+  % N
+  NAS & Naval Air Stations \\
+  % P
+  $p$ & Pressure in soil \\
+  $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
+  PP & Preferential pathway \\
+  % R
+  $\rho$ & Density \\
+  % S
+  $\mathrm{Se}$ & Soil water saturation \\
+  % T
+  $t$ & time \\
+  $\theta_g$ & Vapor/gas filled porosity \\
+  $\theta_w$ & Water filled porosity \\
+  $\theta_r$ & Residual water filled porosity \\
+  $\theta_t$ & Total porosity \\
+  % U
+  $\vec{u}$ & Soil-gas velocity (vector quantity) \\
+  % V
+  VI & Vapor intrusion \\
+  $V_\mathrm{base}$ & Basement volume \\
+  % Z
+  $z$ & Elevation above groundwater \\
+  \bottomrule
+\end{tabular}
+\end{table}
+
 \bibliography{library}

### 2019-04-02 tex/main.tex
Implemented Eric's suggestions

@@
 Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-Such variations make it difficult for those charged with protecting human health to formulate a response should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
+Such variations make it difficult for those charged with protecting human health to formulate a response - should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
 There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
@@
 
-To frame the question of just how much variability in indoor air contaminant concentrations is actually seen, field datasets have been analyzed.
+To frame the question of just how much variability in indoor air contaminant concentrations is actually observed, field datasets have been analyzed.
 For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
@@
 
-\subsection{Variability Of Attenuation From Subslab}
+\subsection{Variability Of Attenuation to Subslab Concentrations}
 
@@
  \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution, and diamonds are points that are considered outliers.}\label{fig:attenuation_subslab}
+The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
  \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
@@
 This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}.\par
+The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
 
 It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
+Thus during the period when the preferential pathway was closed, large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
 
@@
  \centering
- \caption{Simulated preferential pathway scenarios compared to ASU house field data (top panel).
-No preferential pathway scenarios in bottom panel.
-Field data is binned in 40 evenly spaced bins, with the dot representing the mean and errors bars represent the 95\% confidence interval.
-Shaded blue represent the predicted variability due to air exchange rate (using 5th and 95th percentile values).}\label{fig:land_drain_scenarios}
+ \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
+Field data are binned in 40 evenly spaced pressure bins, with the dot representing the mean and errors bars the 95\% confidence interval of data at a particular pressure range.
+Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
+Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".
+}\label{fig:land_drain_scenarios}
  \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
@@
 
-In the calculation whose results are shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
+In the calculation results shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
 Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
 Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater  cgw, giving the attenuation from groundwater $\alpha_\mathrm{gw}$ .
+The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
 The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
@@
 The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
-These results are shown as the bottom panel in Figure \ref{fig:land_drain_scenarios}.  
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed.\par
+These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
+These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e.
+that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par
 
-What the above simulations fail to predict is the actual observed variability of $\alpha_\mathrm{gw}$ at any given $p_\mathrm{in/out}$.
-This variability is captured by the error bars around the data points in the two panels (denoting one standard deviation).
-This variability is roughly an order of magnitude.
-Examining Figure \ref{fig:model} for the case in which the preferential pathway is closed, the “cloud” of data similarly reflect that there is still not captured in the modeling a parameter that determines this variation, independently of the pressure effects.\par
-
-One other feature of this simulation that warrants noting is that the examined range from $p_\mathrm{in/out}$ from -5 to +5 Pa is larger than the “usual” range from -2 to +2 Pa observed at the ASU House (as may be seen from Figure \ref{fig:model}).
-The data at larger and smaller values of $p_\mathrm{in/out}$ were rather sparse, though there were enough points so that the data for the higher range of pressures could be presented in Figure \ref{fig:land_drain_scenarios}.
-But focusing on the range from -2 to + 2 Pa alone, the pressure dependence of variability is much less visible, just as the data of Figure \ref{fig:model} suggested.
-This again means that there is a need to explore other reasons for the order of magnitude spread seen in concentration data, when the preferential pathway is closed.\par
+The above simulations capture the trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ but do not yet capture the full variability of the concentration results over the "most probable" portion of observed pressure distributions shown in Figure \ref{fig:kde} (which tend to be from -2 to +2 Pa).
+The results of Figure \ref{fig:land_drain_scenarios} show a spread of almost an order of magnitude over this pressure range for the case of the "open" preferential pathway, and almost no spread at all when the preferential pathway is "closed".
+Hence the predicted variability is roughly an order of magnitude too low, when considering only the influence of pressure.
+There is a factor that tends to increase the spread of the data one additional order of magnitude beyond what was predicted by the base calculations of Figure \ref{fig:land_drain_scenarios}.
+We believe that it is variations in air exchange rate, operating in concert with the natural variations in pressure differential, that explain the remaining variability.\par
 
@@
 Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the model of VI.
+Examination of these distributions point in a clear direction for modifying the above model.
 Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
@@
 
-To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are given by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
+To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
+These bounds are indicated by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
 
@@
 We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
+Thus, we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
 
@@
 The above analyses have been conducted under simulated steady state conditions.
-The final test of the conclusions reached above comes in applying them in an actual transient simulations.
-We run the model described in Figure \ref{fig:model} and run 24-hour simulations to see how $c_\mathrm{in}$ fluctuates over a "typical" day, and how much of these fluctuations are attributed to air exchange rate versus indoor/outdoor pressure difference.
-This is achieved by using varying $p_\mathrm{in/out}$ as one model input, and then running a scenario where $A_e$ also fluctuates, and one where $A_e$ is constant throughout the entire simulation period.
-Just like before, we consider this effect for both the preferential pathway open and closed cases.\par
+The conclusions regarding the importance of the different parameters are now examined in actual transient simulations.
+The model configuration of Figure \ref{fig:model} is run in 24-hour transient simulations to examine how $c_\mathrm{in}$ fluctuates over the course  of a "typical" day.
+The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
+The ASU House dataset was again the source of the ”typical” $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
+The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
+The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
+The orange line is the air exchange rate value assumed for the calculations at constant air exchange rate.\par
 
-Using the ASU House dataset, the "typical" daily $p_\mathrm{in/out}$ is found by querying the data what the median, hourly, diurnal $p_\mathrm{in/out}$ was during the non-CPM periods, giving us the median $p_\mathrm{in/out}$ at 00:00, ..., 13:00, ..., 23:00.
-This "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly medians are interpolated using cubic splines).
-The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the constant air exchange rate value used.\par
+The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
+The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e.
+a value of two indicate that the maximum daily concentration is twice as high as the lowest.
+This quantity may be compared with the value that is plotted for “one day” in Figure \ref{fig:resampling}.
+When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
+The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$.
+Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
+When the preferential pathway is closed the story is quite different.
+When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
+This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small (and the variation in this pressure differential is not large in any case).
+When the air exchange rate is allowed to fluctuate the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
+This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
 
-The result of these simulations may be seen in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the preferential pathway open and closed respectively scenarios.
-The "max change" value in the legends is the quotient between the lowest and highest concentrations, i.e. a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This is the same property that is plotted in Figure \ref{fig:resampling}.\par
-
-When the preferential pathway is open, we see that there is a maximum change of a factor by around 5, irrespective if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the diurnal and constant $A_e$ cases indicate that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$.
-This is to be expected as we have already established that the preferential pathway increased the "advective potential", i.e. increasing the influence of $p_\mathrm{in/out}$ on contaminant entry rate.
-Even for the small fluctuations in $p_\mathrm{in/out}$, under the influence of this type of preferential pathway, the contaminant entry rate is a more significant than the contaminant exit rate, which determined by air exchange rate.\par
-
-When the preferential pathway is closed the story is quite different.
-Here we see that when air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrate that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small (which is further compounded by the small changes in $p_\mathrm{in/out}$).
-When the air exchange rate is fluctuating the maximum change in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}, showing that for a "typical" day, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
-
-All of this show the complicated nature of temporal variability in $c_\mathrm{in}$.
-It is important to remember that we have only considered the effect of indoor/outdoor pressure difference and air exchange rate in this artcile, but slower processes, e.g.
-changes in groundwater contaminant concentration or various seasonal effects has a significant impact on VI over time, which we will address in future articles, are too of importance.
-But for shorter time periods, temporal variability in indoor contaminant concentration is dominated by indoor/outdoor pressure difference and/or air exchange rate.\par
-
-For a site where the communication between the subsurface and the indoor is good i.e.
-advective transport is significant, $p_\mathrm{in/out}$ is or may be the most significant determinant of $c_\mathrm{in}$ and its temporal variability.
-In this paper we show that that such a scenario may arise due to a preferential pathway, such as the one at the ASU House.
-But these conditions may also exist even in the absence of a preferential pathway as the NAS North Island demonstrate.
-At sites where advective transport is not as significant much of the (especially shorter-term) temporal variability in $c_\mathrm{in}$ may be attributed to fluctuations in air exchange rate.\par
+These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
+It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g.
+changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
+For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
+For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
+We have shown that that such a scenario may arise due to a preferential pathway, entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
+At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
 

### 2019-04-02 code/figures.py

@@ -82,12 +82,20 @@ class PressureKDE:
 
 def get_log_ticks(start, stop, style='e'):
 
+    def smart_strip(str):
+        for i, x in enumerate(str):
+            str[i] = x.rstrip('0')
+            if str[i][-1] == '.':
+
+                str[i] += '0'
+
+        return str
     ticks = np.array([])
     ints = np.arange(np.floor(start),np.ceil(stop)+1)
     for int_now in ints:
-        ticks = np.append(ticks, np.arange(0.1,1.1,0.1)*10.0**int_now)
-        ticks = np.unique(ticks)
-
+        ticks = np.append(ticks, np.arange(0.1,1.0,0.1)*10.0**int_now)
+        #ticks = np.unique(ticks)
+    ticks = np.append(ticks, 1.0*10.0**ints[-1])
 
     if style=='e':
         labels = ['%1.1e' % tick for tick in ticks]
@@ -96,10 +104,10 @@ def get_log_ticks(start, stop, style='e'):
         labels = ['%1.12f' % tick for tick in ticks]
         ticks_to_keep = ['%1.12f' % 10**int for int in ints]
 
+    ticks_to_keep = np.unique(ticks_to_keep)
+
     ticks = np.log10(ticks)
 
-    labels = list(map(lambda x: x.rstrip('0'), labels))
-    ticks_to_keep = list(map(lambda x: x.rstrip('0'), ticks_to_keep))
 
     for i, label in enumerate(labels):
 
@@ -109,6 +117,8 @@ def get_log_ticks(start, stop, style='e'):
         else:
             #print('Removing label')
             labels[i] = ' '
+
+    labels = smart_strip(labels)
     return ticks, labels
 
 
@@ -117,14 +127,20 @@ class AttenuationSubslab:
         asu = pd.read_csv('./data/asu_house.csv')
         asu = asu.loc[asu['Phase']!='CPM']
 
-        ax = sns.boxplot(x="Phase", y="logAttenuationSubslab", data=asu)
+        ax = sns.boxplot(
+            x="Phase",
+            y="logAttenuationSubslab",
+            data=asu,
+            whis=10,
+            )
 
-        ticks, labels = get_log_ticks(-4,2, style='e')
+
+        ticks, labels = get_log_ticks(-3,2, style='f')
         ax.set(
             xlabel='Preferential pathway status',
[…]

### 2019-04-11 tex/cover-letter/cover_letter.tex
Implemented Yuanmings suggestions (among other)

@@
 
-I am pleased to submit this manuscript entitled "Modeling Variation in Indoor Air Contaminant Concentration \& Impact of Preferential Contaminant Pathways on Vapor Intrusion" for review and possible publication in Environmental Science \& Technology.
+I am pleased to submit this manuscript entitled "Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Environmental Science \& Technology.
 This is original work, and has not been previously submitted to any other journal or conference proceedings.
 
-Using numerical models and published field data from two vapor intrusion (VI) sites, we explore the causes of the temporal variation in indoor air contaminant concentrations in VI impacted houses, a poorly understood phenomena, and especially the role of preferential contaminant pathways in this.
+Using numerical models and published field data from three vapor intrusion (VI) sites, we explore the causes of the temporal variation in indoor air contaminant concentrations in VI impacted houses, a poorly understood phenomena.
+We especially investigate the role of preferential contaminant pathways have in VI, and how air exchange rates contribute to variability in indoor air contaminant concentration.
 
 We found under which conditions one may expect to observe significant changes in indoor air concentrations, and when one should be especially be concerned with preferential pathways.
+Furthermore, we found how changes in indoor outdoor pressure difference and air exchange rate independently affect temporal variability in VI under different conditions. 
 These finding help understand and explain the temporal variation that has been observed at many VI sites, and giving practical guidance for assessment of these sites, something that we expect will be of great interest to your readership.

### 2019-04-11 tex/main.tex
Implemented Yuanmings suggestions (among other)

@@
 
-% TODO: Fix all the citation issues
-
 % authors
@@
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
+\author{Yuanming Guo}
+\affiliation[Arizona State University]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
 \author{Yijun Yao}
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
-%\author{Yuanming Guo}
-%\affiliation[Arizona State University]{Arizona State University}
 \author{Eric M. Suuberg}
@@
 % title
+% TODO: Change title to reflect the inclusion of preferential pathway analysis
 \title{Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
@@
 \abbreviations{VI}
-\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Factor analysis, Modeling}
+\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate}
 
@@
 
+% TODO: Add abstract
 \begin{abstract}
@@
 There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-What represents a reasonable sampling strategy for a particular site a single 8 hr sample?
-Repeated 8 hr samples?
+What represents a reasonable sampling strategy for a particular site a single 8-hour sample?
+Repeated 8-hour samples?
 Month-long samples?
@@
 
-High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$ , such as those in Figure 1, took place at both the ASU House and the Indianapolis House over significant periods\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
+High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
 Furthermore, at the Indianapolis site the $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration, and is therefore not examined in this portion of the analysis.\par
+The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
+It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par
 
@@
 
-If the maximum observed variability is quite small across only a few days as these data suggest, a sampling strategy that involves taking daily samples for several days in a row will likely yield little extra useful information beyond what can be obtained with a single sample.\par
+Multiple samples taken over a short time period, e.g.
+a few days, are unlikely to uncover significant variation in indoor air contaminant concentration; the larger transient variations typically manifest after longer time periods.
+\par
 
@@
 The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.\par
-
-It should be noted that control of indoor depressurization is the principle that underlies the controlled pressure method (CPM) for assessing the potential severity of a VI situation.
+In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
 These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
@@
 
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in “normal” VI scenarios.\par
+What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in "normal" VI scenarios.\par
 
@@
 Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.\par
+As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
+Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
 
@@
 
+% TODO: Simplify or clarify the figure, in particular how you use the legend to highlight the max change.
 \begin{figure}
@@
 The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
-The ASU House dataset was again the source of the ”typical” $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
+The ASU House dataset was again the source of the "typical" $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
 The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
@@
 a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This quantity may be compared with the value that is plotted for “one day” in Figure \ref{fig:resampling}.
+This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
 When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$.
+The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
+the contaminant transport into the modeled structure is advection dominated.
 Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
@@
 When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small (and the variation in this pressure differential is not large in any case).
+This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
+Combined with the small $p_\mathrm{in/out}$ show that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
 When the air exchange rate is allowed to fluctuate the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.

### 2019-04-11 code/figures.py

@@ -302,11 +302,18 @@ class IndianapolisTime:
             ax=ax,
         )
 
+
         ax.set_yscale('log')
+
+        start_date, stop_date = datetime.date(2011, 8, 11), datetime.date(2011, 10, 15)
+
+        custom_tick_locs = np.arange(start_date, stop_date, np.timedelta64(1,'W'))
         ax.set(
             title='Indoor TCE concentration at the Indianapolis site',
             ylabel='$c_\\mathrm{in} \; \\mathrm{(\\mu g/m^3)}$',
-            xlim=([datetime.date(2011, 8, 11), datetime.date(2011, 10, 15)]),
+            xticks=custom_tick_locs,
+            xticklabels=custom_tick_locs,
+            xlim=([start_date, stop_date]),
         )
         plt.xticks(rotation=45)
         plt.tight_layout()
@@ -468,9 +475,9 @@ class Diurnal:
         return
 
 
-#Diurnal()
+Diurnal()
 #PressureKDE(y_data_log=True,norm_conc=True)
-AttenuationSubslab()
+#AttenuationSubslab()
 #Modeling()
 #IndianapolisTime()
 #AirExchangeRateKDE()

### 2019-04-12 tex/main.tex
Wrote abstract and references new figure directory

@@
 \graphicspath{
-  {./../figures/},
-  {./../figures/2d_kde/},
-  {./../figures/simulation_predictions/},
-  {./../figures/pair_grids/},
-  {./../figures/temporal_variability/},
+  {./figures/},
 }
@@
 % title
-% TODO: Change title to reflect the inclusion of preferential pathway analysis
 \title{Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
@@
 \abbreviations{VI}
-\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate}
+\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate, Indoor/outdoor pressure difference}
 
@@
 
-% TODO: Add abstract
 \begin{abstract}
-
+Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
+We consider the source of the variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California and in particular how the indoor/outdoor pressure difference and air exchange rate affected indoor air contaminant concentrations at these sites.
+We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and a more "typical" VI site through three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
+For such a preferential pathway to impact a VI site, a medium for effective communication between the preferential pathway and the indoor air space must exist, e.g.
+a permeable region in the subslab (gravel layer).
+Sites characterized by significant advective transport from the subslab to the indoor air space, much of short-term variability in indoor air contaminant concentration may be explained by fluctuations in indoor/outdoor pressure differences; air exchange rate drives most of the short-term variability for sites characterized by little advective transport.
 \end{abstract}
@@
   over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
- \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
+ \includegraphics[width=\textwidth]{resampling.pdf}
 \end{figure}
@@
 
-% TODO: Simplify or clarify the figure, in particular how you use the legend to highlight the max change.
 \begin{figure}

### 2019-04-15 tex/main.tex
Last changes before submission

--- 2019-04-11 tex/main.tex
@@
 
-% TODO: Add abstract
 \begin{abstract}
-
+Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
+We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
+We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
+We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
+For a preferential pathway to impact a VI site, there must exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
+a permeable subslab space that may be provided by a gravel layer.
+At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
+Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
 \end{abstract}
@@
 
-\begin{figure}
- \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
-\end{figure}
-
 \section{Methods}
@@
 The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
+During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
 The period of CPM testing is thus excluded from the analysis.
@@
 
+\begin{figure}
+ \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
+ \includegraphics[width=\textwidth]{time_indianapolis.pdf}
+\end{figure}
+
 Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
@@
 
+A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
+In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
+As in the earlier studies, only the vadose zone soil domain is directly modeled.
+Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
+
 \begin{figure}[htb!]
@@
 \end{figure}
-
-A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
-In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
 
@@
 The ground surface and the pipe are both sources of air to the soil domain.
-Both are assumed to exist be at reference atmospheric pressure.
+Both are assumed to exist at reference atmospheric pressure.
 Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
@@
 
-\begin{figure}[htb!]
- \centering
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
-  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
- \includegraphics[width=\textwidth]{temporal_variability/resampling.pdf}
-\end{figure}
-
 High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at the Indianapolis site the $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
+Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
 The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
@@
 Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
+
+\begin{figure}[htb!]
+ \centering
+ \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
+  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
+ \includegraphics[width=\textwidth]{resampling.pdf}
+\end{figure}
 
@@
 
-\begin{figure}[htb!]
- \centering
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}
- \label{fig:kde}
- \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
-\end{figure}
-
-\begin{table}[htb!]
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
- \begin{tabular}{c c c c c c c}
-  \toprule
-                                     & \multicolumn{2}{c|}{\textbf{North Island NAS}} & \multicolumn{2}{c|}{\textbf{ASU house PP Open}} & \multicolumn{2}{c}{\textbf{ASU House PP Closed}}                      \\
-  Percentile                         & 5th                                            & 95th                                            & 5th                                              & 95th & 5th  & 95th \\
-  $p_\mathrm{in/out}$ (Pa)           & -19.9                                          & 7.4                                             & -1.4                                             & 2.1  & -2.1 & 2.27 \\
-  $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1                                            & 0.2                                             & 13.5                                             & 0.2  & 3.3  & 0.4  \\
-  \bottomrule
- \end{tabular}
-\end{table}
-
 The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
@@
 For this analysis we considered two VI sites, NAS North Island and the ASU House.
-The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  ere on) had been closed.
+The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
 By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
+
+\begin{figure}[htb!]
+ \centering
+ \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}
+ \label{fig:kde}
+ \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
+\end{figure}
 
@@
 
+\begin{table}[htb!]
+ \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
+ \begin{tabular}{c c c c c c c}
+  \toprule
+                                     & \multicolumn{2}{c|}{\textbf{North Island NAS}} & \multicolumn{2}{c|}{\textbf{ASU house PP Open}} & \multicolumn{2}{c}{\textbf{ASU House PP Closed}}                      \\
+  Percentile                         & 5th                                            & 95th                                            & 5th                                              & 95th & 5th  & 95th \\
+  $p_\mathrm{in/out}$ (Pa)           & -19.9                                          & 7.4                                             & -1.4                                             & 2.1  & -2.1 & 2.27 \\
+  $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1                                            & 0.2                                             & 13.5                                             & 0.2  & 3.3  & 0.4  \\
+  \bottomrule
+ \end{tabular}
+\end{table}
+
 \subsection{Variability Of Attenuation to Subslab Concentrations}
-
-\begin{figure}[htb!]
- \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
-\end{figure}
 
@@
 The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
+
+\begin{figure}[htb!]
+ \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
+The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
+ \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
+\end{figure}
 
@@
 
+Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
+The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
+This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
+The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
+The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
+
 \begin{figure}[htb!]
@@
 \end{figure}
-
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
 
@@
 This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway way actually not impact a VI site very much.
+This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
 In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
@@
 
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, when there exists a permeable subslab region.
+The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
 These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
@@
 \subsubsection{Air Exchange Rate Effects}
+
+Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
+Examination of these distributions point in a clear direction for modifying the above model.
+Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
+A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
+Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
+Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
+As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
+Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
 
@@
 
-Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the above model.
-Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
-A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
-Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
-Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
-Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
-
 To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
@@
 \subsubsection{Results of Transient Simulations}
-
-% TODO: Simplify or clarify the figure, in particular how you use the legend to highlight the max change.
-\begin{figure}
- \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
- \includegraphics[width=\textwidth]{diurnal.pdf}
-\end{figure}
 
@@
 
+\begin{figure}
+ \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
+ \includegraphics[width=\textwidth]{diurnal.pdf}
+\end{figure}
+
 The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
@@
 This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
-When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of if $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
+When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
 The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
@@
 This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
-Combined with the small $p_\mathrm{in/out}$ show that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
-When the air exchange rate is allowed to fluctuate the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
+Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
+When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
 This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
@@
 For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
-We have shown that that such a scenario may arise due to a preferential pathway, entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
+We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
 At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par

### 2019-04-15 tex/supporting-information/esthag_supporting.tex
Removed Guo (for now)
Implemented Yuanmings suggestions (among other)
Wrote abstract and references new figure directory
Last changes before submission

@@
 \usepackage{array,multirow,graphicx}
-\graphicspath{{../../figures/2d_kde/}}
+\graphicspath{{./../figures/}}
 % macros
@@
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
+\author{Yuanming Guo}
+\affiliation[Arizona State University]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
 \author{Yijun Yao}
 \affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
-\author{Yuanming Guo}
-\affiliation[Arizona State University]{Arizona State University}
 \author{Eric M. Suuberg}
@@
 \abbreviations{VI}
-\keywords{Vapor intrusion, Preferential pathways}
+\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate, Indoor/outdoor pressure difference}
 
@@
   \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
-The seasonal median $p_\mathrm{in/out}$ and $A_e$ shown.
+The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
 Only non-CPM period considered.}

### 2019-04-29 tex/main.tex
B&E reformating

@@
-\documentclass[journal=esthag,manuscript=article]{achemso}
+\documentclass[preprint,12pt]{elsarticle}
 
@@
 
-% authors
-\author{Jonathan G. V. Ström}
-\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
-\author{Yuanming Guo}
-\affiliation[Arizona State University]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
-\author{Yijun Yao}
-\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
-\author{Eric M. Suuberg}
-\email{eric_suuberg@brown.edu}
-\affiliation[Brown University]{Brown University, School of Engineering, Providence, RI, USA}
+% keywords
+%\abbreviations{VI}
+%\keywords{}
 
+\journal{Building \& Environment}
+
+\begin{document}
+
+\begin{frontmatter}
 % title
@@
 
-% keywords
-\abbreviations{VI}
-\keywords{Vapor intrusion, Preferential pathways, Temporal variability, Finite element modeling, Air exchange rate, Indoor/outdoor pressure difference}
+% authors
+\author[Brown]{Jonathan G. V. Ström}
+\author[ASU]{Yuanming Guo}
+\author[ASU]{Yijun Yao}
+\author[Brown]{Eric M. Suuberg}
 
-\begin{document}
+\ead{eric\_suuberg@brown.edu}
+
+\address[Brown]{Brown University, School of Engineering, Providence, RI, USA}
+\address[ASU]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
 
@@
 
-\section{Introduction}
+\begin{keyword}
+  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Finite element modeling \sep Air exchange rate \sep Indoor/outdoor pressure difference
+\end{keyword}
+
+\end{frontmatter}
+
+\section{Introduction}\label{s:intro}
 
@@
 \begin{figure}
+  \centering
+ \includegraphics[width=\textwidth]{time_indianapolis.pdf}
  \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
 \end{figure}
@@
 \begin{figure}[htb!]
+  \centering
+ \includegraphics[width=\textwidth]{model.png}
  \caption{Foundation and vadose zone soil represented in the modeling.
@@
 In some cases, the preferential pathway has been "turned off".}\label{fig:model}
- \includegraphics[width=\textwidth]{model.png}
 \end{figure}
@@
 
+% TODO: Restructure entire table
+\clearpage
 \begin{table}[htb!]
  \centering
- \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}
- \label{tbl:eqns_bc_parameters}
  \bigskip
@@
   \toprule
-  \textbf{Boundary}            & \textbf{Richard's equation}              & \textbf{Advection-diffusion equation}                                      \\
+  \textbf{Boundary}            & \textbf{Richard's equation}              & \textbf{Transport equation}                                      \\
   % Foundation crack
@@
  \end{tabular}
+ \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}\label{tbl:eqns_bc_parameters}
 \end{table}
 
-\section{Results \& Discussion}
+\section{Results \& Discussion}\label{s:results}
 
-\subsection{Variation In Indoor Air Contaminant Concentration Over Time}
+\subsection{Variation In Indoor Air Contaminant Concentration Over Time}\label{s:results_variation_time}
 
@@
  \centering
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected
-  over a given time period. (e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months). The error bars are the 95\% confidence intervals.}\label{fig:resampling}
  \includegraphics[width=\textwidth]{resampling.pdf}
+ \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected over a given time period.
+(e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months).
+The error bars are the 95\% confidence intervals.}\label{fig:resampling}
 \end{figure}
@@
 
-\subsection{Statistical Analysis of Field Data}
+\subsection{Statistical Analysis of Field Data}\label{s:results_pressure_concentration}
 
@@
  \centering
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}
- \label{fig:kde}
  \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
+ \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}\label{fig:kde}
 \end{figure}
@@
 \begin{table}[htb!]
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
  \begin{tabular}{c c c c c c c}
   \toprule
-                                     & \multicolumn{2}{c|}{\textbf{North Island NAS}} & \multicolumn{2}{c|}{\textbf{ASU house PP Open}} & \multicolumn{2}{c}{\textbf{ASU House PP Closed}}                      \\
+                                     & \multicolumn{2}{c|}{\textbf{North Island}} & \multicolumn{2}{c|}{\textbf{ASU, PP Open}} & \multicolumn{2}{c}{\textbf{ASU, PP Closed}}                      \\
   Percentile                         & 5th                                            & 95th                                            & 5th                                              & 95th & 5th  & 95th \\
@@
  \end{tabular}
+ \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
 \end{table}
 
-\subsection{Variability Of Attenuation to Subslab Concentrations}
+\subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
 
@@
 \begin{figure}[htb!]
+  \centering
+ \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
  \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
 The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
 \end{figure}
@@
 
-\subsection{Modeling Results}
+\subsection{Modeling Results}\label{s:results_modeling}
 
-\subsubsection{Pressure Effects}
+\subsubsection{Pressure Effects}\label{s:results_modeling_pressure}
 
@@
  \centering
+ \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
  \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
@@
 Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
-Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".
-}\label{fig:land_drain_scenarios}
- \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
+Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".}\label{fig:land_drain_scenarios}
 \end{figure}
@@
 
-\subsubsection{Air Exchange Rate Effects}
+\subsubsection{Air Exchange Rate Effects}\label{s:results_modeling_air_exchange}
 
@@
 \begin{table}[htb!]
- \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
  \begin{tabular}{l c c c}
@@
  \end{tabular}
+ \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
 \end{table}
@@
 
-\subsubsection{Results of Transient Simulations}
+\subsubsection{Results of Transient Simulations}\label{s:results_modeling_transient}
 
@@
 \begin{figure}
+  \centering
+ \includegraphics[width=\textwidth]{diurnal.pdf}
  \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
- \includegraphics[width=\textwidth]{diurnal.pdf}
 \end{figure}
@@
 
-\begin{acknowledgement}
+\section*{Acknowledgements}
 This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).
-\end{acknowledgement}
 
+\appendix
+
+\clearpage
 \begin{table}[htb!]
@@
   $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-  $\chi$ & Preferential pathway contaminant concentration scaling parameter \\
+  $\chi$ & PP contaminant concentration scaling parameter \\
   % D
@@
 
-\bibliography{library}
+% TODO: Fix the natbib issues...
+%\bibliographystyle{model1-num-names}
+\bibliography{./library.bib}
 

### 2019-04-30 tex/main.tex
Working on more formatting

@@
 \usepackage[utf8]{inputenc}
+%\bibliographystyle{abbrvnat}
 \usepackage{amsmath}
@@
 \begin{table}[htb!]
+  \centering
  \begin{tabular}{c c c c c c c}
@@
 \begin{table}[htb!]
+  \centering
  \begin{tabular}{l c c c}
@@
 At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
+
+\section{Conclusions}
+
+Based on the statistical analysis of the field data presented, as well as the modeling efforts the following conclusions may be drawn.
+\begin{itemize}
+  \item Indoor air contaminant concentrations are unlikely to vary by more than a factor of three over a week-long period if the site is not characterized by a preferential pathway, such as the one found at the ASU house site; else more than an order magnitude variability within the same period may be expected.
+  \item Preferential pathways can cause significant spatial variability in the subslab.
+Subsequently, subslab vapor samples may reveal that the attenuation from subslab may significantly exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.
+  \item Preferential pathways such as the one found at the ASU house significantly increase the advective transport potential through the foundation breaches at a site.
+This relies on effective communication between the indoor air space and the preferential pathway, e.g.
+a permeable subslab region (gravel layer) must exist.
+  \item Sites characterized by significant advective potential, such as North Island NAS or the ASU house with the preferential pathway open, most of the short-term indoor air contaminant concentration variability may be attributed to fluctuations in indoor/outdoor pressure difference; at sites where the advective potential is low, e.g.
+ASU house with the preferential pathway closed, short-term variability is dominated by fluctuations in air exchange rate.
+\end{itemize}
 
@@
 % TODO: Fix the natbib issues...
-%\bibliographystyle{model1-num-names}
-\bibliography{./library.bib}
+\bibliographystyle{model1-num-names}
+\bibliography{library}
 

### 2019-04-30 tex/main.tex
Fixing tables

@@
 \usepackage{tabularx}
-\usepackage{array,multirow,graphicx}
+\usepackage{array,multirow,multicol,graphicx}
 \usepackage{comment}
@@
 The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation, thus increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
+During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation: increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
 The period of CPM testing is thus excluded from the analysis.
@@
 
-% TODO: Restructure entire table
+% new table
+% TODO: Fix the too wide issue
+\clearpage
+\begin{table}
+  \centering
+  \begin{tabular}{l c c c c c c}
+    \toprule
+    \multicolumn{7}{c}{\textbf{Governing Equations}} \\
+    \midrule
+    % CSTR
+    Unsteady-CST & \multicolumn{6}{c}{$V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$} \\
+    % RICHARDS
+    Richard's & \multicolumn{6}{c}{$\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$} \\
+    % TRANSPORT
+    Transport & \multicolumn{6}{c}{$\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$} \\
+    % MILLINGTON-QUIRK
+    Millington-Quirk & \multicolumn{6}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
+    % VAN GENUCHTEN
+    \multirow{4}{*}{van Genuchten's} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
+     & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
+     & \multicolumn{6}{c}{$k_r = (1-\mathrm{Se})^{l}[1(\mathrm{Se}^{m})^m]^2$} \\
+     & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
+    \midrule
+    \multicolumn{7}{c}{\textbf{Boundary Conditions}} \\
+    \midrule
+    \textbf{Boundary} & \multicolumn{3}{c}{\textbf{Richard's Eqn.}} & \multicolumn{3}{c}{\textbf{Transport Eqn.}} \\
+    % Foundation crack
+    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
+    % Groundwater source
+    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
+    % Ground surface
+    Groundsurface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
+    % Preferential pathway
+    Preferential & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$} \\
+    \midrule
+    \multicolumn{7}{c}{\textbf{Soil \& gravel properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
+    \midrule
+    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
+    % GRAVEL
+    Gravel & $1.3 \cdot 10^{-9}$ & 1680 & 0.42 & 0.005 & 100 & 3.1 \\
+    % SANDY LOAM
+    Sandy Loam & $5.9 \cdot 10^{-13}$ & 1460 & 0.39 & 0.039 & 2.7 & 1.4 \\
+    \midrule
+    \multicolumn{7}{c}{\textbf{Building Properties}} \\
+    \midrule
+     & \textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
+     & 300 & 15 & 0.5 \\
+    \bottomrule
+  \end{tabular}
+  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}\label{tbl:eqns_bc_parameters}
+\end{table}
+
+\begin{comment}
 \clearpage
@@
   % Indoor air space equation
-  Unsteady-CST                            & $V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$                                                                  \\
+  Unsteady-CST                            &                                                                   \\
   % Richard's equation
@@
  \end{tabular}
- \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}\label{tbl:eqns_bc_parameters}
+
 \end{table}
+\end{comment}
+
+% TODO: Restructure entire table
 
@@
 \begin{table}[htb!]
+  \newcommand{\NameEntry}[1]{
+    \multicolumn{2}{c}{
+      \multirow{2}{*}{
+        \begin{minipage}{0.2\textwidth}
+          \centering
+          \textbf{#1}
+        \end{minipage}
+      }
+    }
+  }
   \centering
@@
     \toprule
-                                     & \multicolumn{2}{c|}{\textbf{North Island}} & \multicolumn{2}{c|}{\textbf{ASU, PP Open}} & \multicolumn{2}{c}{\textbf{ASU, PP Closed}}                      \\
+    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
+    \midrule
     Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
@@
 
-% TODO: Fix the natbib issues...
 \bibliographystyle{model1-num-names}

### 2019-04-30 tex/main.tex
Working more on the tables

@@
     % Foundation crack
-    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
+    (ck) & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
     % Groundwater source
-    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
+    (gw) & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
     % Ground surface
-    Groundsurface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
+    (atm) & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
     % Preferential pathway
-    Preferential & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$} \\
+    (pp) & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$} \\
     \midrule
-    \multicolumn{7}{c}{\textbf{Soil \& gravel properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
+    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
     \midrule
@@
     Sandy Loam & $5.9 \cdot 10^{-13}$ & 1460 & 0.39 & 0.039 & 2.7 & 1.4 \\
+    % TCE PROPERTIES
+    \midrule
+    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
+    \midrule
+     & \textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} & \textbf{$M \; \mathrm{(g/mol)}$} \\
+     & $2.47 \cdot 10^{-2}$ & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 & 131.39 \\
     \midrule

### 2019-05-01 tex/cover-letter/cover_letter.tex
Ticking off more things..

@@
 
-\begin{letter}{Editor of Environmental Science \& Technology \\
-1155 Sixteenth Street N.W. \\
-Washington, DC 20036}
+\begin{letter}{Editorial Department of Building and Environment}
 \opening{Dear Sir or Madam:}
 
-I am pleased to submit this manuscript entitled "Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Environmental Science \& Technology.
-This is original work, and has not been previously submitted to any other journal or conference proceedings.
+I am pleased to submit this manuscript entitled "Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Building and Environment.
+This is original work, and has not been previously published by any other journal or conference proceedings.
 
@@
 
-We found under which conditions one may expect to observe significant changes in indoor air concentrations, and when one should be especially be concerned with preferential pathways.
-Furthermore, we found how changes in indoor outdoor pressure difference and air exchange rate independently affect temporal variability in VI under different conditions. 
-These finding help understand and explain the temporal variation that has been observed at many VI sites, and giving practical guidance for assessment of these sites, something that we expect will be of great interest to your readership.
-
-All of the work presented is either created by us or properly referenced, including the TOC art which adhere to the TOC art policy and is the prescribed size of 240 by 135 points.
+Some of our conclusions that we believe contribute to the state-of-the-art:
+\begin{itemize}
+  \item Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
+  \item Preferential pathways can dramatically increase a VI site's sensitivity to building pressurization.
+  \item Air exchange rate is poorly correlated with normal range of building pressurization and may significantly contribute to indoor contaminant variability.
+  \item Indoor contaminant concentration vary roughly by up to a factor of two across a few days; up to an order magnitude variation may be expected over a week.
+\end{itemize}
+We expect the topics explored and conclusions reached in our work will be of great interest to your readership.
 

### 2019-05-01 tex/main.tex
Ticking off more things..

--- 2019-04-30 tex/main.tex
@@
 
-% keywords
-%\abbreviations{VI}
-%\keywords{}
-
-\journal{Building \& Environment}
+\journal{Building and Environment}
 
@@
 We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
-For a preferential pathway to impact a VI site, there must exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
+A preferential pathway can dramatically increase a VI site's sensitivity to build pressurization, provided there exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
 a permeable subslab space that may be provided by a gravel layer.
+Preferential pathways may also erroneously indicate the presence of indoor contaminant sources.
 At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
@@
 \begin{keyword}
-  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Finite element modeling \sep Air exchange rate \sep Indoor/outdoor pressure difference
+  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Attenuation factor \sep Air exchange rate \sep Indoor/outdoor pressure difference
 \end{keyword}
@@
 
-\section{Methods}
+\section{Methods}\label{s:methods}
 
-\subsection{Statistical Analysis Of Field Data}
+\subsection{Statistical Analysis Of Field Data}\label{s:methods_stats}
 
@@
 
-\begin{figure}
+\begin{figure}[htb!]
   \centering
@@
 
-\subsection{Modeling Work}
+\subsection{Modeling Work}\label{s:methods_model}
 
@@
 
-% new table
-% TODO: Fix the too wide issue
-\clearpage
-\begin{table}
+\begin{table}[htb!]
+  \setlength{\tabcolsep}{1pt}
   \centering
   \begin{tabular}{l c c c c c c}
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % GOVERNING EQUATIONS
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \toprule
@@
     % VAN GENUCHTEN
-    \multirow{4}{*}{van Genuchten's} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
+    \multirow{4}{*}{van Genuchten} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
      & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
@@
      & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
+     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+     % BOUNDARY CONDITIONS
+     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \midrule
@@
     % Ground surface
-    Groundsurface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
+    Ground surface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
     % Preferential pathway
-    Preferential & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$} \\
+    \multirow{2}{*}{\begin{minipage}{0.2\textwidth}Preferential\\pathway\end{minipage}} & \multicolumn{3}{c}{\multirow{2}{*}{$p = 0 \; \mathrm{(Pa)}$}} & \multicolumn{3}{c}{\multirow{2}{*}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$}} \\ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % SOIL PROPERTIES
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \midrule
-    \multicolumn{7}{c}{\textbf{Soil \& gravel properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
+    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
     \midrule
-    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
+    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
     % GRAVEL
-    Gravel & $1.3 \cdot 10^{-9}$ & 1680 & 0.42 & 0.005 & 100 & 3.1 \\
+    Gravel & $1.3 \cdot 10^{-9}$ & 0.42 & 0.005 & 100 & 3.1 \\
     % SANDY LOAM
-    Sandy Loam & $5.9 \cdot 10^{-13}$ & 1460 & 0.39 & 0.039 & 2.7 & 1.4 \\
+    Sandy Loam & $5.9 \cdot 10^{-13}$ & 0.39 & 0.039 & 2.7 & 1.4 \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % TCE PROPERTIES
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \midrule
+    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
+    \midrule
+     & \multicolumn{1}{c}{\textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$}} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} \\
+     & \multicolumn{1}{c}{$2.47 \cdot 10^{-2}$} & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    % BUILDING PROPERTIES
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \midrule
@@
     \midrule
-     & \textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
-     & 300 & 15 & 0.5 \\
+     & \multicolumn{1}{c}{\textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$}} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
+     & \multicolumn{1}{c}{300} & 15 & 0.5 \\
     \bottomrule
@@
 
-\begin{comment}
-\clearpage
-\begin{table}[htb!]
- \centering
- \bigskip
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- % Governing equations
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subcaption{Governing equations}
- \begin{tabular}{l l}
-  \toprule
-  % Indoor air space equation
-  Unsteady-CST                            &                                                                   \\
-  % Richard's equation
-  Richard's equation                       & $\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$                                                               \\
-  % Millington-Quirk equation
-  Millington-Quirk                         & $D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$ \\
-  % Advection-diffusion equation
-  Advection-diffusion equation             & $\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$ \\
-  % van Genuchten's equations
-  \multirow{3}{*}{van Genuchten equations} & $\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$                                             \\
-                                           & $\theta_g = \theta_t - \theta_w$                                                                                                      \\
-                                           & $k_r = (1 - \mathrm{Se})^{l} [1 - (\mathrm{Se}^{-m})^m]^2$                                                                            \\
-                                           & $m = 1 - 1/n$                                                                                                                         \\
-  \bottomrule
- \end{tabular}
- \bigskip
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- % Boundary conditions
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subcaption{Boundary conditions}
- \begin{tabular}{l l l}
-  \toprule
-  \textbf{Boundary}            & \textbf{Richard's equation}              & \textbf{Transport equation}                                      \\
-  % Foundation crack
-  Foundation crack          & $p = p_\mathrm{in/out} \; \mathrm{(Pa)}$ & $j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$ \\
-  % Groundwater source
-  Groundwater source        & \textit{No flow}                                      & $c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$                            \\
-  % Ground surface
-  Groundsurface            & $p = 0 \; \mathrm{(Pa)}$                 & $c = 0 \; \mathrm{(\mu g/m^3)}$                                            \\
-  % Preferential pathway
-  Preferential pathway & $p = 0 \; \mathrm{(Pa)}$                 & $c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$                       \\
-  \bottomrule
- \end{tabular}
- \bigskip
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- % Soil input parameters
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subcaption{Soil \& gravel properties\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
- \begin{tabular}{l l l l l l l}
-  \toprule
-  % Descriptions
-  Soil       & $\kappa_s \; \mathrm{(m^2)}$ & $\rho \; \mathrm{(kg/m^3)}$ & $\theta_s$ & $\theta_r$ & $\alpha \; \mathrm{(1/m)}$ & $n$ \\
-  % Gravel
-  Gravel     & $1.3 \cdot 10^{-9}$                     & 1680                                    & 0.42       & 0.005      & 100                        & 3.1 \\
-  % sandy loam
-  Sandy Loam & $5.9 \cdot 10^{-13}$                    & 1460                                    & 0.39       & 0.039      & 2.7                        & 1.4 \\
-  \bottomrule
- \end{tabular}
- \bigskip
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- % TCE input parameters
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subcaption{Trichloroethylene (diluted in air) properties\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}}
- \begin{tabular}{l l l l l l}
-  \toprule
-  $D_\mathrm{air} \; \mathrm{(m^2/h)}$ & $D_\mathrm{water} \; \mathrm{(m^2/h)}$ & $\rho \; \mathrm{(kg/m^3)}$ & $\mu \; \mathrm{(Pa \cdot s)}$ & $K_H$ & $M \; \mathrm{(g/mol)}$ \\
-  $2.47 \cdot 10^{-2}$                 & $3.67 \cdot 10^{-6}$                   & 1.614                                   & $1.86 \cdot 10^{-5}$                          & 0.403 & 131.39                  \\
-  \bottomrule
- \end{tabular}
- \bigskip
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- % Building input parameters
- %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
- \subcaption{Building properties}
- \begin{tabular}{l l l}
-  \toprule
-  $V_\mathrm{base} \; \mathrm{(m^3)}$ & $L_\mathrm{slab} \; \mathrm{(cm)}$ & $A_e \; \mathrm{(1/hr)}$ \\
-  %
-  300                                 & 15                                 & 0.5                      \\
-  \bottomrule
- \end{tabular}
-
-\end{table}
-\end{comment}
-
-% TODO: Restructure entire table
-
+\newpage
 \section{Results \& Discussion}\label{s:results}
@@
 
+\begin{table}[htb!]
+  \newcommand{\NameEntry}[1]{
+    \multicolumn{2}{c}{
+      \multirow{2}{*}{
+        \begin{minipage}{0.2\textwidth}
+          \centering
+          \textbf{#1}
+        \end{minipage}
+      }
+    }
+  }
+  \centering
+  \begin{tabular}{c c c c c c c}
+    \toprule
+    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
+    \midrule
+    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
+    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
+    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
+    \bottomrule
+  \end{tabular}
+ \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
+\end{table}
+
 In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
@@
 
-\begin{table}[htb!]
-  \newcommand{\NameEntry}[1]{
-    \multicolumn{2}{c}{
-      \multirow{2}{*}{
-        \begin{minipage}{0.2\textwidth}
-          \centering
-          \textbf{#1}
-        \end{minipage}
-      }
-    }
-  }
-  \centering
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
-    \midrule
-    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
-    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
-    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
-    \bottomrule
-  \end{tabular}
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
-\end{table}
-
 \subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
@@
 
-\subsection{Modeling Results}\label{s:results_modeling}
+\subsection{Modeling Results}\label{s:results_model}
 
-\subsubsection{Pressure Effects}\label{s:results_modeling_pressure}
+\subsubsection{Pressure Effects}\label{s:results_model_pressure}
 
@@
 
-\subsubsection{Air Exchange Rate Effects}\label{s:results_modeling_air_exchange}
+\subsubsection{Air Exchange Rate Effects}\label{s:results_model_air_exchange}
 
@@
 Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in the supplementary data show (Figure S1), there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
+As the data in Figure \ref{fig:air_exchange_rate} show, there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
 Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
+
+\begin{figure}[htb!]
+  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
+  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
+The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
+Only non-CPM period considered.}\label{fig:air_exchange_rate}
+\end{figure}
 
@@
 
-\appendix
-
 \clearpage
 \begin{table}[htb!]
-  \caption{List of abbreviations}\label{tbl:abbreviations}
+  \centering
   \begin{tabular}{l l}
@@
   \end{tabular}
+  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
 \end{table}

### 2019-05-02 tex/manuscript.tex
Rename

--- 2019-05-01 tex/main.tex
@@
 % title
-\title{Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
+\title{Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
 
@@
   \end{tabular}
-  \caption{Governing equations, boundary conditions \& model input parameters. (See below for table of nomenclature).}\label{tbl:eqns_bc_parameters}
+  \caption{Governing equations, boundary conditions \& model input parameters. See Table \ref{tbl:abbreviations} for nomenclature.}\label{tbl:eqns_bc_parameters}
 \end{table}
@@
 
-Based on the statistical analysis of the field data presented, as well as the modeling efforts the following conclusions may be drawn.
-\begin{itemize}
-  \item Indoor air contaminant concentrations are unlikely to vary by more than a factor of three over a week-long period if the site is not characterized by a preferential pathway, such as the one found at the ASU house site; else more than an order magnitude variability within the same period may be expected.
-  \item Preferential pathways can cause significant spatial variability in the subslab.
-Subsequently, subslab vapor samples may reveal that the attenuation from subslab may significantly exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.
-  \item Preferential pathways such as the one found at the ASU house significantly increase the advective transport potential through the foundation breaches at a site.
-This relies on effective communication between the indoor air space and the preferential pathway, e.g.
-a permeable subslab region (gravel layer) must exist.
-  \item Sites characterized by significant advective potential, such as North Island NAS or the ASU house with the preferential pathway open, most of the short-term indoor air contaminant concentration variability may be attributed to fluctuations in indoor/outdoor pressure difference; at sites where the advective potential is low, e.g.
-ASU house with the preferential pathway closed, short-term variability is dominated by fluctuations in air exchange rate.
-\end{itemize}
+Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
+
+There can be a significant role of advective transport in determining observed variability.
+For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
+In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
+In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
+Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
+
+Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
+In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
+This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
+This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
+
+In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
+Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
+The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
+It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
+These results were consistent with seasonal timescale variations in the intrusion processes.\par
 

### 2019-05-02 tex/cover-letter/cover_letter.tex
New submission

@@
 \usepackage[utf8]{inputenc}
+\usepackage[textwidth=6in,textheight=14in]{geometry}
 \signature{Jonathan Ström}
@@
 
-I am pleased to submit this manuscript entitled "Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Building and Environment.
-This is original work, and has not been previously published by any other journal or conference proceedings.
+I am pleased to submit this manuscript entitled "Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Building and Environment.
+This is original work and has not been previously published by any other journal or conference proceedings; the submission declaration has been complied with.
+We have no interests to declare.
 
@@
   \item Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
-  \item Preferential pathways can dramatically increase a VI site's sensitivity to building pressurization.
-  \item Air exchange rate is poorly correlated with normal range of building pressurization and may significantly contribute to indoor contaminant variability.
-  \item Indoor contaminant concentration vary roughly by up to a factor of two across a few days; up to an order magnitude variation may be expected over a week.
+  \item Preferential pathways can dramatically increase the role of advective transport, provided there is a permeable subslab region to facilitate this.
+  \item At VI sites where advective transport potential is high, much of the observed variation may be attributed to fluctuations in building pressurization; where this potential is much lower, fluctations in air exchange rate may significantly contribute to indoor contaminant variability.
+  \item In the absence of a preferential pathway, indoor contaminant concentration vary roughly by up to a factor of three across a few days; up to an order magnitude variation may be expected over a week.
 \end{itemize}

### 2019-05-02 tex/highlights/highlights.tex
New submission

--- 2019-05-02 tex/cover-letter/cover_letter.tex
@@
-\documentclass{letter}
-\usepackage{hyperref}
+\documentclass{article}
+
 \usepackage[utf8]{inputenc}
-\usepackage[textwidth=6in,textheight=14in]{geometry}
-\signature{Jonathan Ström}
-\address{School of Engineering \\ Brown University \\ Providence 02912}
+
 \begin{document}
+\section*{Highlights}
 
-\begin{letter}{Editorial Department of Building and Environment}
-\opening{Dear Sir or Madam:}
-
-I am pleased to submit this manuscript entitled "Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations" for review and possible publication in Building and Environment.
-This is original work and has not been previously published by any other journal or conference proceedings; the submission declaration has been complied with.
-We have no interests to declare.
-
-Using numerical models and published field data from three vapor intrusion (VI) sites, we explore the causes of the temporal variation in indoor air contaminant concentrations in VI impacted houses, a poorly understood phenomena.
-We especially investigate the role of preferential contaminant pathways have in VI, and how air exchange rates contribute to variability in indoor air contaminant concentration.
-
-Some of our conclusions that we believe contribute to the state-of-the-art:
 \begin{itemize}
@@
 \end{itemize}
-We expect the topics explored and conclusions reached in our work will be of great interest to your readership.
-
-Thank you for considering this manuscript for publication.
-\closing{Sincerely,}
-
-\end{letter}
 \end{document}

### 2019-05-02 tex/manuscript.tex
New submission

@@
 \section*{Acknowledgements}
-This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).
+This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).\par
+
+Declaration of interest: none
 
@@
 
-\bibliographystyle{model1-num-names}
+\bibliographystyle{elsarticle-num}
 \bibliography{library}

### 2019-06-03 tex/manuscript.tex
Added practical implications notes.

@@
 
+% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
+- 24 hr sampling is recommended for a few reasons.
+1. Avoids issues with diurnal fluctations in Ae or dP.
+2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
+- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
+- Passive samples can help capture some of the more longterm variations if these are a concern.
+- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
+Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
+- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
+- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
+1. Even with "good" subslab sampling, hotspots are easily missed.
+2. May mislead that an indoor source is present.
+3. Even without a PP there may still be significant spatial variability (Eklund).
+4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
+5. McAlary said something about this. Look it up.
+6. May not really be worth the effort.
+
 \section*{Acknowledgements}

### 2019-06-04 tex/manuscript.tex
Added implications

@@
 In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
+Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.\par
+
+These findings, as well as those from others reaffirm that screening for preferential pathways should be a routine part of VI investigations.
+Sewers have been shown to play significant roles as preferential pathways and sampling nearby manholes for the presence for contaminants is a simple way to determine if a preferential pathway may play a role at the site.
+Similar and other suggested screening methods have been described by Nielsen et al.\cite{nielsen_remediation_2017}, Pennell et al.\cite{pennell_sewer_2013}, and McHugh et al.\cite{mchugh_evidence_2017}.\par
+
+Even in the absence of a preferential pathway, monitoring the indoor/outdoor pressure difference at a site may be we wise.
+As was observed at North Island NAS, the significant fluctuations in pressurization lead to significant fluctuations in indoor air contaminant concentrations.
+Knowing how the pressurization fluctuates may help inform how much temporal variability one may expect at a VI site; larger fluctuations combined with relatively permeable soils and/or leaky buildings lead to more significant temporal variability in indoor air contaminant concentrations, requiring indoor samples to be taken with a finer time-resolution than usual.
+Under these circumstances, there may be need for continuous monitoring of indoor air contaminant concentration.\par
 
@@
 This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
+
+This adds to the concerns of overly relying on subslab contaminant samples in VI investigations due to the significant spatial variability of vapor contaminants at VI site like the ASU house and others\cite{chow_concentration_2007,folkes_observed_2009,u.s._environmental_protection_agency_assessment_2015}.
+This is further complicated by the fact that the building itself can influence the subslab and create plumes of contaminant vapor\cite{holton_creation_2018}.
+Collecting subslab samples can yield important information, but site investigators should be aware of the potential issues and ask themselves if it is worth the effort to collect these samples.\par
 
@@
 
+This suggests that 24-hour indoor air samples may be the most appropriate choice of sample collection time, avoiding diurnal concerns as well as capturing the variability of a few factors over a day.
+Samples collected over longer time-periods, i.e. 48- or 72-hours, are unlikely to yield a different concentration than a 24-hour samples.
+For the same reason, multiple samples should be spaced several days or few weeks apart to avoid duplicate samples.
+Passive samples could be a cost effective way in conjunction with 24-hour samples to understand the long-term indoor air contaminant concentration trends.
+The analysis in Figure \ref{fig:resampling} also implies that if a 24-hour sample gives a indoor air contaminant concentration that deceeds some regulatory limit by an order of magnitude or more, it is unlikely that the indoor air contaminant concentration will exceed this value at a later date; further sampling may not be necessary.\par
+
+\begin{comment}
 % PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
@@
 6. May not really be worth the effort.
+\end{comment}
 

### 2019-06-04 tex/manuscript.tex
Fixed overfull hbox
Added Eric's edits

--- 2019-06-03 tex/manuscript.tex
@@
 
-Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
-
+Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.
 There can be a significant role of advective transport in determining observed variability.
@@
 In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
+Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.\par
 
-Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
-In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
-This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
-This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
+These findings, as well as those from others reaffirm that screening for preferential pathways should be a routine part of VI investigations.
+Sewers have been shown in several studies to play significant roles as potential preferential pathways.
+Thus, sampling nearby manholes for the presence for contaminants appears prudent in terms of establishing whether such a preferential pathway may play a role at a site.
+This is consistent with what has been suggested by Nielsen et al.\cite{nielsen_remediation_2017}, Pennell et al.\cite{pennell_sewer_2013}, and McHugh et al.\cite{mchugh_evidence_2017}.\par
 
-In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
-Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
-The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
+Even in the absence of a preferential pathway, the present results suggest that some temporal monitoring of indoor/outdoor pressure differences can provide insights regarding whether an unusually large advective driving force for contaminant entry exists.
+As was observed at NAS North Island, significant fluctuations in pressure driving force was what lead to significant temporal fluctuations in indoor air contaminant concentrations.
+Generally speaking, larger pressure fluctuations on sites with relatively permeable soils and/or leaky buildings would be expected to lead to more significant short-term temporal variability in indoor air contaminant concentrations, suggesting monitoring that can track such variations.\par
+
+Modeling work has shown that the existence of preferential pathways that transport contaminant to a permeable sub-slab zone can lead to significant spatial variability in contaminant concentration in the subslab, even if the subslab is reasonably permeable (such as a gravel sub-base).
+The presumption of a well-mixed sub-base is unlikely to be appropriate in the presence of a preferential pathway into the sub-base.
+This warns that taking spatially limited subslab vapor samples may provide a misleading picture of the potential for contaminant entry.
+Such limited sampling in such situations could even lead to apparent to apparent subslab to indoor air attenuation factors exceeding the typical EPA recommended value of 0.03 (and they can even exceed unity, potentially leading to an erroneous conclusion that an indoor source is present).
+This is in addition to other warnings that have been previously presented regarding over-reliance on subslab samples in VI investigations\cite{chow_concentration_2007,folkes_observed_2009,u.s._environmental_protection_agency_assessment_2015,pennell_development_2009}.
+The fact that the building itself can influence the subslab contaminant concentrations has also been the subject of caution\cite{holton_creation_2018}.
+Thus, the list of reasons for exercising caution in relying upon subslab vapor measurements continues to grow.\par
+
+In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of the observed modest stack effect pressure driving force, a temporal variation in indoor air contaminant concentration of roughly an order of magnitude was still observed.
+Statistical analysis of that dataset showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
+The small variability that was seen over these short times could be explained by the inherent variability in air exchange rates.
 It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
-These results were consistent with seasonal timescale variations in the intrusion processes.\par
+These results were consistent with seasonal timescale variations in the intrusion processes.
+The above results suggest that 24-hour indoor air sampling may be appropriate for capturing the variability that actually exists over the timescale of days.
+Considering sampling over longer time-periods, i.e.
+48- or 72-hours, the results seemed to indicate that they would not give different concentration than a 24-hour sample.
+For the same reason, multiple rounds of sampling over the time period of a few days or weeks cannot normally be expected to reveal significant differences, unless one is dealing with a preferential pathway situation which can lead to variations on shorter timescales.\par
 
+The analysis in Figure \ref{fig:resampling} also implies that if a 24-hour sample gives an indoor air contaminant concentration that is an order of magnitude lower than a relevant regulatory limit, it is unlikely that the indoor air contaminant concentration will exceed this value at a later date unless there exists a preferential pathway whose contribution varies significantly with time.\par
+
+\begin{comment}
 % PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)

### 2019-06-04 tex/reviewer_response/reponse.tex
Revision

\documentclass{letter}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[textwidth=6in,textheight=14in]{geometry}
\signature{Jonathan Ström}
\address{School of Engineering \\ Brown University \\ Providence 02912}
\begin{document}

\begin{letter}{ }
\opening{Dear reviewers:}

We thank the reviewers for their efforts in reviewing our paper, and have
responded to their questions and comments as outlined below.

Reviewer \#1: This paper studied temporal variations of indoor contaminant concentrations
due to Vapor Intrusion by analyzing the field data and performing simulations. Some
interesting conclusions were obtained from the analysis of data. The method seems
appropriate. Therefore, I suggest the paper to be accepted by addressing the following
concerns:

\textit{We thank the reviewer for the positive assessment of our paper.}

Comment 1. The graphic abstract should be more self-contained. Necessary explanations
are suggested.

\textit{Since the graphical abstract is not an element required by this journal, we decided that since
it seemed to raise more questions than it answered, it would be better to remove it entirely,
rather than add even more complexity to it.}

Comment 2. Please provide practical implications of the obtained conclusions. For example, how to
detect or reduce peak concentration to protect human health?

\textit{We have done so as part of a significantly expanded "Conclusions" section, and thank the
reviewer for suggesting this useful addition.}

Reviewer \#2: In my opinion it is a remarkably prepared material. During almost of 10 years of
my experience as a reviewer I have never reviewed such an outstanding manuscript, both in
terms of the structure and balanced proportion between the complexity and the way of
presentation of material as well as the clarity of narration. It should be particularly
appreciated due to the fact that presented material relates to quite complex issue like a long-
term data elaboration and interpretation, which is even more difficult taking into account all
possible factors influencing obtained results.
I could not find any lingual, grammatical or punctuation slip. I strongly recommend to accept
this manuscript for publication.

\textit{What can we say but thank you to this reviewer!}

\closing{Sincerely,}

\end{letter}
\end{document}

### 2019-06-14 tex/highlights/highlights.tex
Added new highlights

@@
 \end{document}
+
+Preferential pathways may lead to the false conclusion that there are indoor sources present at a VI site.
+Indoor concentrations vary roughly by a factor of three over a few days; up to an order of magnitude over a week.
+Preferential pathways can dramatically increase the role of advective transport, provided there is a permeable subslab zone.
+At sites with high advective transport potential, fluctuations in pressurization cause much of the observed variability.
+Air exchange rate fluctuations significantly contribute with variability at sites with low advective transport potential.

### 2019-09-24 doc/conclusions.tex
Reorganizing

--- 2019-06-04 tex/manuscript.tex
@@
-\documentclass[preprint,12pt]{elsarticle}
-
-% additional packages
-\usepackage[utf8]{inputenc}
-%\bibliographystyle{abbrvnat}
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{subcaption}
-\usepackage{tabularx}
-\usepackage{array,multirow,multicol,graphicx}
-\usepackage{comment}
-\graphicspath{
-  {./figures/},
-}
-\usepackage{lineno}
-\linenumbers
-% macros
-
-\journal{Building and Environment}
-
-\begin{document}
-
-\begin{frontmatter}
-% title
-\title{Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
-
-% authors
-\author[Brown]{Jonathan G. V. Ström}
-\author[ASU]{Yuanming Guo}
-\author[ASU]{Yijun Yao}
-\author[Brown]{Eric M. Suuberg}
-
-\ead{eric\_suuberg@brown.edu}
-
-\address[Brown]{Brown University, School of Engineering, Providence, RI, USA}
-\address[ASU]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
-
-\begin{abstract}
-Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
-We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
-We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
-A preferential pathway can dramatically increase a VI site's sensitivity to build pressurization, provided there exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
-a permeable subslab space that may be provided by a gravel layer.
-Preferential pathways may also erroneously indicate the presence of indoor contaminant sources.
-At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
-Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
-\end{abstract}
-
-\begin{keyword}
-  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Attenuation factor \sep Air exchange rate \sep Indoor/outdoor pressure difference
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{s:intro}
-
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-Such variations make it difficult for those charged with protecting human health to formulate a response - should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
-There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-What represents a reasonable sampling strategy for a particular site a single 8-hour sample?
-Repeated 8-hour samples?
-Month-long samples?
-Continuous monitoring?\par
-
-VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-The basic nature of VI has been understood for some time and it has been the subject of much study, but some aspects remain poorly understood, such as the causes of the sometimes observed large temporal transients in indoor air concentrations.
-Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
-All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.\par
-
-In one instance the source of the variation was clearly established during the study of the site.
-At the ASU house a drain pipe (or “land drain”) connected to a sewer system was discovered beneath the house.
-Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.\par
-
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
-That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been previously implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.\par
-
-In some of the above cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many instances, this scenario is not further considered here where the focus is on pathways that deliver contaminant via the soil beneath a structure.
-It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is the site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.\par
-
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as “normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at such normal VI sites vs.
-those characterized by preferential pathways within the soil beneath the site.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
-
-\section{Methods}\label{s:methods}
-
-\subsection{Statistical Analysis Of Field Data}\label{s:methods_stats}
-
-To frame the question of just how much variability in indoor air contaminant concentrations is actually observed, field datasets have been analyzed.
-For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.\par
-
-The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation: increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-The period of CPM testing is thus excluded from the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par
-
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to “large” internal pressure fluctuations.
-By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
-The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.\par
-
-The Indianapolis site investigation also spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD) for VI mitigation.
-Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}.
-Unlike at the ASU house, this preferential pathway was never removed or blocked, making it impossible to isolate the role of the preferential pathway at this site.
-It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
-Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
-There is almost a two order of magnitude variation in the concentration data.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
- \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
-\end{figure}
-
-Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
-KDE is a technique used for estimating the probability distribution of a random variable(s) by using multiple kernels, or weighting functions to characterize the data sets.
-In this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values and that there are statistical fluctuations associated with each sampling event.
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.\par
-
-\subsection{Modeling Work}\label{s:methods_model}
-
-A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
-In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{model.png}
- \caption{Foundation and vadose zone soil represented in the modeling.
-Note that here a gravel sub-base material is shown, but in certain simulations, that material is absent and the surrounding soil directly contacts the foundation slab.
-Different assumptions are made regarding the preferential pathway, here shown as a pipe entering the gravel sub-base.
-In some cases, the preferential pathway has been "turned off".}\label{fig:model}
-\end{figure}
-
-The modeled VI impacted structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the  perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is here assumed to consist of only of the basement, having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of some critical parameters is.\par
-
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house and is assumed to consist of sandy loam, except as noted otherwise.
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases here this sub-base material is assumed to be the same as the surrounding soil (termed a "uniform" soil scenario).
-The groundwater beneath the structure is assumed to be homogeneously contaminated with TCE selected as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-Where relevant, the preferential pathway is modeled as a 10 cm (4”) pipe that opens into the gravel sub-base beneath the structure.
-The air in the pipe is also assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$ (allowing the contaminant concentration in the pipe to be parameterized).
-This model illustrates the concept of a "preferential pathway", as the pipe carries contaminant vapor to the immediate vicinity of the foundation, by a path that circumvents the usual soil diffusion pathway.\par
-
-The ground surface and the pipe are both sources of air to the soil domain.
-Both are assumed to exist at reference atmospheric pressure.
-Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
-The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
-The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par
-
-\begin{table}[htb!]
-  \setlength{\tabcolsep}{1pt}
-  \centering
-  \begin{tabular}{l c c c c c c}
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % GOVERNING EQUATIONS
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \toprule
-    \multicolumn{7}{c}{\textbf{Governing Equations}} \\
-    \midrule
-    % CSTR
-    Unsteady-CST & \multicolumn{6}{c}{$V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$} \\
-    % RICHARDS
-    Richard's & \multicolumn{6}{c}{$\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$} \\
-    % TRANSPORT
-    Transport & \multicolumn{6}{c}{$\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$} \\
-    % MILLINGTON-QUIRK
-    Millington-Quirk & \multicolumn{6}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
-    % VAN GENUCHTEN
-    \multirow{4}{*}{van Genuchten} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
-     & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
-     & \multicolumn{6}{c}{$k_r = (1-\mathrm{Se})^{l}[1(\mathrm{Se}^{m})^m]^2$} \\
-     & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-     % BOUNDARY CONDITIONS
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Boundary Conditions}} \\
-    \midrule
-    \textbf{Boundary} & \multicolumn{3}{c}{\textbf{Richard's Eqn.}} & \multicolumn{3}{c}{\textbf{Transport Eqn.}} \\
-    % Foundation crack
-    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
-    % Groundwater source
-    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
-    % Ground surface
-    Ground surface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
-    % Preferential pathway
-    \multirow{2}{*}{\begin{minipage}{0.2\textwidth}Preferential\\pathway\end{minipage}} & \multicolumn{3}{c}{\multirow{2}{*}{$p = 0 \; \mathrm{(Pa)}$}} & \multicolumn{3}{c}{\multirow{2}{*}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$}} \\ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % SOIL PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
-    % GRAVEL
-    Gravel & $1.3 \cdot 10^{-9}$ & 0.42 & 0.005 & 100 & 3.1 \\
-    % SANDY LOAM
-    Sandy Loam & $5.9 \cdot 10^{-13}$ & 0.39 & 0.039 & 2.7 & 1.4 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % TCE PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$}} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} \\
-     & \multicolumn{1}{c}{$2.47 \cdot 10^{-2}$} & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % BUILDING PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Building Properties}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$}} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
-     & \multicolumn{1}{c}{300} & 15 & 0.5 \\
-    \bottomrule
-  \end{tabular}
-  \caption{Governing equations, boundary conditions \& model input parameters. See Table \ref{tbl:abbreviations} for nomenclature.}\label{tbl:eqns_bc_parameters}
-\end{table}
-
-\newpage
-\section{Results \& Discussion}\label{s:results}
-
-\subsection{Variation In Indoor Air Contaminant Concentration Over Time}\label{s:results_variation_time}
-
-High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
-It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par
-
-Figure \ref{fig:indianapolis} showed a large degree of temporal variation in one of the components, and the data for the other components were quite similar.
-What is apparent upon closer examination of such data is that the actual day-to-day variations are typically not nearly as large as those observed when tracking the data for a longer time.
-To demonstrate this point, the quotient of the maximum and minimum $c_\mathrm{in}$ values (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are shown as a function of time in Figure \ref{fig:resampling}.
-The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
-Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{resampling.pdf}
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected over a given time period.
-(e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months).
-The error bars are the 95\% confidence intervals.}\label{fig:resampling}
-\end{figure}
-
-For example, if the data are examined in terms of the mean maximum variation observable over the course of 24 hours (one day) the variation is no greater than about a factor of two for any of the contaminants at the Indianapolis house or for TCE at the ASU house (when the preferential pathway was closed).
-The mean variability at the latter was only a bit higher (about a factor of 3) when the preferential pathway was open.
-In other words, a sampling protocol that involves sampling on two consecutive days would typically not uncover the large temporal variations that characterize the site over longer periods of time.
-As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
-So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.\par
-
-Weeks of temporal separation in sampling events are required to observe the large variations of concern.
-Orders of magnitude differences begin to manifest themselves over the course of months.
-This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
-This would be consistent with requiring months to see the more significant variations.\par
-
-This analysis also suggests that certain types of preferential pathways contribute to larger variations on shorter timescales (ASU House).
-Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
-This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.\par
-
-The longer the resampling period, the larger the maximum variability in observed indoor air contaminant concentrations.
-In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
-Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.\par
-
-Multiple samples taken over a short time period, e.g.
-a few days, are unlikely to uncover significant variation in indoor air contaminant concentration; the larger transient variations typically manifest after longer time periods.
-\par
-
-\subsection{Statistical Analysis of Field Data}\label{s:results_pressure_concentration}
-
-The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
-The rate of advective entry of soil gas into a structure is frequently cited as playing an important role in determining entry rate of contaminant.
-This advective entry rate is closely linked to the indoor-outdoor pressure difference, as can be caused by the “stack effect”, for example.
-Thus we first consider how much variability there might be in the pressure driving force for advection, and if this can explain the observed variability in observed indoor air contaminant concentrations.\par
-
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure. 
-Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).\par
-
-We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
-For this analysis we considered two VI sites, NAS North Island and the ASU House.
-The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
-By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}\label{fig:kde}
-\end{figure}
-
-\begin{table}[htb!]
-  \newcommand{\NameEntry}[1]{
-    \multicolumn{2}{c}{
-      \multirow{2}{*}{
-        \begin{minipage}{0.2\textwidth}
-          \centering
-          \textbf{#1}
-        \end{minipage}
-      }
-    }
-  }
-  \centering
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
-    \midrule
-    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
-    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
-    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
-    \bottomrule
-  \end{tabular}
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
-\end{table}
-
-In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
-A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.\par
-
-Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
-Such issues have of course already been pointed out by the investigators who obtained the data.\par
-
-The power of this KDE representation is that it permits evaluation of the relationship of two independently measured data - the indoor air contaminant concentration and the indoor-outdoor pressure difference.
-Examining the data in this manner immediately points to an important difference between the data from the ASU House and those from NAS North Island.
-At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).\par
-
-The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
-The Pearson’s r-value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
-But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.\par
-
-The ASU house datasets offer a different picture.
-The variability of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
-The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
-These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
-
-The data for the ASU House also offer an insight into the role of the preferential pathway.
-At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed are relatively comparable.
-However, the 5th and 95th percentiles values of $c_\mathrm{in}$/$c_\mathrm{in,mean}$ differ significantly as may be seen in Table \ref{tbl:percentiles}.
-It is clear that existence of the preferential pathway dramatically increases the variability in indoor air contaminant concentration.
-This again is entirely consistent with what the investigators of that site have already reported\cite{guo_identification_2015}.
-The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
-These will be explored with the help of a modeling analysis presented below.\par
-
-\subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
-
-Observed temporal variations in indoor air contaminant concentrations might be explained by temporal variations in subslab contaminant concentrations.
-To examine how variability in subslab contaminant concentration might contribute to variability in indoor air contaminant concentration, data on the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) were examined.
-The dataset utilized for this was that from the ASU House.
-The $c_\mathrm{subslab}$ values were taken from a soil gas probe labeled as "6" at the ASU house.
-This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
- \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
-\end{figure}
-
-It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed, large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
-
-When the PP was open, there was considerably more variability in the subslab concentration values, and the mean value was higher than in the case where the preferential pathway was closed.
-It was also not uncommon for the observed $\alpha_\mathrm{subslab}$ to exceed unity.
-While large $\alpha_\mathrm{subslab}$ values may sometimes indicate indoor sources at a site, there were none at the ASU house.
-A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
-This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.\par
-
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in "normal" VI scenarios.\par
-
-\subsection{Modeling Results}\label{s:results_model}
-
-\subsubsection{Pressure Effects}\label{s:results_model_pressure}
-
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
- \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
-Field data are binned in 40 evenly spaced pressure bins, with the dot representing the mean and errors bars the 95\% confidence interval of data at a particular pressure range.
-Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
-Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".}\label{fig:land_drain_scenarios}
-\end{figure}
-
-In the calculation results shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
-Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
-Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
-The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
-These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
-
-The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.\par
-
-The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
-First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
-Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.\par
-
-The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
-This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
-In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
-These requirements were met at the ASU House.\par
-
-A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
-In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
-The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
-The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.\par
-
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
-These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e.
-that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par
-
-The above simulations capture the trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ but do not yet capture the full variability of the concentration results over the "most probable" portion of observed pressure distributions shown in Figure \ref{fig:kde} (which tend to be from -2 to +2 Pa).
-The results of Figure \ref{fig:land_drain_scenarios} show a spread of almost an order of magnitude over this pressure range for the case of the "open" preferential pathway, and almost no spread at all when the preferential pathway is "closed".
-Hence the predicted variability is roughly an order of magnitude too low, when considering only the influence of pressure.
-There is a factor that tends to increase the spread of the data one additional order of magnitude beyond what was predicted by the base calculations of Figure \ref{fig:land_drain_scenarios}.
-We believe that it is variations in air exchange rate, operating in concert with the natural variations in pressure differential, that explain the remaining variability.\par
-
-\subsubsection{Air Exchange Rate Effects}\label{s:results_model_air_exchange}
-
-Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the above model.
-Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
-A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
-Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
-Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in Figure \ref{fig:air_exchange_rate} show, there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
-Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
-  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
-The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
-Only non-CPM period considered.}\label{fig:air_exchange_rate}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l c c c}
-    \toprule
-    \textbf{Percentile}                                                     & \textbf{10th} & \textbf{50th} & \textbf{90th} \\
-    EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}           & 0.16-0.2      & 0.35-0.49     & 1.21-1.49     \\
-    ASU house\cite{holton_temporal_2013,guo_identification_2015}            & 0.21          & 0.43          & 0.78          \\
-    Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34          & 0.74          & 1.27          \\
-    \bottomrule
-  \end{tabular}
-  \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
-\end{table}
-
-To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are indicated by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
-
-It is apparent that assuming variability in air exchange rate allows capturing most of the observed variability in $\alpha_\mathrm{gw}$.
-We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus, we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
-
-\subsubsection{Results of Transient Simulations}\label{s:results_modeling_transient}
-
-The above analyses have been conducted under simulated steady state conditions.
-The conclusions regarding the importance of the different parameters are now examined in actual transient simulations.
-The model configuration of Figure \ref{fig:model} is run in 24-hour transient simulations to examine how $c_\mathrm{in}$ fluctuates over the course  of a "typical" day.
-The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
-The ASU House dataset was again the source of the "typical" $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
-The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
-The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the air exchange rate value assumed for the calculations at constant air exchange rate.\par
-
-\begin{figure}
-  \centering
- \includegraphics[width=\textwidth]{diurnal.pdf}
- \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
-\end{figure}
-
-The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
-The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e.
-a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
-When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
-the contaminant transport into the modeled structure is advection dominated.
-Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
-When the preferential pathway is closed the story is quite different.
-When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
-Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
-When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
-This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
-
-These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
-It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g.
-changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
-For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
-For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
-We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
-At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
-
 \section{Conclusions}
@@
 The analysis in Figure \ref{fig:resampling} also implies that if a 24-hour sample gives an indoor air contaminant concentration that is an order of magnitude lower than a relevant regulatory limit, it is unlikely that the indoor air contaminant concentration will exceed this value at a later date unless there exists a preferential pathway whose contribution varies significantly with time.\par
-
-\begin{comment}
-% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
-- 24 hr sampling is recommended for a few reasons.
-1. Avoids issues with diurnal fluctations in Ae or dP.
-2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
-- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
-- Passive samples can help capture some of the more longterm variations if these are a concern.
-- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
-Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
-- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
-- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
-1. Even with "good" subslab sampling, hotspots are easily missed.
-2. May mislead that an indoor source is present.
-3. Even without a PP there may still be significant spatial variability (Eklund).
-4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
-5. McAlary said something about this. Look it up.
-6. May not really be worth the effort.
-\end{comment}
-
-\section*{Acknowledgements}
-This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).\par
-
-Declaration of interest: none
-
-\clearpage
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l l}
-    \toprule
-    % A/ALPHA
-    $A_\mathrm{ck}$ & Crack area \\
-    $A_e$ & Air exchange rate \\
-    $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
-    $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
-    % C
-    $c_\mathrm{in}$ & Indoor air contaminant concentration \\
-    $c$ & Soil-gas contaminant concentration \\
-    $c_w$ & Soil-water contaminant concentration \\
-    $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-    $\chi$ & PP contaminant concentration scaling parameter \\
-    % D
-    $D_\mathrm{eff}$ & Effective diffusion coefficient \\
-    $D_\mathrm{air}$ & Diffusion coefficient in air \\
-    $D_\mathrm{water}$ & Diffusion coefficient in water \\
-    % J
-    $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
-    % K
-    $\kappa_s$ & Saturated soil permeability \\
-    $K_H$ & Dimensionless Henry's law constant \\
-    $k_r$ & Relative permeability \\
-    % L
-    $L_\mathrm{slab}$ & Thickness of the foundation slab \\
-    % M
-    $M$ & Molar mass \\
-    $\mu$ & Contaminant vapor viscosity \\
-    % N
-    NAS & Naval Air Stations \\
-    % P
-    $p$ & Pressure in soil \\
-    $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
-    PP & Preferential pathway \\
-    % R
-    $\rho$ & Density \\
-    % S
-    $\mathrm{Se}$ & Soil water saturation \\
-    % T
-    $t$ & time \\
-    $\theta_g$ & Vapor/gas filled porosity \\
-    $\theta_w$ & Water filled porosity \\
-    $\theta_r$ & Residual water filled porosity \\
-    $\theta_t$ & Total porosity \\
-    % U
-    $\vec{u}$ & Soil-gas velocity (vector quantity) \\
-    % V
-    VI & Vapor intrusion \\
-    $V_\mathrm{base}$ & Basement volume \\
-    % Z
-    $z$ & Elevation above groundwater \\
-    \bottomrule
-  \end{tabular}
-  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
-\end{table}
-
-\bibliographystyle{elsarticle-num}
-\bibliography{library}
-
-\end{document}

### 2019-09-24 doc/intro.tex
Reorganizing

--- 2019-06-03 tex/manuscript.tex
@@
-\documentclass[preprint,12pt]{elsarticle}
-
-% additional packages
-\usepackage[utf8]{inputenc}
-%\bibliographystyle{abbrvnat}
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{subcaption}
-\usepackage{tabularx}
-\usepackage{array,multirow,multicol,graphicx}
-\usepackage{comment}
-\graphicspath{
-  {./figures/},
-}
-\usepackage{lineno}
-\linenumbers
-% macros
-
-\journal{Building and Environment}
-
-\begin{document}
-
-\begin{frontmatter}
-% title
-\title{Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
-
-% authors
-\author[Brown]{Jonathan G. V. Ström}
-\author[ASU]{Yuanming Guo}
-\author[ASU]{Yijun Yao}
-\author[Brown]{Eric M. Suuberg}
-
-\ead{eric\_suuberg@brown.edu}
-
-\address[Brown]{Brown University, School of Engineering, Providence, RI, USA}
-\address[ASU]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
-
-\begin{abstract}
-Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
-We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
-We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
-A preferential pathway can dramatically increase a VI site's sensitivity to build pressurization, provided there exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
-a permeable subslab space that may be provided by a gravel layer.
-Preferential pathways may also erroneously indicate the presence of indoor contaminant sources.
-At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
-Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
-\end{abstract}
-
-\begin{keyword}
-  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Attenuation factor \sep Air exchange rate \sep Indoor/outdoor pressure difference
-\end{keyword}
-
-\end{frontmatter}
-
 \section{Introduction}\label{s:intro}
@@
 The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
-
-\section{Methods}\label{s:methods}
-
-\subsection{Statistical Analysis Of Field Data}\label{s:methods_stats}
-
-To frame the question of just how much variability in indoor air contaminant concentrations is actually observed, field datasets have been analyzed.
-For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.\par
-
-The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation: increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-The period of CPM testing is thus excluded from the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par
-
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to “large” internal pressure fluctuations.
-By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
-The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.\par
-
-The Indianapolis site investigation also spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD) for VI mitigation.
-Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}.
-Unlike at the ASU house, this preferential pathway was never removed or blocked, making it impossible to isolate the role of the preferential pathway at this site.
-It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
-Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
-There is almost a two order of magnitude variation in the concentration data.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
- \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
-\end{figure}
-
-Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
-KDE is a technique used for estimating the probability distribution of a random variable(s) by using multiple kernels, or weighting functions to characterize the data sets.
-In this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values and that there are statistical fluctuations associated with each sampling event.
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.\par
-
-\subsection{Modeling Work}\label{s:methods_model}
-
-A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
-In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{model.png}
- \caption{Foundation and vadose zone soil represented in the modeling.
-Note that here a gravel sub-base material is shown, but in certain simulations, that material is absent and the surrounding soil directly contacts the foundation slab.
-Different assumptions are made regarding the preferential pathway, here shown as a pipe entering the gravel sub-base.
-In some cases, the preferential pathway has been "turned off".}\label{fig:model}
-\end{figure}
-
-The modeled VI impacted structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the  perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is here assumed to consist of only of the basement, having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of some critical parameters is.\par
-
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house and is assumed to consist of sandy loam, except as noted otherwise.
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases here this sub-base material is assumed to be the same as the surrounding soil (termed a "uniform" soil scenario).
-The groundwater beneath the structure is assumed to be homogeneously contaminated with TCE selected as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-Where relevant, the preferential pathway is modeled as a 10 cm (4”) pipe that opens into the gravel sub-base beneath the structure.
-The air in the pipe is also assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$ (allowing the contaminant concentration in the pipe to be parameterized).
-This model illustrates the concept of a "preferential pathway", as the pipe carries contaminant vapor to the immediate vicinity of the foundation, by a path that circumvents the usual soil diffusion pathway.\par
-
-The ground surface and the pipe are both sources of air to the soil domain.
-Both are assumed to exist at reference atmospheric pressure.
-Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
-The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
-The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par
-
-\begin{table}[htb!]
-  \setlength{\tabcolsep}{1pt}
-  \centering
-  \begin{tabular}{l c c c c c c}
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % GOVERNING EQUATIONS
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \toprule
-    \multicolumn{7}{c}{\textbf{Governing Equations}} \\
-    \midrule
-    % CSTR
-    Unsteady-CST & \multicolumn{6}{c}{$V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$} \\
-    % RICHARDS
-    Richard's & \multicolumn{6}{c}{$\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$} \\
-    % TRANSPORT
-    Transport & \multicolumn{6}{c}{$\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$} \\
-    % MILLINGTON-QUIRK
-    Millington-Quirk & \multicolumn{6}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
-    % VAN GENUCHTEN
-    \multirow{4}{*}{van Genuchten} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
-     & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
-     & \multicolumn{6}{c}{$k_r = (1-\mathrm{Se})^{l}[1(\mathrm{Se}^{m})^m]^2$} \\
-     & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-     % BOUNDARY CONDITIONS
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Boundary Conditions}} \\
-    \midrule
-    \textbf{Boundary} & \multicolumn{3}{c}{\textbf{Richard's Eqn.}} & \multicolumn{3}{c}{\textbf{Transport Eqn.}} \\
-    % Foundation crack
-    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
-    % Groundwater source
-    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
-    % Ground surface
-    Ground surface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
-    % Preferential pathway
-    \multirow{2}{*}{\begin{minipage}{0.2\textwidth}Preferential\\pathway\end{minipage}} & \multicolumn{3}{c}{\multirow{2}{*}{$p = 0 \; \mathrm{(Pa)}$}} & \multicolumn{3}{c}{\multirow{2}{*}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$}} \\ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % SOIL PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
-    % GRAVEL
-    Gravel & $1.3 \cdot 10^{-9}$ & 0.42 & 0.005 & 100 & 3.1 \\
-    % SANDY LOAM
-    Sandy Loam & $5.9 \cdot 10^{-13}$ & 0.39 & 0.039 & 2.7 & 1.4 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % TCE PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$}} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} \\
-     & \multicolumn{1}{c}{$2.47 \cdot 10^{-2}$} & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % BUILDING PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Building Properties}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$}} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
-     & \multicolumn{1}{c}{300} & 15 & 0.5 \\
-    \bottomrule
-  \end{tabular}
-  \caption{Governing equations, boundary conditions \& model input parameters. See Table \ref{tbl:abbreviations} for nomenclature.}\label{tbl:eqns_bc_parameters}
-\end{table}
-
-\newpage
-\section{Results \& Discussion}\label{s:results}
-
-\subsection{Variation In Indoor Air Contaminant Concentration Over Time}\label{s:results_variation_time}
-
-High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
-It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par
-
-Figure \ref{fig:indianapolis} showed a large degree of temporal variation in one of the components, and the data for the other components were quite similar.
-What is apparent upon closer examination of such data is that the actual day-to-day variations are typically not nearly as large as those observed when tracking the data for a longer time.
-To demonstrate this point, the quotient of the maximum and minimum $c_\mathrm{in}$ values (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are shown as a function of time in Figure \ref{fig:resampling}.
-The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
-Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{resampling.pdf}
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected over a given time period.
-(e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months).
-The error bars are the 95\% confidence intervals.}\label{fig:resampling}
-\end{figure}
-
-For example, if the data are examined in terms of the mean maximum variation observable over the course of 24 hours (one day) the variation is no greater than about a factor of two for any of the contaminants at the Indianapolis house or for TCE at the ASU house (when the preferential pathway was closed).
-The mean variability at the latter was only a bit higher (about a factor of 3) when the preferential pathway was open.
-In other words, a sampling protocol that involves sampling on two consecutive days would typically not uncover the large temporal variations that characterize the site over longer periods of time.
-As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
-So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.\par
-
-Weeks of temporal separation in sampling events are required to observe the large variations of concern.
-Orders of magnitude differences begin to manifest themselves over the course of months.
-This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
-This would be consistent with requiring months to see the more significant variations.\par
-
-This analysis also suggests that certain types of preferential pathways contribute to larger variations on shorter timescales (ASU House).
-Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
-This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.\par
-
-The longer the resampling period, the larger the maximum variability in observed indoor air contaminant concentrations.
-In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
-Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.\par
-
-Multiple samples taken over a short time period, e.g.
-a few days, are unlikely to uncover significant variation in indoor air contaminant concentration; the larger transient variations typically manifest after longer time periods.
-\par
-
-\subsection{Statistical Analysis of Field Data}\label{s:results_pressure_concentration}
-
-The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
-The rate of advective entry of soil gas into a structure is frequently cited as playing an important role in determining entry rate of contaminant.
-This advective entry rate is closely linked to the indoor-outdoor pressure difference, as can be caused by the “stack effect”, for example.
-Thus we first consider how much variability there might be in the pressure driving force for advection, and if this can explain the observed variability in observed indoor air contaminant concentrations.\par
-
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure. 
-Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).\par
-
-We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
-For this analysis we considered two VI sites, NAS North Island and the ASU House.
-The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
-By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}\label{fig:kde}
-\end{figure}
-
-\begin{table}[htb!]
-  \newcommand{\NameEntry}[1]{
-    \multicolumn{2}{c}{
-      \multirow{2}{*}{
-        \begin{minipage}{0.2\textwidth}
-          \centering
-          \textbf{#1}
-        \end{minipage}
-      }
-    }
-  }
-  \centering
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
-    \midrule
-    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
-    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
-    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
-    \bottomrule
-  \end{tabular}
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
-\end{table}
-
-In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
-A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.\par
-
-Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
-Such issues have of course already been pointed out by the investigators who obtained the data.\par
-
-The power of this KDE representation is that it permits evaluation of the relationship of two independently measured data - the indoor air contaminant concentration and the indoor-outdoor pressure difference.
-Examining the data in this manner immediately points to an important difference between the data from the ASU House and those from NAS North Island.
-At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).\par
-
-The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
-The Pearson’s r-value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
-But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.\par
-
-The ASU house datasets offer a different picture.
-The variability of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
-The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
-These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
-
-The data for the ASU House also offer an insight into the role of the preferential pathway.
-At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed are relatively comparable.
-However, the 5th and 95th percentiles values of $c_\mathrm{in}$/$c_\mathrm{in,mean}$ differ significantly as may be seen in Table \ref{tbl:percentiles}.
-It is clear that existence of the preferential pathway dramatically increases the variability in indoor air contaminant concentration.
-This again is entirely consistent with what the investigators of that site have already reported\cite{guo_identification_2015}.
-The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
-These will be explored with the help of a modeling analysis presented below.\par
-
-\subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
-
-Observed temporal variations in indoor air contaminant concentrations might be explained by temporal variations in subslab contaminant concentrations.
-To examine how variability in subslab contaminant concentration might contribute to variability in indoor air contaminant concentration, data on the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) were examined.
-The dataset utilized for this was that from the ASU House.
-The $c_\mathrm{subslab}$ values were taken from a soil gas probe labeled as "6" at the ASU house.
-This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
- \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
-\end{figure}
-
-It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed, large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
-
-When the PP was open, there was considerably more variability in the subslab concentration values, and the mean value was higher than in the case where the preferential pathway was closed.
-It was also not uncommon for the observed $\alpha_\mathrm{subslab}$ to exceed unity.
-While large $\alpha_\mathrm{subslab}$ values may sometimes indicate indoor sources at a site, there were none at the ASU house.
-A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
-This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.\par
-
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in "normal" VI scenarios.\par
-
-\subsection{Modeling Results}\label{s:results_model}
-
-\subsubsection{Pressure Effects}\label{s:results_model_pressure}
-
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
- \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
-Field data are binned in 40 evenly spaced pressure bins, with the dot representing the mean and errors bars the 95\% confidence interval of data at a particular pressure range.
-Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
-Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".}\label{fig:land_drain_scenarios}
-\end{figure}
-
-In the calculation results shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
-Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
-Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
-The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
-These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
-
-The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.\par
-
-The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
-First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
-Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.\par
-
-The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
-This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
-In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
-These requirements were met at the ASU House.\par
-
-A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
-In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
-The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
-The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.\par
-
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
-These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e.
-that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par
-
-The above simulations capture the trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ but do not yet capture the full variability of the concentration results over the "most probable" portion of observed pressure distributions shown in Figure \ref{fig:kde} (which tend to be from -2 to +2 Pa).
-The results of Figure \ref{fig:land_drain_scenarios} show a spread of almost an order of magnitude over this pressure range for the case of the "open" preferential pathway, and almost no spread at all when the preferential pathway is "closed".
-Hence the predicted variability is roughly an order of magnitude too low, when considering only the influence of pressure.
-There is a factor that tends to increase the spread of the data one additional order of magnitude beyond what was predicted by the base calculations of Figure \ref{fig:land_drain_scenarios}.
-We believe that it is variations in air exchange rate, operating in concert with the natural variations in pressure differential, that explain the remaining variability.\par
-
-\subsubsection{Air Exchange Rate Effects}\label{s:results_model_air_exchange}
-
-Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the above model.
-Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
-A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
-Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
-Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in Figure \ref{fig:air_exchange_rate} show, there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
-Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
-  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
-The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
-Only non-CPM period considered.}\label{fig:air_exchange_rate}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l c c c}
-    \toprule
-    \textbf{Percentile}                                                     & \textbf{10th} & \textbf{50th} & \textbf{90th} \\
-    EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}           & 0.16-0.2      & 0.35-0.49     & 1.21-1.49     \\
-    ASU house\cite{holton_temporal_2013,guo_identification_2015}            & 0.21          & 0.43          & 0.78          \\
-    Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34          & 0.74          & 1.27          \\
-    \bottomrule
-  \end{tabular}
-  \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
-\end{table}
-
-To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are indicated by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
-
-It is apparent that assuming variability in air exchange rate allows capturing most of the observed variability in $\alpha_\mathrm{gw}$.
-We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus, we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
-
-\subsubsection{Results of Transient Simulations}\label{s:results_modeling_transient}
-
-The above analyses have been conducted under simulated steady state conditions.
-The conclusions regarding the importance of the different parameters are now examined in actual transient simulations.
-The model configuration of Figure \ref{fig:model} is run in 24-hour transient simulations to examine how $c_\mathrm{in}$ fluctuates over the course  of a "typical" day.
-The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
-The ASU House dataset was again the source of the "typical" $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
-The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
-The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the air exchange rate value assumed for the calculations at constant air exchange rate.\par
-
-\begin{figure}
-  \centering
- \includegraphics[width=\textwidth]{diurnal.pdf}
- \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
-\end{figure}
-
-The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
-The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e.
-a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
-When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
-the contaminant transport into the modeled structure is advection dominated.
-Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
-When the preferential pathway is closed the story is quite different.
-When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
-Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
-When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
-This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
-
-These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
-It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g.
-changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
-For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
-For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
-We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
-At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
-
-\section{Conclusions}
-
-Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
-
-There can be a significant role of advective transport in determining observed variability.
-For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
-In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
-In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
-
-Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
-In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
-This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
-This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
-
-In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
-Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
-The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
-It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
-These results were consistent with seasonal timescale variations in the intrusion processes.\par
-
-% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
-- 24 hr sampling is recommended for a few reasons.
-1. Avoids issues with diurnal fluctations in Ae or dP.
-2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
-- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
-- Passive samples can help capture some of the more longterm variations if these are a concern.
-- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
-Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
-- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
-- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
-1. Even with "good" subslab sampling, hotspots are easily missed.
-2. May mislead that an indoor source is present.
-3. Even without a PP there may still be significant spatial variability (Eklund).
-4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
-5. McAlary said something about this. Look it up.
-6. May not really be worth the effort.
-
-\section*{Acknowledgements}
-This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).\par
-
-Declaration of interest: none
-
-\clearpage
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l l}
-    \toprule
-    % A/ALPHA
-    $A_\mathrm{ck}$ & Crack area \\
-    $A_e$ & Air exchange rate \\
-    $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
-    $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
-    % C
-    $c_\mathrm{in}$ & Indoor air contaminant concentration \\
-    $c$ & Soil-gas contaminant concentration \\
-    $c_w$ & Soil-water contaminant concentration \\
-    $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-    $\chi$ & PP contaminant concentration scaling parameter \\
-    % D
-    $D_\mathrm{eff}$ & Effective diffusion coefficient \\
-    $D_\mathrm{air}$ & Diffusion coefficient in air \\
-    $D_\mathrm{water}$ & Diffusion coefficient in water \\
-    % J
-    $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
-    % K
-    $\kappa_s$ & Saturated soil permeability \\
-    $K_H$ & Dimensionless Henry's law constant \\
-    $k_r$ & Relative permeability \\
-    % L
-    $L_\mathrm{slab}$ & Thickness of the foundation slab \\
-    % M
-    $M$ & Molar mass \\
-    $\mu$ & Contaminant vapor viscosity \\
-    % N
-    NAS & Naval Air Stations \\
-    % P
-    $p$ & Pressure in soil \\
-    $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
-    PP & Preferential pathway \\
-    % R
-    $\rho$ & Density \\
-    % S
-    $\mathrm{Se}$ & Soil water saturation \\
-    % T
-    $t$ & time \\
-    $\theta_g$ & Vapor/gas filled porosity \\
-    $\theta_w$ & Water filled porosity \\
-    $\theta_r$ & Residual water filled porosity \\
-    $\theta_t$ & Total porosity \\
-    % U
-    $\vec{u}$ & Soil-gas velocity (vector quantity) \\
-    % V
-    VI & Vapor intrusion \\
-    $V_\mathrm{base}$ & Basement volume \\
-    % Z
-    $z$ & Elevation above groundwater \\
-    \bottomrule
-  \end{tabular}
-  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
-\end{table}
-
-\bibliographystyle{elsarticle-num}
-\bibliography{library}
-
-\end{document}

### 2019-09-24 doc/manuscript.tex
Reorganizing

--- 2019-06-03 tex/manuscript.tex
@@
 
-\section{Introduction}\label{s:intro}
-
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-Such variations make it difficult for those charged with protecting human health to formulate a response - should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
-There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-What represents a reasonable sampling strategy for a particular site a single 8-hour sample?
-Repeated 8-hour samples?
-Month-long samples?
-Continuous monitoring?\par
-
-VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-The basic nature of VI has been understood for some time and it has been the subject of much study, but some aspects remain poorly understood, such as the causes of the sometimes observed large temporal transients in indoor air concentrations.
-Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
-All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.\par
-
-In one instance the source of the variation was clearly established during the study of the site.
-At the ASU house a drain pipe (or “land drain”) connected to a sewer system was discovered beneath the house.
-Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.\par
-
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
-That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been previously implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.\par
-
-In some of the above cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many instances, this scenario is not further considered here where the focus is on pathways that deliver contaminant via the soil beneath a structure.
-It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is the site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.\par
-
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as “normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at such normal VI sites vs.
-those characterized by preferential pathways within the soil beneath the site.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
-
-\section{Methods}\label{s:methods}
-
-\subsection{Statistical Analysis Of Field Data}\label{s:methods_stats}
-
-To frame the question of just how much variability in indoor air contaminant concentrations is actually observed, field datasets have been analyzed.
-For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.\par
-
-The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation: increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-The period of CPM testing is thus excluded from the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par
-
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to “large” internal pressure fluctuations.
-By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
-The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.\par
-
-The Indianapolis site investigation also spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD) for VI mitigation.
-Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}.
-Unlike at the ASU house, this preferential pathway was never removed or blocked, making it impossible to isolate the role of the preferential pathway at this site.
-It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
-Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
-There is almost a two order of magnitude variation in the concentration data.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
- \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
-\end{figure}
-
-Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
-KDE is a technique used for estimating the probability distribution of a random variable(s) by using multiple kernels, or weighting functions to characterize the data sets.
-In this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values and that there are statistical fluctuations associated with each sampling event.
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.\par
-
-\subsection{Modeling Work}\label{s:methods_model}
-
-A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
-In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{model.png}
- \caption{Foundation and vadose zone soil represented in the modeling.
-Note that here a gravel sub-base material is shown, but in certain simulations, that material is absent and the surrounding soil directly contacts the foundation slab.
-Different assumptions are made regarding the preferential pathway, here shown as a pipe entering the gravel sub-base.
-In some cases, the preferential pathway has been "turned off".}\label{fig:model}
-\end{figure}
-
-The modeled VI impacted structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the  perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is here assumed to consist of only of the basement, having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of some critical parameters is.\par
-
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house and is assumed to consist of sandy loam, except as noted otherwise.
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases here this sub-base material is assumed to be the same as the surrounding soil (termed a "uniform" soil scenario).
-The groundwater beneath the structure is assumed to be homogeneously contaminated with TCE selected as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-Where relevant, the preferential pathway is modeled as a 10 cm (4”) pipe that opens into the gravel sub-base beneath the structure.
-The air in the pipe is also assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$ (allowing the contaminant concentration in the pipe to be parameterized).
-This model illustrates the concept of a "preferential pathway", as the pipe carries contaminant vapor to the immediate vicinity of the foundation, by a path that circumvents the usual soil diffusion pathway.\par
-
-The ground surface and the pipe are both sources of air to the soil domain.
-Both are assumed to exist at reference atmospheric pressure.
-Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
-The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
-The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par
-
-\begin{table}[htb!]
-  \setlength{\tabcolsep}{1pt}
-  \centering
-  \begin{tabular}{l c c c c c c}
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % GOVERNING EQUATIONS
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \toprule
-    \multicolumn{7}{c}{\textbf{Governing Equations}} \\
-    \midrule
-    % CSTR
-    Unsteady-CST & \multicolumn{6}{c}{$V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$} \\
-    % RICHARDS
-    Richard's & \multicolumn{6}{c}{$\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$} \\
-    % TRANSPORT
-    Transport & \multicolumn{6}{c}{$\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$} \\
-    % MILLINGTON-QUIRK
-    Millington-Quirk & \multicolumn{6}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
-    % VAN GENUCHTEN
-    \multirow{4}{*}{van Genuchten} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
-     & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
-     & \multicolumn{6}{c}{$k_r = (1-\mathrm{Se})^{l}[1(\mathrm{Se}^{m})^m]^2$} \\
-     & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-     % BOUNDARY CONDITIONS
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Boundary Conditions}} \\
-    \midrule
-    \textbf{Boundary} & \multicolumn{3}{c}{\textbf{Richard's Eqn.}} & \multicolumn{3}{c}{\textbf{Transport Eqn.}} \\
-    % Foundation crack
-    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
-    % Groundwater source
-    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
-    % Ground surface
-    Ground surface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
-    % Preferential pathway
-    \multirow{2}{*}{\begin{minipage}{0.2\textwidth}Preferential\\pathway\end{minipage}} & \multicolumn{3}{c}{\multirow{2}{*}{$p = 0 \; \mathrm{(Pa)}$}} & \multicolumn{3}{c}{\multirow{2}{*}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$}} \\ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % SOIL PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
-    % GRAVEL
-    Gravel & $1.3 \cdot 10^{-9}$ & 0.42 & 0.005 & 100 & 3.1 \\
-    % SANDY LOAM
-    Sandy Loam & $5.9 \cdot 10^{-13}$ & 0.39 & 0.039 & 2.7 & 1.4 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % TCE PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$}} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} \\
-     & \multicolumn{1}{c}{$2.47 \cdot 10^{-2}$} & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % BUILDING PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Building Properties}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$}} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
-     & \multicolumn{1}{c}{300} & 15 & 0.5 \\
-    \bottomrule
-  \end{tabular}
-  \caption{Governing equations, boundary conditions \& model input parameters. See Table \ref{tbl:abbreviations} for nomenclature.}\label{tbl:eqns_bc_parameters}
-\end{table}
-
+\input{intro.tex}
+\input{methods.tex}
 \newpage
-\section{Results \& Discussion}\label{s:results}
-
-\subsection{Variation In Indoor Air Contaminant Concentration Over Time}\label{s:results_variation_time}
-
-High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
-It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par
-
-Figure \ref{fig:indianapolis} showed a large degree of temporal variation in one of the components, and the data for the other components were quite similar.
-What is apparent upon closer examination of such data is that the actual day-to-day variations are typically not nearly as large as those observed when tracking the data for a longer time.
-To demonstrate this point, the quotient of the maximum and minimum $c_\mathrm{in}$ values (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are shown as a function of time in Figure \ref{fig:resampling}.
-The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
-Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{resampling.pdf}
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected over a given time period.
-(e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months).
-The error bars are the 95\% confidence intervals.}\label{fig:resampling}
-\end{figure}
-
-For example, if the data are examined in terms of the mean maximum variation observable over the course of 24 hours (one day) the variation is no greater than about a factor of two for any of the contaminants at the Indianapolis house or for TCE at the ASU house (when the preferential pathway was closed).
-The mean variability at the latter was only a bit higher (about a factor of 3) when the preferential pathway was open.
-In other words, a sampling protocol that involves sampling on two consecutive days would typically not uncover the large temporal variations that characterize the site over longer periods of time.
-As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
-So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.\par
-
-Weeks of temporal separation in sampling events are required to observe the large variations of concern.
-Orders of magnitude differences begin to manifest themselves over the course of months.
-This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
-This would be consistent with requiring months to see the more significant variations.\par
-
-This analysis also suggests that certain types of preferential pathways contribute to larger variations on shorter timescales (ASU House).
-Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
-This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.\par
-
-The longer the resampling period, the larger the maximum variability in observed indoor air contaminant concentrations.
-In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
-Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.\par
-
-Multiple samples taken over a short time period, e.g.
-a few days, are unlikely to uncover significant variation in indoor air contaminant concentration; the larger transient variations typically manifest after longer time periods.
-\par
-
-\subsection{Statistical Analysis of Field Data}\label{s:results_pressure_concentration}
-
-The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
-The rate of advective entry of soil gas into a structure is frequently cited as playing an important role in determining entry rate of contaminant.
-This advective entry rate is closely linked to the indoor-outdoor pressure difference, as can be caused by the “stack effect”, for example.
-Thus we first consider how much variability there might be in the pressure driving force for advection, and if this can explain the observed variability in observed indoor air contaminant concentrations.\par
-
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure. 
-Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).\par
-
-We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
-For this analysis we considered two VI sites, NAS North Island and the ASU House.
-The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
-By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}\label{fig:kde}
-\end{figure}
-
-\begin{table}[htb!]
-  \newcommand{\NameEntry}[1]{
-    \multicolumn{2}{c}{
-      \multirow{2}{*}{
-        \begin{minipage}{0.2\textwidth}
-          \centering
-          \textbf{#1}
-        \end{minipage}
-      }
-    }
-  }
-  \centering
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
-    \midrule
-    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
-    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
-    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
-    \bottomrule
-  \end{tabular}
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
-\end{table}
-
-In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
-A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.\par
-
-Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
-Such issues have of course already been pointed out by the investigators who obtained the data.\par
-
-The power of this KDE representation is that it permits evaluation of the relationship of two independently measured data - the indoor air contaminant concentration and the indoor-outdoor pressure difference.
-Examining the data in this manner immediately points to an important difference between the data from the ASU House and those from NAS North Island.
-At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).\par
-
-The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
-The Pearson’s r-value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
-But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.\par
-
-The ASU house datasets offer a different picture.
-The variability of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
-The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
-These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
-
-The data for the ASU House also offer an insight into the role of the preferential pathway.
-At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed are relatively comparable.
-However, the 5th and 95th percentiles values of $c_\mathrm{in}$/$c_\mathrm{in,mean}$ differ significantly as may be seen in Table \ref{tbl:percentiles}.
-It is clear that existence of the preferential pathway dramatically increases the variability in indoor air contaminant concentration.
-This again is entirely consistent with what the investigators of that site have already reported\cite{guo_identification_2015}.
-The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
-These will be explored with the help of a modeling analysis presented below.\par
-
-\subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
-
-Observed temporal variations in indoor air contaminant concentrations might be explained by temporal variations in subslab contaminant concentrations.
-To examine how variability in subslab contaminant concentration might contribute to variability in indoor air contaminant concentration, data on the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) were examined.
-The dataset utilized for this was that from the ASU House.
-The $c_\mathrm{subslab}$ values were taken from a soil gas probe labeled as "6" at the ASU house.
-This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
- \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
-\end{figure}
-
-It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed, large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
-
-When the PP was open, there was considerably more variability in the subslab concentration values, and the mean value was higher than in the case where the preferential pathway was closed.
-It was also not uncommon for the observed $\alpha_\mathrm{subslab}$ to exceed unity.
-While large $\alpha_\mathrm{subslab}$ values may sometimes indicate indoor sources at a site, there were none at the ASU house.
-A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
-This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.\par
-
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in "normal" VI scenarios.\par
-
-\subsection{Modeling Results}\label{s:results_model}
-
-\subsubsection{Pressure Effects}\label{s:results_model_pressure}
-
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
- \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
-Field data are binned in 40 evenly spaced pressure bins, with the dot representing the mean and errors bars the 95\% confidence interval of data at a particular pressure range.
-Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
-Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".}\label{fig:land_drain_scenarios}
-\end{figure}
-
-In the calculation results shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
-Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
-Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
-The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
-These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
-
-The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.\par
-
-The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
-First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
-Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.\par
-
-The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
-This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
-In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
-These requirements were met at the ASU House.\par
-
-A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
-In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
-The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
-The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.\par
-
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
-These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e.
-that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par
-
-The above simulations capture the trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ but do not yet capture the full variability of the concentration results over the "most probable" portion of observed pressure distributions shown in Figure \ref{fig:kde} (which tend to be from -2 to +2 Pa).
-The results of Figure \ref{fig:land_drain_scenarios} show a spread of almost an order of magnitude over this pressure range for the case of the "open" preferential pathway, and almost no spread at all when the preferential pathway is "closed".
-Hence the predicted variability is roughly an order of magnitude too low, when considering only the influence of pressure.
-There is a factor that tends to increase the spread of the data one additional order of magnitude beyond what was predicted by the base calculations of Figure \ref{fig:land_drain_scenarios}.
-We believe that it is variations in air exchange rate, operating in concert with the natural variations in pressure differential, that explain the remaining variability.\par
-
-\subsubsection{Air Exchange Rate Effects}\label{s:results_model_air_exchange}
-
-Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the above model.
-Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
-A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
-Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
-Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in Figure \ref{fig:air_exchange_rate} show, there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
-Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
-  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
-The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
-Only non-CPM period considered.}\label{fig:air_exchange_rate}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l c c c}
-    \toprule
-    \textbf{Percentile}                                                     & \textbf{10th} & \textbf{50th} & \textbf{90th} \\
-    EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}           & 0.16-0.2      & 0.35-0.49     & 1.21-1.49     \\
-    ASU house\cite{holton_temporal_2013,guo_identification_2015}            & 0.21          & 0.43          & 0.78          \\
-    Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34          & 0.74          & 1.27          \\
-    \bottomrule
-  \end{tabular}
-  \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
-\end{table}
-
-To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are indicated by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
-
-It is apparent that assuming variability in air exchange rate allows capturing most of the observed variability in $\alpha_\mathrm{gw}$.
-We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus, we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
-
-\subsubsection{Results of Transient Simulations}\label{s:results_modeling_transient}
-
-The above analyses have been conducted under simulated steady state conditions.
-The conclusions regarding the importance of the different parameters are now examined in actual transient simulations.
-The model configuration of Figure \ref{fig:model} is run in 24-hour transient simulations to examine how $c_\mathrm{in}$ fluctuates over the course  of a "typical" day.
-The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
-The ASU House dataset was again the source of the "typical" $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
-The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
-The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the air exchange rate value assumed for the calculations at constant air exchange rate.\par
-
-\begin{figure}
-  \centering
- \includegraphics[width=\textwidth]{diurnal.pdf}
- \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
-\end{figure}
-
-The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
-The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e.
-a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
-When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
-the contaminant transport into the modeled structure is advection dominated.
-Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
-When the preferential pathway is closed the story is quite different.
-When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
-Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
-When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
-This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
-
-These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
-It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g.
-changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
-For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
-For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
-We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
-At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
-
-\section{Conclusions}
-
-Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
-
-There can be a significant role of advective transport in determining observed variability.
-For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
-In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
-In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
-
-Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
-In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
-This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
-This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
-
-In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
-Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
-The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
-It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
-These results were consistent with seasonal timescale variations in the intrusion processes.\par
-
-% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
-- 24 hr sampling is recommended for a few reasons.
-1. Avoids issues with diurnal fluctations in Ae or dP.
-2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
-- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
-- Passive samples can help capture some of the more longterm variations if these are a concern.
-- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
-Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
-- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
-- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
-1. Even with "good" subslab sampling, hotspots are easily missed.
-2. May mislead that an indoor source is present.
-3. Even without a PP there may still be significant spatial variability (Eklund).
-4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
-5. McAlary said something about this. Look it up.
-6. May not really be worth the effort.
+\input{results.tex}
+\input{conclusions.tex}
 
@@
 \clearpage
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l l}
-    \toprule
-    % A/ALPHA
-    $A_\mathrm{ck}$ & Crack area \\
-    $A_e$ & Air exchange rate \\
-    $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
-    $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
-    % C
-    $c_\mathrm{in}$ & Indoor air contaminant concentration \\
-    $c$ & Soil-gas contaminant concentration \\
-    $c_w$ & Soil-water contaminant concentration \\
-    $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-    $\chi$ & PP contaminant concentration scaling parameter \\
-    % D
-    $D_\mathrm{eff}$ & Effective diffusion coefficient \\
-    $D_\mathrm{air}$ & Diffusion coefficient in air \\
-    $D_\mathrm{water}$ & Diffusion coefficient in water \\
-    % J
-    $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
-    % K
-    $\kappa_s$ & Saturated soil permeability \\
-    $K_H$ & Dimensionless Henry's law constant \\
-    $k_r$ & Relative permeability \\
-    % L
-    $L_\mathrm{slab}$ & Thickness of the foundation slab \\
-    % M
-    $M$ & Molar mass \\
-    $\mu$ & Contaminant vapor viscosity \\
-    % N
-    NAS & Naval Air Stations \\
-    % P
-    $p$ & Pressure in soil \\
-    $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
-    PP & Preferential pathway \\
-    % R
-    $\rho$ & Density \\
-    % S
-    $\mathrm{Se}$ & Soil water saturation \\
-    % T
-    $t$ & time \\
-    $\theta_g$ & Vapor/gas filled porosity \\
-    $\theta_w$ & Water filled porosity \\
-    $\theta_r$ & Residual water filled porosity \\
-    $\theta_t$ & Total porosity \\
-    % U
-    $\vec{u}$ & Soil-gas velocity (vector quantity) \\
-    % V
-    VI & Vapor intrusion \\
-    $V_\mathrm{base}$ & Basement volume \\
-    % Z
-    $z$ & Elevation above groundwater \\
-    \bottomrule
-  \end{tabular}
-  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
-\end{table}
+\input{appendix.tex}
 

### 2019-09-24 doc/methods.tex
Reorganizing

--- 2019-06-03 tex/manuscript.tex
@@
-\documentclass[preprint,12pt]{elsarticle}
-
-% additional packages
-\usepackage[utf8]{inputenc}
-%\bibliographystyle{abbrvnat}
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{subcaption}
-\usepackage{tabularx}
-\usepackage{array,multirow,multicol,graphicx}
-\usepackage{comment}
-\graphicspath{
-  {./figures/},
-}
-\usepackage{lineno}
-\linenumbers
-% macros
-
-\journal{Building and Environment}
-
-\begin{document}
-
-\begin{frontmatter}
-% title
-\title{Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
-
-% authors
-\author[Brown]{Jonathan G. V. Ström}
-\author[ASU]{Yuanming Guo}
-\author[ASU]{Yijun Yao}
-\author[Brown]{Eric M. Suuberg}
-
-\ead{eric\_suuberg@brown.edu}
-
-\address[Brown]{Brown University, School of Engineering, Providence, RI, USA}
-\address[ASU]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
-
-\begin{abstract}
-Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
-We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
-We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
-A preferential pathway can dramatically increase a VI site's sensitivity to build pressurization, provided there exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
-a permeable subslab space that may be provided by a gravel layer.
-Preferential pathways may also erroneously indicate the presence of indoor contaminant sources.
-At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
-Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
-\end{abstract}
-
-\begin{keyword}
-  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Attenuation factor \sep Air exchange rate \sep Indoor/outdoor pressure difference
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{s:intro}
-
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-Such variations make it difficult for those charged with protecting human health to formulate a response - should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
-There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-What represents a reasonable sampling strategy for a particular site a single 8-hour sample?
-Repeated 8-hour samples?
-Month-long samples?
-Continuous monitoring?\par
-
-VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-The basic nature of VI has been understood for some time and it has been the subject of much study, but some aspects remain poorly understood, such as the causes of the sometimes observed large temporal transients in indoor air concentrations.
-Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
-All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.\par
-
-In one instance the source of the variation was clearly established during the study of the site.
-At the ASU house a drain pipe (or “land drain”) connected to a sewer system was discovered beneath the house.
-Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.\par
-
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
-That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been previously implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.\par
-
-In some of the above cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many instances, this scenario is not further considered here where the focus is on pathways that deliver contaminant via the soil beneath a structure.
-It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is the site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.\par
-
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as “normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at such normal VI sites vs.
-those characterized by preferential pathways within the soil beneath the site.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
-
 \section{Methods}\label{s:methods}
@@
 \end{table}
-
-\newpage
-\section{Results \& Discussion}\label{s:results}
-
-\subsection{Variation In Indoor Air Contaminant Concentration Over Time}\label{s:results_variation_time}
-
-High frequency measurement of indoor air contaminant concentrations, $c_\mathrm{in}$, such as those in Figure \ref{fig:indianapolis}, took place at both the ASU House and the Indianapolis House over significant periods (Indianapolis: ca 1.7 years, ASU house: ca 3.5 years)\cite{u.s._environmental_protection_agency_assessment_2015,holton_temporal_2013}.
-Furthermore, at the Indianapolis site $c_\mathrm{in}$ for three different contaminants, chloroform, TCE, and tetrachloroethylene (PCE) were all collected, allowing examination of the variability of each VI contaminant.
-The NAS North Island NAS dataset was obtained over a much shorter duration (9 days), and is therefore not examined in this portion of the analysis.
-It should also be noted that the ASU house used 4-hour sorbent tubes, while Indianapolis took instantaneous "grab" samples.\par
-
-Figure \ref{fig:indianapolis} showed a large degree of temporal variation in one of the components, and the data for the other components were quite similar.
-What is apparent upon closer examination of such data is that the actual day-to-day variations are typically not nearly as large as those observed when tracking the data for a longer time.
-To demonstrate this point, the quotient of the maximum and minimum $c_\mathrm{in}$ values (denoted as $c_\mathrm{max}/c_\mathrm{min}$) are shown as a function of time in Figure \ref{fig:resampling}.
-The values shown in Figure \ref{fig:resampling} are the means of the quotients calculated for samples separated by the indicated times and the error bars indicate the 95th percentile of all the data points.
-Hypothetical resampling periods of one, two, three days, and the same number of weeks, and months were chosen.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{resampling.pdf}
- \caption{Mean values of the maximum change in indoor air contaminant concentration that may be expected over a given time period.
-(e.g., 1D is 1 day, 2W is 2 weeks, and 3M is 3 months).
-The error bars are the 95\% confidence intervals.}\label{fig:resampling}
-\end{figure}
-
-For example, if the data are examined in terms of the mean maximum variation observable over the course of 24 hours (one day) the variation is no greater than about a factor of two for any of the contaminants at the Indianapolis house or for TCE at the ASU house (when the preferential pathway was closed).
-The mean variability at the latter was only a bit higher (about a factor of 3) when the preferential pathway was open.
-In other words, a sampling protocol that involves sampling on two consecutive days would typically not uncover the large temporal variations that characterize the site over longer periods of time.
-As Figure \ref{fig:indianapolis} shows, there are certainly isolated days in which a larger daily change was observed, but these were not typical, to the extent that they fall outside of the 95\% criteria used in defining the error bars.
-So while such unusual jumps might be seen (for unknown reasons) in a very small percentage of cases, the expectation is much more represented by what is shown in Figure \ref{fig:resampling}.\par
-
-Weeks of temporal separation in sampling events are required to observe the large variations of concern.
-Orders of magnitude differences begin to manifest themselves over the course of months.
-This is not surprising, since those who performed the measurements have already reported that there were seasonal aspects to the values obtained.
-This would be consistent with requiring months to see the more significant variations.\par
-
-This analysis also suggests that certain types of preferential pathways contribute to larger variations on shorter timescales (ASU House).
-Even though there was a preferential pathway present at the Indianapolis House, the transients associated with its presence were of a slower nature and the behavior was not unlike what was observed at ASU House when the preferential pathway was closed.
-This warns that the mere existence of a preferential pathway is not by itself sufficient to create a situation of large variations over short sampling times.\par
-
-The longer the resampling period, the larger the maximum variability in observed indoor air contaminant concentrations.
-In the case of the ASU House with the preferential pathway open, the variability went from less than a threefold difference on the timescale of a day, to two to three orders of magnitude over the course of weeks.
-Thus there are different timescales that characterize different extents of variation, again pointing to the existence of more than a single factor that determines variability.\par
-
-Multiple samples taken over a short time period, e.g.
-a few days, are unlikely to uncover significant variation in indoor air contaminant concentration; the larger transient variations typically manifest after longer time periods.
-\par
-
-\subsection{Statistical Analysis of Field Data}\label{s:results_pressure_concentration}
-
-The data in Figure \ref{fig:indianapolis} and Figure \ref{fig:resampling} raise the question of what then actually determines the large degree of temporal variation sometimes reported.
-The rate of advective entry of soil gas into a structure is frequently cited as playing an important role in determining entry rate of contaminant.
-This advective entry rate is closely linked to the indoor-outdoor pressure difference, as can be caused by the “stack effect”, for example.
-Thus we first consider how much variability there might be in the pressure driving force for advection, and if this can explain the observed variability in observed indoor air contaminant concentrations.\par
-
-The pressure difference between the indoor and outdoor/ambient ($p_\mathrm{in/out}$) leads to advection, by which contaminants are drawn into (or prevented from) entering a structure. 
-Changes in $p_\mathrm{in/out}$ can take place quickly, leaving open the possibility of their impacting VI far more rapidly than can fluctuations in say groundwater depth or contaminant concentration (these latter processes take weeks or even months to impact the overlying structure).\par
-
-We examine the relationship between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ by constructing the two-dimensional kernel density estimation (KDE) plots seen in Figure \ref{fig:kde}.
-The KDE plots allow us to view the measured distributions of $p_\mathrm{in/out}$ and $c_\mathrm{in}$, and develop a visual impression of how well these distributions correlate with one another.
-For this analysis we considered two VI sites, NAS North Island and the ASU House.
-The ASU House dataset was divided into two periods, one before and the other after the land drain (called the preferential pathway (PP) from  here on) had been closed.
-By comparing these two periods on a single plot, the impact of the preferential pathway becomes clearer.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{nas_asu_pp.pdf}
- \caption{2D-KDE plot showing the distributions of indoor air contaminant concentration, the indoor/outdoor pressure difference, and how they correlate to each other.}\label{fig:kde}
-\end{figure}
-
-\begin{table}[htb!]
-  \newcommand{\NameEntry}[1]{
-    \multicolumn{2}{c}{
-      \multirow{2}{*}{
-        \begin{minipage}{0.2\textwidth}
-          \centering
-          \textbf{#1}
-        \end{minipage}
-      }
-    }
-  }
-  \centering
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    \multirow{2}{*}{ } & \NameEntry{North Island NAS} & \NameEntry{ASU House PP Open} & \NameEntry{ASU House PP Closed} \\ \\
-    \midrule
-    Percentile & 5th & 95th & 5th & 95th & 5th  & 95th \\
-    $p_\mathrm{in/out}$ (Pa) & -19.9 & 7.4 & -1.4 & 2.1  & -2.1 & 2.27 \\
-    $c_\mathrm{in}/c_\mathrm{in,mean}$ & 4.1 & 0.2 & 13.5 & 0.2  & 3.3  & 0.4  \\
-    \bottomrule
-  \end{tabular}
- \caption{5th and 95th percentile values of $p_\mathrm{in/out}$ and $c_\mathrm{in}/c_\mathrm{in,mean}$ in Figure \ref{fig:kde}.}\label{tbl:percentiles}
-\end{table}
-
-In Figure \ref{fig:kde}, the indoor air contaminant concentration $c_\mathrm{in}$ is normalized to the mean $c_\mathrm{in,mean}$ of each dataset, allowing comparison of the impact $p_\mathrm{in/out}$ on $c_\mathrm{in}$ independently from the large differences in absolute values of indoor air concentrations at the different sites.
-A value of 10 on the y-axis indicates that the corresponding plotted value of $c_\mathrm{in}$ is 10 times greater than the mean for the dataset, and 0.1 indicate that it is one tenth of the mean.\par
-
-Inspection of the range of normalized $c_\mathrm{in}$ values in Figure \ref{fig:kde} again shows the two order of magnitude spread in observed values, implying a sampling at one particular time might give a value that is two orders of magnitude different than a result from a different time.
-Such issues have of course already been pointed out by the investigators who obtained the data.\par
-
-The power of this KDE representation is that it permits evaluation of the relationship of two independently measured data - the indoor air contaminant concentration and the indoor-outdoor pressure difference.
-Examining the data in this manner immediately points to an important difference between the data from the ASU House and those from NAS North Island.
-At NAS North Island site $p_\mathrm{in/out}$ varies significantly; the 5th and 95th percentile of $p_\mathrm{in/out}$ are -19.9 and 7.4 Pa respectively.
-This may be contrasted with 5th and 95th percentile $p_\mathrm{in/out}$ at the ASU house: -1.4 and 2.1 Pa (with the PP open), and -2.1 and 2.27 Pa (PP closed).\par
-
-The much larger under- and overpressurization of the NAS North Island site compared to the ASU House makes the pressure dependence of indoor air concentration much more visible at the former site.
-The Pearson’s r-value for the correlation between $p_\mathrm{in/out}$ and $c_\mathrm{in}$ for each dataset is shown in the legend, and confirms what is apparent to the eye; the pressure driving force is a determining factor for observed contamination at NAS North Island.
-But the broadness of the band of the NAS North Island concentration data set suggests that there is still a source of variability in $c_\mathrm{in}$ that has not been fully captured - this will be addressed below.\par
-
-The ASU house datasets offer a different picture.
-The variability of $c_\mathrm{in}$ is just as large, or even larger than at NAS North Island, yet the $p_\mathrm{in/out}$ varied far less.
-The weaker dependence of $c_\mathrm{in}$ on the pressure difference is confirmed by the much lower r-values for the correlations between the variables.
-In other words, there is not nearly as strong a correlation between variation in indoor air contaminant concentration and pressure difference for the ASU House as there was for NAS North Island.
-These results strongly suggest that there are other factors besides indoor pressure determining indoor air contaminant concentrations, and their variations, that may not be accounted for in applying this method.\par
-
-The data for the ASU House also offer an insight into the role of the preferential pathway.
-At first glance it may seem like the $c_\mathrm{in}$ values for the periods when the PP is open and closed are relatively comparable.
-However, the 5th and 95th percentiles values of $c_\mathrm{in}$/$c_\mathrm{in,mean}$ differ significantly as may be seen in Table \ref{tbl:percentiles}.
-It is clear that existence of the preferential pathway dramatically increases the variability in indoor air contaminant concentration.
-This again is entirely consistent with what the investigators of that site have already reported\cite{guo_identification_2015}.
-The correlation with indoor-outdoor pressure difference is weak in the ASU house cases, so there are clearly factors other than pressure difference that determine the variability in each.
-These will be explored with the help of a modeling analysis presented below.\par
-
-\subsection{Variability Of Attenuation to Subslab Concentrations}\label{s:results_attenuation}
-
-Observed temporal variations in indoor air contaminant concentrations might be explained by temporal variations in subslab contaminant concentrations.
-To examine how variability in subslab contaminant concentration might contribute to variability in indoor air contaminant concentration, data on the attenuation from subslab ($\alpha_\mathrm{subslab} = c_\mathrm{in}/c_\mathrm{subslab}$) were examined.
-The dataset utilized for this was that from the ASU House.
-The $c_\mathrm{subslab}$ values were taken from a soil gas probe labeled as "6" at the ASU house.
-This probe was located closest to both the exit of the preferential pathway pipe, and to a reported breach in the foundation that served as a key entry pathway for contaminant getting into the house\cite{guo_identification_2015}.
-The results are shown in Figure \ref{fig:attenuation_subslab}, which shows the full distributions for both the case in which the preferential pathway was "open" and when it "closed".\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=0.7\textwidth]{asu_attenuation_subslab.pdf}
- \caption{Boxplot of $\log_{10}$ (subslab to indoor air contaminant attenuation) at the ASU house site.
-The box shows the quartiles of the distribution, the whiskers the extent of the distribution.}\label{fig:attenuation_subslab}
-\end{figure}
-
-It is apparent that during the period when the preferential pathway was closed, $\alpha_\mathrm{subslab}$ did not vary significantly, and was quite close to the EPA recommended $\alpha_\mathrm{subslab}$ value of 0.03\cite{u.s._environmental_protection_agency_oswer_2015}.
-Thus during the period when the preferential pathway was closed, large temporal variations in subslab concentrations could not have been driving the variations in indoor air contaminant concentrations.\par
-
-When the PP was open, there was considerably more variability in the subslab concentration values, and the mean value was higher than in the case where the preferential pathway was closed.
-It was also not uncommon for the observed $\alpha_\mathrm{subslab}$ to exceed unity.
-While large $\alpha_\mathrm{subslab}$ values may sometimes indicate indoor sources at a site, there were none at the ASU house.
-A more likely explanation is that even though probe "6" was located in close proximity to the exit of the preferential pathway, there might have still existed significant spatial variability in $c_\mathrm{subslab}$ that could not be captured with a single measurement.
-This suggests caution is needed in profiling subslab contaminant concentrations in the presence of preferential pathways - significant variations are possible.\par
-
-What the results of Figure \ref{fig:attenuation_subslab} do clearly show is that the existence of a preferential pathway of the kind at ASU House (and idealized in Figure \ref{fig:indianapolis}) can influence the temporal variation of subslab concentrations in a much less predictable way than those observed in "normal" VI scenarios.\par
-
-\subsection{Modeling Results}\label{s:results_model}
-
-\subsubsection{Pressure Effects}\label{s:results_model_pressure}
-
-Having established the potential impacts of certain inputs on determining variability in indoor air contaminant concentrations, the mathematical model of VI can help further elucidate other key aspects.
-The results of calculations on a scenario corresponding to Figure \ref{fig:model} are presented in Figure \ref{fig:land_drain_scenarios}.
-This scenario is not intended to exactly represent the situation at ASU House, but it is similar in the key aspect of having a preferential pathway delivering contaminant to a gravel sub-base.
-The full, complex geometry of the ASU House has not been represented, but the modeled structure is of comparable size, and will be subject to operational parameters based upon what were measured at that site.
-The general modeling conditions are those shown in Table \ref{fig:indianapolis}.\par
-
-\begin{figure}[htb!]
- \centering
- \includegraphics[width=\textwidth]{land_drain_scenarios_combo.pdf}
- \caption{Simulated preferential pathway scenarios compared to actual ASU house field data.
-Field data are binned in 40 evenly spaced pressure bins, with the dot representing the mean and errors bars the 95\% confidence interval of data at a particular pressure range.
-Shaded blue represent the range of model predictions for the indicated pressure difference, due to air exchange rate variability (using 5th and 95th percentile values of measured exchange rates).
-Top panel is for various cases representing an "open" preferential pathway, the lower panel with the pathway "closed".}\label{fig:land_drain_scenarios}
-\end{figure}
-
-In the calculation results shown in the top panel of Figure \ref{fig:land_drain_scenarios}, a preferential pathway is assumed to provide air containing contaminant vapor at a concentration equivalent to the vapor in equilibrium with the underlying groundwater source.
-Here, the indoor air exchange rate $A_e$ was assumed to be a constant 0.5 per hour, and $p_\mathrm{in/out}$ was varied from -5 to 5 Pa.
-Values of predicted indoor air contaminant concentrations, $c_\mathrm{in}$ were obtained from steady state calculations.
-The predicted $c_\mathrm{in}$ values were then normalized by the assumed vapor concentration in equilibrium with groundwater $c_\mathrm{gw}$, giving the attenuation from groundwater $\alpha_\mathrm{gw}$.
-The predicted values of $\alpha_\mathrm{gw}$ as a function of $p_\mathrm{in/out}$ are given by the central blue line in the upper panel of Figure \ref{fig:land_drain_scenarios}.
-These predicted values are compared to actual measured $\alpha_\mathrm{gw}$ values from the ASU House for the period during which the preferential pathway was open (blue points).\par
-
-The model successfully predicts the observed trends in $\alpha_\mathrm{gw}$ as $p_\mathrm{in/out}$ decreases (increased depressurization) but somewhat underpredicts $\alpha_\mathrm{gw}$ as the house is overpressurized.
-Most significantly, the model captures that even for a small increase in depressurization (0 to -5 Pa) a very large increase in $\alpha_\mathrm{gw}$ (two order of magnitude) can occur.\par
-
-The asymmetry relative to the predictions for depressurization and overpressurization is due to two factors.
-First, the preferential pathway acts not only as a source of contaminant vapor, but also  as a source of air to the subslab.
-Because of the large resistance to soil gas flow in the surrounding soil, having a local source of air to support the increase of advective flow into the structure from the subslab region makes a large difference.\par
-
-The above was proven by a second simulation, where the model was rerun with the preferential pathway present, but with the permeable (gravel) layer in the subslab removed and replaced by the surrounding soil (sandy loam).
-This gave a ”uniform soil” scenario the results of which are shown as an orange line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This simulation demonstrates that without a permeable subslab to effectively allow the ”advective potential” to be realized, existence of preferential pathway will actually not impact a VI site very much.
-In order for a preferential pathway to significantly contribute to VI, this requires a scenario involving good advective communication between it the indoor environment.
-These requirements were met at the ASU House.\par
-
-A perhaps obvious second requirement is that the preferential pathway must deliver contaminant vapors to be impactful.
-In another simulation, the permeable (gravel) subslab region was included, but the preferential pathway merely delivered clean air to the subslab.
-The result of this simulation is shown as the green line in the top panel of Figure \ref{fig:land_drain_scenarios}.
-This shows that while there was a lightly larger $\alpha_\mathrm{gw}$ compared to the ”uniform soil” scenario, it is nowhere near as significant as when the preferential pathway delivers contaminant vapors.
-The contaminated and uncontaminated preferential  pathway scenarios (blue and green lines respectively) thus bound the range of $\alpha_\mathrm{gw}$ that would be observed for a given $p_\mathrm{in/out}$ depending on the contaminant vapor concentration in the preferential pathway.\par
-
-The model is also able to capture the weak trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ when a preferential pathway is absent, but when there still exists a permeable subslab region.
-These results are shown in the bottom panel of Figure \ref{fig:land_drain_scenarios}.  
-These results are again in agreement with what was observed at the ASU House when the preferential pathway was closed, i.e.
-that there was a much more modest variation in indoor air concentration, irrespective of pressure, when the preferential pathway was cut off.\par
-
-The above simulations capture the trend in $\alpha_\mathrm{gw}$ with $p_\mathrm{in/out}$ but do not yet capture the full variability of the concentration results over the "most probable" portion of observed pressure distributions shown in Figure \ref{fig:kde} (which tend to be from -2 to +2 Pa).
-The results of Figure \ref{fig:land_drain_scenarios} show a spread of almost an order of magnitude over this pressure range for the case of the "open" preferential pathway, and almost no spread at all when the preferential pathway is "closed".
-Hence the predicted variability is roughly an order of magnitude too low, when considering only the influence of pressure.
-There is a factor that tends to increase the spread of the data one additional order of magnitude beyond what was predicted by the base calculations of Figure \ref{fig:land_drain_scenarios}.
-We believe that it is variations in air exchange rate, operating in concert with the natural variations in pressure differential, that explain the remaining variability.\par
-
-\subsubsection{Air Exchange Rate Effects}\label{s:results_model_air_exchange}
-
-Table \ref{tbl:air_exchange_rate} shows the observed variations in air exchange rates for the ASU House and Indianapolis House, compared with EPA’s summary of the distribution of typical residential air exchange rates\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}.
-Examination of these distributions point in a clear direction for modifying the above model.
-Instead of using a constant value of air exchange rate, as is customary, its values should be parameterized.
-A higher air exchange would of course be associated with lower $c_\mathrm{in}$ and vice versa.
-Moreover, $A_e$ may sometimes be correlated with $p_\mathrm{in/out}$.
-Determining any general relationship between $A_e$ and $p_\mathrm{in/out}$ is difficult: the structure itself and weather phenomena have a significant effect on air exchange.
-As the data in Figure \ref{fig:air_exchange_rate} show, there is no easily discernable correlation between these variables at the ASU site, though there is a hint of slight seasonal dependence.
-Note: a relationship between $A_e$ and $p_\mathrm{in/out}$ may be established for larger $p_\mathrm{in/out}$ via the building leakage curves, which are widely used for heating, ventilation and air conditioning systems in construction.\par
-
-\begin{figure}[htb!]
-  \includegraphics[width=\textwidth]{pressure_air_exchange_rate.pdf}
-  \caption{2D KDE figure showing distributions and relationship between indoor/outdoor pressure difference and air exchange rate.
-The seasonal median $p_\mathrm{in/out}$ and $A_e$ are indicated by the location of the respective arrow tips.
-Only non-CPM period considered.}\label{fig:air_exchange_rate}
-\end{figure}
-
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l c c c}
-    \toprule
-    \textbf{Percentile}                                                     & \textbf{10th} & \textbf{50th} & \textbf{90th} \\
-    EPA\cite{u.s._epa_exposure_2011,m._d._koontz_estimation_1995}           & 0.16-0.2      & 0.35-0.49     & 1.21-1.49     \\
-    ASU house\cite{holton_temporal_2013,guo_identification_2015}            & 0.21          & 0.43          & 0.78          \\
-    Indianapolis\cite{u.s._environmental_protection_agency_assessment_2015} & 0.34          & 0.74          & 1.27          \\
-    \bottomrule
-  \end{tabular}
-  \caption{Air exchange rate values (1/hr)}\label{tbl:air_exchange_rate}
-\end{table}
-
-To show the influence of possible statistical fluctuations of air exchange rate on the predictions of $\alpha_\mathrm{gw}$ values, the scenarios of Figure \ref{fig:land_drain_scenarios} were rerun calculated using the 5th and 95th percentile measured $A_e$ values, 0.17 and 0.90 respectively (based upon the actual distributions in Figure S1), providing predicted upper and lower bounds for $\alpha_\mathrm{gw}$.
-These bounds are indicated by the shaded blue regions around the center line calculated for an assumed constant $A_e$ of 0.5 per hour.\par
-
-It is apparent that assuming variability in air exchange rate allows capturing most of the observed variability in $\alpha_\mathrm{gw}$.
-We believe that this explains the portion of the variation in indoor air contaminant concentration data that cannot be explained by either existence of preferential pathways or by the range in indoor depressurization.
-Thus, we believe that it is the interplay of preferential pathway conditions, with indoor pressure variations and normal air exchange rates that help to explain the observations of significant variations in reported indoor air contaminant concentrations.\par
-
-\subsubsection{Results of Transient Simulations}\label{s:results_modeling_transient}
-
-The above analyses have been conducted under simulated steady state conditions.
-The conclusions regarding the importance of the different parameters are now examined in actual transient simulations.
-The model configuration of Figure \ref{fig:model} is run in 24-hour transient simulations to examine how $c_\mathrm{in}$ fluctuates over the course  of a "typical" day.
-The simulations vary $p_\mathrm{in/out}$ as one model input, and then assume either a constant or time-varying air exchange rate, $A_e$.
-The ASU House dataset was again the source of the "typical" $p_\mathrm{in/out}$ temporal variation, obtained by examining the median, hourly, diurnal $p_\mathrm{in/out}$ during the non-CPM periods.
-The statistically "typical" $p_\mathrm{in/out}$ cycle may be seen in the upper left panel of Figure \ref{fig:diurnal} (note that values between the hourly median values are interpolated using cubic splines).
-The "typical" air exchange rate is calculated in exactly the same way and is shown by the blue line in the upper right panel of Figure \ref{fig:diurnal}.
-The orange line is the air exchange rate value assumed for the calculations at constant air exchange rate.\par
-
-\begin{figure}
-  \centering
- \includegraphics[width=\textwidth]{diurnal.pdf}
- \caption{Transient simulation of a "typical" VI day, using diurnal indoor/outdoor pressure difference and air exchange rate as inputs. Effect of preferential pathway considered.}\label{fig:diurnal}
-\end{figure}
-
-The result of these simulations are shown in the bottom two panels of Figure \ref{fig:diurnal}, where the left and right panels show the results of open and closed preferential pathways, respectively.
-The "max change" value in the legends is the quotient of the lowest and highest predicted concentrations, i.e.
-a value of two indicate that the maximum daily concentration is twice as high as the lowest.
-This quantity may be compared with the value that is plotted for "one day" in Figure \ref{fig:resampling}.
-When the preferential pathway is open, there is a maximum daily variation of roughly a factor of 5, irrespective of whether $A_e$ fluctuates or not, which is somewhat more than the maximum daily variation shown in Figure \ref{fig:resampling}.
-The relatively small difference between the variable and constant $A_e$ cases indicates that most of the variability during a "typical" day is here attributable to fluctuations in $p_\mathrm{in/out}$, i.e.
-the contaminant transport into the modeled structure is advection dominated.
-Even for the small fluctuations in $p_\mathrm{in/out}$ the contaminant entry rate fluctuation drives the observed indoor concentration.
-When the preferential pathway is closed the story is quite different.
-When air exchange rate is held constant, there is essentially no variation in $c_\mathrm{in}$.
-This is again not surprising, as Figure \ref{fig:land_drain_scenarios} demonstrated that when the preferential pathway is closed, the influence of $p_\mathrm{in/out}$ on contaminant entry rate (and subsequently $c_\mathrm{in}$) is small.
-Combined with the small $p_\mathrm{in/out}$ this indicates that the contaminant transport into the modeled structure in this scenario is dominated by diffusion.
-When the air exchange rate is allowed to fluctuate, the maximum daily variation in $c_\mathrm{in}$ is 1.68, which is in line with what is shown in Figure \ref{fig:resampling}.
-This shows that for a "typical" day, when the preferential pathway is closed off, much of the daily variation in $c_\mathrm{in}$ is due to daily fluctuations in air exchange rate.\par
-
-These results demonstrate the complicated nature of temporal variability in $c_\mathrm{in}$.
-It is important to recall that only the effects of indoor/outdoor pressure difference and air exchange rate have been considered here, but slower processes, e.g.
-changes in groundwater contaminant concentration or various seasonal effects can also have a significant impact on VI over time.
-For the shorter time periods of concern in recent studies of temporal variability in indoor contaminant concentrations we believe that these are dominated by combinations of indoor/outdoor pressure differentials and air exchange rate.
-For a site where advective communication between the subsurface and the indoor is good, $p_\mathrm{in/out}$ is likely a significant determinant of $c_\mathrm{in}$ and its temporal variability.
-We have shown that that such a scenario may arise due to a preferential pathway entering a permeable sub-base, but may also exist even in the absence of a preferential pathway just as the results from NAS North Island demonstrate.
-At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
-
-\section{Conclusions}
-
-Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
-
-There can be a significant role of advective transport in determining observed variability.
-For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
-In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
-In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
-
-Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
-In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
-This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
-This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
-
-In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
-Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
-The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
-It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
-These results were consistent with seasonal timescale variations in the intrusion processes.\par
-
-% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
-- 24 hr sampling is recommended for a few reasons.
-1. Avoids issues with diurnal fluctations in Ae or dP.
-2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
-- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
-- Passive samples can help capture some of the more longterm variations if these are a concern.
-- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
-Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
-- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
-- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
-1. Even with "good" subslab sampling, hotspots are easily missed.
-2. May mislead that an indoor source is present.
-3. Even without a PP there may still be significant spatial variability (Eklund).
-4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
-5. McAlary said something about this. Look it up.
-6. May not really be worth the effort.
-
-\section*{Acknowledgements}
-This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).\par
-
-Declaration of interest: none
-
-\clearpage
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l l}
-    \toprule
-    % A/ALPHA
-    $A_\mathrm{ck}$ & Crack area \\
-    $A_e$ & Air exchange rate \\
-    $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
-    $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
-    % C
-    $c_\mathrm{in}$ & Indoor air contaminant concentration \\
-    $c$ & Soil-gas contaminant concentration \\
-    $c_w$ & Soil-water contaminant concentration \\
-    $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-    $\chi$ & PP contaminant concentration scaling parameter \\
-    % D
-    $D_\mathrm{eff}$ & Effective diffusion coefficient \\
-    $D_\mathrm{air}$ & Diffusion coefficient in air \\
-    $D_\mathrm{water}$ & Diffusion coefficient in water \\
-    % J
-    $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
-    % K
-    $\kappa_s$ & Saturated soil permeability \\
-    $K_H$ & Dimensionless Henry's law constant \\
-    $k_r$ & Relative permeability \\
-    % L
-    $L_\mathrm{slab}$ & Thickness of the foundation slab \\
-    % M
-    $M$ & Molar mass \\
-    $\mu$ & Contaminant vapor viscosity \\
-    % N
-    NAS & Naval Air Stations \\
-    % P
-    $p$ & Pressure in soil \\
-    $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
-    PP & Preferential pathway \\
-    % R
-    $\rho$ & Density \\
-    % S
-    $\mathrm{Se}$ & Soil water saturation \\
-    % T
-    $t$ & time \\
-    $\theta_g$ & Vapor/gas filled porosity \\
-    $\theta_w$ & Water filled porosity \\
-    $\theta_r$ & Residual water filled porosity \\
-    $\theta_t$ & Total porosity \\
-    % U
-    $\vec{u}$ & Soil-gas velocity (vector quantity) \\
-    % V
-    VI & Vapor intrusion \\
-    $V_\mathrm{base}$ & Basement volume \\
-    % Z
-    $z$ & Elevation above groundwater \\
-    \bottomrule
-  \end{tabular}
-  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
-\end{table}
-
-\bibliographystyle{elsarticle-num}
-\bibliography{library}
-
-\end{document}

### 2019-09-24 doc/results.tex
Reorganizing

--- 2019-06-03 tex/manuscript.tex
@@
-\documentclass[preprint,12pt]{elsarticle}
-
-% additional packages
-\usepackage[utf8]{inputenc}
-%\bibliographystyle{abbrvnat}
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{subcaption}
-\usepackage{tabularx}
-\usepackage{array,multirow,multicol,graphicx}
-\usepackage{comment}
-\graphicspath{
-  {./figures/},
-}
-\usepackage{lineno}
-\linenumbers
-% macros
-
-\journal{Building and Environment}
-
-\begin{document}
-
-\begin{frontmatter}
-% title
-\title{Factors Affecting Temporal Variations In Vapor Intrusion-Induced Indoor Air Contaminant Concentrations}
-
-% authors
-\author[Brown]{Jonathan G. V. Ström}
-\author[ASU]{Yuanming Guo}
-\author[ASU]{Yijun Yao}
-\author[Brown]{Eric M. Suuberg}
-
-\ead{eric\_suuberg@brown.edu}
-
-\address[Brown]{Brown University, School of Engineering, Providence, RI, USA}
-\address[ASU]{Arizona State University, School of Sustainable Engineering and the Building Environment, Tempe, AZ, USA}
-
-\begin{abstract}
-Temporal variability in indoor air contaminant concentrations at vapor intrusion (VI) sites has been a concern for some time.
-We consider the source of the reported variability at VI sites located near Hill Air Force Base (AFB) in Utah, an EPA experimental house in Indiana, and Naval Air Station North Island in California.
-We focus in particular on how the indoor/outdoor pressure differences and air exchange rates affected indoor air contaminant concentrations at these sites.
-We investigate how these dynamics differ for a site that is characterized by a preferential pathway (like Hill AFB) and VI sites that are not influenced by such pathways, using three-dimensional fluid dynamics models and statistical analysis of the aforementioned sites.
-A preferential pathway can dramatically increase a VI site's sensitivity to build pressurization, provided there exist a medium allowing effective communication between a contaminant-delivering preferential pathway and the indoor air space, e.g.
-a permeable subslab space that may be provided by a gravel layer.
-Preferential pathways may also erroneously indicate the presence of indoor contaminant sources.
-At sites characterized by significant advective transport from the subslab to the indoor air space, much of the short-term variability in indoor air contaminant concentration can be explained by an impact of fluctuations in indoor/outdoor pressure differences.
-Meanwhile, air exchange rate variation drives most of the short-term variability at sites characterized by minor variations in advective transport.
-\end{abstract}
-
-\begin{keyword}
-  Vapor intrusion \sep Preferential pathways \sep Temporal variability \sep Attenuation factor \sep Air exchange rate \sep Indoor/outdoor pressure difference
-\end{keyword}
-
-\end{frontmatter}
-
-\section{Introduction}\label{s:intro}
-
-Long term vapor intrusion (VI) studies in both residential and larger commercial structures have raised concerns regarding significant observed transient behavior in indoor air contaminant concentrations\cite{u.s._environmental_protection_agency_oswer_2015,folkes_observed_2009,holton_temporal_2013,johnston_spatiotemporal_2014,hosangadi_high-frequency_2017,mchugh_recent_2017,u.s._environmental_protection_agency_assessment_2015}.
-Such variations make it difficult for those charged with protecting human health to formulate a response - should evaluation of the risk of exposure be based upon observed peak concentrations, or long-term averages, or something else?
-There is even uncertainty within the VI community regarding how to best develop sampling strategies to address this problem\cite{u.s._environmental_protection_agency_oswer_2015,holton_temporal_2013,johnson_integrated_2016}.
-What represents a reasonable sampling strategy for a particular site a single 8-hour sample?
-Repeated 8-hour samples?
-Month-long samples?
-Continuous monitoring?\par
-
-VI involves the migration of volatilizing contaminants from soil, groundwater or other subsurface sources into overlying structures.
-The basic nature of VI has been understood for some time and it has been the subject of much study, but some aspects remain poorly understood, such as the causes of the sometimes observed large temporal transients in indoor air concentrations.
-Results from a house operated by Arizona State University (ASU) near Hill AFB in Utah, an EPA experimental house in Indianapolis, IN and a large warehouse at the Naval Air Station (NAS) North Island, CA have all shown significant transient variations in indoor air contaminant concentrations.
-All were outfitted with sampling and monitoring equipment that allowed tracking temporal variation in indoor air contaminant concentrations on time scales of hours.
-All have shown that these concentrations vary significantly with time - orders of magnitude on the timescale of a day or days\cite{holton_evaluation_2015,guo_vapor_2015,hosangadi_high-frequency_2017}.\par
-
-In one instance the source of the variation was clearly established during the study of the site.
-At the ASU house a drain pipe (or “land drain”) connected to a sewer system was discovered beneath the house.
-Careful isolation of this source led to a clear conclusion that this “preferential pathway” for contaminant vapor migration significantly contributed to observed indoor air contaminant levels and their fluctuations\cite{guo_vapor_2015,guo_identification_2015}.
-While in this case the issue of a contribution from a preferential pathway was clearly resolved, what it left open was a question of whether existence of such a preferential pathway would always be expected to lead to large fluctuations in indoor air contaminant concentrations.\par
-
-Similarly, a sewer pipe has recently been suggested to be a source of the contaminants found in the EPA Indianapolis house.
-That site was also characterized by large indoor air contaminant concentration fluctuations\cite{mchugh_evidence_2017,u.s._environmental_protection_agency_assessment_2015}.
-Sewer lines have been previously implicated as VI sources at several sites\cite{pennell_sewer_2013,mchugh_evidence_2017,roghani_occurrence_2018,riis_vapor_2010}.
-A Danish study has estimated that roughly 20\% of all VI sites in central Denmark involve significant sewer VI pathways\cite{nielsen_remediation_2017}.
-Thus while consideration of sewer or other preferential pathways is now part of normal good practice in VI site investigation\cite{u.s._environmental_protection_agency_oswer_2015}, it is still not known whether the existence of such pathways automatically means that large temporal fluctuations are necessarily to be expected.\par
-
-In some of the above cited cases\cite{pennell_sewer_2013,riis_vapor_2010}, a sewer provided a pathway for direct entry of contaminant into the living space.
-While potentially important in many instances, this scenario is not further considered here where the focus is on pathways that deliver contaminant via the soil beneath a structure.
-It is, however, now known that even absent a preferential pathway, there may be significant transient variation in indoor air contaminant concentrations at VI sites\cite{folkes_observed_2009,brenner_results_2010,johnston_spatiotemporal_2014}.
-One example is the site at NAS North Island at which no preferential pathways have been identified.
-Instead, a building at this site is characterized by significant temporal variations in indoor-outdoor pressure differential\cite{hosangadi_high-frequency_2017}.
-It is believed that this is the origin of the observed indoor air contaminant concentration fluctuations at that site.\par
-
-This paper investigates the sources of the temporal variation in indoor air contaminant concentrations in both the presence and absence of preferential pathways.
-In this work, the latter scenarios are referred to as “normal” VI scenarios, in which there is typically a groundwater source of the contaminant.
-Specifically, we pose the question of just how much variation in indoor air contaminant concentration may be expected at such normal VI sites vs.
-those characterized by preferential pathways within the soil beneath the site.
-The conditions required for preferential pathways to become significant contributors to temporal variations in indoor air contaminant concentrations are also explored, and the consequences for sampling strategies are discussed.\par
-
-\section{Methods}\label{s:methods}
-
-\subsection{Statistical Analysis Of Field Data}\label{s:methods_stats}
-
-To frame the question of just how much variability in indoor air contaminant concentrations is actually observed, field datasets have been analyzed.
-For this purpose, datasets from the ASU house in Utah, the EPA Indianapolis site and North Island NAS were chosen for analysis.
-Readers are referred to the original published works for details regarding data acquisition\cite{holton_evaluation_2015,guo_vapor_2015,holton_temporal_2013,hosangadi_high-frequency_2017,u.s._environmental_protection_agency_assessment_2015}.\par
-
-The ASU house data were obtained over a period of several years.
-During part of this time, controlled pressure method (CPM) tests were being conducted, in which the house was underpressurized to an extent greater than that characterizing “normal” house operation: increasing VI potential\cite{mchugh_evaluation_2012,mchugh_recent_2017,holton_evaluation_2015}.
-The period of CPM testing is thus excluded from the analysis.
-Likewise, the existence of a preferential pathway at the ASU house needs to be considered in examining that dataset; during some of the testing at that site, this pathway was cut off, resulting in “normal” VI conditions in which the main source of contaminant was diffusion of contaminant vapor from an underlying groundwater source.\par
-
-The NAS North Island dataset has not (as far as is known) been influenced by a preferential pathway, but the structure there was subject to “large” internal pressure fluctuations.
-By “large” is meant still only of order 10-20 Pa, but these were greater than those generally recorded at the ASU house during normal operations.
-The underlying soil at NAS North Island is sandy\cite{hosangadi_high-frequency_2017} and more permeable than that at the ASU site, which will be shown to lead to greater pressure sensitivity in the former case.\par
-
-The Indianapolis site investigation also spanned a number of years and periodically included the testing of a sub-slab depressurziation system (SSD) for VI mitigation.
-Only the period before the installation of this system was considered in the present analysis.
-It is likely a sewer line beneath the structure acted as a preferential pathway\cite{mchugh_evidence_2017}.
-Unlike at the ASU house, this preferential pathway was never removed or blocked, making it impossible to isolate the role of the preferential pathway at this site.
-It is still of interest to consider the data from this site because of the completeness and extensiveness of the data collection.
-Figure \ref{fig:indianapolis} illustrates a typical reported series of indoor air trichloroethylene (TCE) concentration measurements from this site.
-There is almost a two order of magnitude variation in the concentration data.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{time_indianapolis.pdf}
- \caption{Typical data on indoor air TCE contaminant concentrations at the Indianapolis site\cite{u.s._environmental_protection_agency_assessment_2015}.}\label{fig:indianapolis}
-\end{figure}
-
-Some of the analysis of the above three field data sets relies on a probability density estimation technique called ”kernel density estimation” (KDE).
-KDE is a technique used for estimating the probability distribution of a random variable(s) by using multiple kernels, or weighting functions to characterize the data sets.
-In this case, Gaussian kernels are used to create the KDEs.
-This means that it is presumed that the variables of interest (i.e., indoor air contaminant concentrations and indoor-outdoor pressure differentials, as sampled) are normally distributed around mean values and that there are statistical fluctuations associated with each sampling event.
-In this instance, the scipy statistical package was used to construct the KDEs, assuming a bandwidth parameter determined by Scott’s rue.
-The SciPy Python library was used to conduct all statistical analysis and data processing\cite{jones_scipy_2011}.\par
-
-\subsection{Modeling Work}\label{s:methods_model}
-
-A previously described three-dimensional computational fluid dynamics model of a generic VI impacted house has been used to elucidate certain aspects of transient VI processes.
-In the present work, there has been an addition of a preferential pathway to the “standard” model that has been described before in publications by this group\cite{shen_influence_2013,yao_investigating_2017,yao_three-dimensional_2017}.
-As in the earlier studies, only the vadose zone soil domain is directly modeled.
-Figure \ref{fig:model} shows a cutaway view of the relevant modeling domain.\par
-
-\begin{figure}[htb!]
-  \centering
- \includegraphics[width=\textwidth]{model.png}
- \caption{Foundation and vadose zone soil represented in the modeling.
-Note that here a gravel sub-base material is shown, but in certain simulations, that material is absent and the surrounding soil directly contacts the foundation slab.
-Different assumptions are made regarding the preferential pathway, here shown as a pipe entering the gravel sub-base.
-In some cases, the preferential pathway has been "turned off".}\label{fig:model}
-\end{figure}
-
-The modeled VI impacted structure is assumed to have a 10x10 m foundation footprint, with the bottom of the foundation slab lying 1 m below ground surface (bgs), simulating a house with a basement.
-The indoor air space is modeled as a continuously stirred tank (CST)\cite{u.s._environmental_protection_agency_oswer_2015} and all of the contaminant entering the house is assumed to enter with soil gas through a 1 cm wide crack located between the foundation walls and the foundation slab around the  perimeter of the house.
-All of the contaminant leaving the indoor air space is assumed to do so via air exchange with the ambient.
-The indoor control volume is here assumed to consist of only of the basement, having a total volume of $300 \; \mathrm{m^3}$.
-Clearly different assumptions could be made regarding the structural features and the size of the crack entry route, but for present purposes, this is unimportant as the intent is only to show for “typical” values what the influence of some critical parameters is.\par
-
-The modeled surrounding soil domain extends 5 meters from the perimeter of the house and is assumed to consist of sandy loam, except as noted otherwise.
-Directly beneath the foundation slab, there is assumed to be a 30 cm (one foot) thick gravel layer, except in certain cases here this sub-base material is assumed to be the same as the surrounding soil (termed a "uniform" soil scenario).
-The groundwater beneath the structure is assumed to be homogeneously contaminated with TCE selected as a prototypical contaminant.
-The groundwater itself is not modeled, as the bottom of the model domain is defined by the top of the water table.
-Where relevant, the preferential pathway is modeled as a 10 cm (4”) pipe that opens into the gravel sub-base beneath the structure.
-The air in the pipe is also assumed to be contaminated with TCE at a vapor concentration equal to the vapor in equilibrium with the groundwater contaminant concentration below the structure, modified by a scaling factor $\chi$ (allowing the contaminant concentration in the pipe to be parameterized).
-This model illustrates the concept of a "preferential pathway", as the pipe carries contaminant vapor to the immediate vicinity of the foundation, by a path that circumvents the usual soil diffusion pathway.\par
-
-The ground surface and the pipe are both sources of air to the soil domain.
-Both are assumed to exist at reference atmospheric pressure.
-Soil gas transport is governed by Richard’s equation, a modified version of Darcy’s Law, taking the variability of soil moisture in the vadose zone into account\cite{richards_capillary_1931}.
-The van Genuchten equations are used to predict the soil moisture content and thus the effective permeability of the soil\cite{van_genuchten_closed-form_1980}.
-The effective diffusivity of contaminant in soil is calculated using the Millington-Quirk model\cite{millington_permeability_1961}.
-The transport of contaminant vapor in the soil is assumed to be governed by the advection-diffusion equation, in which either advection or diffusion may dominate depending upon position and particular circumstances.
-The key working equations and the boundary conditions are summarized in Table \ref{tbl:eqns_bc_parameters}.\par
-
-\begin{table}[htb!]
-  \setlength{\tabcolsep}{1pt}
-  \centering
-  \begin{tabular}{l c c c c c c}
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % GOVERNING EQUATIONS
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \toprule
-    \multicolumn{7}{c}{\textbf{Governing Equations}} \\
-    \midrule
-    % CSTR
-    Unsteady-CST & \multicolumn{6}{c}{$V\frac{d c_\mathrm{in}}{d t} = \int_{A_\mathrm{ck}} j_\mathrm{ck} dA - c_\mathrm{in} A_e V_\mathrm{slab}$} \\
-    % RICHARDS
-    Richard's & \multicolumn{6}{c}{$\nabla \cdot \rho \Big( - \frac{\kappa_s}{\mu} k_r \nabla p \Big) = 0$} \\
-    % TRANSPORT
-    Transport & \multicolumn{6}{c}{$\frac{\partial}{\partial t} \Big( \theta_w c_w + \theta_g c \Big) = \nabla (D_\mathrm{eff} \cdot \nabla c) - \vec{u} \cdot \nabla c$} \\
-    % MILLINGTON-QUIRK
-    Millington-Quirk & \multicolumn{6}{c}{$D_\mathrm{eff} = D_\mathrm{air}\frac{\theta_g^{10/3}}{\theta_t^2} + \frac{D_\mathrm{water}}{K_H} \frac{\theta_w^{10/3}}{\theta_t^2}$} \\
-    % VAN GENUCHTEN
-    \multirow{4}{*}{van Genuchten} & \multicolumn{6}{c}{$\mathrm{Se} = \frac{\theta_w - \theta_r}{\theta_t - \theta_r} = [1 + |\alpha z|^n]^{-m}$} \\
-     & \multicolumn{6}{c}{$\theta_g = \theta_t - \theta_w$}\\
-     & \multicolumn{6}{c}{$k_r = (1-\mathrm{Se})^{l}[1(\mathrm{Se}^{m})^m]^2$} \\
-     & \multicolumn{6}{c}{$m = 1 - 1/n$} \\
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-     % BOUNDARY CONDITIONS
-     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Boundary Conditions}} \\
-    \midrule
-    \textbf{Boundary} & \multicolumn{3}{c}{\textbf{Richard's Eqn.}} & \multicolumn{3}{c}{\textbf{Transport Eqn.}} \\
-    % Foundation crack
-    Foundation crack & \multicolumn{3}{c}{$p = p_\mathrm{in/out} \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$j_\mathrm{ck} = \frac{u c}{1 - \exp{(u L_\mathrm{slab}/D_\mathrm{air})}}$} \\
-    % Groundwater source
-    Groundwater & \multicolumn{3}{c}{\textit{No flow}} & \multicolumn{3}{c}{$c = c_\mathrm{gw} K_H \; \mathrm{(\mu g/m^3)}$} \\
-    % Ground surface
-    Ground surface & \multicolumn{3}{c}{$p = 0 \; \mathrm{(Pa)}$} & \multicolumn{3}{c}{$c = 0 \; \mathrm{(\mu g/m^3)}$} \\
-    % Preferential pathway
-    \multirow{2}{*}{\begin{minipage}{0.2\textwidth}Preferential\\pathway\end{minipage}} & \multicolumn{3}{c}{\multirow{2}{*}{$p = 0 \; \mathrm{(Pa)}$}} & \multicolumn{3}{c}{\multirow{2}{*}{$c = c_\mathrm{gw} K_H \chi \; \mathrm{(\mu g/m^3)}$}} \\ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % SOIL PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Soil Properties}\cite{dan_capillary_2012,abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-    \textbf{Soil} & \textbf{$\kappa_s \; \mathrm{(m^2)}$} & \textbf{$\theta_s$} & \textbf{$\theta_r$} & \textbf{$\alpha \; \mathrm{(1/m)}$} & \textbf{$n$} \\
-    % GRAVEL
-    Gravel & $1.3 \cdot 10^{-9}$ & 0.42 & 0.005 & 100 & 3.1 \\
-    % SANDY LOAM
-    Sandy Loam & $5.9 \cdot 10^{-13}$ & 0.39 & 0.039 & 2.7 & 1.4 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % TCE PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Trichloroethylene (diluted in air) Properties}\cite{abreu_conceptual_2012,u.s._environmental_protection_agency_userss_2004}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$D_\mathrm{air} \; \mathrm{(m^2/h)}$}} & \textbf{$D_\mathrm{water} \; \mathrm{(m^2/h)}$} & \textbf{$\rho \; \mathrm{(kg/m^3)}$} & \textbf{$\mu \; \mathrm{(Pa \cdot s)}$} & \textbf{$K_H$} \\
-     & \multicolumn{1}{c}{$2.47 \cdot 10^{-2}$} & $3.67 \cdot 10^{-6}$ & 1.614 & $1.86 \cdot 10^{-5}$ & 0.403 \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    % BUILDING PROPERTIES
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \midrule
-    \multicolumn{7}{c}{\textbf{Building Properties}} \\
-    \midrule
-     & \multicolumn{1}{c}{\textbf{$V_\mathrm{base} \; \mathrm{(m^3)}$}} & \textbf{$L_\mathrm{slab} \; \mathrm{(cm)}$} & \textbf{$A_e \; \mathrm{(1/hr)}$} \\
-     & \multicolumn{1}{c}{300} & 15 & 0.5 \\
-    \bottomrule
-  \end{tabular}
-  \caption{Governing equations, boundary conditions \& model input parameters. See Table \ref{tbl:abbreviations} for nomenclature.}\label{tbl:eqns_bc_parameters}
-\end{table}
-
-\newpage
 \section{Results \& Discussion}\label{s:results}
@@
 At sites where advective transport into the structure is limited, much of the temporal variability in $c_\mathrm{in}$ may be attributed to natural fluctuations in air exchange rate.\par
-
-\section{Conclusions}
-
-Based on statistical analysis of field data from vapor intrusion sites showing significant temporal variations in indoor air contaminant concentrations, supplemented by computational fluid dynamics modeling, it is concluded that several different factors can play a role in determining these variations.\par
-
-There can be a significant role of advective transport in determining observed variability.
-For advective transport to play a decisive role, the present results show that there must exist a sub-foundation zone that permits significant flow of soil gas.
-In the case of NAS North Island, a relatively permeable sub-foundation soil permitted fairly high variations in building indoor pressure to cause significant fluctuations in contaminant entry rate.
-In the case of the ASU house, a permeable gravel sub-base allowed good communication of the indoor air environment with the contents of a preferential contaminant pathway that entered that sub-base, even though the pressure driving force was not nearly as large as that at NAS North Island.
-Both situations led to well-documented large variations (two or three orders of magnitude) in indoor air contaminant concentrations.
-
-Modeling work has shown that the existence of preferential pathways that communicate with a permeable sub-slab zone can create significant spatial variability in contaminant concentration in the subslab.
-In other words, the existence of a permeable sub-slab layer (such as gravel) does not assure that there will exist a well-mixed sub-slab zone.
-This warns that taking isolated subslab vapor samples may be misleading, as a result of spatial variability of contaminant concentrations.
-This might in turn lead to apparent attenuation from the sub-slab to indoors that appears to exceed the EPA recommended value of 0.03 and even exceed unity - potentially leading to the erroneous conclusion that an indoor source is present.\par
-
-In the absence of a preferential pathway (a situation that existed at the ASU house when the pathway was closed off) and in the presence of only a modest observed pressure driving force, there was still observed a significant temporal variation in indoor air contaminant concentration of around an order of magnitude.
-Statistical analysis of that data set showed, however, that indoor air contaminant concentrations were actually unlikely to vary by more than a factor of three over a week-long or shorter sampling period.
-The small variability that was seen over these short times could be explained by inherent variability in air exchange rates.
-It was when indoor air samples taken over longer periods were compared that an almost order of magnitude variation became apparent.
-These results were consistent with seasonal timescale variations in the intrusion processes.\par
-
-% PRACTICAL IMPLICATIONS (WEAVE THESE INTO THE TEXT SOMEHOW)
-- 24 hr sampling is recommended for a few reasons.
-1. Avoids issues with diurnal fluctations in Ae or dP.
-2. Variation of a few factors is expected over 24 hrs, while not much else in subsequent days. Thus most short-term variability may be captured in this timeperiod.
-- If you are more than (say an order of magnitude) from some regulatory limit in a 24 hr sample, additional sampling at a later date may not be necessary.
-- Passive samples can help capture some of the more longterm variations if these are a concern.
-- Use manometers to determine dP, if these are relatively large, and if the structure is relatively leaky/the soil is quite permeable, more significant flucations in IACC may be to expected and a "tighter" sampling scheme may be warranted.
-Of course PP may make even small dP cause significant changes in IACC (which should be screened for).
-- PPs should be screened for. Danish study has strategies for this, as well as Pennell. Sampling nearby manholes recommended as these are likely pathways.
-- Subslab samples may be missleading. At ASU we saw that alpha > 1 existed, which shows that:
-1. Even with "good" subslab sampling, hotspots are easily missed.
-2. May mislead that an indoor source is present.
-3. Even without a PP there may still be significant spatial variability (Eklund).
-4. Holton's tracer study shows that the indoor itself may cause persistent plumes (especially in the deep subslab area).
-5. McAlary said something about this. Look it up.
-6. May not really be worth the effort.
-
-\section*{Acknowledgements}
-This project was supported by grant ES-201502 from the Strategic Environmental Research and Development Program and Environmental Security Technology Certification Program (SERDP-ESTCP).\par
-
-Declaration of interest: none
-
-\clearpage
-\begin{table}[htb!]
-  \centering
-  \begin{tabular}{l l}
-    \toprule
-    % A/ALPHA
-    $A_\mathrm{ck}$ & Crack area \\
-    $A_e$ & Air exchange rate \\
-    $\alpha, \; n, \; m, \; l$ & van Genuchten parameters \\
-    $\alpha_\mathrm{gw}$ & Attenuation from groundwater contaminant vapor source \\
-    % C
-    $c_\mathrm{in}$ & Indoor air contaminant concentration \\
-    $c$ & Soil-gas contaminant concentration \\
-    $c_w$ & Soil-water contaminant concentration \\
-    $c_\mathrm{gw}$ & Contaminant groundwater concentration \\
-    $\chi$ & PP contaminant concentration scaling parameter \\
-    % D
-    $D_\mathrm{eff}$ & Effective diffusion coefficient \\
-    $D_\mathrm{air}$ & Diffusion coefficient in air \\
-    $D_\mathrm{water}$ & Diffusion coefficient in water \\
-    % J
-    $j_\mathrm{ck}$ & Contaminant molar flux through the foundation crack \\
-    % K
-    $\kappa_s$ & Saturated soil permeability \\
-    $K_H$ & Dimensionless Henry's law constant \\
-    $k_r$ & Relative permeability \\
-    % L
-    $L_\mathrm{slab}$ & Thickness of the foundation slab \\
-    % M
-    $M$ & Molar mass \\
-    $\mu$ & Contaminant vapor viscosity \\
-    % N
-    NAS & Naval Air Stations \\
-    % P
-    $p$ & Pressure in soil \\
-    $p_\mathrm{in/out}$ & Indoor/outdoor pressure difference \\
-    PP & Preferential pathway \\
-    % R
-    $\rho$ & Density \\
-    % S
-    $\mathrm{Se}$ & Soil water saturation \\
-    % T
-    $t$ & time \\
-    $\theta_g$ & Vapor/gas filled porosity \\
-    $\theta_w$ & Water filled porosity \\
-    $\theta_r$ & Residual water filled porosity \\
-    $\theta_t$ & Total porosity \\
-    % U
-    $\vec{u}$ & Soil-gas velocity (vector quantity) \\
-    % V
-    VI & Vapor intrusion \\
-    $V_\mathrm{base}$ & Basement volume \\
-    % Z
-    $z$ & Elevation above groundwater \\
-    \bottomrule
-  \end{tabular}
-  \caption{List of abbreviations and nomenclature}\label{tbl:abbreviations}
-\end{table}
-
-\bibliographystyle{elsarticle-num}
-\bibliography{library}
-
-\end{document}

### 2019-09-25 code/analysis.py
Working on more reorganization

@@ -1,0 +1,70 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+from scipy.integrate import odeint, ode, solve_ivp
+from scipy.optimize import curve_fit, leastsq
+from scipy.interpolate import interp1d
+
+
+class Data:
+    def __init__(self, file):
+        self.path = file
+        return
+
+    def get_path(self):
+        return self.path
+
+    def get_data(self):
+        return self.data
+
+
+class COMSOL(Data):
+    """
+    Class to load, process, and return the COMSOL simulation data
+    """
+
+    def __init__(self, file):
+        Data.__init__(self, file)
+        self.process_raw_data()
+        self.A_ck = 0.1 * 4
+        return
+
+    def get_crack_area(self):
+        return self.A_ck
+
+    def get_raw_data(self):
+        path = self.get_path()
+        return pd.read_csv(path, header=4)
+
+    def get_renaming_scheme(self):
+        renaming = {'% Time (h)': 'time', 'p_in (Pa)': 'p_in', 'alpha (1)': 'alpha',
+                    'c_in (ug/m^3)': 'c_in', 'j_ck (ug/(m^2*s))': 'j_ck', 'm_ads (g)': 'm_ads',
+                    'c_ads (mol/kg)': 'c_ads',
+                    'c_ads/c_gas (1)': 'c_ads/c_gas', 'alpha_ck (1)': 'alpha_ck',
+                    'n_ck (ug/s)': 'n_ck', 'Pe (1)': 'Pe', 'c_gas (ug/m^3)': 'c_gas',
+                    'u_ck (cm/h)': 'u_ck', '% K_ads (m^3/kg)': 'K_ads',
+                    't (h)': 'time', 'c_ads_vol (ug/m^3)': 'c_ads_vol', 'c_liq (ug/m^3)': 'c_liq',
+                    '% Pressurization cycles index': 'p_cycle',
+                    'Pressurization cycles index': 'p_cycle'}
+        return renaming
+
+    def process_raw_data(self):
+        raw_df = self.get_raw_data()
+        self.data = raw_df.rename(columns=self.get_renaming_scheme())
+        return
+
+    def get_time_data(self):
+        df = self.get_data()
+        return df['time'].values
+
+    def get_entry_flux_data(self):
+        df = self.get_data()
+        return df['j_ck'].values
+
+    def get_entry_rate_data(self):
+        df = self.get_data()
+        return df['n_ck'].values
+
[…]

### 2019-09-25 code/figure_old.py

@@ -1,0 +1,483 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import matplotlib.gridspec as gridspec
+from scipy import stats
+from get_simulation_data import Soil, PreferentialPathway
+import datetime
+
+"""
+This figure shows a case (North Island NAS) where indoor/outdoor pressure
+difference the key driver at a VI site, and how are are able to model this
+by assuming a permeable soil type (sand).
+"""
+class PressureKDE:
+    def __init__(self,y_data_log=False , norm_conc=True):
+        nas = pd.read_csv('./data/north_island.csv').dropna()
+        asu = pd.read_csv('./data/asu_house.csv').dropna()
+        asu_open = asu.loc[asu['Phase']=='Open']
+        asu_closed = asu.loc[asu['Phase']=='Closed']
+
+
+        x='IndoorOutdoorPressure'
+
+        if y_data_log is True:
+            y='logIndoorConcentration'
+        else:
+            y='IndoorConcentration'
+
+
+        datasets = (nas, asu_open, asu_closed,)
+        labels = ('North Island NAS', 'ASU house, PP open', 'ASU house, PP closed',)
+
+        fig, ax = plt.subplots()
+
+        for data, label in zip(datasets, labels):
+            if norm_conc is True:
+                if y_data_log is True:
+                    data2 = data[y]-data[y].mean()
+                else:
+                    data2 = data[y]/data[y].mean()
+
+            else:
+                data2 = data[y]
+
+
+            r, p = stats.pearsonr(data[x], data2)
+
+            print(label)
+            print(data[x].describe(percentiles=[0.05,0.95]), data['IndoorConcentration'].describe(percentiles=[0.05,0.95]), data2.describe(percentiles=[0.05,0.95]))
+            sns.kdeplot(
+                data=data[x],
+                data2=data2,
+                shade_lowest=False,
+                shade=True,
+                label=label + ', r = %1.2f' % r,
+                ax=ax,
+            )
+
+
+
+        yticks, yticklabels = get_log_ticks(-1, 1.5)
+        # formatting options
+        ax.set(
+            xlim=[-30,15],
+            ylim=[-1,1.5],
[…]

### 2019-09-25 code/figures.py

@@ -1,483 +1,12 @@
 import numpy as np
 import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-import matplotlib.gridspec as gridspec
-from scipy import stats
-from get_simulation_data import Soil, PreferentialPathway
-import datetime
 
-"""
-This figure shows a case (North Island NAS) where indoor/outdoor pressure
-difference the key driver at a VI site, and how are are able to model this
-by assuming a permeable soil type (sand).
-"""
-class PressureKDE:
-    def __init__(self,y_data_log=False , norm_conc=True):
-        nas = pd.read_csv('./data/north_island.csv').dropna()
-        asu = pd.read_csv('./data/asu_house.csv').dropna()
-        asu_open = asu.loc[asu['Phase']=='Open']
-        asu_closed = asu.loc[asu['Phase']=='Closed']
 
+def sub_slab_attenuation():
+    df = get_asu_house_data().set_index('CPM')
 
-        x='IndoorOutdoorPressure'
+    sns.boxplot(x="CPM", y="alpha_slab", data=df)
 
-        if y_data_log is True:
-            y='logIndoorConcentration'
-        else:
-            y='IndoorConcentration'
 
-
-        datasets = (nas, asu_open, asu_closed,)
-        labels = ('North Island NAS', 'ASU house, PP open', 'ASU house, PP closed',)
-
-        fig, ax = plt.subplots()
-
-        for data, label in zip(datasets, labels):
-            if norm_conc is True:
-                if y_data_log is True:
-                    data2 = data[y]-data[y].mean()
-                else:
-                    data2 = data[y]/data[y].mean()
-
-            else:
-                data2 = data[y]
-
-
-            r, p = stats.pearsonr(data[x], data2)
-
-            print(label)
-            print(data[x].describe(percentiles=[0.05,0.95]), data['IndoorConcentration'].describe(percentiles=[0.05,0.95]), data2.describe(percentiles=[0.05,0.95]))
-            sns.kdeplot(
-                data=data[x],
-                data2=data2,
-                shade_lowest=False,
-                shade=True,
-                label=label + ', r = %1.2f' % r,
-                ax=ax,
-            )
-
-
-
-        yticks, yticklabels = get_log_ticks(-1, 1.5)
[…]

### 2019-09-25 code/simulation_predictions.py

@@ -44,7 +44,7 @@ def get_interp_func(df):
     Ae = df['AirExchangeRate']
     alpha = df['logAttenuationGroundwater']
 
-    
+
     # 2d interpolation function
     interp_func = interp2d(
         p_in,
@@ -117,51 +117,6 @@ df['AirExchangeRate'] *= 3600 # convert from 1/s to 1/hr
 df['logIndoorConcentration'] = df['IndoorConcentration'].apply(np.log10)
 df['logAttenuationSubslab'] = df['AttenuationSubslab'].apply(np.log10)
 df['logAttenuationGroundwater'] = df['AttenuationGroundwater'].apply(np.log10)
-
-
-
-"""
-sim_cases = ('Pp','No Pp',)
-phases = ('Open','Closed',)
-num_samples = 15
-for sim_case, phase in zip(sim_cases, phases):
-    df_sort = df.loc[df['Simulation']==sim_case] # sorts simulation types
-    asu_sort = asu.loc[asu['Phase']==phase]
-
-    interp_func = get_interp_func(df_sort)
-    samp_p, samp_ae, prediction = get_prediction(asu_sort)
-    make_plots(asu_sort)
-    plt.savefig(fig_dir+'sampling_simulation_pp_'+phase.lower()+ext,dpi=dpi)
-
-
-
-# seasonal analysis
-seasons = asu['Season'].unique()
-sim_cases = ('Pp','No Pp',)
-phases = ('Open','Closed',)
-num_samples = 20
-for season in seasons:
-    for sim_case, phase in zip(sim_cases, phases):
-        try:
-            df_sort = df.loc[df['Simulation']==sim_case] # sorts simulation types
-            asu_sort = asu.loc[(asu['Phase']==phase) & (asu['Season']==season)]
-
-            interp_func = get_interp_func(df_sort)
-            samp_p, samp_ae, prediction = get_prediction(asu_sort)
-            make_plots(asu_sort)
-            titles = {
-                'Pp': 'Preferential pathway open',
-                'No Pp': 'Preferential pathway closed',
-            }
-
-            plt.suptitle(titles[sim_case]+' - '+season.lower(),y=1.0)
-            plt.savefig(fig_dir+'sampling_simulation_pp_'+phase.lower()+'_'+season.lower()+ext,dpi=dpi)
-        except:
-            continue
-
-"""
-# "Ae-span" figure
-
 
 class SimPrediction:
     def __init__(self,status='Closed'):

### 2019-09-25 code/utils.py

@@ -1,0 +1,22 @@
+import os
+dropbox_folder = None
+
+def get_dropbox_path():
+    for dirname, dirnames, filenames in os.walk(os.path.expanduser('~')):
+        for subdirname in dirnames:
+            if(subdirname == 'Dropbox'):
+                dropbox_folder = os.path.join(dirname, subdirname)
+                break
+        if dropbox_folder:
+            break
+    return dropbox_folder
+
+
+def get_epa_house_db_path(path='/var/Indianapolis.db'):
+    return get_dropbox_path()+path
+
+
+def get_asu_house_db_path(path='/var/HillAFB.db'):
+    return get_dropbox_path()+path
+
+print(get_asu_house_db_path())
